### 2021-02-12 chill-yield-vignette.Rmd
First stab with the pasitR tools (we used these to estimates for chill and yield, different topic but similar problem)

---
title: "GHG and milk yield"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

<!-- Print pdf and word versions -->
<!-- rmarkdown::render("chill-yield-vignette.Rmd", output_format = "pdf_document") -->
<!-- rmarkdown::render("chill-yield-vignette.Rmd", output_format = "word_document") -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# devtools::install_github("hortibonn/pasitR")
library(pasitR)

# Add Unit Tests with https://testthat.r-lib.org/
# Install the released version from CRAN
# install.packages("testthat")

#The easiest way to get started is with usethis. Assuming you’re in a package directory, just run usethis::use_test("name") to create a test file, and set up all the other infrastructure you need. If you’re using RStudio, press Cmd/Ctrl + Shift + T (or run devtools::test() if not) to run all the tests in a package..

```

## Load data

```{r load_data}
ghg_variables<- read.csv("MC results/mcSimulationResults.csv")

# select the appropriate variables from the MC outputs names(ghg_variables)

#assuming 'on-farm' is the annual GHG 
chill <- ghg_variables$on_farm
#assuming 'kg_FPCM_year-farm' is the annual milk yield
yield <- ghg_variables$kg_FPCM_year

```

## Scatter plot

Using `chillscatter()` create a chill portions (x) and yield (y) scatter plot with associated and estimated densities with loess smooth linear fits density curves. Plot made with the scatter.hist() function in the plyr() package (Wickham 2011) in R (R Core Team 2017).

```{r, fig.width=7, fig.height=7}
pasitR::chillscatter(chill, yield)
```

## Kernel plot

Use `chillkernel()` or another type of graphical representation of the same relationship. `chillkernel()` performs a two-dimensional kernel density estimation for yield and chill using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002). The function produces a matrix of the estimated density (z) of yield (y) and chill (x). As the density function restricts the shape of the kernel to a bivariate normal kernel, it looks slightly different compared to the scatter plot estimates above.

```{r, fig.width=7, fig.height=7}
pasitR::chillkernel(chill, yield)
```

`chillkernel()` shows a density surface plot of chill portions (x) and yield (y). The legend shows the value for the estimated density (z). The plot is made with the `filled.contour()` function of the `graphics()` package.

In `chillkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (yield along y-axis) given a specific chill portion (along x-axis).

## Estimated yield given the expected chill

`chillkernelslice()` calculates the estimated yield given the expected chill, based on a slice of 'z' from the Kernel density calcualted with `chillkernel()`. The `expectedchill` parameter is set to 30.

```{r, fig.width=7, fig.height=7}
pasitR::chillkernelslice(chill, yield, expectedchill = 30)
```

`chillkernelslice()` plots the probabilities (shown along the y-axis) for the expected yield (shown along the x-axis). Since this is a cut through the density kernel `chillkernel()`, which integrates to 1, the probability values are relative, not absolute measures.

## Chill portion intervals

`chillviolin()` determines different possible chill portion intervals by calculating the optimal interval width for chill portions using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).
`Optimal interval width for our sample = 2 * interquartile range for our sample / (total number of observations for the interquartile range for our sample)^(1/3)`

```{r, fig.width=7, fig.height=7}
pasitR::chillviolin(chill, yield)
```

`chillviolin()` shows violin plots of chill portions (x) and yield (y) with six different intervals of chill portions. Plot made with `ggplot2()` (Wickham 2016).

## Probability of yield given chill

`chillkernelslicerange()` The chill portion intervals, optimized interquartile ranges shown in `chillviolin()` can be used to select a range to slice from the density kernel `chillkernel()` as was doen for a single chill value in `chillkernelslice()`. Here we set the maximum chill to 20 and the minimum to 10.

```{r, fig.width=7, fig.height=7}

# min_chill	
# is a value of the lowest expected amount of chill for which the yield should be estimated (must be < expectedhighchill).

# 
# max_chill	
# is a value of the highest expected amount of chill for which the yield should be estimated (must be > expectedlowchill).

pasitR::chillkernelslicerange(chill, yield, min_chill = min(chill), max_chill = max(chill))
```

`chillkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected yield (shown along the x-axis). As with `chillkernelslice()` the probability values shown are relative, not absolute measures. They are the result of cuts through the density kernel `chillkernel()`, which integrates to 1.

## References

R Core Team (2017) R: A Language and Environment for Statistical Computing [R Version 3.4.1 (2017-06-30) “Single Candle”] 3.4.1.

Venables, W. N., and B. D. Ripley (2002) Modern Applied Statistics with S. Fourth Edition. New York: Springer.

Wickham, Hadley (2011) Plyr: Tools for Splitting, Applying and Combining Data, Version 1.4. Available Online at Cran. r-Project. Elegant Graphics for Data Analysis. Springer.

### 2021-06-17 210617_GHG_model.Rmd
review of code related to direct emissions from cattle

---
title: "GHG model"
author: "Cory and Eike"
date: "6/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Retracing the equations of the GHG emissions model

We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport, and to possibly optimize the model.

Here's the original model

```{r}

library(decisionSupport)
#CW Note ####
# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
# This project will look in relative file paths (i.e. same or lower folders)
input_table <- "livestock_ghg_input_table.csv"
results_folder <- "MC results"

# Generate a function for testing the model function 'line by line' 
# by taking a single random sample of the provided estimates
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
# run the function on the input_table
make_variables(estimate_read_csv(input_table))

# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

# Enteric fermentation emissions ($CH_4$)
  #CW Note ####
  #in these code sections I am missing the overview. 
  # it is not immediately clear what this is meant to do... 
  # from a specific publication?

# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
# We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.

# We start with the net energy for maintenance (NE_m; in MJ/day)
# This includes a maintenance coefficient, for which different values are recommended depending on animal type
NE_m<-maintenaince_coefficient_Cfi*(live_weight_LW^0.75)

# The energy use for specific activities is calculated by multiplying NE_m by an activity coefficient (C_activity).

NE_activity<-C_activity*NE_m

#Net energy for growth (NE_growth; MJ/day):

NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*(0.96*SWG)^1.097)

# here, C is an animal type coefficient
# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)
# SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.

#This used to be the following equation, which we didn't find a source for:
#NE_growth<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))

#The code doesn't include net energy mobilization by animals that lose weight (p. 314)

#Net energy for lactation

NE_lactation<-milk_yield*(1.47+0.40*milk_fat)

#variables:
#milk_yield (kg milk per day)
#milk_fat (percent)

#Draft animal services are not accounted for

#Energy need during pregnancy 
NE_pregnancy<-0.10*NE_m

#in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.

#We've now covered all equations up to eq 10 (p. 315)

#Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere

#Now we calculate the ratio between net energy consumed and the digestible energy consumed, for maintenance (NE_DE) and growth (NEg_DE) (eqs. 12 and 13)

NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+(1.126*(10^-5)*(dig_energy^2))-(25.4/dig_energy))

NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+(1.308*(10^-5)*(dig_energy^2))-(37.4/dig_energy))

#the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
NE_mobilized<-0
#same for extra energy spent by working animals
NE_w<-0

gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+(NE_growth/NEg_DE))/(dig_energy/100)

enteric_CH4<-(gross_energy_intake*Y_m*365)/55.65

#Y_m here is the methane conversion rate - Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
#This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2=enteric_CH4*25

# Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
volsol<-(gross_energy_intake*(1-(dig_energy/100))+(0.04*gross_energy_intake))*((1-0.08)/18.45) 
mm_ch4<-(volsol*365)*(bo*0.67*((mms_pasture*0.015)+(mms_spread*0.005)+(mms_drylot*0.015)+(mms_solidstore*0.04)+
        (mms_composted*0.005)+(mms_liquid*0.2)+(mms_biogas*0.1)+(mms_burn*0.1)))

# Estimate CO2eq for manure methane
mm_CO2<-mm_ch4*25*365

# Calculate Nitrous oxide emissions from manure and urine on pasture
n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
n_retension<-((milk_yield*(3.3/100))/6.38)
n_excretion<-n_intake*(1-n_retension)*365
mm_direct_n2o<-n_excretion*((mms_pasture*0.02)+(mms_spread*0.005)+(mms_drylot*0.02)+(mms_solidstore*0.005)+
               (mms_composted*0.006)+(mms_liquid*0.005)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28)

# Nitrous oxide (N2O) from volatization 
mm_vol_n2o<-(n_excretion*((mms_pasture*0.02)+(mms_spread*0.07)+(mms_drylot*0.2)+(mms_solidstore*0.3)+
            (mms_composted*0.3)+(mms_liquid*0.4)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28))

# Nitrous oxide (N2O) from leaching 
mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

# Estimate CO2eq for manure nitrous
mm_CO2_N2Od<-mm_direct_n2o*298*365
mm_CO2_N2Ovol<-mm_vol_n2o*298*365
mm_CO2_N2Oleach<-mm_leach_n2o*298*365
mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol+mm_CO2_N2Oleach

# Emissions from feed production and transport. 
feed_CO2<-feed_prod_CO2*feed_trans_CO2

# Total emissions
on_farm<-enteric_CH4+enteric_CO2+mm_CO2+mm_CO2_N2O+feed_CO2

#CW Note ####
# Not clear what this part does (through to the ggplot)

# Relationship between greenhouse gas intensity and milk yields. 
# Fat adjusts
ann_adj <-kg_FPCM_year*365 
df <- data.frame(
  GHGi = on_farm/ann_adj, 
  FPCM = ann_adj)

# plot

#library(ggplot2)
#ggplot(df, aes(x = FPCM, y = GHGi)) +geom_point() + ylim(0,50)

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            #working backwards from 'enteric_CO2' I see that 
            # this is the product of something twice multiplied by 365
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

```

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-06-17 Uncertainty in Kenyan Livestock Emission Reductions.R

@@ -88,8 +88,8 @@ df <- data.frame(
 
 # plot
 
-library(ggplot2)
-ggplot(df, aes(x = FPCM, y = GHGi)) +geom_point() + ylim(0,50) 
+#library(ggplot2)
+#ggplot(df, aes(x = FPCM, y = GHGi)) +geom_point() + ylim(0,50) 
 
 return(list(enteric_CH4=enteric_CH4,
             #CW NOTE ####
@@ -103,10 +103,10 @@ return(list(enteric_CH4=enteric_CH4,
 }
 
 # Run the Monte Carlo using the decisionSupport function ####
-decisionSupport(inputFilePath = input_table, #input file with estimates
+mcResults<-decisionSupport(inputFilePath = input_table, #input file with estimates
                 outputPath = results_folder, #output folder
                 welfareFunction = ghg_emissions, #the function created above
                 numberOfModelRuns = 100, #make 100 for now
                 functionSyntax = "plainNames", 
                 write_table = TRUE,)
-
+plot(mcResults)

### 2021-07-13 210617_GHG_model.Rmd
edits 13.07.21

@@
 # Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
-volsol<-(gross_energy_intake*(1-(dig_energy/100))+(0.04*gross_energy_intake))*((1-0.08)/18.45) 
-mm_ch4<-(volsol*365)*(bo*0.67*((mms_pasture*0.015)+(mms_spread*0.005)+(mms_drylot*0.015)+(mms_solidstore*0.04)+
-        (mms_composted*0.005)+(mms_liquid*0.2)+(mms_biogas*0.1)+(mms_burn*0.1)))
+# daily volatile solid excreted (page 10.42 of 2006 IPCC Guidelines for National Greenhouse Gas Inventories)
+
+volsol<-(gross_energy_intake*
+           (1-(dig_energy/100))+
+           (urinary_energy*gross_energy_intake))*
+             ((1-ash_content)/18.45)
+
+# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
+# treat these as constants for each animal type. The following line assigns the corresponding Bo value
+# to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
+# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.
+
+Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))
+
+# the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
+mm_ch4<-(volsol*365)*
+        (Bo*0.67*
+           ((mms_pasture*MCFpasture)+
+            (mms_spread*MCFspread)+
+            (mms_drylot*MCFdrylot)+
+            (mms_solidstore*MCFsolidstore)+
+            (mms_composted*MCFcomposted)+
+            (mms_liquid*MCFliquid)+
+            (mms_biogas*MCFbiogas)+
+            (mms_burn*MCFburn)+
+              # the original code was lacking the 'sold' fraction - we added this for completeness
+            (mms_sold*MCFsold)))
 
 # Estimate CO2eq for manure methane
-mm_CO2<-mm_ch4*25*365
+### this one also contained a multiplication by 365 that seemed erroneous
+mm_CO2<-mm_ch4*25
+
+##### WE STOPPED HERE
 

### 2021-07-13 210617_GHG_model.Rmd
Add references and format coed with notes

@@
 ---
-title: "GHG model"
+title: "Greenhouse gas emissions model"
 author: "Cory and Eike"
-date: "6/17/2021"
+bibliography:
+ - references/packages.bib
+ - references/papers.bib
 output: html_document
@@
 
-## Retracing the equations of the GHG emissions model
+```{r, warning = F, include = F}
+#install and load packages
+library(decisionSupport)
+library(rmarkdown)
 
-We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport, and to possibly optimize the model.
+#Automatically write R package citation entries to a .bib file
+knitr::write_bib(c(.packages(), 
+                   'decisionSupport'), 'references/packages.bib')
 
-Here's the original model
+```
+
+## Generating the GHG emissions model
+
+We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport [@R-decisionSupport], and to possibly optimize the model.
+
+The formulas come from Chapter 10 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories [@dong_ipcc_2006].
+The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
+Many variables are based on the work done in 'Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya' [@wilkes_variation_2020] and in 'Methods and guidance to support MRV of livestock emissions' [@wilkes_methods_2019].
+
+*Here are our changes and additions to the original model:*
 
@@
 library(decisionSupport)
-#CW Note ####
+###CW Note ####
 # remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
-# This project will look in relative file paths (i.e. same or lower folders)
+###CW Note ####
+### # This project will be stored in relative file paths (i.e. same or lower folders)
 input_table <- "livestock_ghg_input_table.csv"
-results_folder <- "MC results"
 
@@
 # Enteric fermentation emissions ($CH_4$)
-  #CW Note ####
+  ####CW Note ####
   #in these code sections I am missing the overview. 
@@
 # This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
-# We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.
+#####EL CW Note ####
+## We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.
 
@@
 # FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)
+
+##### EL CW Note ####
 # SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.
 
+##### EL CW Note #### 
 #This used to be the following equation, which we didn't find a source for:
@@
 
+##### EL CW Note #### 
 #The code doesn't include net energy mobilization by animals that lose weight (p. 314)
@@
 
+##### EL CW Note #### 
 #in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.
@@
 
+##### EL CW Note #### 
 #Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere
@@
 
+##### EL CW Note #### 
 #the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
@@
 
+##### EL CW Note #### 
 #Y_m here is the methane conversion rate - Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)
@@
 # Compute the CO2 equation for enteric methane
-#This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
+##### EL CW Note #### 
+##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
 #This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
@@
 # treat these as constants for each animal type. The following line assigns the corresponding Bo value
-# to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
+##### EL CW Note #### 
+##### # to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
 # animal types 4-6 and 0.135 for animal types 1-3 and 7-12.
@@
 # Estimate CO2eq for manure methane
+### ##### EL CW Note #### 
 ### this one also contained a multiplication by 365 that seemed erroneous
@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+This document was generated using R Markdown [@R-rmarkdown].
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+# References

### 2021-07-13 BN_attempt.R
remove BN attempt (no longer valid)

@@ -1,17 +1,0 @@
-#test of the hill climbing 'greedy' bn with bnlearn
-#Tsamardinos, Ioannis, Laura E. Brown, and Constantin F. Aliferis. “The max-min hill-climbing Bayesian network structure learning algorithm.” Machine learning 65.1 (2006): 31-78.
-
-library(bnlearn)
-
-ghg_variables<- read.csv("Kenya_baseline_individual_cow_wNotes.csv")
-
-# all functions in bnlearn require
-# complete data so that error message is expected. However, you can
-# estimate the CPTs from incomplete data using table() and prop.table()
-# and assemble them in a fitted BN with custom.fit(). On the other hand,
-# maybe it would be better to write an EM wrapper around bn.fit() to
-# make the best of the dependence structure of the data?
-table(ghg_variables)
-prop.table(ghg_variables)
-# overcome the Error in check.data(x) : 
-# the data set contains NaN/NA values.

### 2021-07-13 210617_GHG_model.Rmd
clean and add citaitons/ uncertainty for Ch4 and milk calculations (placeholders for now)

@@
 library(decisionSupport)
+library(ggplot2)
+library(plyr)
 library(rmarkdown)
+# devtools::install_github("CWWhitney/uncertainty")
+library(uncertainty)
 
@@
 
-```{r}
+```{r load_data}
 
@@
 ### # This project will be stored in relative file paths (i.e. same or lower folders)
-input_table <- "livestock_ghg_input_table.csv"
+input_table <- "data/livestock_ghg_input_table.csv"
+```
 
+```{r make_variables}
 # Generate a function for testing the model function 'line by line' 
@@
 for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
+
 # run the function on the input_table
 make_variables(estimate_read_csv(input_table))
+```
 
+```{r ghg_emissions_function}
 # Generate the GHG emissions function 
@@
   ####CW Note ####
-  #in these code sections I am missing the overview. 
-  # it is not immediately clear what this is meant to do... 
-  # from a specific publication?
+  #####in these code sections I am missing the overview. 
+  ##### it is not immediately clear what this is meant to do... 
+  ##### from a specific publication?
 
@@
 #Net energy for growth (NE_growth; MJ/day):
+# here, C is an animal type coefficient
+# ### CW Note #### Assuming 'C' is C_activity or should it have been animal_type
+C <- C_activity
+# C <- animal_type
+NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*
+                   (0.96*SWG)^1.097)
 
-NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*(0.96*SWG)^1.097)
-
-# here, C is an animal type coefficient
 # FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)
@@
 
-NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+(1.126*(10^-5)*(dig_energy^2))-(25.4/dig_energy))
+NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+
+            (1.126*(10^-5)*(dig_energy^2))-
+              (25.4/dig_energy))
 
-NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+(1.308*(10^-5)*(dig_energy^2))-(37.4/dig_energy))
+NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+
+            (1.308*(10^-5)*(dig_energy^2))-
+              (37.4/dig_energy))
 
@@
 
-gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+(NE_growth/NEg_DE))/(dig_energy/100)
+gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+
+                          (NE_growth/NEg_DE))/(dig_energy/100)
 
@@
 ##### EL CW Note #### 
-#Y_m here is the methane conversion rate - Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)
+######Y_m  here is the methane conversion rate -
+###### Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)
 
@@
 #This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
-enteric_CO2=enteric_CH4*25
+enteric_CO2<-enteric_CH4*25
 
@@
 
+##### #####
+##### 
 ##### WE STOPPED HERE
+#####
+##### #####
 
 # Calculate Nitrous oxide emissions from manure and urine on pasture
+# ### ##### CW Note #### 
+# ### ##### To get this to run I am assuming a temp 10% pregnancy_coefficient_Cp
+pregnancy_coefficient_Cp <- 0.1
+### #####
+### ####
 n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
@@
 
+# ### ##### CW Note
+# Removing this plot
 # plot
@@
             #CW NOTE ####
-            #working backwards from 'enteric_CO2' I see that 
-            # this is the product of something twice multiplied by 365
+            # ####working backwards from 'enteric_CO2' I see that 
+            # #### this is the product of something twice multiplied by 365
+            # ####### 
             enteric_CO2=enteric_CO2,
@@
 
+Apply the `decisionSupport` tools for running the Monte Carlo [@R-decisionSupport].
+See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.
+
+```{r monte_carlo_simulation}
+
+GHG_simulation <- mcSimulation(
+  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+  model_function = ghg_emissions,
+  numberOfModelRuns = 1e2, #100 for now
+  functionSyntax = "plainNames"
+)
+
+```
+
+```{r plot_distributions}
+
+plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "enteric_CH4",
+                   method = "boxplot_density",
+                   old_names = "enteric_CH4",
+                   new_names = "Methane Emissions from Enteric Fermentation")
+
+```
+
+# uncertainty
+
+Here we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`.
+
+Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest {in_var} (x) and an outcome variable {out_var} (y) with associated and estimated densities with loess smooth linear fits density curves.
+Plot made with the scatter.hist() function in the plyr() package [@R-plyr] in R [@R-base].
+
+Check the `milk_yield` effect on `enteric_CH4`.
+
+```{r uncertainty_package}
+uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4)
+```
+
+Use `varkernel()` for another type of graphical representation of the same relationship.
+`varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable {out_var} and the influence variable of interest {in_var} using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002).
+The function produces a matrix of the estimated density (z) of the expected value of the outcome variable {out_var} (y) and the influence variable of interest {in_var} (x).
+It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a bivariate normal kernel.
+
+Check the effect of {in_var} `milk_yield` on the {out_var} `enteric_CH4`.
+
+```{r name_in_out_vars}
+in_var  <- GHG_simulation$x$milk_yield
+out_var <- GHG_simulation$y$enteric_CH4
+```
+
+```{r varkernel}
+uncertainty::varkernel(in_var = in_var, out_var = out_var)
+```
+
+`varkernel()` shows a density surface plot of a given influence variable of interest {in_var} (x) and the outcome variable {out_var} (y).
+The legend shows the value for the estimated density (z).
+The plot is made with the `filled.contour()` function of the `graphics()` package.
+
+In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable {out_var} along y-axis) given a specific value for a given influence variable of interest {in_var} (along x-axis).
+
+## Estimated outcome variable {out_var} given the expected variable
+
+`varkernelslice()` calculates the estimated outcome variable {out_var} given the expected value of the influence variable of interest {in_var}, based on a slice of 'z' from the Kernel density calculated with `varkernel()`.
+The `expectedvariable` parameter is set to 30.
+
+```{r varkernelslice}
+uncertainty::varkernelslice(in_var = in_var, out_var = out_var, expectedin_var = 30)
+```
+
+`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis).
+Since this is a cut through the density kernel `varkernel()`, which integrates to 1, the probability values are relative, not absolute measures.
+
+## Influencing variable intervals
+
+`varviolin()` determines different possible influencing variable intervals {in_var} by calculating the optimal interval width for the given variable values.
+This is done using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).
+
+$Os = 2 * IQRs / (obs_IQRs)^{(1/3)}$
+
+Where $Os$ is The optimal interval width for our sample, $IQRs$ is the interquartile range for our sample and $obs_IQRs$ is the total number of observations for the interquartile range for our sample.
+
+```{r varviolin}
+uncertainty::varviolin(in_var = in_var, out_var = out_var)
+```
+
+`varviolin()` shows violin plots of influencing variable {in_var} values (x) and outcome variable {out_var} (y) values, with six different intervals of influencing variable {in_var} values.
+Plot made with `ggplot2()` [@R-ggplot2].
+
+## Probability of outcome variable given variable
+
+`varkernelslicerange()` shows the influencing variable {in_var} value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`.
+Here we set the maximum variable value to 4 and the minimum to 1.
+
+```{r varkernelslicerange}
+uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
+                                 min_in_var = 1, max_in_var = 4)
+```
+
+`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis).
+As with `varkernelslice()` the probability values shown are relative, not absolute measures.
+They are the result of cuts through the density kernel `varkernel()`, which integrates to 1.
+
 This document was generated using R Markdown [@R-rmarkdown].

### 2021-07-13 210617_GHG_model.Rmd
uodate the variable names

@@
 
-Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest {in_var} (x) and an outcome variable {out_var} (y) with associated and estimated densities with loess smooth linear fits density curves.
+Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest `in_var` (x) and an outcome variable `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves.
 Plot made with the scatter.hist() function in the plyr() package [@R-plyr] in R [@R-base].
@@
 Use `varkernel()` for another type of graphical representation of the same relationship.
-`varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable {out_var} and the influence variable of interest {in_var} using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002).
-The function produces a matrix of the estimated density (z) of the expected value of the outcome variable {out_var} (y) and the influence variable of interest {in_var} (x).
+`varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable `out_var` and the influence variable of interest `in_var` using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002).
+The function produces a matrix of the estimated density (z) of the expected value of the outcome variable `out_var` (y) and the influence variable of interest `in_var` (x).
 It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a bivariate normal kernel.
 
-Check the effect of {in_var} `milk_yield` on the {out_var} `enteric_CH4`.
+Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4`.
 
@@
 
-`varkernel()` shows a density surface plot of a given influence variable of interest {in_var} (x) and the outcome variable {out_var} (y).
+`varkernel()` shows a density surface plot of a given influence variable of interest `in_var` (x) and the outcome variable `out_var` (y).
 The legend shows the value for the estimated density (z).
@@
 
-In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable {out_var} along y-axis) given a specific value for a given influence variable of interest {in_var} (along x-axis).
+In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable `out_var` along y-axis) given a specific value for a given influence variable of interest `in_var` (along x-axis).
 
-## Estimated outcome variable {out_var} given the expected variable
+## Estimated outcome variable `out_var` given the expected variable
 
-`varkernelslice()` calculates the estimated outcome variable {out_var} given the expected value of the influence variable of interest {in_var}, based on a slice of 'z' from the Kernel density calculated with `varkernel()`.
+`varkernelslice()` calculates the estimated outcome variable `out_var` given the expected value of the influence variable of interest `in_var`, based on a slice of 'z' from the Kernel density calculated with `varkernel()`.
 The `expectedvariable` parameter is set to 30.
@@
 
-`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis).
+`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable `out_var` (shown along the x-axis).
 Since this is a cut through the density kernel `varkernel()`, which integrates to 1, the probability values are relative, not absolute measures.
@@
 
-`varviolin()` determines different possible influencing variable intervals {in_var} by calculating the optimal interval width for the given variable values.
+`varviolin()` determines different possible influencing variable intervals `in_var` by calculating the optimal interval width for the given variable values.
 This is done using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).
@@
 
-`varviolin()` shows violin plots of influencing variable {in_var} values (x) and outcome variable {out_var} (y) values, with six different intervals of influencing variable {in_var} values.
+`varviolin()` shows violin plots of influencing variable `in_var` values (x) and outcome variable `out_var` (y) values, with six different intervals of influencing variable `in_var` values.
 Plot made with `ggplot2()` [@R-ggplot2].
@@
 
-`varkernelslicerange()` shows the influencing variable {in_var} value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`.
-Here we set the maximum variable value to 4 and the minimum to 1.
+`varkernelslicerange()` shows the influencing variable `in_var` value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`.
+Here we set the maximum expected value for `in_var` to 4 and the minimum expected value to 1.
 
@@
 
-`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis).
+`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable `out_var` (shown along the x-axis).
 As with `varkernelslice()` the probability values shown are relative, not absolute measures.

### 2021-07-13 210617_GHG_model.Rmd
update references

@@
 
-The formulas come from Chapter 10 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories [@dong_ipcc_2006].
+The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006].
 The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
-Many variables are based on the work done in 'Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya' [@wilkes_variation_2020] and in 'Methods and guidance to support MRV of livestock emissions' [@wilkes_methods_2019].
+The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021] , also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
+Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].
 
-*Here are our changes and additions to the original model:*
+**Here are our changes and additions to the original model:**
 
@@
 ###CW Note ####
-# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
+# #### remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' folder locations throughout - 
+#### these are not useful with RMD and RStudio / Git integrations
 ###CW Note ####
@@
 
+Here we generate a `make_variables` function for testing the model function 'line by line'. We use it to take a single random sample of the provided estimates.
+
 ```{r make_variables}
-# Generate a function for testing the model function 'line by line' 
-# by taking a single random sample of the provided estimates
+###CW Note ####
+### We will need to clean up the warnings later
+### Reasonable ranges still missing for many variables
 make_variables <- function(est,n=1){ x <- random(rho=est, n=n)

### 2021-08-19 210617_GHG_model.Rmd
first running model (needs quality checks)

@@
 
+# decisions on manure management
+# the management options we're considering here are the ones from the Excel table. The IPCC guidelines
+# list more and partly different ones.
+
+# We're deciding on manure management according to a decision tree (Cory will build a figure):
+
+# question 1: how much of the manure is centrally collected? share_centrally
+# question 2: of manure that's not centrally collected, how much is burned? share_central_burn
+mms_burn<-(1-share_centrally)*share_central_burn
+mms_pasture<-(1-share_centrally)*(1-share_central_burn)
+
+# question 3: of centrally collected manure, how much is spread daily on cropland or pasture? share_spread
+mms_spread<-share_centrally*share_spread
+not_spread<-share_centrally*(1-share_spread)
+# question 4: how much of the manure that's not spread is sold? share_sold
+mms_sold<-not_spread*share_sold
+not_sold<-not_spread*(1-share_sold)
+# question 5: how much of the stored manure that's not sold is stored in liquid form? share_liquid
+liquid<-not_sold*share_liquid
+solid<-not_sold*(1-share_liquid)
+# question 6: how much of the liquid manure is converted to biogas? share_biogas
+mms_biogas<-liquid*share_biogas
+mms_liquid_slurry<-liquid*(1-share_biogas)
+# question 7: how much of the solid manure is composted? share_compost
+mms_composted<-solid*share_compost
+remain_solid<-solid*(1-share_compost)
+# question 8: how much of the solid manure is stored in dry lots? share_drylot
+mms_drylot<-remain_solid*share_drylot
+mms_solidstore<-remain_solid*(1-share_drylot)
+
 # the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
@@
             (mms_composted*MCFcomposted)+
-            (mms_liquid*MCFliquid)+
+            (mms_liquid_slurry*MCFliquid)+
             (mms_biogas*MCFbiogas)+
@@
 
-##### #####
-##### 
-##### WE STOPPED HERE
-#####
-##### #####
+# Calculate Nitrous oxide emissions from manure and urine on pasture
 
-# Calculate Nitrous oxide emissions from manure and urine on pasture
-# ### ##### CW Note #### 
-# ### ##### To get this to run I am assuming a temp 10% pregnancy_coefficient_Cp
-pregnancy_coefficient_Cp <- 0.1
-### #####
 ### ####
-n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
-n_retension<-((milk_yield*(3.3/100))/6.38)
-n_excretion<-n_intake*(1-n_retension)*365
-mm_direct_n2o<-n_excretion*((mms_pasture*0.02)+(mms_spread*0.005)+(mms_drylot*0.02)+(mms_solidstore*0.005)+
-               (mms_composted*0.006)+(mms_liquid*0.005)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28)
+n_intake<-(gross_energy_intake/18.45)*((perc_crude_protein_diet/100)/6.25) # equation 10.32
 
-# Nitrous oxide (N2O) from volatization 
-mm_vol_n2o<-(n_excretion*((mms_pasture*0.02)+(mms_spread*0.07)+(mms_drylot*0.2)+(mms_solidstore*0.3)+
-            (mms_composted*0.3)+(mms_liquid*0.4)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28))
+# There was some code here on N retention in milk.
+That's apparently from equation 10.33, but it only had the first part of the equation, missing the part on weight gain.
+We decided to go with the alternative way of estimating N retention (figures given in Table 10.20)
+# n_retention<-((milk_yield*(3.3/100))/6.38)
+
+# Table 10.20 contains N retention coefficients:
+# Dairy cows: 0.20 N_retention_dairy_cow
+# Other cattle: 0.07 N_retention_other_cattle
+
+animal_type_dairy_cow=1 ###ASSUMPTION
+
+if(animal_type==animal_type_dairy_cow)
+  n_excretion<-n_intake*(1-N_retention_dairy_cow)*365 else
+    n_excretion<-n_intake*(1-N_retention_other_cattle)*365 #equation 10.31
+
+# Direct nitrous oxide emissions (Eq 10.25 from IPCC guidelines)
+mm_direct_n2o<-n_excretion*((mms_pasture*mndec_pasture)+
+                           (mms_spread*mndec_spread)+
+                           (mms_drylot*mndec_drylot)+
+                           (mms_solidstore*mndec_solidstore)+
+                           (mms_composted*mndec_composted)+
+                           (mms_liquid*mndec_liquid)+
+                           (mms_biogas*mndec_biogas)+
+                           (mms_burn*mndec_burn)+
+                           (mms_sold*mndec_sold))*(44/28)
+
+# Nitrous oxide (N2O) from volatization (Eq 10.26 from IPCC guidelines)
+mm_vol_n2o<-n_excretion*((mms_pasture*mnvc_pasture)+
+                           (mms_spread*mnvc_spread)+
+                           (mms_drylot*mnvc_drylot)+
+                           (mms_solidstore*mnvc_solidstore)+
+                           (mms_composted*mnvc_composted)+
+                           (mms_liquid*mnvc_liquid)+
+                           (mms_biogas*mnvc_biogas)+
+                           (mms_burn*mnvc_burn)+
+                           (mms_sold*mnvc_sold))*(44/28)
 
 # Nitrous oxide (N2O) from leaching
-mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
-              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)
+# comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
+#mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
+#              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)
 
 # Estimate CO2eq for manure nitrous
-mm_CO2_N2Od<-mm_direct_n2o*298*365
+mm_CO2_N2Od<-mm_direct_n2o*298*365 # (298 is the CO2-equivalent of N2O)
 mm_CO2_N2Ovol<-mm_vol_n2o*298*365
-mm_CO2_N2Oleach<-mm_leach_n2o*298*365
-mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol+mm_CO2_N2Oleach
+# mm_CO2_N2Oleach<-mm_leach_n2o*298*365
+mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol #+mm_CO2_N2Oleach
 
 # Emissions from feed production and transport. 
+# comment: the numbers in the table seem spuriously precise.
 feed_CO2<-feed_prod_CO2*feed_trans_CO2
@@
 # Fat adjusts
-ann_adj <-kg_FPCM_year*365 
-df <- data.frame(
-  GHGi = on_farm/ann_adj, 
-  FPCM = ann_adj)
+#ann_adj <-kg_FPCM_year*365 
+#df <- data.frame(
+#  GHGi = on_farm/ann_adj, 
+#  FPCM = ann_adj)
 
@@
             # ####### 
+            milk_yield=milk_yield,
             enteric_CO2=enteric_CO2,

### 2021-08-23 210617_GHG_model.Rmd
should be ok now for individual animals - next up is modeling an entire population. This will probably just require a loop within the function over all the animal types. In the input table, we'd have to specify the number of animals per type. The feeding system may have to stay fixed (i.e. we'll then be doing a farm-scale analysis, assuming that all animals are in the same feeding system).

@@
 
-The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006].
+The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
 The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
@@
 # This includes a maintenance coefficient, for which different values are recommended depending on animal type
-NE_m<-maintenaince_coefficient_Cfi*(live_weight_LW^0.75)
 
-# The energy use for specific activities is calculated by multiplying NE_m by an activity coefficient (C_activity).
+# We're not sure what the animal types stand for, but this is important for calculating emissions.
+# We need to know, which types are dairy cattle, which ones are lactating cows, which ones are bulls, and which ones are 'others'. We have 12 animal types.
+
+animal_types_lact_cow<-c(4,5,6) # lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant.
+This resulted in a weighted mean of the various maintenance coefficients.
+I'm not sure where we could get such data.
+
+# animal type 1 - 'intact' grown-up males >=36 months
+# animal type 2 - castrated grown-up males >=36 months
+# animal type 3 - males 12-36 months
+# animal type 4 - cows
+# animal type 5 - cows
+# animal type 6 - cows
+# animal type 7 - females <= 12 months
+# animal type 8 - males <= 12 months
+# animal type 9 - females 12-36 months
+# animal type 10 - female calves (<12 months)
+# animal type 11 - male calves (<12 months)
+# animal type 12 - first-time pregnant cows
+
+# define C and Cfi based on sex (Page 10.17, below Equation 10.6)
+if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
+if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
+if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)
+
+if(animal_type %in% c(4,5,6)) Cfi<-maintenance_coefficient_Cfi_lact_cow # lactating cows (0.386)
+if(animal_type %in% c(1)) Cfi<-maintenance_coefficient_Cfi_bulls # bulls (0.37)
+if(animal_type %in% c(2,3,7,8,9,10,11,12)) Cfi<-maintenance_coefficient_Cfi_other # others (0.322)
+
+# define mature body weight (assuming this needs to be disaggregated by sex, for Eqation 10.6)
+if(animal_type %in% c(1,2,3,8,11)) mature_weight_MW<-mature_weight_MW_male
+if(animal_type %in% c(4,5,6,7,9,10,12)) mature_weight_MW<-mature_weight_MW_female
+
+# define milk yield
+if(animal_type %in% c(1,2,3,7,8,9,10,11,12)) milk_yield<-0
+if(animal_type %in% c(4,5,6)) milk_yield<-eval(parse(text=paste0("milk_yield_",animal_type)))
+
+#define pregnancy coefficient (full Cp for first-time pregnant, lower for later in most cases)
+if(animal_type==12) Cp<-Cp_0
+if(animal_type==4) Cp<-Cp_0*Cp_rate_4
+if(animal_type==5) Cp<-Cp_0*Cp_rate_5
+if(animal_type==6) Cp<-Cp_0*Cp_rate_6
+if(animal_type %in% c(1,2,3,7,8,9,10,11)) Cp<-0
+
+# define activity coefficient Ca (Table 10.5)
+# Stall 0.00 - Animals are confined to a small area (i.e., tethered, pen, barn) with the result that they expend very little or no energy to acquire feed.
+# Pasture 0.17 - Animals are confined in areas with sufficient forage requiring modest energy expense to acquire feed.
+# Grazing large areas 0.36 - Animals graze in open range land or hilly terrain and expend significant energy to acquire feed.
+
+# I don't quite understand how Ca was assigned in the original data table.
+There are many 0s for 'System 1' (Stall?), but also lots of values between 0 and 0.17 or 0.36.
+I can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table.
+
+if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
+if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
+if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)
+
+# Mean live weight (kg) of animals surely varies by animal type, so we need separate estimates for them
+live_weight_LW<-eval(parse(text=paste0("live_weight_LW_",animal_type)))
+
+# Mean weight gain of animal type (kg/day)
+weight_gain_WG<-eval(parse(text=paste0("weight_gain_WG_",animal_type)))
+
+# Calculate the Net Energy for Maintenance (NE_m) (MJ/day) (Equation 10.3)
+
+NE_m<-Cfi*(live_weight_LW^0.75)
+
+# Calculate the Net Energy for Activity (NE_activity) (MJ/day) (Equation 10.4)
 
@@
 
-#Net energy for growth (NE_growth; MJ/day):
-# here, C is an animal type coefficient
-# ### CW Note #### Assuming 'C' is C_activity or should it have been animal_type
-C <- C_activity
-# C <- animal_type
-NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*
-                   (0.96*SWG)^1.097)
+#Net energy for growth (NE_growth) (MJ/day) (Equation 10.6)
 
-# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)
-
-##### EL CW Note ####
-# SWG is the "Shrunk Weight Gain" in kg per day - we're not sure what exactly this means.
-
-##### EL CW Note #### 
-#This used to be the following equation, which we didn't find a source for:
-#NE_growth<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))
-
-##### EL CW Note #### 
-#The code doesn't include net energy mobilization by animals that lose weight (p. 314)
+NE_growth<-22.02*
+  ((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*
+  (weight_gain_WG^1.097)
 
@@
 #Draft animal services are not accounted for
+#NE_work <- 0.1 * NE_m * hours_per_day
 
-#Energy need during pregnancy 
-NE_pregnancy<-0.10*NE_m
+#Energy need during pregnancy (Equation 10.13)
+NE_pregnancy<-Cp*NE_m
 
-##### EL CW Note #### 
-#in Joshua's equation the 0.10 was a variable called pregnancy_coefficient_Cp with a value around 0.03. We're not sure where this came from.
+# the Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant).
+For all other lactating cows, this is reduced - not clear how this was done.
 
-#We've now covered all equations up to eq 10 (p. 315)
+# Ratio of net energy available in a diet for maintenance to digestible energy consumed (Equation 10.14)
+REM<-1.123-(4.092/1000*DE_perc)+(1.126/100000*DE_perc^2)-(25.4/DE_perc)
 
-##### EL CW Note #### 
-#Equation 11 contains calculation of the metabolizable energy, but that doesn't seem to be used anywhere
+# Ratio of net energy available for growth in a diet to digestable energy consumed (Equation 10.15)
+REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)
 
-#Now we calculate the ratio between net energy consumed and the digestible energy consumed, for maintenance (NE_DE) and growth (NEg_DE) (eqs. 12 and 13)
+# Gross Energy (GE) MJ/day
+NE_work<-0 # work animals aren't covered here, so we set this to zero
+GE<-(((NE_m+NE_activity+NE_lactation+NE_work+NE_pregnancy)/REM)+(NE_growth/REG))/(DE_perc/100)
 
-NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+
-            (1.126*(10^-5)*(dig_energy^2))-
-              (25.4/dig_energy))
+# CH4 emission factors for enteric fermentation from a livestock category (Equation 10.21)
 
-NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+
-            (1.308*(10^-5)*(dig_energy^2))-
-              (37.4/dig_energy))
+enteric_CH4<-(GE*Y_m*365)/55.65 # kg CH4 per animal per year
 
-##### EL CW Note #### 
-#the following equation usually contains the energy mobilized by animals losing weight. This wasn't calculated, so we set this to 0
-NE_mobilized<-0
-#same for extra energy spent by working animals
-NE_w<-0
-
-gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+
-                          (NE_growth/NEg_DE))/(dig_energy/100)
-
-enteric_CH4<-(gross_energy_intake*Y_m*365)/55.65
-
-##### EL CW Note #### 
-######Y_m  here is the methane conversion rate -
-###### Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)
-
-# Compute the CO2 equation for enteric methane
-##### EL CW Note #### 
-##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
+# Compute the CO2-equivalent of methane emissions for enteric methane
 #This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
-enteric_CO2<-enteric_CH4*25
+enteric_CH4_CO2eq<-enteric_CH4*25
 
 # Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
-# daily volatile solid excreted (page 10.42 of 2006 IPCC Guidelines for National Greenhouse Gas Inventories)
+# daily volatile solid excreted (Equation 10.24)
 
-volsol<-(gross_energy_intake*
-           (1-(dig_energy/100))+
-           (urinary_energy*gross_energy_intake))*
+volsol<-(GE*
+           (1-(DE_perc/100))+
+           (urinary_energy*GE))*
              ((1-ash_content)/18.45)
@@
 
-# the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
+# CH4 emission factor from manure management (Equation 10.23)
 mm_ch4<-(volsol*365)*
@@
 # Estimate CO2eq for manure methane
-### ##### EL CW Note #### 
-### this one also contained a multiplication by 365 that seemed erroneous
-mm_CO2<-mm_ch4*25
+mm_CH4_CO2eq<-mm_ch4*25
 
@@
 
-### ####
-n_intake<-(gross_energy_intake/18.45)*((perc_crude_protein_diet/100)/6.25) # equation 10.32
+# estimate N intake (kg/animal/day) (Equation 10.32)
+N_intake<-(GE/18.45)*((perc_crude_protein_diet/100)/6.25)
 
-# There was some code here on N retention in milk.
-That's apparently from equation 10.33, but it only had the first part of the equation, missing the part on weight gain.
-We decided to go with the alternative way of estimating N retention (figures given in Table 10.20)
-# n_retention<-((milk_yield*(3.3/100))/6.38)
+# estimate N retention (kg/animal/day) (Equation 10.33)
+N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
+  ((weight_gain_WG*(268-(7.03*NE_growth/weight_gain_WG)))/(1000*6.25))
 
-# Table 10.20 contains N retention coefficients:
-# Dairy cows: 0.20 N_retention_dairy_cow
-# Other cattle: 0.07 N_retention_other_cattle
+# estimate N excretion (kg N/animal/yr) (Equation 10.31)
+N_excretion<-N_intake*(1-N_retention)*365
 
-animal_type_dairy_cow=1 ###ASSUMPTION
-
-if(animal_type==animal_type_dairy_cow)
-  n_excretion<-n_intake*(1-N_retention_dairy_cow)*365 else
-    n_excretion<-n_intake*(1-N_retention_other_cattle)*365 #equation 10.31
-
-# Direct nitrous oxide emissions (Eq 10.25 from IPCC guidelines)
-mm_direct_n2o<-n_excretion*((mms_pasture*mndec_pasture)+
+# Direct nitrous oxide emissions (kg N2O/animal/yr) (Eq 10.25 from IPCC guidelines)
+mm_direct_n2o<-N_excretion*((mms_pasture*mndec_pasture)+
                            (mms_spread*mndec_spread)+
@@
                            (mms_composted*mndec_composted)+
-                           (mms_liquid*mndec_liquid)+
+                           (mms_liquid_slurry*mndec_liquid)+
                            (mms_biogas*mndec_biogas)+
@@
 
-# Nitrous oxide (N2O) from volatization (Eq 10.26 from IPCC guidelines)
-mm_vol_n2o<-n_excretion*((mms_pasture*mnvc_pasture)+
+# N losses from volatilization (kg N/animal/yr) (Eq 10.26 from IPCC guidelines)
+mm_vol_n<-N_excretion*((mms_pasture*mnvc_pasture)+
                            (mms_spread*mnvc_spread)+
@@
                            (mms_composted*mnvc_composted)+
-                           (mms_liquid*mnvc_liquid)+
+                           (mms_liquid_slurry*mnvc_liquid)+
                            (mms_biogas*mnvc_biogas)+
                            (mms_burn*mnvc_burn)+
-                           (mms_sold*mnvc_sold))*(44/28)
+                           (mms_sold*mnvc_sold))
 
-# Nitrous oxide (N2O) from leaching
+# indirect N2O emissions due to volatilization of N from manure management (kg N2O/animal/year) (Equation 10.27)
+mm_vol_n2o<-mm_vol_n*EF4*(44/28)
+
+# Nitrous oxide (N2O) from leaching (kg N/animal/year) (equation 10.28)
+mm_leach_n<-N_excretion*((mms_pasture*frac_leach_pasture)+
+                           (mms_spread*frac_leach_spread)+
+                           (mms_drylot*frac_leach_drylot)+
+                           (mms_solidstore*frac_leach_solidstore)+
+                           (mms_composted*frac_leach_composted)+
+                           (mms_liquid_slurry*frac_leach_liquid)+
+                           (mms_biogas*frac_leach_biogas)+
+                           (mms_burn*frac_leach_burn)+
+                           (mms_sold*frac_leach_sold))
+
+# indirect N2O emissions due to leaching of N from manure management (kg N2O/animal/year) (Equation 10.29)
+mm_leach_n2o<-mm_leach_n*EF5*(44/28)
+
 # comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
@@
 # Estimate CO2eq for manure nitrous
-mm_CO2_N2Od<-mm_direct_n2o*298*365 # (298 is the CO2-equivalent of N2O)
-mm_CO2_N2Ovol<-mm_vol_n2o*298*365
-# mm_CO2_N2Oleach<-mm_leach_n2o*298*365
-mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol #+mm_CO2_N2Oleach
+mm_N2O_CO2eq<-(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)
 
@@
 # Total emissions
-on_farm<-enteric_CH4+enteric_CO2+mm_CO2+mm_CO2_N2O+feed_CO2
+on_farm<-enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq+feed_CO2
 
@@
             milk_yield=milk_yield,
-            enteric_CO2=enteric_CO2,
-            mm_CO2=mm_CO2,
-            mm_CO2_N2O=mm_CO2_N2O,
+            enteric_CH4_CO2eq=enteric_CH4_CO2eq,
+            mm_CH4_CO2eq=mm_CH4_CO2eq,
+            mm_N2O_CO2eq=mm_N2O_CO2eq,
             feed_CO2=feed_CO2,
@@
 ```{r uncertainty_package}
-uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4)
+uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4_CO2eq)
 ```

### 2021-08-23 210617_GHG_model.Rmd
added herd functionality

@@
 ```{r make_variables}
-###CW Note ####
-### We will need to clean up the warnings later
 ### Reasonable ranges still missing for many variables
@@
 # Enteric fermentation emissions ($CH_4$)
-  ####CW Note ####
-  #####in these code sections I am missing the overview. 
-  ##### it is not immediately clear what this is meant to do... 
-  ##### from a specific publication?
 
-# This equation is documented in "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" p. 313
-#####EL CW Note ####
-## We note that these equations are full of formatting mistakes that don't exactly instill confidence that all the numbers are correct.
+# The main source of all this is Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
+The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
 
@@
 
-# We're not sure what the animal types stand for, but this is important for calculating emissions.
-# We need to know, which types are dairy cattle, which ones are lactating cows, which ones are bulls, and which ones are 'others'. We have 12 animal types.
-
-animal_types_lact_cow<-c(4,5,6) # lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant.
-This resulted in a weighted mean of the various maintenance coefficients.
-I'm not sure where we could get such data.
+# We're not sure what the animal types stand for, but we inferred some information from the numbers in the sample dataset.
+We further looked for guidance in the "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" (https://www.ipcc-nggip.iges.or.jp/public/gp/bgp/4_1_CH4_Enteric_Fermentation.pdf)
 
@@
 # animal type 3 - males 12-36 months
-# animal type 4 - cows
-# animal type 5 - cows
-# animal type 6 - cows
-# animal type 7 - females <= 12 months
-# animal type 8 - males <= 12 months
+# animal type 4 - lactating cows
+# animal type 5 - lactating cows
+# animal type 6 - lactating cows
+# animal type 7 - weaned young females <= 12 months
+# animal type 8 - weaned young males <= 12 months
 # animal type 9 - females 12-36 months
-# animal type 10 - female calves (<12 months)
-# animal type 11 - male calves (<12 months)
+# animal type 10 - pre-weaned female calves (<12 months)
+# animal type 11 - pre-weaned male calves (<12 months)
 # animal type 12 - first-time pregnant cows
 
+# animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant.
+This resulted in a weighted mean of the various maintenance coefficients.
+We're not sure where we could get such data.
+
+  mm_N2O_CO2eq<-c()
+  mm_CH4_CO2eq<-c()
+  enteric_CH4_CO2eq<-c()
+  total_milk<-c()
+
+for (animal_type in 1:12)
+{
 # define C and Cfi based on sex (Page 10.17, below Equation 10.6)
@@
 
+N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))
+
 # define activity coefficient Ca (Table 10.5)
@@
 
-# I don't quite understand how Ca was assigned in the original data table.
+# We don't quite understand how Ca was assigned in the original data table.
 There are many 0s for 'System 1' (Stall?), but also lots of values between 0 and 0.17 or 0.36.
-I can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table.
+We can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table.
+For now, we're restricting these Ca values to only those that are given as 'pure' categories in the IPCC guidelines (Table 10.5).
 
@@
 
-#variables:
 #milk_yield (kg milk per day)
@@
 
-# Ratio of net energy available for growth in a diet to digestable energy consumed (Equation 10.15)
+# Ratio of net energy available for growth in a diet to digestible energy consumed (Equation 10.15)
 REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)
@@
 #This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
-enteric_CH4_CO2eq<-enteric_CH4*25
+enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25
 
@@
 # Estimate CO2eq for manure methane
-mm_CH4_CO2eq<-mm_ch4*25
+mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25
 
@@
 # Estimate CO2eq for manure nitrous
-mm_N2O_CO2eq<-(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)
+mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)
+
+total_milk[animal_type]<-N_animal_type*milk_yield*365
+
+}
 
@@
             # ####### 
-            milk_yield=milk_yield,
-            enteric_CH4_CO2eq=enteric_CH4_CO2eq,
-            mm_CH4_CO2eq=mm_CH4_CO2eq,
-            mm_N2O_CO2eq=mm_N2O_CO2eq,
+            milk_yield=sum(total_milk),
+            enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
+            mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
+            mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
             feed_CO2=feed_CO2,
-            on_farm=on_farm))
+            on_farm=sum(on_farm)))
 }
@@
   model_function = ghg_emissions,
-  numberOfModelRuns = 1e2, #100 for now
+  numberOfModelRuns = 1e3, #1000 for now
   functionSyntax = "plainNames"
@@
 plot_distributions(mcSimulation_object = GHG_simulation,
-                   vars = "enteric_CH4",
+                   vars = "enteric_CH4_CO2eq",
                    method = "boxplot_density",
-                   old_names = "enteric_CH4",
+                   old_names = "enteric_CH4_CO2eq",
                    new_names = "Methane Emissions from Enteric Fermentation")
@@
 ```{r uncertainty_package}
-uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4_CO2eq)
+uncertainty::varscatter(in_var = GHG_simulation$y$milk_yield,
+                        out_var = GHG_simulation$y$enteric_CH4_CO2eq)
 ```
@@
 
-Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4`.
+Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4_CO2eq`.
 
 ```{r name_in_out_vars}
-in_var  <- GHG_simulation$x$milk_yield
-out_var <- GHG_simulation$y$enteric_CH4
+in_var  <- GHG_simulation$y$milk_yield
+out_var <- GHG_simulation$y$enteric_CH4_CO2eq
 ```
@@
 uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
-                                 min_in_var = 1, max_in_var = 4)
+                                 min_in_var = round(min(in_var)+(max(in_var)-min(in_var))/4),
+                                 max_in_var = round(min(in_var)+(max(in_var)-min(in_var))/4*3))
 ```

### 2021-08-25 210617_GHG_model.Rmd
attempt at cleaning up the code, added PLS, plot all distributions (all the things in the return), use the uncertainty tools on the total C02 and milk yield.. Still a lot to do.. @eikeluedeling PTAL (probably enough just to open the html)

@@
 knitr::write_bib(c(.packages(), 
-                   'decisionSupport'), 'references/packages.bib')
-
+                   'DiagrammeR',
+                   'decisionSupport',
+                   'MASS'), 'references/packages.bib')
 ```
@@
 
-We're looking at the equation put together by Joshua Wafula to check if this can easily be used by decisionSupport [@R-decisionSupport], and to possibly optimize the model.
-
 The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
 The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
-The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021] , also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
+The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021], also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
 Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].
+We run the model with the decisionSupport package [@R-decisionSupport].
 
-**Here are our changes and additions to the original model:**
+We load the library and the data.
+The data contains ranges of possible values for the simulation.
+Each variable is represented with a confidence interval range and a specification of the distribution shape.
 
 ```{r load_data}
-
+library(DiagrammeR)
 library(decisionSupport)
-###CW Note ####
-# #### remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' folder locations throughout - 
-#### these are not useful with RMD and RStudio / Git integrations
-###CW Note ####
+library(tidyverse)
 ### # This project will be stored in relative file paths (i.e. same or lower folders)
 input_table <- "data/livestock_ghg_input_table.csv"
+```
+
+The manure management is based on a decision tree.
+
+## Decision tree
+
+Decision tree describing the options for manure management.
+Each step in the decision tree (each node) represents a share of manure that is managed according to the name of the node.
+Green arrows indicate 'Yes' and red arrows 'No'.
+The descriptions for the uses can be found in the IPCC report on manure management chapter 10 [@dong_ipcc_2006].
+
+```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1}
+mermaid("
+        graph LR
+        C(Central collection)-->B(Burned); linkStyle 0 stroke: red, stroke-width:1px
+                                B-->P(Pasture); linkStyle 1 stroke: red, stroke-width:1px
+        C-->D(Daily spread); linkStyle 2 stroke: green, stroke-width:1px
+            D-->S(Sold); linkStyle 3 stroke: red, stroke-width:1px
+                S-->L(Liquid storage); linkStyle 4 stroke: red, stroke-width:1px
+                    L-->Co(Composted); linkStyle 5 stroke: red, stroke-width:1px
+                        Co-->Dl(Dry lot); linkStyle 6 stroke: red, stroke-width:1px
+                             Dl-->St(Solid storage); linkStyle 7 stroke: red, stroke-width:1px
+                    L-->Bi(Biogas); linkStyle 8 stroke: green, stroke-width:1px
+                        Bi-->Ls(Liquid slurry); linkStyle 9 stroke: red, stroke-width:1px
+        ")
 ```
@@
 ```{r make_variables}
-### Reasonable ranges still missing for many variables
+
 make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
@@
             (mms_burn*MCFburn)+
+              # CW EL Comment ####
               # the original code was lacking the 'sold' fraction - we added this for completeness
@@
 
-# comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
+# CW EL Comment ####
+# # comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
 #mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
@@
 # Emissions from feed production and transport. 
+# CW EL Comment ####
   # comment: the numbers in the table seem spuriously precise.
+  ####
+
 feed_CO2<-feed_prod_CO2*feed_trans_CO2
@@
 
-#CW Note ####
-# Not clear what this part does (through to the ggplot)
-
-# Relationship between greenhouse gas intensity and milk yields. 
-# Fat adjusts
-#ann_adj <-kg_FPCM_year*365 
-#df <- data.frame(
-#  GHGi = on_farm/ann_adj, 
-#  FPCM = ann_adj)
-
-# ### ##### CW Note
-# Removing this plot
-# plot
-
-#library(ggplot2)
-#ggplot(df, aes(x = FPCM, y = GHGi)) +geom_point() + ylim(0,50)
-
 return(list(enteric_CH4=enteric_CH4,
-            #CW NOTE ####
-            # ####working backwards from 'enteric_CO2' I see that 
-            # #### this is the product of something twice multiplied by 365
-            # ####### 
             milk_yield=sum(total_milk),
@@
 }
-
 ```
@@
 ```{r monte_carlo_simulation}
-
 GHG_simulation <- mcSimulation(
@@
   model_function = ghg_emissions,
-  numberOfModelRuns = 1e3, #1000 for now
+  numberOfModelRuns = 1e2, #100 for now
   functionSyntax = "plainNames"
 )
-
 ```
 
-```{r plot_distributions}
+Here we plot the distributions of the various outputs from the model.
 
-plot_distributions(mcSimulation_object = GHG_simulation,
+```{r plot_distributions, fig.width=10, fig.height=10}
+# enteric_CH4=enteric_CH4,
+enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "enteric_CH4",
+                   method = "boxplot_density",
+                   old_names = "enteric_CH4",
+                   new_names = "Total Methane Emissions from Livestock by Enteric Fermentation") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+# milk_yield=sum(total_milk),
+milk_yield_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "milk_yield",
+                   method = "boxplot_density",
+                   old_names = "milk_yield",
+                   new_names = "Total Milk Yield") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+# enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
+enteric_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
                    vars = "enteric_CH4_CO2eq",
@@
                    old_names = "enteric_CH4_CO2eq",
-                   new_names = "Methane Emissions from Enteric Fermentation")
+                   new_names = "Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock") + 
+  theme(axis.title.x=element_blank())
 
+# mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
+mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "mm_CH4_CO2eq",
+                   method = "boxplot_density",
+                   old_names = "mm_CH4_CO2eq",
+                   new_names = "Carbon Dioxide Equivalent Methane Emissions from Livestock Manure") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+# mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
+mm_N2O_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "mm_N2O_CO2eq",
+                   method = "boxplot_density",
+                   old_names = "mm_N2O_CO2eq",
+                   new_names = "Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+# feed_CO2=feed_CO2,
+feed_CO2_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "feed_CO2",
+                   method = "boxplot_density",
+                   old_names = "feed_CO2",
+                   new_names = "Carbon Dioxide Emissions from Livestock Feed Production and Transport") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+# on_farm=sum(on_farm)
+on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "on_farm",
+                   method = "boxplot_density",
+                   old_names = "on_farm",
+                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
+  theme(axis.title.y=element_blank())
+
+library(patchwork)
+# remove the axis marks from the two lefthand-most plots and the x-axis title from the left and right plots
+
+# plot with patchwork
+layout <- "
+           AAABBB
+           CCCDDD
+           EEEFFF
+           GGGGGG
+          "
+
+enteric_CH4_dist_plot + milk_yield_dist_plot + 
+  enteric_CH4_CO2eq_dist_plot + mm_CH4_CO2eq_dist_plot + 
+  mm_N2O_CO2eq_dist_plot + feed_CO2_dist_plot + 
+  on_farm_dist_plot + 
+plot_layout(guides = "collect", design = layout) 
 ```
 
-# uncertainty
+# Model sensitivity
 
-Here we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`.
+#### Perform a Partial Least Squares Regression (PLSR) of Monte Carlo simulation results
+
+We apply a post-hoc analysis to the `mcSimulation()` outputs with `plsr.mcSimulation()` to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model.
+This function uses the outputs of the `mcSimulation()` selecting all the input variables from the decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`.
+We use the code `names(GHG_simulation$y)[7]` to select the outcome variable `r names(GHG_simulation$y)[7]`, which is an element of the list `y` in our `GHG_simulation` outputs (this must be a character element).
+
+```{r}
+pls_result <- plsr.mcSimulation(object = GHG_simulation,
+                  resultName = names(GHG_simulation$y)[7], ncomp = 1)
+```
+
+We run the `plot_pls()` on the results from `plsr.mcSimulation()` with a number of standard settings.
+The length of the bars is equal to VIP with a vertical line at '1' on the x-axis indicating a standard cut-off for VIP used for variable selection [@Whitney2017].
+The overall plot only shows those variables with a VIP > 0.8, which is the common threshold for variable selection [@Lanzanova2019, @Luedeling2016].
+The colors of the bars represent the positive or negative coefficient of the given input variable with the output variable.
+
+Here we import the input table again to replace the labels for the variables on the y-axis.
+The input table can include a 'label' and 'variable' column.
+The standard labels (from the 'variable' column) are usually computer readable and not very nice for a plot.
+The `plot_pls()` function uses the text in the 'label' column as replacement for the default text in the 'variable' column.
+
+```{r}
+input_table <- read.csv("data/livestock_ghg_input_table.csv")
+
+plot_pls(pls_result, input_table = input_table, threshold = 0)
+```
+
+# Uncertainty
+
+In the following sections we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`.
 
@@
 uncertainty::varscatter(in_var = GHG_simulation$y$milk_yield,
-                        out_var = GHG_simulation$y$enteric_CH4_CO2eq)
+                        out_var = GHG_simulation$y$on_farm)
 ```
@@
 Use `varkernel()` for another type of graphical representation of the same relationship.
-`varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable `out_var` and the influence variable of interest `in_var` using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002).
+`varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable `out_var` and the influence variable of interest `in_var` using the `kde2d()` function in the `MASS()` package [@R-MASS].
 The function produces a matrix of the estimated density (z) of the expected value of the outcome variable `out_var` (y) and the influence variable of interest `in_var` (x).
@@
 
-Check the effect of `in_var` `milk_yield` on the `out_var` `enteric_CH4_CO2eq`.
+Check the effect of `in_var` `milk_yield` on the `out_var` `on_farm`.
 
@@
 in_var  <- GHG_simulation$y$milk_yield
-out_var <- GHG_simulation$y$enteric_CH4_CO2eq
+out_var <- GHG_simulation$y$on_farm
 ```
@@
 
-```{r varviolin}
+```{r varviolin, fig.width=17, fig.height=7}
 uncertainty::varviolin(in_var = in_var, out_var = out_var)
@@
 `varkernelslicerange()` shows the influencing variable `in_var` value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`.
-Here we set the maximum expected value for `in_var` to 4 and the minimum expected value to 1.
+Here we set the maximum expected value for `in_var` to `r (min(in_var)+(max(in_var)-min(in_var))/4*3)` and the minimum expected value to `r (min(in_var)+(max(in_var)-min(in_var))/4)`.
 

### 2021-09-01 210617_GHG_model.Rmd

@@
 title: "Greenhouse gas emissions model"
-author: "Cory and Eike"
+author: "Cory Whitney and Eike Luedeling"
 bibliography:
@@
                    'MASS'), 'references/packages.bib')
+#resolves an issue with the scientific style of numbers,answer from @Paul Hiemstra: https://stackoverflow.com/a/25947542/4061993 
+#from the documentation of ?options:
+options(scipen=999)
 ```
@@
 
-We apply a post-hoc analysis to the `mcSimulation()` outputs with `plsr.mcSimulation()` to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model.
-This function uses the outputs of the `mcSimulation()` selecting all the input variables from the decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`.
-We use the code `names(GHG_simulation$y)[7]` to select the outcome variable `r names(GHG_simulation$y)[7]`, which is an element of the list `y` in our `GHG_simulation` outputs (this must be a character element).
+Below we apply a post-hoc analysis to the `mcSimulation()` outputs with the `plsr.mcSimulation()` function in the `decisoonSupport` package.
+We use the function to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model.
+This function uses the outputs of the `mcSimulation()` selecting all the input variables from the `GHG_simulation` decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`.
+We use the code `names(GHG_simulation$y)[7]` to select the outcome variable `r names(GHG_simulation$y)[7]`, which is an element of the list `y` in our `GHG_simulation` outputs.
 
@@
 
-Here we import the input table again to replace the labels for the variables on the y-axis.
+Here we import the input table again (the same table we used for the Monte Carlo model) to replace the labels for the variables on the y-axis.
 The input table can include a 'label' and 'variable' column.
@@
 
-In the following sections we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`.
+In the following sections we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package.
+Install using `devtools::install_github("CWWhitney/uncertainty")`.
+We apply the following uncertainty function on the `milk_yield` and `on_farm` variables from the `r nrow(GHG_simulation$y)` of `GHG_simulation` model results.
+
+```{r name_in_out_vars}
+in_var  <- GHG_simulation$y$milk_yield
+out_var <- GHG_simulation$y$on_farm
+```
 
 Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest `in_var` (x) and an outcome variable `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves.
-Plot made with the scatter.hist() function in the plyr() package [@R-plyr] in R [@R-base].
+Plot made with the scatter.hist() function in the `plyr()` package [@R-plyr] in R [@R-base].
 
@@
 ```{r uncertainty_package}
-uncertainty::varscatter(in_var = GHG_simulation$y$milk_yield,
-                        out_var = GHG_simulation$y$on_farm)
+uncertainty::varscatter(in_var = in_var,
+                        out_var = out_var)
 ```
@@
 Check the effect of `in_var` `milk_yield` on the `out_var` `on_farm`.
-
-```{r name_in_out_vars}
-in_var  <- GHG_simulation$y$milk_yield
-out_var <- GHG_simulation$y$on_farm
-```
 
@@
 ```{r varviolin, fig.width=17, fig.height=7}
-uncertainty::varviolin(in_var = in_var, out_var = out_var)
+uncertainty::varviolin(in_var = in_var, out_var = out_var, legend.position = "left")
 ```
@@
 `varkernelslicerange()` shows the influencing variable `in_var` value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`.
-Here we set the maximum expected value for `in_var` to `r (min(in_var)+(max(in_var)-min(in_var))/4*3)` and the minimum expected value to `r (min(in_var)+(max(in_var)-min(in_var))/4)`.
+Here we set the maximum expected value for `in_var` to `r round(max(in_var), digits = 0)` and the minimum expected value to `r round(min(in_var), digits = 0)`.
 
@@
 uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
-                                 min_in_var = round(min(in_var)+(max(in_var)-min(in_var))/4),
-                                 max_in_var = round(min(in_var)+(max(in_var)-min(in_var))/4*3))
+                                 min_in_var = min(in_var),
+                                 max_in_var = max(in_var))
 ```

### 2021-09-01 210617_GHG_model.Rmd
review edits

@@
 input_table <- "data/livestock_ghg_input_table.csv"
+
+options(knitr.kable.NA = '')
+read.csv(input_table)[,c(1:2,4:6)] %>%
+  kableExtra::kbl(digits =2) %>%
+  kableExtra::kable_classic_2(full_width = F)
+
 ```
@@
 
-## Decision tree
+## Decision tree for manure management
 
-Decision tree describing the options for manure management.
+It is difficult to know with precision how much manure is managed according to each of the management categories.
+Capturing the corresponding uncertainty is difficult, because estimating each fraction separately is unlikely to add up to 100%.
+We address this problem with a decision tree.
+
 Each step in the decision tree (each node) represents a share of manure that is managed according to the name of the node.
@@
 
-Here we generate a `make_variables` function for testing the model function 'line by line'. We use it to take a single random sample of the provided estimates.
+## Making variables for model development
+
+Here we generate a `make_variables` function for testing the model function 'line by line'.
+We use it to take a single random sample of the provided estimates (this isn't required for running the model later, but it helps when developing the code).
 
@@
 ```
+
+## The model
+
+The following is our draft of the emissions model, with explanations of all steps (we can format this better once we agree on the final model).
 
@@
 
-# We're not sure what the animal types stand for, but we inferred some information from the numbers in the sample dataset.
-We further looked for guidance in the "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" (https://www.ipcc-nggip.iges.or.jp/public/gp/bgp/4_1_CH4_Enteric_Fermentation.pdf)
+# The following animal types were used in the example:
 
@@
 We're not sure where we could get such data.
+In the original dataset, types 4, 5 and 6 referred to different pregnancy/lactation states, but it doesn't seem like this scheme was rigorously followed.
 
+# We make placeholder vectors to collect emissions data for all animal types
 mm_N2O_CO2eq<-c()
@@
 total_milk<-c()
+
+# Now we start a long loop, where emissions are calculated for each animal type
 
@@
 
-#Net energy for lactation
+# Net energy for lactation
 
@@
 
-#milk_yield (kg milk per day)
-#milk_fat (percent)
-
-#Draft animal services are not accounted for
-#NE_work <- 0.1 * NE_m * hours_per_day
+# Draft animal services are not accounted for
+# NE_work <- 0.1 * NE_m * hours_per_day
 
@@
 
-# the Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant).
+# The Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant).
 For all other lactating cows, this is reduced - not clear how this was done.
@@
 
-# We're deciding on manure management according to a decision tree (Cory will build a figure):
+# We're deciding on manure management according to a decision tree (see figure above):
 
@@
 
-Apply the `decisionSupport` tools for running the Monte Carlo [@R-decisionSupport].
+## Running the model
+
+We apply the `decisionSupport` tools for running the Monte Carlo simulation [@R-decisionSupport].
 See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.
@@
 
-# Model sensitivity
+## Model sensitivity
 
@@
 Install using `devtools::install_github("CWWhitney/uncertainty")`.
-We apply the following uncertainty function on the `milk_yield` and `on_farm` variables from the `r nrow(GHG_simulation$y)` of `GHG_simulation` model results.
+We apply the following uncertainty function on the `milk_yield` and `on_farm` variables from the `r nrow(GHG_simulation$y)` `GHG_simulation` model results.
 
@@
 
-Check the `milk_yield` effect on `enteric_CH4`.
+Check the `milk_yield` (x-axis) effect on `on_farm` emissions (y-axis).
 
@@
 
-## Probability of outcome variable given variable
+## Probability of outcome variable for a given range of the input variable
 
@@
 uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
-                                 min_in_var = min(in_var),
-                                 max_in_var = max(in_var))
+                                 min_in_var = round(min(in_var)),
+                                 max_in_var = round(max(in_var)))
 ```

### 2021-09-02 210617_GHG_model.Rmd
cleaning up the uncertainty sections. Stil trying to gt to that golden simple end point of clear descriptions of GHG given milk yields (nearly there)

@@
 #install and load packages
+library(DiagrammeR)
 library(decisionSupport)
@@
 library(decisionSupport)
-library(tidyverse)
+library(plyr)
+library(magrittr)
+library(kableExtra)
 ### # This project will be stored in relative file paths (i.e. same or lower folders)
+
  input_table <- "data/livestock_ghg_input_table.csv"
 
-options(knitr.kable.NA = '')
+options(knitr.kable.NA = '', encoding = "Windows-1252")
  read.csv(input_table)[,c(1:2,4:6)]   %>%
-  kableExtra::kbl(digits =2) %>%
+    kableExtra::kbl(digits = 2)  %>%
     kableExtra::kable_classic_2(full_width = F)
@@
 Here we import the input table again (the same table we used for the Monte Carlo model) to replace the labels for the variables on the y-axis.
-The input table can include a 'label' and 'variable' column.
+The input table includes a 'label' and 'variable' column.
 The standard labels (from the 'variable' column) are usually computer readable and not very nice for a plot.
@@
 
-In the following sections we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package.
-Install using `devtools::install_github("CWWhitney/uncertainty")`.
-We apply the following uncertainty function on the `milk_yield` and `on_farm` variables from the `r nrow(GHG_simulation$y)` `GHG_simulation` model results.
+We apply a collection of functions for use in assessing uncertainty in the relationship between GHG emissions and milk yields in livestock production.
+We use functions from the `uncertainty` package.
+Install the package using `devtools::install_github("CWWhitney/uncertainty")`.
+
+We apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
+These come from the model results (`r nrow(GHG_simulation$y)`) of the `GHG_simulation` model results to the `in_var` (x) and an outcome variable `out_var` (y).
+in the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
 
@@
 
-Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest `in_var` (x) and an outcome variable `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves.
+## Milk_yield (x-axis) effect on on farm GHG emissions (y-axis)
+
+Here we use the `varscatter()` function to create a scatter plot of milk yields `in_var` (x) and GHG emissions `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves.
 Plot made with the scatter.hist() function in the `plyr()` package [@R-plyr] in R [@R-base].
-
-Check the `milk_yield` (x-axis) effect on `on_farm` emissions (y-axis).
+The result is a scatter plot with associated and estimated densities (using a loess smooth) given by the red dot (mean) and red ellipses (1 and 2 sigma from mean).
+The red line going across the plot shows the linear fit.
+Histograms are shown with smooth lines (loess smooth linear fits) density curves.
+The numeric value in the upper right gives the Spearman correlation coefficient between the influencing variable and the outcome variable.
 
@@
 
-Use `varkernel()` for another type of graphical representation of the same relationship.
-`varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable `out_var` and the influence variable of interest `in_var` using the `kde2d()` function in the `MASS()` package [@R-MASS].
-The function produces a matrix of the estimated density (z) of the expected value of the outcome variable `out_var` (y) and the influence variable of interest `in_var` (x).
-It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a bivariate normal kernel.
-
-Check the effect of `in_var` `milk_yield` on the `out_var` `on_farm`.
+To verify the finding we use the `varkernel()` function for another type of graphical representation of the same relationship.
+We use the `varkernel()` function to perform a two-dimensional kernel density estimation for the expected value of GHG emissions and milk yield.
+The core of the process involves kernel density estimation.
+This is a nonparametric technique for estimating probability density functions (similar to histogram density estimation but with more dimensions).
+We apply this with the `kde2d()` function in the `MASS()` package [@R-MASS].
+The function produces a matrix of the estimated density (z) of the expected value of the GHG emissions and milk yield.
+It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a normal distribution (a bivariate normal kernel).
 
@@
 
-`varkernel()` shows a density surface plot of a given influence variable of interest `in_var` (x) and the outcome variable `out_var` (y).
+The result of the `varkernel()` function shows a density surface plot of GHG emissions `in_var` (x) and milk yield `out_var` (y).
+In these `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the GHG emissions `out_var` along the y-axis) given a specific value for milk yield `in_var` (along the x-axis).
 The legend shows the value for the estimated density (z).
-The plot is made with the `filled.contour()` function of the `graphics()` package.
-
-In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable `out_var` along y-axis) given a specific value for a given influence variable of interest `in_var` (along x-axis).
-
-## Estimated outcome variable `out_var` given the expected variable
-
-`varkernelslice()` calculates the estimated outcome variable `out_var` given the expected value of the influence variable of interest `in_var`, based on a slice of 'z' from the Kernel density calculated with `varkernel()`.
-The `expectedvariable` parameter is set to 30.
-
-```{r varkernelslice}
-uncertainty::varkernelslice(in_var = in_var, out_var = out_var, expectedin_var = 30)
-```
-
-`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable `out_var` (shown along the x-axis).
-Since this is a cut through the density kernel `varkernel()`, which integrates to 1, the probability values are relative, not absolute measures.
+The plot is made with the `filled.contour()` function of the `graphics()` package [@R-base].
 
@@
 
-`varviolin()` determines different possible influencing variable intervals `in_var` by calculating the optimal interval width for the given variable values.
-This is done using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).
+Here we use the `varviolin()` function to determine different milk yield variable intervals by calculating the optimal interval width.
+This is done using the interquartile range `IQR()` function in the `stats()` package.
+We follow the Freedman-Diaconis rule to select the width of the bins.
 
@@
 
-`varviolin()` shows violin plots of influencing variable `in_var` values (x) and outcome variable `out_var` (y) values, with six different intervals of influencing variable `in_var` values.
-Plot made with `ggplot2()` [@R-ggplot2].
+The result of the `varviolin()` function shows violin plots of milk yield `in_var` values (x) and our outcome variable GHG emissions `out_var` (y) values.
+The plot was made with `ggplot2()` [@R-ggplot2].
 

### 2021-09-03 210617_GHG_model.Rmd

@@
 
-## Influencing variable intervals
+## Probability of GHG emissions for a given range of milk yield
 
-Here we use the `varviolin()` function to determine different milk yield variable intervals by calculating the optimal interval width.
-This is done using the interquartile range `IQR()` function in the `stats()` package.
-We follow the Freedman-Diaconis rule to select the width of the bins.
-
-$Os = 2 * IQRs / (obs_IQRs)^{(1/3)}$
-
-Where $Os$ is The optimal interval width for our sample, $IQRs$ is the interquartile range for our sample and $obs_IQRs$ is the total number of observations for the interquartile range for our sample.
-
-```{r varviolin, fig.width=17, fig.height=7}
-uncertainty::varviolin(in_var = in_var, out_var = out_var, legend.position = "left")
-```
-
-The result of the `varviolin()` function shows violin plots of milk yield `in_var` values (x) and our outcome variable GHG emissions `out_var` (y) values.
-The plot was made with `ggplot2()` [@R-ggplot2].
-
-## Probability of outcome variable for a given range of the input variable
-
-`varkernelslicerange()` shows the influencing variable `in_var` value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`.
-Here we set the maximum expected value for `in_var` to `r round(max(in_var), digits = 0)` and the minimum expected value to `r round(min(in_var), digits = 0)`.
+We use the `varkernelslicerange()` function to show the milk yield (influencing variable `in_var` value) intervals.
+Here we choose a very wide distribution (the whole range of milk yields from the Monte Carlo simulation).
+Here we set the maximum expected value for milk yield `in_var` to `r round(max(in_var), digits = 0)` and the minimum expected value to `r round(min(in_var), digits = 0)`.
 
@@
 
+## Influencing variable intervals
+
+For a more specific estimate we can choose from those optimized interquartile ranges to select a range to slice from the density kernel `varkernel()`.
+we do that here using the `varviolin()` function to determine meaningful milk yield variable intervals and how they interact with the outcome variable GHG emissions.
+We do this by calculating an 'optimal' interval width using the interquartile range `IQR()` function in the `stats()` package.
+
+```{r varviolin, fig.width=17, fig.height=7}
+uncertainty::varviolin(in_var = in_var, out_var = out_var, legend.position = "left")
+```
+
+The result of the `varviolin()` function shows violin plots of milk yield `in_var` values (x) and our outcome variable GHG emissions `out_var` (y) values.
+The plot was made with `ggplot2()` [@R-ggplot2].
+These offer a possible selection of milk yields and show the resulting expected GHG emissions.
+
 This document was generated using R Markdown [@R-rmarkdown].

### 2021-09-06 210617_GHG_model.Rmd
Draft for discussion with Todd and Andy

@@
   model_function = ghg_emissions,
-  numberOfModelRuns = 1e2, #100 for now
+  numberOfModelRuns = 1e4, # 10,000 runs
   functionSyntax = "plainNames"
@@
 
-## Milk_yield (x-axis) effect on on farm GHG emissions (y-axis)
+## Milk_yield (x-axis) effect on on-farm GHG emissions (y-axis)
 

### 2021-10-03 211003_scenarios.Rmd
scenario calculations

--- 2021-09-06 210617_GHG_model.Rmd
@@
 ---
-title: "Greenhouse gas emissions model"
+title: "Greenhouse gas emissions model - making scenarios"
 author: "Cory Whitney and Eike Luedeling"
@@
 ---
+
+We're defining the function as in 210617_GHG_model.Rmd, but there's no need to show this again here.
 
@@
 
-```{r load_data}
+```{r load_data, include=FALSE}
 library(DiagrammeR)
@@
 
-The manure management is based on a decision tree.
-
-## Decision tree for manure management
-
-It is difficult to know with precision how much manure is managed according to each of the management categories.
-Capturing the corresponding uncertainty is difficult, because estimating each fraction separately is unlikely to add up to 100%.
-We address this problem with a decision tree.
-
-Each step in the decision tree (each node) represents a share of manure that is managed according to the name of the node.
-Green arrows indicate 'Yes' and red arrows 'No'.
-The descriptions for the uses can be found in the IPCC report on manure management chapter 10 [@dong_ipcc_2006].
-
-```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1}
-mermaid("
-        graph LR
-        C(Central collection)-->B(Burned); linkStyle 0 stroke: red, stroke-width:1px
-                                B-->P(Pasture); linkStyle 1 stroke: red, stroke-width:1px
-        C-->D(Daily spread); linkStyle 2 stroke: green, stroke-width:1px
-            D-->S(Sold); linkStyle 3 stroke: red, stroke-width:1px
-                S-->L(Liquid storage); linkStyle 4 stroke: red, stroke-width:1px
-                    L-->Co(Composted); linkStyle 5 stroke: red, stroke-width:1px
-                        Co-->Dl(Dry lot); linkStyle 6 stroke: red, stroke-width:1px
-                             Dl-->St(Solid storage); linkStyle 7 stroke: red, stroke-width:1px
-                    L-->Bi(Biogas); linkStyle 8 stroke: green, stroke-width:1px
-                        Bi-->Ls(Liquid slurry); linkStyle 9 stroke: red, stroke-width:1px
-        ")
-```
-
-## Making variables for model development
-
-Here we generate a `make_variables` function for testing the model function 'line by line'.
-We use it to take a single random sample of the provided estimates (this isn't required for running the model later, but it helps when developing the code).
-
-```{r make_variables}
+```{r make_variables, include=FALSE}
 
@@
 
-## The model
-
-The following is our draft of the emissions model, with explanations of all steps (we can format this better once we agree on the final model).
-
-```{r ghg_emissions_function}
+```{r ghg_emissions_function, include=FALSE}
 # Generate the GHG emissions function 
@@
 
-## Running the model
+## Running the model with scenarios
 
-We apply the `decisionSupport` tools for running the Monte Carlo simulation [@R-decisionSupport].
-See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.
+The normal mcSimulation function takes random draws for each of the input variables from the distributions specified in the input table.
+The distributions that are used are the same for every run of the Monte Carlo simulation.
+Sometimes it's desirable to specify particular scenarios *a priori*, e.g.
+to simulate outcomes for particular farm types, climate scenarios etc.
+To do this, we would normally have to define separate input tables and run separate simulations.
+This is quite inconvenient.
+Here, we'll develop a function that can do this job automatically.
+
+The new function will consist of a wrapper around the mcSimulation function, which will be able to modify the input table according to a scenario file we provide.
+This file will contain information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
+
+```{r scenario_sampler}
+scenario_sampler<-function(base_estimate, scenarios, model_function, randomMethod = "calculate", functionSyntax="data.frameNames",
+  relativeTolerance = 0.05,
+  verbosity = 0)
+{
+  initial_estimate = estimate_read_csv(base_estimate)
+  scenarios<-read.csv(scenarios)
+
+  scenario_names<-colnames(scenarios[3:ncol(scenarios)])
+
+  for(scens in scenario_names)
+  {
+    estim<-initial_estimate
+    for (i in 2:nrow(scenarios))
+    {if(scenarios$param[i]=="both")
+      {estim$marginal[scenarios[i,1],"lower"]<-as.numeric(scenarios[i,scens])
+       estim$marginal[scenarios[i,1],"upper"]<-as.numeric(scenarios[i,scens])}
+     if(scenarios$param[i]=="lower")
+       estim$marginal[scenarios[i,1],"lower"]<-as.numeric(scenarios[i,scens])
+     if(scenarios$param[i]=="upper")
+       estim$marginal[scenarios[i,1],"upper"]<-as.numeric(scenarios[i,scens])
+     if(scenarios$param[i]=="distribution")
+       estim$marginal[scenarios[i,1],"distribution"]<-scenarios[i,scens]
+     }
+    scen_results<-mcSimulation(
+      estimate = estim,
+      model_function = model_function,
+      numberOfModelRuns = as.numeric(scenarios[which(scenarios[1,]=="Runs"),scens]),
+      functionSyntax = functionSyntax)
+    scen_results$x[,"Scenario"]<-scens
+    scen_results$y[,"Scenario"]<-scens
+    if(scens==scenario_names[1])
+      end_results<-scen_results
+    if(!scens==scenario_names[1])
+    {end_results$x<-rbind(end_results$x,scen_results$x)
+     end_results$y<-rbind(end_results$y,scen_results$y)
+     end_results$call<-"multi-scenario call"
+     }
+
+  }
+  class(end_results) <- cbind("mcSimulation", class(end_results))
+  return(end_results)
+}
+
+```
+
+This function does the same as the mcSimulation function, but with an additional scenario_file that can be supplied.
+This file specifies which variables should be modified for each of the scenarios, as well as the number of model runs for for it.
+Here's an example file:
+
+```{r}
+knitr::kable(read.csv("data/scenarios.csv"))
+
+```
+
+Each column beyond the first two columns contains information for a particular scenario.
+This is restricted to the parameters that should be *changed* for this scenario.
+All variables that aren't specified in this table remain at their original values given in the usual input table (which is called `base_estimate` here).
+
+The first row ("Runs" in the first column) indicates the number of runs.
+
+Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table.
+The second column indicates which value should be changes, with options being "lower", "upper", "distribution" and "both".
+If "both" is selected, both the lower and upper bounds are modified (this should only be done for constants).
 
 ```{r monte_carlo_simulation}
-GHG_simulation <- mcSimulation(
-  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+
+GHG_simulation<-scenario_sampler(base_estimate = "data/livestock_ghg_input_table.csv",
+                                 scenarios = "data/scenarios.csv",
                                  model_function = ghg_emissions,
-  numberOfModelRuns = 1e4, # 10,000 runs
-  functionSyntax = "plainNames"
-)
+                                 functionSyntax = "plainNames")
+
+#GHG_simulation <- mcSimulation(
+#  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+#  model_function = ghg_emissions,
+#  numberOfModelRuns = 1e4, # 10,000 runs
+#  functionSyntax = "plainNames"
+#)
 ```
@@
 
-Below we apply a post-hoc analysis to the `mcSimulation()` outputs with the `plsr.mcSimulation()` function in the `decisoonSupport` package.
+Below we apply a post-hoc analysis to the `mcSimulation()` outputs with the `plsr.mcSimulation()` function in the `decisionSupport` package.
 We use the function to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model.
@@
 ```{r}
-pls_result <- plsr.mcSimulation(object = GHG_simulation,
-                  resultName = names(GHG_simulation$y)[7], ncomp = 1)
+GHG_simulation_PLS<-GHG_simulation
+GHG_simulation_PLS$x<-GHG_simulation_PLS$x[,1:ncol(GHG_simulation_PLS$x)-1]
+GHG_simulation_PLS$y<-GHG_simulation_PLS$y[,1:ncol(GHG_simulation_PLS$y)-1]
+
+pls_result <- plsr.mcSimulation(object = GHG_simulation_PLS,
+                  resultName = names(GHG_simulation_PLS$y)[7], ncomp = 1)
 ```

### 2021-10-04 210617_GHG_model.Rmd
Add updates from the uncertainty package (mostly just labels for the plots)

@@
 
-```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1}
+```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1, include = F}
 mermaid("
@@
 
-```{r plot_distributions, fig.width=10, fig.height=10}
+```{r plot_distributions, fig.width=10, fig.height=10, include = F}
+# add , include = F to keep this distribution figure building step out of the RMD
 # enteric_CH4=enteric_CH4,
@@
 pls_result <- plsr.mcSimulation(object = GHG_simulation,
-                  resultName = names(GHG_simulation$y)[7], ncomp = 1)
+                  resultName = names(GHG_simulation$y)[7], 
+                  ncomp = 1)
 ```
@@
 
-plot_pls(pls_result, input_table = input_table, threshold = 0)
+plot_pls(pls_result, 
+         input_table = input_table, 
+         threshold = 0)
 ```
@@
 
-We apply a collection of functions for use in assessing uncertainty in the relationship between GHG emissions and milk yields in livestock production.
+Here we apply a collection of functions for use in assessing uncertainty in the relationship between GHG emissions and milk yields in livestock production.
 We use functions from the `uncertainty` package.
@@
 These come from the model results (`r nrow(GHG_simulation$y)`) of the `GHG_simulation` model results to the `in_var` (x) and an outcome variable `out_var` (y).
-in the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
+In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
 
 ```{r name_in_out_vars}
-in_var  <- GHG_simulation$y$milk_yield
-out_var <- GHG_simulation$y$on_farm
+in_var  <- GHG_simulation$y$milk_yield # "Total Milk Yield"
+out_var <- GHG_simulation$y$on_farm # "Total emissions"
 ```
@@
 Here we use the `varscatter()` function to create a scatter plot of milk yields `in_var` (x) and GHG emissions `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves.
-Plot made with the scatter.hist() function in the `plyr()` package [@R-plyr] in R [@R-base].
+We created the plot with the `scatter.hist()` function in the `plyr()` package [@R-plyr] in R [@R-base].
 The result is a scatter plot with associated and estimated densities (using a loess smooth) given by the red dot (mean) and red ellipses (1 and 2 sigma from mean).
@@
 uncertainty::varscatter(in_var = in_var,
-                        out_var = out_var)
+                        out_var = out_var,
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
 ```
@@
 ```{r varkernel}
-uncertainty::varkernel(in_var = in_var, out_var = out_var)
+uncertainty::varkernel(in_var = in_var, 
+                       out_var = out_var,
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
 ```
@@
 
-We use the `varkernelslicerange()` function to show the milk yield (influencing variable `in_var` value) intervals.
+We use the `varkernelslicerange()` function to show the total emissions given a specified milk yield (influencing variable `in_var` value) interval.
 Here we choose a very wide distribution (the whole range of milk yields from the Monte Carlo simulation).
-Here we set the maximum expected value for milk yield `in_var` to `r round(max(in_var), digits = 0)` and the minimum expected value to `r round(min(in_var), digits = 0)`.
+Here we set the maximum expected value for milk yield `in_var` to the maximum (`r round(max(in_var), digits = 0)`) and the minimum expected value (`r round(min(in_var), digits = 0)`) from our simulation.
+We can also choose another range by supplying milk yield values to `max_in_var` and `min_in_var`.
 
 ```{r varkernelslicerange}
-uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
+uncertainty::varkernelslicerange(in_var = in_var, 
+                                 out_var = out_var, 
                                  min_in_var = round(min(in_var)),
-                                 max_in_var = round(max(in_var)))
+                                 max_in_var = round(max(in_var)), 
+                                 xlab_vars = "Total emissions given milk yield range")
 ```
@@
 For a more specific estimate we can choose from those optimized interquartile ranges to select a range to slice from the density kernel `varkernel()`.
-we do that here using the `varviolin()` function to determine meaningful milk yield variable intervals and how they interact with the outcome variable GHG emissions.
+We do that here using the `varviolin()` function to determine meaningful milk yield variable intervals and how they interact with the outcome variable GHG emissions.
 We do this by calculating an 'optimal' interval width using the interquartile range `IQR()` function in the `stats()` package.
@@
 ```{r varviolin, fig.width=17, fig.height=7}
-uncertainty::varviolin(in_var = in_var, out_var = out_var, legend.position = "left")
+uncertainty::varviolin(in_var = in_var, 
+                       out_var = out_var, 
+                       legend.position = "left",
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
 ```

### 2021-10-17 211003_scenarios.Rmd
farm scenario update

@@
 ```{r scenario_sampler}
-scenario_sampler<-function(base_estimate, scenarios, model_function, randomMethod = "calculate", functionSyntax="data.frameNames",
+scenario_sampler<-function(base_estimate, scenarios, model_function, ..., numberOfModelRuns=NA, randomMethod = "calculate", functionSyntax="data.frameNames",
   relativeTolerance = 0.05,
@@
     }
+
+    if("Runs" %in% scenarios[1,])
+      n_Runs<-as.numeric(scenarios[which(scenarios[1,]=="Runs"),scens]) else
+        if(!is.na(numberOfModelRuns))
+          n_Runs<-numberOfModelRuns else
+            stop(paste("Not clear how many times scenario", scens, "should be run"))
+      if(!(n_Runs==round(n_Runs))&(n_Runs>0))
+        stop(paste("Number of runs for scenario", scens, "is not a positive integer"))
     scen_results<-mcSimulation(
@@
       model_function = model_function,
-      numberOfModelRuns = as.numeric(scenarios[which(scenarios[1,]=="Runs"),scens]),
+      ...,
+      numberOfModelRuns = n_Runs,
       functionSyntax = functionSyntax)

### 2021-10-17 211016_GHG_model.Rmd
farm scenario update

--- 2021-09-06 210617_GHG_model.Rmd
@@
 library(rmarkdown)
-# devtools::install_github("CWWhitney/uncertainty")
+#devtools::install_github("CWWhitney/uncertainty")
 library(uncertainty)
@@
                     L-->Bi(Biogas); linkStyle 8 stroke: green, stroke-width:1px
-                        Bi-->Ls(Liquid slurry); linkStyle 9 stroke: red, stroke-width:1px
+                        Bi-->Ls(Liquid); linkStyle 9 stroke: red, stroke-width:1px
         ")
@@
 
-```{r make_variables}
+```{r make_variables, warning=FALSE}
 
@@
 # Generate the GHG emissions function 
-ghg_emissions<-function(x, varnames){
+ghg_emissions<-function(x){
 
@@
 # question 2: of manure that's not centrally collected, how much is burned? share_central_burn
-mms_burn<-(1-share_centrally)*share_central_burn
-mms_pasture<-(1-share_centrally)*(1-share_central_burn)
+#mms_burn<-(1-share_centrally)*share_central_burn
+#mms_pasture<-(1-share_centrally)*(1-share_central_burn)
 
 # question 3: of centrally collected manure, how much is spread daily on cropland or pasture? share_spread
-mms_spread<-share_centrally*share_spread
-not_spread<-share_centrally*(1-share_spread)
+#mms_spread<-share_centrally*share_spread
+#not_spread<-share_centrally*(1-share_spread)
 # question 4: how much of the manure that's not spread is sold? share_sold
-mms_sold<-not_spread*share_sold
-not_sold<-not_spread*(1-share_sold)
+#mms_sold<-not_spread*share_sold
+#not_sold<-not_spread*(1-share_sold)
 # question 5: how much of the stored manure that's not sold is stored in liquid form? share_liquid
-liquid<-not_sold*share_liquid
-solid<-not_sold*(1-share_liquid)
+#liquid<-not_sold*share_liquid
+#solid<-not_sold*(1-share_liquid)
 # question 6: how much of the liquid manure is converted to biogas? share_biogas
-mms_biogas<-liquid*share_biogas
-mms_liquid_slurry<-liquid*(1-share_biogas)
+#mms_biogas<-liquid*share_biogas
+#mms_liquid<-liquid*(1-share_biogas)
 # question 7: how much of the solid manure is composted? share_compost
-mms_composted<-solid*share_compost
-remain_solid<-solid*(1-share_compost)
+#mms_composted<-solid*share_compost
+#remain_solid<-solid*(1-share_compost)
 # question 8: how much of the solid manure is stored in dry lots? share_drylot
-mms_drylot<-remain_solid*share_drylot
-mms_solidstore<-remain_solid*(1-share_drylot)
+#mms_drylot<-remain_solid*share_drylot
+#mms_solidstore<-remain_solid*(1-share_drylot)
 
@@
             (mms_composted*MCFcomposted)+
-            (mms_liquid_slurry*MCFliquid)+
+            (mms_liquid*MCFliquid)+
             (mms_biogas*MCFbiogas)+
@@
 # estimate N retention (kg/animal/day) (Equation 10.33)
+if(weight_gain_WG==0)
+  N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38) else
     N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
@@
                            (mms_composted*mndec_composted)+
-                           (mms_liquid_slurry*mndec_liquid)+
+                           (mms_liquid*mndec_liquid)+
                            (mms_biogas*mndec_biogas)+
@@
                            (mms_composted*mnvc_composted)+
-                           (mms_liquid_slurry*mnvc_liquid)+
+                           (mms_liquid*mnvc_liquid)+
                            (mms_biogas*mnvc_biogas)+
@@
                            (mms_composted*frac_leach_composted)+
-                           (mms_liquid_slurry*frac_leach_liquid)+
+                           (mms_liquid*frac_leach_liquid)+
                            (mms_biogas*frac_leach_biogas)+
@@
 
-feed_CO2<-feed_prod_CO2*feed_trans_CO2
+feed_CO2<-feed_prod_CO2+feed_trans_CO2
 
@@
 
-```{r monte_carlo_simulation}
+```{r monte_carlo_simulation, warning=FALSE}
 GHG_simulation <- mcSimulation(
@@
 
+As of version 1.107 of `decisionSupport`, the package contains the `scenario_mc` function, which allows specifying different distributions of the input variables for specific subsets of the total number of model runs.
+We'll use this functionality here to simulate results for actual herd compositions encountered in Kenya.
+
+```{r scenario_making}
+Kenya_herds<-read.csv("data/Kenya_baseline_individual_cow_wNotes.csv")
+
+herds<-table(Kenya_herds[,c("QQNO","animal_type")])
+
+mms<-t(aggregate(Kenya_herds[,c("mms_pasture","mms_spread","mms_drylot","mms_solidstore","mms_composted","mms_liquid","mms_biogas","mms_burn","system","feed_kgCO2","feed_transport_kgCO2")],by=list(Kenya_herds$QQNO), FUN=median))
+mms<-mms[2:nrow(mms),]
+rownames(mms)[which(rownames(mms)=="system")]<-"feeding_system"
+rownames(mms)[which(rownames(mms)=="feed_kgCO2")]<-"feed_prod_CO2"
+rownames(mms)[which(rownames(mms)=="feed_transport_kgCO2")]<-"feed_trans_CO2"
+
+herd_scenarios<-cbind(data.frame(Variable=paste0("N_animal_type_",1:12),param=rep("both",12)),
+                      as.data.frame.matrix(t(herds)))
+colnames(herd_scenarios)[3:ncol(herd_scenarios)]<-paste0("farm_",1:(ncol(herd_scenarios)-2))
+
+mms_scenarios<-cbind(data.frame(Variable=rownames(mms),param=rep("both",nrow(mms))),
+                     as.data.frame.matrix(mms))
+colnames(mms_scenarios)[3:ncol(mms_scenarios)]<-paste0("farm_",1:(ncol(mms_scenarios)-2))
+
+scenarios<-rbind(herd_scenarios,mms_scenarios)
+
+scenarios[,1:7]   %>%
+    kableExtra::kbl(digits = 2)  %>%
+    kableExtra::kable_classic_2(full_width = F)
+
+```
+
+With these scenarios, we can now feed the Monte Carlo simulation with realistic farm compositions.
+To adequately consider uncertainty, we can run a user-specified number of random scenarios for each farm.
+
+```{r scenario_runs, warning=FALSE}
+
+GHG_simulation_real_farms<-scenario_mc(base_estimate=estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                     scenarios=scenarios,
+                     model_function = ghg_emissions,
+                     numberOfModelRuns = 10,
+                       functionSyntax="plainNames")
+
+```
+
 Here we plot the distributions of the various outputs from the model.
@@
 enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "enteric_CH4",
+                   method = "boxplot_density",
+                   old_names = "enteric_CH4",
+                   new_names = "Total Methane Emissions from Livestock by Enteric Fermentation") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+enteric_CH4_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
                    vars = "enteric_CH4",
@@
 
+milk_yield_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
+                   vars = "milk_yield",
+                   method = "boxplot_density",
+                   old_names = "milk_yield",
+                   new_names = "Total Milk Yield") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
 # enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
@@
 
+enteric_CH4_CO2eq_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
+                   vars = "enteric_CH4_CO2eq",
+                   method = "boxplot_density",
+                   old_names = "enteric_CH4_CO2eq",
+                   new_names = "Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock") + 
+  theme(axis.title.x=element_blank())
+
 # mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
 mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "mm_CH4_CO2eq",
+                   method = "boxplot_density",
+                   old_names = "mm_CH4_CO2eq",
+                   new_names = "Carbon Dioxide Equivalent Methane Emissions from Livestock Manure") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
+mm_CH4_CO2eq_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
                    vars = "mm_CH4_CO2eq",
@@
 
+mm_N2O_CO2eq_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
+                   vars = "mm_N2O_CO2eq",
+                   method = "boxplot_density",
+                   old_names = "mm_N2O_CO2eq",
+                   new_names = "Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
 # feed_CO2=feed_CO2,
@@
 
+feed_CO2_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
+                   vars = "feed_CO2",
+                   method = "boxplot_density",
+                   old_names = "feed_CO2",
+                   new_names = "Carbon Dioxide Emissions from Livestock Feed Production and Transport") + 
+  theme(axis.title.x=element_blank(),
+        axis.title.y=element_blank())
+
 # on_farm=sum(on_farm)
 on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+                   vars = "on_farm",
+                   method = "boxplot_density",
+                   old_names = "on_farm",
+                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
+  theme(axis.title.y=element_blank())
+
+on_farm_dist_plot_farms <- plot_distributions(mcSimulation_object = GHG_simulation_real_farms,
                    vars = "on_farm",
@@
           "
+```
 
+Here's the plot that results from using the same parameters for all farms:
+
+```{r}
 enteric_CH4_dist_plot + milk_yield_dist_plot + 
@@
 plot_layout(guides = "collect", design = layout)
+
+```
+
+And here's what happens when we run 10 random scenarios for each real herd composition:
+
+```{r}
+enteric_CH4_dist_plot_farms + milk_yield_dist_plot_farms + 
+  enteric_CH4_CO2eq_dist_plot_farms + mm_CH4_CO2eq_dist_plot_farms + 
+  mm_N2O_CO2eq_dist_plot_farms + feed_CO2_dist_plot_farms + 
+  on_farm_dist_plot_farms + 
+plot_layout(guides = "collect", design = layout)
+
 ```
@@
 ```{r}
-pls_result <- plsr.mcSimulation(object = GHG_simulation,
-                  resultName = names(GHG_simulation$y)[7], ncomp = 1)
+pls_input<-GHG_simulation_real_farms
+pls_input$x<-pls_input$x[,which(!("Scenario" == colnames(pls_input$x)))]
+pls_result <- plsr.mcSimulation(object = pls_input,
+                  resultName = names(pls_input$y)[7], ncomp = 1)
 ```
@@
 We run the `plot_pls()` on the results from `plsr.mcSimulation()` with a number of standard settings.
-The length of the bars is equal to VIP with a vertical line at '1' on the x-axis indicating a standard cut-off for VIP used for variable selection [@Whitney2017].
-The overall plot only shows those variables with a VIP > 0.8, which is the common threshold for variable selection [@Lanzanova2019, @Luedeling2016].
+The length of the bars is equal to VIP with a vertical line at '1' on the x-axis indicating a standard cut-off for VIP used for variable selection [@whitney_simulating_2017].
+The overall plot only shows those variables with a VIP > 0.8, which is the common threshold for variable selection [@lanzanovaImprovingDevelopmentEfficiency2019b, @luedeling2016decision].
 The colors of the bars represent the positive or negative coefficient of the given input variable with the output variable.
@@
 We apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
-These come from the model results (`r nrow(GHG_simulation$y)`) of the `GHG_simulation` model results to the `in_var` (x) and an outcome variable `out_var` (y).
+These come from the model results (`r nrow(GHG_simulation_real_farms$y)`) of the `GHG_simulation_real_farms` model results to the `in_var` (x) and an outcome variable `out_var` (y).
 in the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
@@
 ```{r name_in_out_vars}
-in_var  <- GHG_simulation$y$milk_yield
-out_var <- GHG_simulation$y$on_farm
+in_var  <- GHG_simulation_real_farms$y$milk_yield
+out_var <- GHG_simulation_real_farms$y$on_farm
 ```

### 2021-10-31 GHG_model.Rmd
Renamed and rearranged the files, add a single R file with the function and 'source' that in each markdown file (now we have just one for the scenarios and one for the overall model). Found a few issues with the code - We are still getting some 'NaN' values from the loop in the .R function file. We may need to take a closer look at this soon.

--- 2021-10-17 210617_GHG_model.Rmd
@@
 # Generate the GHG emissions function 
-ghg_emissions<-function(x, varnames){
-
-# Enteric fermentation emissions ($CH_4$)
-
-# The main source of all this is Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
-The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
-
-# We start with the net energy for maintenance (NE_m; in MJ/day)
-# This includes a maintenance coefficient, for which different values are recommended depending on animal type
-
-# The following animal types were used in the example:
-
-# animal type 1 - 'intact' grown-up males >=36 months
-# animal type 2 - castrated grown-up males >=36 months
-# animal type 3 - males 12-36 months
-# animal type 4 - lactating cows
-# animal type 5 - lactating cows
-# animal type 6 - lactating cows
-# animal type 7 - weaned young females <= 12 months
-# animal type 8 - weaned young males <= 12 months
-# animal type 9 - females 12-36 months
-# animal type 10 - pre-weaned female calves (<12 months)
-# animal type 11 - pre-weaned male calves (<12 months)
-# animal type 12 - first-time pregnant cows
-
-# animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant.
-This resulted in a weighted mean of the various maintenance coefficients.
-We're not sure where we could get such data.
-In the original dataset, types 4, 5 and 6 referred to different pregnancy/lactation states, but it doesn't seem like this scheme was rigorously followed.
-
-# We make placeholder vectors to collect emissions data for all animal types
-mm_N2O_CO2eq<-c()
-mm_CH4_CO2eq<-c()
-enteric_CH4_CO2eq<-c()
-total_milk<-c()
-
-# Now we start a long loop, where emissions are calculated for each animal type
-
-for (animal_type in 1:12)
-{
-# define C and Cfi based on sex (Page 10.17, below Equation 10.6)
-if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
-if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
-if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)
-
-if(animal_type %in% c(4,5,6)) Cfi<-maintenance_coefficient_Cfi_lact_cow # lactating cows (0.386)
-if(animal_type %in% c(1)) Cfi<-maintenance_coefficient_Cfi_bulls # bulls (0.37)
-if(animal_type %in% c(2,3,7,8,9,10,11,12)) Cfi<-maintenance_coefficient_Cfi_other # others (0.322)
-
-# define mature body weight (assuming this needs to be disaggregated by sex, for Eqation 10.6)
-if(animal_type %in% c(1,2,3,8,11)) mature_weight_MW<-mature_weight_MW_male
-if(animal_type %in% c(4,5,6,7,9,10,12)) mature_weight_MW<-mature_weight_MW_female
-
-# define milk yield
-if(animal_type %in% c(1,2,3,7,8,9,10,11,12)) milk_yield<-0
-if(animal_type %in% c(4,5,6)) milk_yield<-eval(parse(text=paste0("milk_yield_",animal_type)))
-
-#define pregnancy coefficient (full Cp for first-time pregnant, lower for later in most cases)
-if(animal_type==12) Cp<-Cp_0
-if(animal_type==4) Cp<-Cp_0*Cp_rate_4
-if(animal_type==5) Cp<-Cp_0*Cp_rate_5
-if(animal_type==6) Cp<-Cp_0*Cp_rate_6
-if(animal_type %in% c(1,2,3,7,8,9,10,11)) Cp<-0
-
-N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))
-
-# define activity coefficient Ca (Table 10.5)
-# Stall 0.00 - Animals are confined to a small area (i.e., tethered, pen, barn) with the result that they expend very little or no energy to acquire feed.
-# Pasture 0.17 - Animals are confined in areas with sufficient forage requiring modest energy expense to acquire feed.
-# Grazing large areas 0.36 - Animals graze in open range land or hilly terrain and expend significant energy to acquire feed.
-
-# We don't quite understand how Ca was assigned in the original data table.
-There are many 0s for 'System 1' (Stall?), but also lots of values between 0 and 0.17 or 0.36.
-We can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table.
-For now, we're restricting these Ca values to only those that are given as 'pure' categories in the IPCC guidelines (Table 10.5).
-
-if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
-if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
-if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)
-
-# Mean live weight (kg) of animals surely varies by animal type, so we need separate estimates for them
-live_weight_LW<-eval(parse(text=paste0("live_weight_LW_",animal_type)))
-
-# Mean weight gain of animal type (kg/day)
-weight_gain_WG<-eval(parse(text=paste0("weight_gain_WG_",animal_type)))
-
-# Calculate the Net Energy for Maintenance (NE_m) (MJ/day) (Equation 10.3)
-
-NE_m<-Cfi*(live_weight_LW^0.75)
-
-# Calculate the Net Energy for Activity (NE_activity) (MJ/day) (Equation 10.4)
-
-NE_activity<-C_activity*NE_m
-
-#Net energy for growth (NE_growth) (MJ/day) (Equation 10.6)
-
-NE_growth<-22.02*
-  ((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*
-  (weight_gain_WG^1.097)
-
-# Net energy for lactation
-
-NE_lactation<-milk_yield*(1.47+0.40*milk_fat)
-
-# Draft animal services are not accounted for
-# NE_work <- 0.1 * NE_m * hours_per_day
-
-#Energy need during pregnancy (Equation 10.13)
-NE_pregnancy<-Cp*NE_m
-
-# The Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant).
-For all other lactating cows, this is reduced - not clear how this was done.
-
-# Ratio of net energy available in a diet for maintenance to digestible energy consumed (Equation 10.14)
-REM<-1.123-(4.092/1000*DE_perc)+(1.126/100000*DE_perc^2)-(25.4/DE_perc)
-
-# Ratio of net energy available for growth in a diet to digestible energy consumed (Equation 10.15)
-REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)
-
-# Gross Energy (GE) MJ/day
-NE_work<-0 # work animals aren't covered here, so we set this to zero
-GE<-(((NE_m+NE_activity+NE_lactation+NE_work+NE_pregnancy)/REM)+(NE_growth/REG))/(DE_perc/100)
-
-# CH4 emission factors for enteric fermentation from a livestock category (Equation 10.21)
-
-enteric_CH4<-(GE*Y_m*365)/55.65 # kg CH4 per animal per year
-
-# Compute the CO2-equivalent of methane emissions for enteric methane
-#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
-enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25
-
-# Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
-# daily volatile solid excreted (Equation 10.24)
-
-volsol<-(GE*
-           (1-(DE_perc/100))+
-           (urinary_energy*GE))*
-             ((1-ash_content)/18.45)
-
-# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
-# treat these as constants for each animal type. The following line assigns the corresponding Bo value
-##### EL CW Note #### 
-##### # to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
-# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.
-
-Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))
-
-# decisions on manure management
-# the management options we're considering here are the ones from the Excel table. The IPCC guidelines
-# list more and partly different ones.
-
-# We're deciding on manure management according to a decision tree (see figure above):
-
-# question 1: how much of the manure is centrally collected? share_centrally
-# question 2: of manure that's not centrally collected, how much is burned? share_central_burn
-mms_burn<-(1-share_centrally)*share_central_burn
-mms_pasture<-(1-share_centrally)*(1-share_central_burn)
-
-# question 3: of centrally collected manure, how much is spread daily on cropland or pasture? share_spread
-mms_spread<-share_centrally*share_spread
-not_spread<-share_centrally*(1-share_spread)
-# question 4: how much of the manure that's not spread is sold? share_sold
-mms_sold<-not_spread*share_sold
-not_sold<-not_spread*(1-share_sold)
-# question 5: how much of the stored manure that's not sold is stored in liquid form? share_liquid
-liquid<-not_sold*share_liquid
-solid<-not_sold*(1-share_liquid)
-# question 6: how much of the liquid manure is converted to biogas? share_biogas
-mms_biogas<-liquid*share_biogas
-mms_liquid_slurry<-liquid*(1-share_biogas)
-# question 7: how much of the solid manure is composted? share_compost
-mms_composted<-solid*share_compost
-remain_solid<-solid*(1-share_compost)
-# question 8: how much of the solid manure is stored in dry lots? share_drylot
-mms_drylot<-remain_solid*share_drylot
-mms_solidstore<-remain_solid*(1-share_drylot)
-
-# CH4 emission factor from manure management (Equation 10.23)
-mm_ch4<-(volsol*365)*
-        (Bo*0.67*
-           ((mms_pasture*MCFpasture)+
-            (mms_spread*MCFspread)+
-            (mms_drylot*MCFdrylot)+
-            (mms_solidstore*MCFsolidstore)+
-            (mms_composted*MCFcomposted)+
-            (mms_liquid_slurry*MCFliquid)+
-            (mms_biogas*MCFbiogas)+
-            (mms_burn*MCFburn)+
-              # CW EL Comment ####
-              # the original code was lacking the 'sold' fraction - we added this for completeness
-            (mms_sold*MCFsold)))
-
-# Estimate CO2eq for manure methane
-mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25
-
-# Calculate Nitrous oxide emissions from manure and urine on pasture
-
-# estimate N intake (kg/animal/day) (Equation 10.32)
-N_intake<-(GE/18.45)*((perc_crude_protein_diet/100)/6.25)
-
-# estimate N retention (kg/animal/day) (Equation 10.33)
-N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
-  ((weight_gain_WG*(268-(7.03*NE_growth/weight_gain_WG)))/(1000*6.25))
-
-# estimate N excretion (kg N/animal/yr) (Equation 10.31)
-N_excretion<-N_intake*(1-N_retention)*365
-
-# Direct nitrous oxide emissions (kg N2O/animal/yr) (Eq 10.25 from IPCC guidelines)
-mm_direct_n2o<-N_excretion*((mms_pasture*mndec_pasture)+
-                           (mms_spread*mndec_spread)+
-                           (mms_drylot*mndec_drylot)+
-                           (mms_solidstore*mndec_solidstore)+
-                           (mms_composted*mndec_composted)+
-                           (mms_liquid_slurry*mndec_liquid)+
-                           (mms_biogas*mndec_biogas)+
-                           (mms_burn*mndec_burn)+
-                           (mms_sold*mndec_sold))*(44/28)
-
-# N losses from volatilization (kg N/animal/yr) (Eq 10.26 from IPCC guidelines)
-mm_vol_n<-N_excretion*((mms_pasture*mnvc_pasture)+
-                           (mms_spread*mnvc_spread)+
-                           (mms_drylot*mnvc_drylot)+
-                           (mms_solidstore*mnvc_solidstore)+
-                           (mms_composted*mnvc_composted)+
-                           (mms_liquid_slurry*mnvc_liquid)+
-                           (mms_biogas*mnvc_biogas)+
-                           (mms_burn*mnvc_burn)+
-                           (mms_sold*mnvc_sold))
-
-# indirect N2O emissions due to volatilization of N from manure management (kg N2O/animal/year) (Equation 10.27)
-mm_vol_n2o<-mm_vol_n*EF4*(44/28)
-
-# Nitrous oxide (N2O) from leaching (kg N/animal/year) (equation 10.28)
-mm_leach_n<-N_excretion*((mms_pasture*frac_leach_pasture)+
-                           (mms_spread*frac_leach_spread)+
-                           (mms_drylot*frac_leach_drylot)+
-                           (mms_solidstore*frac_leach_solidstore)+
-                           (mms_composted*frac_leach_composted)+
-                           (mms_liquid_slurry*frac_leach_liquid)+
-                           (mms_biogas*frac_leach_biogas)+
-                           (mms_burn*frac_leach_burn)+
-                           (mms_sold*frac_leach_sold))
-
-# indirect N2O emissions due to leaching of N from manure management (kg N2O/animal/year) (Equation 10.29)
-mm_leach_n2o<-mm_leach_n*EF5*(44/28)
-
-# CW EL Comment ####
-# # comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
-#mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
-#              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)
-
-# Estimate CO2eq for manure nitrous
-mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)
-
-total_milk[animal_type]<-N_animal_type*milk_yield*365
-
-}
-
-# Emissions from feed production and transport. 
-# CW EL Comment ####
-  # comment: the numbers in the table seem spuriously precise.
-  ####
-
-feed_CO2<-feed_prod_CO2*feed_trans_CO2
-
-# Total emissions
-on_farm<-enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq+feed_CO2
-
-return(list(enteric_CH4=enteric_CH4,
-            milk_yield=sum(total_milk),
-            enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
-            mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
-            mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
-            feed_CO2=feed_CO2,
-            on_farm=sum(on_farm)))
-}
+source("GHG_emissions_function.R", local = knitr::knit_global())
 ```

### 2021-10-31 Scenarios.Rmd
Renamed and rearranged the files, add a single R file with the function and 'source' that in each markdown file (now we have just one for the scenarios and one for the overall model). Found a few issues with the code - We are still getting some 'NaN' values from the loop in the .R function file. We may need to take a closer look at this soon.

--- 2021-10-17 211003_scenarios.Rmd
@@
 
-We're defining the function as in 210617_GHG_model.Rmd, but there's no need to show this again here.
+We're defining the function as in the 'Greenhouse gas emissions model' document. Please look there for more inforamtion about the original model design.
 
@@
 
-## Generating the GHG emissions model
-
-The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
-The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
-The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021], also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
-Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].
-We run the model with the decisionSupport package [@R-decisionSupport].
-
 We load the library and the data.
@@
 
-```{r make_variables, include=FALSE}
-
-make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
-for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
-
-# run the function on the input_table
-make_variables(estimate_read_csv(input_table))
-```
-
 ```{r ghg_emissions_function, include=FALSE}
 # Generate the GHG emissions function 
-ghg_emissions<-function(x, varnames){
-
-# Enteric fermentation emissions ($CH_4$)
-
-# The main source of all this is Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
-The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
-
-# We start with the net energy for maintenance (NE_m; in MJ/day)
-# This includes a maintenance coefficient, for which different values are recommended depending on animal type
-
-# The following animal types were used in the example:
-
-# animal type 1 - 'intact' grown-up males >=36 months
-# animal type 2 - castrated grown-up males >=36 months
-# animal type 3 - males 12-36 months
-# animal type 4 - lactating cows
-# animal type 5 - lactating cows
-# animal type 6 - lactating cows
-# animal type 7 - weaned young females <= 12 months
-# animal type 8 - weaned young males <= 12 months
-# animal type 9 - females 12-36 months
-# animal type 10 - pre-weaned female calves (<12 months)
-# animal type 11 - pre-weaned male calves (<12 months)
-# animal type 12 - first-time pregnant cows
-
-# animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant.
-This resulted in a weighted mean of the various maintenance coefficients.
-We're not sure where we could get such data.
-In the original dataset, types 4, 5 and 6 referred to different pregnancy/lactation states, but it doesn't seem like this scheme was rigorously followed.
-
-# We make placeholder vectors to collect emissions data for all animal types
-mm_N2O_CO2eq<-c()
-mm_CH4_CO2eq<-c()
-enteric_CH4_CO2eq<-c()
-total_milk<-c()
-
-# Now we start a long loop, where emissions are calculated for each animal type
-
-for (animal_type in 1:12)
-{
-# define C and Cfi based on sex (Page 10.17, below Equation 10.6)
-if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
-if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
-if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)
-
-if(animal_type %in% c(4,5,6)) Cfi<-maintenance_coefficient_Cfi_lact_cow # lactating cows (0.386)
-if(animal_type %in% c(1)) Cfi<-maintenance_coefficient_Cfi_bulls # bulls (0.37)
-if(animal_type %in% c(2,3,7,8,9,10,11,12)) Cfi<-maintenance_coefficient_Cfi_other # others (0.322)
-
-# define mature body weight (assuming this needs to be disaggregated by sex, for Eqation 10.6)
-if(animal_type %in% c(1,2,3,8,11)) mature_weight_MW<-mature_weight_MW_male
-if(animal_type %in% c(4,5,6,7,9,10,12)) mature_weight_MW<-mature_weight_MW_female
-
-# define milk yield
-if(animal_type %in% c(1,2,3,7,8,9,10,11,12)) milk_yield<-0
-if(animal_type %in% c(4,5,6)) milk_yield<-eval(parse(text=paste0("milk_yield_",animal_type)))
-
-#define pregnancy coefficient (full Cp for first-time pregnant, lower for later in most cases)
-if(animal_type==12) Cp<-Cp_0
-if(animal_type==4) Cp<-Cp_0*Cp_rate_4
-if(animal_type==5) Cp<-Cp_0*Cp_rate_5
-if(animal_type==6) Cp<-Cp_0*Cp_rate_6
-if(animal_type %in% c(1,2,3,7,8,9,10,11)) Cp<-0
-
-N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))
-
-# define activity coefficient Ca (Table 10.5)
-# Stall 0.00 - Animals are confined to a small area (i.e., tethered, pen, barn) with the result that they expend very little or no energy to acquire feed.
-# Pasture 0.17 - Animals are confined in areas with sufficient forage requiring modest energy expense to acquire feed.
-# Grazing large areas 0.36 - Animals graze in open range land or hilly terrain and expend significant energy to acquire feed.
-
-# We don't quite understand how Ca was assigned in the original data table.
-There are many 0s for 'System 1' (Stall?), but also lots of values between 0 and 0.17 or 0.36.
-We can only imagine that these are weighted averages or reflect additional information about the feeding system, but this information doesn't seem to be given in the table.
-For now, we're restricting these Ca values to only those that are given as 'pure' categories in the IPCC guidelines (Table 10.5).
-
-if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
-if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
-if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)
-
-# Mean live weight (kg) of animals surely varies by animal type, so we need separate estimates for them
-live_weight_LW<-eval(parse(text=paste0("live_weight_LW_",animal_type)))
-
-# Mean weight gain of animal type (kg/day)
-weight_gain_WG<-eval(parse(text=paste0("weight_gain_WG_",animal_type)))
-
-# Calculate the Net Energy for Maintenance (NE_m) (MJ/day) (Equation 10.3)
-
-NE_m<-Cfi*(live_weight_LW^0.75)
-
-# Calculate the Net Energy for Activity (NE_activity) (MJ/day) (Equation 10.4)
-
-NE_activity<-C_activity*NE_m
-
-#Net energy for growth (NE_growth) (MJ/day) (Equation 10.6)
-
-NE_growth<-22.02*
-  ((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*
-  (weight_gain_WG^1.097)
-
-# Net energy for lactation
-
-NE_lactation<-milk_yield*(1.47+0.40*milk_fat)
-
-# Draft animal services are not accounted for
-# NE_work <- 0.1 * NE_m * hours_per_day
-
-#Energy need during pregnancy (Equation 10.13)
-NE_pregnancy<-Cp*NE_m
-
-# The Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant).
-For all other lactating cows, this is reduced - not clear how this was done.
-
-# Ratio of net energy available in a diet for maintenance to digestible energy consumed (Equation 10.14)
-REM<-1.123-(4.092/1000*DE_perc)+(1.126/100000*DE_perc^2)-(25.4/DE_perc)
-
-# Ratio of net energy available for growth in a diet to digestible energy consumed (Equation 10.15)
-REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)
-
-# Gross Energy (GE) MJ/day
-NE_work<-0 # work animals aren't covered here, so we set this to zero
-GE<-(((NE_m+NE_activity+NE_lactation+NE_work+NE_pregnancy)/REM)+(NE_growth/REG))/(DE_perc/100)
-
-# CH4 emission factors for enteric fermentation from a livestock category (Equation 10.21)
-
-enteric_CH4<-(GE*Y_m*365)/55.65 # kg CH4 per animal per year
-
-# Compute the CO2-equivalent of methane emissions for enteric methane
-#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
-enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25
-
-# Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
-# daily volatile solid excreted (Equation 10.24)
-
-volsol<-(GE*
-           (1-(DE_perc/100))+
-           (urinary_energy*GE))*
-             ((1-ash_content)/18.45)
-
-# The maximum methane emission from manure (Bo) depends on the animal type. In the input table, we
-# treat these as constants for each animal type. The following line assigns the corresponding Bo value
-##### EL CW Note #### 
-##### # to the Bo variable. In the dataset we're using for Kenya, only two values for Bo are given: 0.17 for
-# animal types 4-6 and 0.135 for animal types 1-3 and 7-12.
-
-Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))
-
-# decisions on manure management
-# the management options we're considering here are the ones from the Excel table. The IPCC guidelines
-# list more and partly different ones.
-
-# We're deciding on manure management according to a decision tree (see figure above):
-
-# question 1: how much of the manure is centrally collected? share_centrally
-# question 2: of manure that's not centrally collected, how much is burned? share_central_burn
-mms_burn<-(1-share_centrally)*share_central_burn
-mms_pasture<-(1-share_centrally)*(1-share_central_burn)
-
-# question 3: of centrally collected manure, how much is spread daily on cropland or pasture? share_spread
-mms_spread<-share_centrally*share_spread
-not_spread<-share_centrally*(1-share_spread)
-# question 4: how much of the manure that's not spread is sold? share_sold
-mms_sold<-not_spread*share_sold
-not_sold<-not_spread*(1-share_sold)
-# question 5: how much of the stored manure that's not sold is stored in liquid form? share_liquid
-liquid<-not_sold*share_liquid
-solid<-not_sold*(1-share_liquid)
-# question 6: how much of the liquid manure is converted to biogas? share_biogas
-mms_biogas<-liquid*share_biogas
-mms_liquid_slurry<-liquid*(1-share_biogas)
-# question 7: how much of the solid manure is composted? share_compost
-mms_composted<-solid*share_compost
-remain_solid<-solid*(1-share_compost)
-# question 8: how much of the solid manure is stored in dry lots? share_drylot
-mms_drylot<-remain_solid*share_drylot
-mms_solidstore<-remain_solid*(1-share_drylot)
-
-# CH4 emission factor from manure management (Equation 10.23)
-mm_ch4<-(volsol*365)*
-        (Bo*0.67*
-           ((mms_pasture*MCFpasture)+
-            (mms_spread*MCFspread)+
-            (mms_drylot*MCFdrylot)+
-            (mms_solidstore*MCFsolidstore)+
-            (mms_composted*MCFcomposted)+
-            (mms_liquid_slurry*MCFliquid)+
-            (mms_biogas*MCFbiogas)+
-            (mms_burn*MCFburn)+
-              # CW EL Comment ####
-              # the original code was lacking the 'sold' fraction - we added this for completeness
-            (mms_sold*MCFsold)))
-
-# Estimate CO2eq for manure methane
-mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25
-
-# Calculate Nitrous oxide emissions from manure and urine on pasture
-
-# estimate N intake (kg/animal/day) (Equation 10.32)
-N_intake<-(GE/18.45)*((perc_crude_protein_diet/100)/6.25)
-
-# estimate N retention (kg/animal/day) (Equation 10.33)
-N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
-  ((weight_gain_WG*(268-(7.03*NE_growth/weight_gain_WG)))/(1000*6.25))
-
-# estimate N excretion (kg N/animal/yr) (Equation 10.31)
-N_excretion<-N_intake*(1-N_retention)*365
-
-# Direct nitrous oxide emissions (kg N2O/animal/yr) (Eq 10.25 from IPCC guidelines)
-mm_direct_n2o<-N_excretion*((mms_pasture*mndec_pasture)+
-                           (mms_spread*mndec_spread)+
-                           (mms_drylot*mndec_drylot)+
-                           (mms_solidstore*mndec_solidstore)+
-                           (mms_composted*mndec_composted)+
-                           (mms_liquid_slurry*mndec_liquid)+
-                           (mms_biogas*mndec_biogas)+
-                           (mms_burn*mndec_burn)+
-                           (mms_sold*mndec_sold))*(44/28)
-
-# N losses from volatilization (kg N/animal/yr) (Eq 10.26 from IPCC guidelines)
-mm_vol_n<-N_excretion*((mms_pasture*mnvc_pasture)+
-                           (mms_spread*mnvc_spread)+
-                           (mms_drylot*mnvc_drylot)+
-                           (mms_solidstore*mnvc_solidstore)+
-                           (mms_composted*mnvc_composted)+
-                           (mms_liquid_slurry*mnvc_liquid)+
-                           (mms_biogas*mnvc_biogas)+
-                           (mms_burn*mnvc_burn)+
-                           (mms_sold*mnvc_sold))
-
-# indirect N2O emissions due to volatilization of N from manure management (kg N2O/animal/year) (Equation 10.27)
-mm_vol_n2o<-mm_vol_n*EF4*(44/28)
-
-# Nitrous oxide (N2O) from leaching (kg N/animal/year) (equation 10.28)
-mm_leach_n<-N_excretion*((mms_pasture*frac_leach_pasture)+
-                           (mms_spread*frac_leach_spread)+
-                           (mms_drylot*frac_leach_drylot)+
-                           (mms_solidstore*frac_leach_solidstore)+
-                           (mms_composted*frac_leach_composted)+
-                           (mms_liquid_slurry*frac_leach_liquid)+
-                           (mms_biogas*frac_leach_biogas)+
-                           (mms_burn*frac_leach_burn)+
-                           (mms_sold*frac_leach_sold))
-
-# indirect N2O emissions due to leaching of N from manure management (kg N2O/animal/year) (Equation 10.29)
-mm_leach_n2o<-mm_leach_n*EF5*(44/28)
-
-# CW EL Comment ####
-# # comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
-#mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
-#              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)
-
-# Estimate CO2eq for manure nitrous
-mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)
-
-total_milk[animal_type]<-N_animal_type*milk_yield*365
-
-}
-
-# Emissions from feed production and transport. 
-# CW EL Comment ####
-  # comment: the numbers in the table seem spuriously precise.
-  ####
-
-feed_CO2<-feed_prod_CO2*feed_trans_CO2
-
-# Total emissions
-on_farm<-enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq+feed_CO2
-
-return(list(enteric_CH4=enteric_CH4,
-            milk_yield=sum(total_milk),
-            enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
-            mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
-            mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
-            feed_CO2=feed_CO2,
-            on_farm=sum(on_farm)))
-}
+source("GHG_emissions_function.R", local = knitr::knit_global())
 ```
@@
 
-The normal mcSimulation function takes random draws for each of the input variables from the distributions specified in the input table.
+The normal `mcSimulation` function takes random draws for each of the input variables from the distributions specified in the input table.
 The distributions that are used are the same for every run of the Monte Carlo simulation.
@@
 
-The new function will consist of a wrapper around the mcSimulation function, which will be able to modify the input table according to a scenario file we provide.
+The new `scenario_sampler` function will consist of a wrapper around the `mcSimulation` function, which will be able to modify the input table according to a scenario file we provide.
 This file will contain information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
@@
 
-This function does the same as the mcSimulation function, but with an additional scenario_file that can be supplied.
+This function does the same as the `mcSimulation` function, but with an additional `scenario_file` that can be supplied.
 This file specifies which variables should be modified for each of the scenarios, as well as the number of model runs for for it.
@@
 
-GHG_simulation<-scenario_sampler(base_estimate = "data/livestock_ghg_input_table.csv",
+GHG_simulation_scenarios<-scenario_sampler(base_estimate = "data/livestock_ghg_input_table.csv",
                                  scenarios = "data/scenarios.csv",
@@
 
-#GHG_simulation <- mcSimulation(
-#  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
-#  model_function = ghg_emissions,
-#  numberOfModelRuns = 1e4, # 10,000 runs
-#  functionSyntax = "plainNames"
-#)
 ```
@@
 # enteric_CH4=enteric_CH4,
-enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "enteric_CH4",
@@
 # milk_yield=sum(total_milk),
-milk_yield_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+milk_yield_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "milk_yield",
@@
 # enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
-enteric_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+enteric_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "enteric_CH4_CO2eq",
@@
 # mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
-mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "mm_CH4_CO2eq",
@@
 # mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
-mm_N2O_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+mm_N2O_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "mm_N2O_CO2eq",
@@
 # feed_CO2=feed_CO2,
-feed_CO2_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+feed_CO2_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "feed_CO2",
@@
 # on_farm=sum(on_farm)
-on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation,
+on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
                    vars = "on_farm",
@@
 
-Below we apply a post-hoc analysis to the `mcSimulation()` outputs with the `plsr.mcSimulation()` function in the `decisionSupport` package.
-We use the function to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model.
-This function uses the outputs of the `mcSimulation()` selecting all the input variables from the `GHG_simulation` decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`.
-We use the code `names(GHG_simulation$y)[7]` to select the outcome variable `r names(GHG_simulation$y)[7]`, which is an element of the list `y` in our `GHG_simulation` outputs.
-
 ```{r}
-GHG_simulation_PLS<-GHG_simulation
-GHG_simulation_PLS$x<-GHG_simulation_PLS$x[,1:ncol(GHG_simulation_PLS$x)-1]
-GHG_simulation_PLS$y<-GHG_simulation_PLS$y[,1:ncol(GHG_simulation_PLS$y)-1]
-
-pls_result <- plsr.mcSimulation(object = GHG_simulation_PLS,
-                  resultName = names(GHG_simulation_PLS$y)[7], ncomp = 1)
+pls_result <- plsr.mcSimulation(object = GHG_simulation_scenarios,
+                  resultName = names(GHG_simulation_scenarios$y)[7], 
+                  ncomp = 1)
 ```
-
-We run the `plot_pls()` on the results from `plsr.mcSimulation()` with a number of standard settings.
-The length of the bars is equal to VIP with a vertical line at '1' on the x-axis indicating a standard cut-off for VIP used for variable selection [@Whitney2017].
-The overall plot only shows those variables with a VIP > 0.8, which is the common threshold for variable selection [@Lanzanova2019, @Luedeling2016].
-The colors of the bars represent the positive or negative coefficient of the given input variable with the output variable.
-
-Here we import the input table again (the same table we used for the Monte Carlo model) to replace the labels for the variables on the y-axis.
-The input table includes a 'label' and 'variable' column.
-The standard labels (from the 'variable' column) are usually computer readable and not very nice for a plot.
-The `plot_pls()` function uses the text in the 'label' column as replacement for the default text in the 'variable' column.
 
@@
 
-We apply a collection of functions for use in assessing uncertainty in the relationship between GHG emissions and milk yields in livestock production.
-We use functions from the `uncertainty` package.
-Install the package using `devtools::install_github("CWWhitney/uncertainty")`.
-
-We apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
-These come from the model results (`r nrow(GHG_simulation$y)`) of the `GHG_simulation` model results to the `in_var` (x) and an outcome variable `out_var` (y).
-in the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
+As with the GHG Model, we apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
+These come from the model results (`r nrow(GHG_simulation_scenarios$y)`) of the `GHG_simulation_scenarios` model results to the `in_var` (x) and an outcome variable `out_var` (y).
+In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
 
 ```{r name_in_out_vars}
-in_var  <- GHG_simulation$y$milk_yield
-out_var <- GHG_simulation$y$on_farm
+in_var  <- GHG_simulation_scenarios$y$milk_yield # "Total Milk Yield"
+out_var <- GHG_simulation_scenarios$y$on_farm # "Total emissions"
 ```
@@
 
-Here we use the `varscatter()` function to create a scatter plot of milk yields `in_var` (x) and GHG emissions `out_var` (y) with associated and estimated densities with loess smooth linear fits density curves.
-Plot made with the scatter.hist() function in the `plyr()` package [@R-plyr] in R [@R-base].
-The result is a scatter plot with associated and estimated densities (using a loess smooth) given by the red dot (mean) and red ellipses (1 and 2 sigma from mean).
-The red line going across the plot shows the linear fit.
-Histograms are shown with smooth lines (loess smooth linear fits) density curves.
-The numeric value in the upper right gives the Spearman correlation coefficient between the influencing variable and the outcome variable.
-
 ```{r uncertainty_package}
 uncertainty::varscatter(in_var = in_var,
-                        out_var = out_var)
+                        out_var = out_var,
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
 ```
-
-To verify the finding we use the `varkernel()` function for another type of graphical representation of the same relationship.
-We use the `varkernel()` function to perform a two-dimensional kernel density estimation for the expected value of GHG emissions and milk yield.
-The core of the process involves kernel density estimation.
-This is a nonparametric technique for estimating probability density functions (similar to histogram density estimation but with more dimensions).
-We apply this with the `kde2d()` function in the `MASS()` package [@R-MASS].
-The function produces a matrix of the estimated density (z) of the expected value of the GHG emissions and milk yield.
-It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a normal distribution (a bivariate normal kernel).
-
 ```{r varkernel}
-uncertainty::varkernel(in_var = in_var, out_var = out_var)
+uncertainty::varkernel(in_var = in_var, 
+                       out_var = out_var,
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
 ```
-
-The result of the `varkernel()` function shows a density surface plot of GHG emissions `in_var` (x) and milk yield `out_var` (y).
-In these `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the GHG emissions `out_var` along the y-axis) given a specific value for milk yield `in_var` (along the x-axis).
-The legend shows the value for the estimated density (z).
-The plot is made with the `filled.contour()` function of the `graphics()` package [@R-base].
 
@@
 
-We use the `varkernelslicerange()` function to show the milk yield (influencing variable `in_var` value) intervals.
-Here we choose a very wide distribution (the whole range of milk yields from the Monte Carlo simulation).
-Here we set the maximum expected value for milk yield `in_var` to `r round(max(in_var), digits = 0)` and the minimum expected value to `r round(min(in_var), digits = 0)`.
-
 ```{r varkernelslicerange}
-uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
+uncertainty::varkernelslicerange(in_var = in_var, 
+                                 out_var = out_var, 
                                  min_in_var = round(min(in_var)),
-                                 max_in_var = round(max(in_var)))
+                                 max_in_var = round(max(in_var)), 
+                                 xlab_vars = "Total emissions given milk yield range")
 ```
-
-`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable `out_var` (shown along the x-axis).
-As with `varkernelslice()` the probability values shown are relative, not absolute measures.
-They are the result of cuts through the density kernel `varkernel()`, which integrates to 1.
 
@@
 
-For a more specific estimate we can choose from those optimized interquartile ranges to select a range to slice from the density kernel `varkernel()`.
-we do that here using the `varviolin()` function to determine meaningful milk yield variable intervals and how they interact with the outcome variable GHG emissions.
-We do this by calculating an 'optimal' interval width using the interquartile range `IQR()` function in the `stats()` package.
-
 ```{r varviolin, fig.width=17, fig.height=7}
-uncertainty::varviolin(in_var = in_var, out_var = out_var, legend.position = "left")
+uncertainty::varviolin(in_var = in_var, 
+                       out_var = out_var, 
+                       legend.position = "left",
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
 ```
-
-The result of the `varviolin()` function shows violin plots of milk yield `in_var` values (x) and our outcome variable GHG emissions `out_var` (y) values.
-The plot was made with `ggplot2()` [@R-ggplot2].
-These offer a possible selection of milk yields and show the resulting expected GHG emissions.
 

### 2021-10-31 GHG_Emissions_Function.R

@@ -1,0 +1,285 @@
+
+library(DiagrammeR)
+library(decisionSupport)
+library(ggplot2)
+library(plyr)
+library(rmarkdown)
+# devtools::install_github("CWWhitney/uncertainty")
+library(uncertainty)
+
+# Generate the GHG emissions function 
+ghg_emissions<-function(x, varnames){
+  
+  # Enteric fermentation emissions ($CH_4$)
+  
+  # The main source of all this is Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use). The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
+  
+  # We start with the net energy for maintenance (NE_m; in MJ/day)
+  # This includes a maintenance coefficient, for which different values are recommended depending on animal type
+  
+  # The following animal types were used in the example:
+  
+  # animal type 1 - 'intact' grown-up males >=36 months
+  # animal type 2 - castrated grown-up males >=36 months
+  # animal type 3 - males 12-36 months
+  # animal type 4 - lactating cows
+  # animal type 5 - lactating cows
+  # animal type 6 - lactating cows
+  # animal type 7 - weaned young females <= 12 months
+  # animal type 8 - weaned young males <= 12 months
+  # animal type 9 - females 12-36 months
+  # animal type 10 - pre-weaned female calves (<12 months)
+  # animal type 11 - pre-weaned male calves (<12 months)
+  # animal type 12 - first-time pregnant cows
+  
+  # animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant. This resulted in a weighted mean of the various maintenance coefficients. We're not sure where we could get such data. In the original dataset, types 4, 5 and 6 referred to different pregnancy/lactation states, but it doesn't seem like this scheme was rigorously followed.
+  
+  # We make placeholder vectors to collect emissions data for all animal types
+  mm_N2O_CO2eq<-c()
+  mm_CH4_CO2eq<-c()
+  enteric_CH4_CO2eq<-c()
+  total_milk<-c()
+  
[…]

### 2021-10-31 Uncertainty in Kenyan Livestock Emission Reductions.R

@@ -1,112 +1,0 @@
-# Estimating Uncertainty in Kenyan Livestock Emission Reductions
-
-# Load data
-# CW Note ####
-# We need this file
-# Also, no need for the whole directory here. 
-# The .Rproj will find the .csv in a relative location (i.e. subfolder)
-
-library(decisionSupport)
-#CW Note ####
-# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
-# This project will look in relative file paths (i.e. same or lower folders)
-input_table <- "livestock_ghg_input_table.csv"
-results_folder <- "MC results"
-
-# Generate a function for testing the model function 'line by line' 
-# by taking a single random sample of the provided estimates
-make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
-for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
-# run the function on the input_table
-make_variables(estimate_read_csv(input_table))
-
-# Generate the GHG emissions function 
-ghg_emissions<-function(x, varnames){
-
-# Enteric fermentation emissions ($CH_4$)
-  #CW Note ####
-  #in these code sections I am missing the overview. 
-  # it is not immediately clear what this is meant to do... 
-  # from a specific publication? 
-nem<-maintenaince_coefficient_Cfi*(live_weight_LW^0.75)
-nea<-activity_coefficient_Ca*nem
-nel<-milk_yield*(1.47+0.4*milk_fat) 
-nep<-pregnancy_coefficient_Cp*nem
-neg<-22.02*(((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*(weight_gain_WG^1.097))
-rem<-(1.123-(4.092*(10^-3)*dig_energy)+(1.126*(10^-5)*(dig_energy^2))-(25.4/dig_energy))
-reg<-(1.164-(5.16*(10^-3)*dig_energy)+(1.308*(10^-5)*(dig_energy^2))-(37.4/dig_energy))
-gross_energy<-((((nem+nea+nel+nep)/rem)+(neg/reg))/dig_energy/100)
-enteric_CH4<-(gross_energy*(6.5/100)*365)/55.65 
-
[…]

### 2021-10-31 Scenarios.Rmd
Cleaned up some of the assessments we really only need for the main model report

@@
 
-## Model sensitivity
-
-#### Perform a Partial Least Squares Regression (PLSR) of Monte Carlo simulation results
-
-```{r}
-pls_result <- plsr.mcSimulation(object = GHG_simulation_scenarios,
-                  resultName = names(GHG_simulation_scenarios$y)[7], 
-                  ncomp = 1)
-```
-
-```{r}
-input_table <- read.csv("data/livestock_ghg_input_table.csv")
-
-plot_pls(pls_result, 
-         input_table = input_table, 
-         threshold = 0)
-```
-
 # Uncertainty
@@
 
-## Influencing variable intervals
-
-```{r varviolin, fig.width=17, fig.height=7}
-uncertainty::varviolin(in_var = in_var, 
-                       out_var = out_var, 
-                       legend.position = "left",
-                        xlab = "Total Milk Yield",
-                        ylab = "Total emissions")
-```
-
 This document was generated using R Markdown [@R-rmarkdown].

### 2021-11-01 GHG_Emissions_Function.R
Add notes and headers Still a small (non trivial) issue with NaN values

@@ -1,21 +1,40 @@
+# Generate the GHG emissions function ####
+# Cory Whitney and Eike Luedeling
 
-library(DiagrammeR)
+# # install packages ####
 library(decisionSupport)
 library(ggplot2)
 library(plyr)
-library(rmarkdown)
-# devtools::install_github("CWWhitney/uncertainty")
-library(uncertainty)
 
-# Generate the GHG emissions function 
+# load data ####
+input_table <- "data/livestock_ghg_input_table.csv"
+
+# make variables function ####
+# Here we generate a `make_variables` function 
+# for testing the model function 'line by line'. 
+# We use it to take a single random sample of the provided estimates 
+# (this isn't required for running the model later, 
+# but it helps when developing the code).
+make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
+for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
+
+# run the function on the input_table
+make_variables(estimate_read_csv(input_table))
+
+# Generate the GHG emissions function ####
 ghg_emissions<-function(x, varnames){
   
   # Enteric fermentation emissions ($CH_4$)
   
-  # The main source of all this is Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use). The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
+  # The main source of all this is Chapter 10 of the *2006 IPCC Guidelines 
+  # for National Greenhouse Gas Inventories* 
+  # (Volume 4 - Agriculture, Forestry and Other Land Use). 
+  # The document can be found on the IPCC web 
+  # https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf
   
   # We start with the net energy for maintenance (NE_m; in MJ/day)
-  # This includes a maintenance coefficient, for which different values are recommended depending on animal type
+  # This includes a maintenance coefficient, 
+  # for which different values are recommended depending on animal type
   
   # The following animal types were used in the example:
   
@@ -32,15 +51,25 @@ ghg_emissions<-function(x, varnames){
   # animal type 11 - pre-weaned male calves (<12 months)
   # animal type 12 - first-time pregnant cows
   
[…]

### 2021-11-01 GHG_model_report.Rmd
Rename and shft files for clarity, update readme with indications about files

--- 2021-10-31 GHG_model.Rmd
@@
 library(ggplot2)
+library(kableExtra)
+library(magrittr)
 library(plyr)
@@
 ```{r load_data}
-library(DiagrammeR)
-library(decisionSupport)
-library(plyr)
-library(magrittr)
-library(kableExtra)
 ### # This project will be stored in relative file paths (i.e. same or lower folders)
@@
 
-```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1, include = F}
+```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1}
 mermaid("

### 2021-11-01 Scenarios_report.Rmd
Rename and shft files for clarity, update readme with indications about files

--- 2021-10-31 Scenarios.Rmd
@@
 
-We're defining the function as in the 'Greenhouse gas emissions model' document. Please look there for more inforamtion about the original model design.
+We're defining the function as in the ['Greenhouse gas emissions model']() document. Please look there for more information about the original model design.
 
@@
 
-```{r, warning = F, include = F}
+```{r, warning=FALSE, include=FALSE}
 #install and load packages
@@
 ```{r load_data, include=FALSE}
-library(DiagrammeR)
-library(decisionSupport)
-library(plyr)
-library(magrittr)
-library(kableExtra)
+
 ### # This project will be stored in relative file paths (i.e. same or lower folders)

### 2021-11-10 Scenarios_report.Rmd
update csv
update with the new scenario_mc function

@@
 
-The new `scenario_sampler` function will consist of a wrapper around the `mcSimulation` function, which will be able to modify the input table according to a scenario file we provide.
-This file will contain information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
-
-```{r scenario_sampler}
-scenario_sampler<-function(base_estimate, scenarios, model_function, ..., numberOfModelRuns=NA, randomMethod = "calculate", functionSyntax="data.frameNames",
-  relativeTolerance = 0.05,
-  verbosity = 0)
-{
-  initial_estimate = estimate_read_csv(base_estimate)
-  scenarios<-read.csv(scenarios)
-
-  scenario_names<-colnames(scenarios[3:ncol(scenarios)])
-
-  for(scens in scenario_names)
-  {
-    estim<-initial_estimate
-    for (i in 2:nrow(scenarios))
-    {if(scenarios$param[i]=="both")
-      {estim$marginal[scenarios[i,1],"lower"]<-as.numeric(scenarios[i,scens])
-       estim$marginal[scenarios[i,1],"upper"]<-as.numeric(scenarios[i,scens])}
-     if(scenarios$param[i]=="lower")
-       estim$marginal[scenarios[i,1],"lower"]<-as.numeric(scenarios[i,scens])
-     if(scenarios$param[i]=="upper")
-       estim$marginal[scenarios[i,1],"upper"]<-as.numeric(scenarios[i,scens])
-     if(scenarios$param[i]=="distribution")
-       estim$marginal[scenarios[i,1],"distribution"]<-scenarios[i,scens]
-    }
-
-    if("Runs" %in% scenarios[1,])
-      n_Runs<-as.numeric(scenarios[which(scenarios[1,]=="Runs"),scens]) else
-        if(!is.na(numberOfModelRuns))
-          n_Runs<-numberOfModelRuns else
-            stop(paste("Not clear how many times scenario", scens, "should be run"))
-      if(!(n_Runs==round(n_Runs))&(n_Runs>0))
-        stop(paste("Number of runs for scenario", scens, "is not a positive integer"))
-    scen_results<-mcSimulation(
-      estimate = estim,
-      model_function = model_function,
-      ...,
-      numberOfModelRuns = n_Runs,
-      functionSyntax = functionSyntax)
-    scen_results$x[,"Scenario"]<-scens
-    scen_results$y[,"Scenario"]<-scens
-    if(scens==scenario_names[1])
-      end_results<-scen_results
-    if(!scens==scenario_names[1])
-    {end_results$x<-rbind(end_results$x,scen_results$x)
-     end_results$y<-rbind(end_results$y,scen_results$y)
-     end_results$call<-"multi-scenario call"
-     }
-
-  }
-  class(end_results) <- cbind("mcSimulation", class(end_results))
-  return(end_results)
-}
-
-```
+The new `scenario_mc` function consists of a wrapper around the `mcSimulation` function, which is able to modify the input table according to a scenario file we provide.
+This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
 
@@
 ```{r}
-knitr::kable(read.csv("data/scenarios.csv"))
+knitr::kable(read.table("data/scenarios.csv", sep = ","))
 
@@
 
-GHG_simulation_scenarios<-scenario_sampler(base_estimate = "data/livestock_ghg_input_table.csv",
-                                 scenarios = "data/scenarios.csv",
+GHG_simulation_scenarios<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                                 scenarios = read.csv("data/scenarios.csv",fileEncoding="UTF-8-BOM"),
                                  model_function = ghg_emissions,

### 2021-11-10 GHG_model_report.Rmd
Update some inputs, update graph,...

@@
 
-```{r graph-mermaid-manure, echo=FALSE, fig.width=10, fig.height=1}
+```{r graph-mermaid-manure, echo=FALSE, fig.width=4, fig.height=1}
 mermaid("

### 2021-11-11 Scenarios_report.Rmd
real farm scenarios

@@
 
+```{r}
+scenarios<-read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM")
+
+Kenyaherds<-read.table("data/Kenya_baseline_individual_cow_wNotes.csv", sep = ",",
+                       fileEncoding = "UTF-8-BOM", header=TRUE)
+
+scenarios<-data.frame(Variable=variables_to_update,param="both")
+
+# count the number of animals for each type
+
+animals<-table(Kenyaherds[,c("QQNO","animal_type")])
+
+for(i in 1:nrow(animals))
+  {coln<-paste0("Farm_",row.names(animals)[i])
+   scenarios[,coln]<-NA
+     scenarios[which(scenarios$Variable %in% paste0("N_animal_type_",1:12)),coln]<-animals[i,]
+}
+
+# for some animal-scale variables (e.g. milk yield), we need averages per farm
+
+# farm-scale variables
+
+Farms<-unique(Kenyaherds$QQNO)
+
+for (f in Farms)
+  {coln<-paste0("Farm_",f)
+   scenarios[which(scenarios$Variable %in% "feeding_system"),coln]<-
+     median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])
+   }
+
+write.csv(scenarios,"data/farm_scenarios.csv")
+
+```
+
 ```{r monte_carlo_simulation}
@@
 GHG_simulation_scenarios<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
-                                 scenarios = read.csv("data/scenarios.csv",fileEncoding="UTF-8-BOM"),
-                                 model_function = ghg_emissions,
+                                 scenarios = read.csv("data/farm_scenarios.csv",fileEncoding="UTF-8-BOM")[1:2,],
+                                 model_function = ghg_emissions,numberOfModelRuns=10,
                                  functionSyntax = "plainNames")

### 2021-11-11 GHG_Emissions_Function.R

@@ -108,16 +108,7 @@ ghg_emissions<-function(x, varnames){
     # Grazing large areas 0.36 - Animals graze in open range land 
     #     or hilly terrain and expend significant energy to acquire feed.
     
-    # EL CW Note ####
-    # We don't quite understand how Ca was assigned in the original data table. 
-    # There are many 0s for 'System 1' (Stall?), 
-    # but also lots of values between 0 and 0.17 or 0.36. 
-    # We can only imagine that these are weighted averages or reflect 
-    # additional information about the feeding system, 
-    # but this information doesn't seem to be given in the table. 
-    # For now, we're restricting these Ca values to only those 
-    # that are given as 'pure' categories in the IPCC guidelines (Table 10.5).
-    
+
     if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
     if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
     if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)
@@ -145,11 +136,6 @@ ghg_emissions<-function(x, varnames){
     
     #Energy need during pregnancy (Equation 10.13)
     NE_pregnancy<-Cp*NE_m
-    
-    # CW EL Note ####
-    # The Cp coefficient given in Equation 10.13 is 0.1, 
-    # but in the sample this is only applied for animal type 12 (first-time pregnant). 
-    # For all other lactating cows, this is reduced - not clear how this was done.
     
     # Ratio of net energy available in a diet for maintenance 
     # to digestible energy consumed (Equation 10.14)
@@ -185,47 +171,7 @@ ghg_emissions<-function(x, varnames){
     # In the input table, we treat these as constants for each animal type. 
     # The following line assigns the corresponding Bo value
     
-    ##### EL CW Note #### 
-    ##### # to the Bo variable. In the dataset we're using for Kenya, 
-    ##### only two values for Bo are given: 0.17 for
-    # animal types 4-6 and 0.135 for animal types 1-3 and 7-12.
-    
     Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))
-    
-    # decisions on manure management
-    # CW EL Note ####
-    # the management options we're considering here are the ones from the Excel table. 
-    # The IPCC guidelines list more and partly different ones.
-    
[…]

### 2021-11-11 Scenarios_report.Rmd
not working yet...

@@
 ```{r}
-scenarios<-read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM")
+scenarios<-read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)
 
@@
 
-scenarios<-data.frame(Variable=variables_to_update,param="both")
-
 # count the number of animals for each type
@@
 animals<-table(Kenyaherds[,c("QQNO","animal_type")])
+
+variables_to_update<-c(paste0("N_animal_type_",1:12),"feeding_system")
+
+scenarios<-data.frame(Variables=variables_to_update,param="both")
 

### 2021-11-11 Scenarios_report.Rmd
fugde scenario_mc

--- 2021-11-11 Scenarios_report.Rmd
@@
 animals<-table(Kenyaherds[,c("QQNO","animal_type")])
+
+variables_to_update<-c(paste0("N_animal_type_",1:12),"feeding_system","feed_prod_CO2","feed_trans_CO2")
+
+scenarios<-data.frame(Variable=variables_to_update,param="both")
 
@@
      median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])
+   scenarios[which(scenarios$Variable %in% "feed_prod_CO2"),coln]<-
+     median(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])
+   scenarios[which(scenarios$Variable %in% "feed_trans_CO2"),coln]<-
+     median(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])
+
    }
 
-write.csv(scenarios,"data/farm_scenarios.csv")
+write.csv(scenarios,"data/farm_scenarios.csv",row.names = FALSE)
 
@@
 
-```{r monte_carlo_simulation}
+```{r monte_carlo_simulation, warning=FALSE}
 
 GHG_simulation_scenarios<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
-                                 scenarios = read.csv("data/farm_scenarios.csv",fileEncoding="UTF-8-BOM")[1:2,],
+                                 scenarios = read.csv("data/farm_scenarios.csv",fileEncoding="UTF-8-BOM"),
                                  model_function = ghg_emissions,numberOfModelRuns=10,

### 2021-11-11 GHG_model_report.Rmd
Add devtools options to load the latest version of decisionSupport

@@
 library(DiagrammeR)
+# devtools::install_github("eikeluedeling/decisionSupport")
 library(decisionSupport)

### 2021-11-12 Scenarios_report.Rmd
model fixed

--- 2021-11-11 Scenarios_report.Rmd
@@
 ```{r}
-knitr::kable(read.table("data/scenarios.csv", sep = ","))
+knitr::kable(read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE))
 
@@
 ```{r}
-scenarios<-read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)
-
 Kenyaherds<-read.table("data/Kenya_baseline_individual_cow_wNotes.csv", sep = ",",

### 2021-11-12 GHG_Emissions_Function.R

@@ -267,7 +267,7 @@ ghg_emissions<-function(x, varnames){
   
   # Emissions from feed production and transport. 
 
-  feed_CO2<-feed_prod_CO2*feed_trans_CO2
+  feed_CO2<-feed_prod_CO2+feed_trans_CO2
   
   # Total emissions 
 

### 2021-11-12 Scenarios_report.Rmd
now extracting milk yield, live weight, weight gain and feeding system from survey table. Digestable energy and crude protein are also extracted, but just as averages per farm for now. Uncertainty analysis now shown per head.

@@
 
-variables_to_update<-c(paste0("N_animal_type_",1:12),"feeding_system","feed_prod_CO2","feed_trans_CO2")
+variables_to_update_const<-c(paste0("N_animal_type_",1:12),
+                             "feeding_system")
 
-scenarios<-data.frame(Variable=variables_to_update,param="both")
+scenarios<-data.frame(Variable=variables_to_update_const,param="both")
+
+variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
+                             "feed_prod_CO2",
+                             "feed_trans_CO2",
+                             paste0("live_weight_LW_",1:12),
+                             paste0("weight_gain_WG_",1:12),
+                             "DE_perc",
+                             "perc_crude_protein_diet")
+
+scenarios<-rbind(scenarios,data.frame(Variable=rep(variables_to_update_posnorm,each=3),param=c("lower","upper","distribution")))
 
@@
 
+### error estimates for variables
+milk_yield_error<-0.25
+feed_trans_error<-0.25
+feed_prod_error<-0.25
+live_weight_error<-0.1
+weight_gain_error<-0.1
+digestible_energy_error<-0.1
+crude_protein_error<-0.1
+
 # farm-scale variables
@@
   {coln<-paste0("Farm_",f)
-   scenarios[which(scenarios$Variable %in% "feeding_system"),coln]<-
+  # update feeding system 
+  scenarios[which(scenarios$Variable == "feeding_system"),coln]<-
      median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])
-   scenarios[which(scenarios$Variable %in% "feed_prod_CO2"),coln]<-
-     median(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])
-   scenarios[which(scenarios$Variable %in% "feed_trans_CO2"),coln]<-
-     median(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])
+
+  # update digestible energy (mean for all animals for now) (with error)
+  scenarios[which(scenarios$Variable == "DE_perc"),coln][1:2]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"dig_energy"])*
+                           c(1-digestible_energy_error,1+digestible_energy_error)
+
+  # update crude protein (mean for all animals for now) (with error)
+  scenarios[which(scenarios$Variable == "perc_crude_protein_diet"),coln][1:2]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"CP."])*
+                           c(1-crude_protein_error,1+crude_protein_error)
+
+  # update feed production CO2 (with error)
+   scenarios[which(scenarios$Variable == "feed_prod_CO2"),coln][1:2]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])*
+                           c(1-feed_prod_error,1+feed_prod_error)
+  # update feed transport CO2 (with error)
+   scenarios[which(scenarios$Variable == "feed_trans_CO2"),coln][1:2]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])*
+                           c(1-feed_trans_error,1+feed_trans_error)
+  # update milk yield for each cow type (animal type 4-6) (with error)
+   for(typ in 4:6)
+     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
+       scenarios[which(scenarios$Variable==paste0("milk_yield_",typ)),coln][1:2]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "milk_yield"])*c(1-milk_yield_error,1+milk_yield_error)
+  # update live weight and weight gain for each animal type (with error)
+   for(typ in 1:12)
+     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
+       {scenarios[which(scenarios$Variable==paste0("live_weight_LW_",typ)),coln][1:2]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "live_weight_LW"])*c(1-live_weight_error,1+live_weight_error)
+       scenarios[which(scenarios$Variable==paste0("weight_gain_WG_",typ)),coln][1:2]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "weight_gain_WG"])*c(1-weight_gain_error,1+weight_gain_error)
+     }
+
+   for(varia in variables_to_update_posnorm)
+   if(sum(is.na(scenarios[which(scenarios$Variable==varia),coln][1:2]))==0)   
+     if(!equals(scenarios[which(scenarios$Variable==varia),coln][1],
+               scenarios[which(scenarios$Variable==varia),coln][2]))
+        scenarios[which(scenarios$Variable==varia&
+                   scenarios$param=="distribution"),coln]<-"posnorm" else
+            scenarios[which(scenarios$Variable==varia&
+                   scenarios$param=="distribution"),coln]<-"const"
 
@@
 
-# Uncertainty
+# Normalizing milk yield and emissions per head
+
+```{r}
+
+animals_per_run<-rowSums(GHG_simulation_scenarios$x[paste0("N_animal_type_",1:12)])
+
+milk_yield_per_head<-GHG_simulation_scenarios$y$milk_yield/animals_per_run
+total_emissions_per_head<-GHG_simulation_scenarios$y$on_farm/animals_per_run
+
+```
+
+# Uncertainty for per-head data
 
@@
 ```{r name_in_out_vars}
-in_var  <- GHG_simulation_scenarios$y$milk_yield # "Total Milk Yield"
-out_var <- GHG_simulation_scenarios$y$on_farm # "Total emissions"
+in_var  <- milk_yield_per_head # "Total Milk Yield"
+out_var <- total_emissions_per_head # "Total emissions"
 ```

### 2021-11-13 Scenarios_report.Rmd
implemented most changes we talked about today. First draft of breed intervention is included, but will definitely need a check.

@@
 
+# Interventions
+
+Here's a wild idea: Can we simply change values in the GHG_simulation_scenarios$x table now and apply the model to the resulting table?
+We may not have to draw new values (and that actually wouldn't even be good, because this would dilute the intervention effect).
+
+To clearly characterize the impact of interventions, we should compare their impact with a counterfactual, i.e.
+with the same farms without the intervention.
+While this is usually impossible in practice, it is easy with a model.
+We just have to run the simulation again with identical inputs except for those that are affected by the intervention.
+
+We already simulated emissions for a large population of farms, and we can now apply our interventions to this population by modifying certain parameters of the input tables.
+What we don't want now, however, is to randomize all values again, since this would add additional noise to the results, which would mask the signal caused by the interventions.
+
+We now first need a function that can apply our greenhouse gas emissions model to a pre-defined set of input values.
+We can find syntax that can achieve this in the mcSimulation function.
+Here it is (with slight modifications):
+
+```{r}
+run_model<-function(x_data,model)
+  {
+  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
+  model_function_ <- function(x) {
+
+    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
+    eval(expr = body(model), envir = e)
+  }
+  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
+                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
+  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
+  returnObject$call <- match.call()
+  class(returnObject) <- cbind("mcSimulation", class(returnObject))
+  return(returnObject)
+}
+
+```
+
+The following code can precisely reproduce the results of the previous analysis:
+
+```{r, eval=FALSE}
+res<-run_model(GHG_simulation_scenarios$x,ghg_emissions)
+```
+
+To simulate interventions, we can now feed this function with modified x_data data.frames.
+
+## Breed
+
+Introduction of higher-performance breeds may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.
+
+To retain the farm-specific performance levels obtained from the farm survey, we simulate these impacts in a relative manner.
+For this, we first need to determine the performance and characteristics of each of the breeds.
+We'll base this on the farm survey data (`Kenyaherds` dataset from above).
+
+```{r}
+aggregate(Kenyaherds$milk_yield, by=list(Kenyaherds$breed,Kenyaherds$animal_type), FUN=mean)
+
+cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
+
+ggplot(cows, aes(factor(breed),milk_yield)) +
+  geom_violin(draw_quantiles=c(0.5),fill="light green") +
+  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
+
+milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
+milk_summary<-milk_summary[,c(1,4,2)]
+colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
+milk_summary
+
+```
+
+This evaluation shows that the difference in median milk yields is relatively small, but that the milk yield distributions show a lot of variation that is apparently not explained by the breed.
+It is also apparent that the number of animals these estimates are based on varies considerably.
+Two breeds make up `r round(sum(milk_summary$n[1:2])/sum(milk_summary$n)*100,1)` % of the entire population, with others having very few animals, which means that the medians may be somewhat unreliable (especially for the single animal of breed 7).
+
+Let's define the intervention as follows:
+
+All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences.
+This means that milk yields are multiplied by the following factors:
+
+```{r}
+milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))
+
+milk_summary             
+```
+
+A change of the breed also causes changes in weight, mature weight and weight gain.
+
+```{r}
+live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(live_weights)<-c("animal_type","breed","median_live_weight")
+live_weights[,"Intervention_factor"]<-NA
+for(i in 1:nrow(live_weights))
+  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
+     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]
+
+# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.
+
+mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
+mature_weights[,"Intervention_factor"]<-NA
+for(i in 1:nrow(mature_weights))
+  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
+     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]
+
+# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1
+
+weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
+weight_gains[,"Intervention_factor"]<-NA
+for(i in 1:nrow(weight_gains))
+  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
+     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
+weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1
+
+```
+
+Now we can modify the input table accordingly.
+
+GHG_simulation_scenarios$x
+
+go through all the farms. Determine the 'improvement' coefficients (based on herd structure)
+
+```{r}
+farms<-unique(Kenyaherds$QQNO)
+breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)
+
+Kenyaherds[,"milk_change"]<-0
+cows<-which(Kenyaherds$animal_type %in% 4:6)
+Kenyaherds$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds$breed[x]]  )
+
+milk_change<-aggregate(Kenyaherds$milk_change[cows], by=list(Kenyaherds$QQNO[cows]), FUN=mean)
+
+breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% milk_change[,1])]<-milk_change[,2]
+breed_intervention$milk_yield_5<-breed_intervention$milk_yield_4
+breed_intervention$milk_yield_6<-breed_intervention$milk_yield_4
+
+for(i in 1:nrow(Kenyaherds))
+  {Kenyaherds[i,"live_weight_change"]<-1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds$animal_type[i]&
+                                                                                live_weights$breed==Kenyaherds$breed[i])]
+#   Kenyaherds[i,"mature_weight_change"]<-1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds$animal_type[i]&
+#                                                                                mature_weights$breed==Kenyaherds$breed[i])]
+   Kenyaherds[i,"weight_gain_change"]<-1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds$animal_type[i]&
+                                                                                weight_gains$breed==Kenyaherds$breed[i])]
+}
+
+live_weight_agg<-aggregate(Kenyaherds$live_weight_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
+#mature_weight_agg<-aggregate(Kenyaherds$mature_weight_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
+weight_gain_agg<-aggregate(Kenyaherds$weight_gain_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
+
+colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
+#colnames(mature_weight_agg)<-c("farm","animal_type","conv_factor")
+colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")
+
+for(i in 1:nrow(live_weight_agg))
+  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
+                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
+    live_weight_agg$conv_factor[i]
+
+#for(i in 1:nrow(mature_weight_agg))
+#  breed_intervention[which(breed_intervention$Farm==mature_weight_agg$farm[i]),
+#                     paste0("mature_weight_LW_",mature_weight_agg$animal_type[i])]<-
+#    mature_weight_agg$conv_factor[i]
+
+for(i in 1:nrow(weight_gain_agg))
+  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
+                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
+    weight_gain_agg$conv_factor[i]
+
+breed_intervention[is.na(breed_intervention)]<-1
+
+```
+
+go through the x file
+for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors
+
+```{r}
+x_breed_intervention<-GHG_simulation_scenarios$x
+
+adoption_rate<-0.6
+
+for(i in 1:nrow(x_breed_intervention))
+{if(rbinom(1,1,adoption_rate))
+    x_breed_intervention[,colnames(breed_intervention)[2:ncol(breed_intervention)]]
+  # implement changes in x file
+}
+  res<-run_model(x_breed_intervention,ghg_emissions)
+
+```
+
+```{r}
+uncertainty::varkernel(in_var = res$x$milk_yield_5, 
+                       out_var = res$y$on_farm,
+                        xlab = "Total Milk Yield",
+                        ylab = "Total emissions")
+
+```
+
+## Feed
+
+## Herd structure
+
 This document was generated using R Markdown [@R-rmarkdown].

### 2021-11-13 Scenarios_report.Rmd
breed intervention - first complete draft

@@
 variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
+                             paste0("Cp_rate_",4:6),
                              "feed_prod_CO2",
@@
 ### error estimates for variables
-milk_yield_error<-0.25
-feed_trans_error<-0.25
-feed_prod_error<-0.25
-live_weight_error<-0.1
-weight_gain_error<-0.1
+milk_yield_error<-0.275 # Migose et al. (2020, https://www.sciencedirect.com/science/article/abs/pii/S1871141319307371 )
+feed_trans_error<-0.25 # Vellinga et al (2013, https://edepot.wur.nl/254098)
+feed_prod_error<-0.25 # Vellinga et al (2013, https://edepot.wur.nl/254098)
+live_weight_error<-0.145 # Goopy et al. (2018, https://www.publish.csiro.au/an/an16577 ) 
+weight_gain_error<-0.145 # Goopy et al. (2018, https://www.publish.csiro.au/an/an16577 ) 
 digestible_energy_error<-0.1
@@
      if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
-       scenarios[which(scenarios$Variable==paste0("milk_yield_",typ)),coln][1:2]<-
+       {scenarios[which(scenarios$Variable==paste0("milk_yield_",typ)),coln][1:2]<-
          mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
                          "milk_yield"])*c(1-milk_yield_error,1+milk_yield_error)
+       scenarios[which(scenarios$Variable==paste0("CP_rate_",typ)),coln][1:2]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "Cp"]/0.1) *c(1,1)
+
+     }
   # update live weight and weight gain for each animal type (with error)
@@
                         out_var = out_var,
-                        xlab = "Total Milk Yield",
-                        ylab = "Total emissions")
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
 ```
@@
                        out_var = out_var,
-                        xlab = "Total Milk Yield",
-                        ylab = "Total emissions")
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
 ```
@@
 
-## Breed
+## Breed intervention
 
@@
 ```{r}
-aggregate(Kenyaherds$milk_yield, by=list(Kenyaherds$breed,Kenyaherds$animal_type), FUN=mean)
-
 cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
@@
 
-GHG_simulation_scenarios$x
-
 go through all the farms. Determine the 'improvement' coefficients (based on herd structure)
@@
 
-Kenyaherds[,"milk_change"]<-0
-cows<-which(Kenyaherds$animal_type %in% 4:6)
-Kenyaherds$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds$breed[x]]  )
+Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
+                                        "system",
+                                        "animal_type",
+                                        "breed",
+                                        "live_weight_LW",
+                                        "mature_weight_MW",
+                                        "weight_gain_WG",
+                                        "milk_yield",
+                                        "dig_energy",
+                                        "CP.")]
 
-milk_change<-aggregate(Kenyaherds$milk_change[cows], by=list(Kenyaherds$QQNO[cows]), FUN=mean)
+Kenyaherds_interventions[,"milk_change"]<-0
+cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
+Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )
+
+milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)
 
@@
 
-for(i in 1:nrow(Kenyaherds))
-  {Kenyaherds[i,"live_weight_change"]<-1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds$animal_type[i]&
-                                                                                live_weights$breed==Kenyaherds$breed[i])]
-#   Kenyaherds[i,"mature_weight_change"]<-1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds$animal_type[i]&
-#                                                                                mature_weights$breed==Kenyaherds$breed[i])]
-   Kenyaherds[i,"weight_gain_change"]<-1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds$animal_type[i]&
-                                                                                weight_gains$breed==Kenyaherds$breed[i])]
+for(i in 1:nrow(Kenyaherds_interventions))
+  {Kenyaherds_interventions[i,"live_weight_change"]<-
+    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
+                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
+#   Kenyaherds_interventions[i,"mature_weight_change"]<-
+#      1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
+#                                                 mature_weights$breed==Kenyaherds_interventions$breed[i])]
+   Kenyaherds_interventions[i,"weight_gain_change"]<-
+     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
+                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
 }
 
-live_weight_agg<-aggregate(Kenyaherds$live_weight_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
-#mature_weight_agg<-aggregate(Kenyaherds$mature_weight_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
-weight_gain_agg<-aggregate(Kenyaherds$weight_gain_change, by=list(Kenyaherds$QQNO,Kenyaherds$animal_type), FUN=mean)
+live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
+                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+#mature_weight_agg<-aggregate(Kenyaherds_interventions$mature_weight_change,
+#by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
+                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
 
@@
 {if(rbinom(1,1,adoption_rate))
-    x_breed_intervention[,colnames(breed_intervention)[2:ncol(breed_intervention)]]
+    x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]]<-
+      x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]] * 
+      breed_intervention[which(paste0("Farm_",breed_intervention$Farm)==x_breed_intervention$Scenario[i]),
+                         2:ncol(breed_intervention)]
   # implement changes in x file
 }
-  res<-run_model(x_breed_intervention,ghg_emissions)
+  breed_res<-run_model(x_breed_intervention,ghg_emissions)
+
+  animals_per_run<-rowSums(breed_res$x[paste0("N_animal_type_",1:12)])
+
+breed_milk_yield_per_head<-breed_res$y$milk_yield/animals_per_run
+breed_total_emissions_per_head<-breed_res$y$on_farm/animals_per_run
+
+```
+
+### Evaluate impact of breed intervention
+
+Assumed adoption rate, based on all farms (even the ones that already have high-performance farms and therefore don't get an 'upgrade'): `r adoption_rate*100` %
+
+```{r, fig.show="hold", out.width="50%", fig.cap="Breed intervention (left) vs. baseline (right)", message=FALSE}
+
+uncertainty::varscatter(in_var = breed_milk_yield_per_head,
+                        out_var = breed_total_emissions_per_head,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run,
+                        out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = breed_milk_yield_per_head, 
+                       out_var = breed_total_emissions_per_head,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
+                       out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernelslicerange(in_var = breed_res$y$milk_yield/animals_per_run, 
+                                 out_var = breed_res$y$on_farm/animals_per_run, 
+                                 min_in_var = round(min(breed_res$y$milk_yield/animals_per_run)),
+                                 max_in_var = round(max(breed_res$y$milk_yield/animals_per_run)), 
+                                 xlab_vars = "Total emissions given milk yield range")
+
+uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
+                                 out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run, 
+                                 min_in_var = round(min(GHG_simulation_scenarios$y$milk_yield/animals_per_run)),
+                                 max_in_var = round(max(GHG_simulation_scenarios$y$milk_yield/animals_per_run)), 
+                                 xlab_vars = "Total emissions given milk yield range")
 
@@
 ```{r}
-uncertainty::varkernel(in_var = res$x$milk_yield_5, 
-                       out_var = res$y$on_farm,
-                        xlab = "Total Milk Yield",
-                        ylab = "Total emissions")
+breed_change_in_CO2_eq_per_animal<-breed_res$y$on_farm/animals_per_run-
+  GHG_simulation_scenarios$y$on_farm/animals_per_run
 
+breed_change_in_milk_yield_per_animal<-breed_res$y$milk_yield/animals_per_run-
+  GHG_simulation_scenarios$y$milk_yield/animals_per_run
 ```

### 2021-11-14 Scenarios_report.Rmd
add link to the repo compress explinations of scenarios add new chunk with full farm_scenarios file add citations

@@
 
-We're defining the function as in the ['Greenhouse gas emissions model']() document. Please look there for more information about the original model design.
+We're defining the function as in the ['Greenhouse gas emissions model'](https://github.com/ToddRosenstock/ghg-uncertainty) document.
+Please look there for more information about the original model design.
 
@@
 
-We load the library and the data.
+Here we load the library and the data.
 The data contains ranges of possible values for the simulation.
@@
 This is quite inconvenient.
-Here, we'll develop a function that can do this job automatically.
+Here, we usep a function that can do this job automatically.
 
-The new `scenario_mc` function consists of a wrapper around the `mcSimulation` function, which is able to modify the input table according to a scenario file we provide.
+The `scenario_mc` function in the `decisionSupport` package (@R-decisionSupport) consists of a wrapper around the `mcSimulation` function.
+The function is able to modify the input table according to a scenario file we provide.
 This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
+The function is essentially the same as the `mcSimulation` function, but with an additional `scenarios` option which imports the contents of a .csv file.
 
-This function does the same as the `mcSimulation` function, but with an additional `scenario_file` that can be supplied.
-This file specifies which variables should be modified for each of the scenarios, as well as the number of model runs for for it.
-Here's an example file:
+The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for for it.
+Each column beyond the first two columns in the file contain information for a particular scenario.
+This is restricted to the parameters that should be *changed* for this scenario.
+Here's an example `scenarios` file:
 
@@
 
-Each column beyond the first two columns contains information for a particular scenario.
-This is restricted to the parameters that should be *changed* for this scenario.
-All variables that aren't specified in this table remain at their original values given in the usual input table (which is called `base_estimate` here).
-
 The first row ("Runs" in the first column) indicates the number of runs.
-
 Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table.
-The second column indicates which value should be changes, with options being "lower", "upper", "distribution" and "both".
+The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both".
 If "both" is selected, both the lower and upper bounds are modified (this should only be done for constants).
@@
 
+Here's the resulting file:
+
+```{r}
+knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE))
+
+```
+
+Here we run the model with the `scenario_mc` function.
+All variables that aren't specified in the file called in the `scenarios` option remain at their original values, given in the usual input table, called with `base_estimate`.
+
 ```{r monte_carlo_simulation, warning=FALSE}
@@
 GHG_simulation_scenarios<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
-                                 scenarios = read.csv("data/farm_scenarios.csv",fileEncoding="UTF-8-BOM"),
-                                 model_function = ghg_emissions,numberOfModelRuns=10,
+                                 scenarios = read.csv("data/farm_scenarios.csv",
+                                                      fileEncoding = "UTF-8-BOM"),
+                                 model_function = ghg_emissions,
+                                 numberOfModelRuns = 10,
                                  functionSyntax = "plainNames")

### 2021-12-20 Scenarios_report.Rmd
added herd structure (fewer) males scenarios and cleaned up a bit

--- 2021-11-13 Scenarios_report.Rmd
@@
 library(uncertainty)
+library(chillR)
 
@@
 
-```{r monte_carlo_simulation, warning=FALSE}
+```{r monte_carlo_simulation, warning=FALSE, eval=FALSE}
 
@@
 
+```
+
+```{r, eval=FALSE, include=FALSE}
+dir.create("results")
+save_temperature_scenarios(GHG_simulation_scenarios,"results","Baseline")
+```
+
+```{r, include=FALSE}
+GHG_simulation_scenarios<-load_temperature_scenarios("results","Baseline")
+class(GHG_simulation_scenarios)<-"mcSimulation"
 ```
@@
 
-# Normalizing milk yield and emissions per head
-
-```{r}
-
-animals_per_run<-rowSums(GHG_simulation_scenarios$x[paste0("N_animal_type_",1:12)])
-
-milk_yield_per_head<-GHG_simulation_scenarios$y$milk_yield/animals_per_run
-total_emissions_per_head<-GHG_simulation_scenarios$y$on_farm/animals_per_run
-
-```
-
 # Uncertainty for per-head data
@@
 
-```{r name_in_out_vars}
-in_var  <- milk_yield_per_head # "Total Milk Yield"
-out_var <- total_emissions_per_head # "Total emissions"
-```
-
 ## Milk_yield (x-axis) effect on on-farm GHG emissions (y-axis)
@@
 ```{r uncertainty_package}
-uncertainty::varscatter(in_var = in_var,
-                        out_var = out_var,
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 ```{r varkernel}
-uncertainty::varkernel(in_var = in_var, 
-                       out_var = out_var,
+uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 ```{r varkernelslicerange}
-uncertainty::varkernelslicerange(in_var = in_var, 
-                                 out_var = out_var, 
-                                 min_in_var = round(min(in_var)),
-                                 max_in_var = round(max(in_var)), 
+uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                                 out_var = GHG_simulation_scenarios$y$pc_on_farm, 
+                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                  xlab_vars = "Total emissions given milk yield range")
@@
 # Interventions
-
-Here's a wild idea: Can we simply change values in the GHG_simulation_scenarios$x table now and apply the model to the resulting table?
-We may not have to draw new values (and that actually wouldn't even be good, because this would dilute the intervention effect).
 
@@
 
-```{r}
-x_breed_intervention<-GHG_simulation_scenarios$x
+```{r, eval=FALSE}
+x_breed_intervention <- GHG_simulation_scenarios$x
 
-adoption_rate<-0.6
+adoption_rate <- 0.6
 
-for(i in 1:nrow(x_breed_intervention))
-{if(rbinom(1,1,adoption_rate))
-    x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]]<-
-      x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]] * 
-      breed_intervention[which(paste0("Farm_",breed_intervention$Farm)==x_breed_intervention$Scenario[i]),
+for (i in 1:nrow(x_breed_intervention))
+{
+  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
+    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
+      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
+      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
+                                 x_breed_intervention$Scenario[i]),
                          2:ncol(breed_intervention)]
-  # implement changes in x file
 }
-  breed_res<-run_model(x_breed_intervention,ghg_emissions)
+# run model for modified x file
+breed_res <- run_model(x_breed_intervention, ghg_emissions)
 
-  animals_per_run<-rowSums(breed_res$x[paste0("N_animal_type_",1:12)])
+```
 
-breed_milk_yield_per_head<-breed_res$y$milk_yield/animals_per_run
-breed_total_emissions_per_head<-breed_res$y$on_farm/animals_per_run
+```{r, include=FALSE, eval=FALSE}
+breed_res$call<-"multi-scenario call"
+save_temperature_scenarios(breed_res,"results","Breed_intervention")
+```
 
+```{r, include=FALSE}
+adoption_rate <- 0.6
+breed_res<-load_temperature_scenarios("results","Breed_intervention")
+class(breed_res)<-"mcSimulation"
 ```
@@
 
-uncertainty::varscatter(in_var = breed_milk_yield_per_head,
-                        out_var = breed_total_emissions_per_head,
+uncertainty::varscatter(in_var = breed_res$y$pc_milk_yield,
+                        out_var = breed_res$y$pc_on_farm,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run,
-                        out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 
-uncertainty::varkernel(in_var = breed_milk_yield_per_head, 
-                       out_var = breed_total_emissions_per_head,
+uncertainty::varkernel(in_var = breed_res$y$pc_milk_yield,
+                        out_var = breed_res$y$pc_on_farm,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 
-uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
-                       out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
+uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 
-uncertainty::varkernelslicerange(in_var = breed_res$y$milk_yield/animals_per_run, 
-                                 out_var = breed_res$y$on_farm/animals_per_run, 
-                                 min_in_var = round(min(breed_res$y$milk_yield/animals_per_run)),
-                                 max_in_var = round(max(breed_res$y$milk_yield/animals_per_run)), 
+uncertainty::varkernelslicerange(in_var = breed_res$y$pc_milk_yield,
+                                 out_var = breed_res$y$pc_on_farm, 
+                                 min_in_var = round(min(breed_res$y$pc_milk_yield)),
+                                 max_in_var = round(max(breed_res$y$pc_milk_yield)), 
                                  xlab_vars = "Total emissions given milk yield range")
 
-uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
-                                 out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run, 
-                                 min_in_var = round(min(GHG_simulation_scenarios$y$milk_yield/animals_per_run)),
-                                 max_in_var = round(max(GHG_simulation_scenarios$y$milk_yield/animals_per_run)), 
+uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                  xlab_vars = "Total emissions given milk yield range")
@@
 ```{r}
-breed_change_in_CO2_eq_per_animal<-breed_res$y$on_farm/animals_per_run-
-  GHG_simulation_scenarios$y$on_farm/animals_per_run
+breed_change_in_CO2_eq_per_animal<-breed_res$y$pc_on_farm-
+  GHG_simulation_scenarios$y$pc_on_farm
 
-breed_change_in_milk_yield_per_animal<-breed_res$y$milk_yield/animals_per_run-
-  GHG_simulation_scenarios$y$milk_yield/animals_per_run
+breed_change_in_milk_yield_per_animal<-breed_res$y$pc_milk_yield-
+  GHG_simulation_scenarios$y$pc_milk_yield
 ```
@@
 
+### Retiring unproductive bulls
+
+To simulate this scenario, we simply have to remove all animals of type 1 from the herds.
+
+```{r , warning=FALSE, eval=FALSE}
+x_nobull_intervention <- GHG_simulation_scenarios$x
+
+nobull_adoption_rate <- 1.0
+
+for (i in 1:nrow(x_nobull_intervention))
+{
+  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
+    x_nobull_intervention$N_animal_type_1[i]<-0
+}
+# run model for modified x file
+nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+nobull_res$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_res,"results","noBull_intervention")
+```
+
+```{r, include=FALSE}
+nobull_res<-load_temperature_scenarios("results","noBull_intervention")
+class(nobull_res)<-"mcSimulation"
+```
+
+#### Evaluate impact of "Retiring unproductive bulls" intervention
+
+We're assuming that all bulls >36 months old have been removed.
+
+```{r, fig.show="hold", out.width="50%", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE}
+
+uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
+                        out_var = nobull_res$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
+                        out_var = nobull_res$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernelslicerange(in_var = nobull_res$y$pc_milk_yield,
+                                 out_var = nobull_res$y$pc_on_farm, 
+                                 min_in_var = round(min(nobull_res$y$pc_milk_yield,na.rm=TRUE)),
+                                 max_in_var = round(max(nobull_res$y$pc_milk_yield,na.rm=TRUE)), 
+                                 xlab_vars = "Total emissions given milk yield range")
+
+uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
+                                 xlab_vars = "Total emissions given milk yield range")
+
+```
+
+### Reducing unproductive replacement males
+
+The idea is to raise the off-take rate of replacement males by 23% (not sure why this number).
+We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate.
+The result should be quite similar.
+
+```{r , warning=FALSE, eval=FALSE}
+x_lessReplace_intervention <- GHG_simulation_scenarios$x
+
+lessReplace_adoption_rate <- 0.23
+
+for (i in 1:nrow(x_lessReplace_intervention))
+{
+  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
+    x_lessReplace_intervention$N_animal_type_3[i]<-0
+}
+# run model for modified x file
+lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+lessReplace_res$call<-"multi-scenario call"
+save_temperature_scenarios(lessReplace_res,"results","lessReplace_intervention")
+```
+
+```{r, include=FALSE}
+lessReplace_res<-load_temperature_scenarios("results","lessReplace_intervention")
+class(lessReplace_res)<-"mcSimulation"
+```
+
+#### Evaluate impact of "Retiring unproductive bulls" intervention
+
+We're assuming that all bulls >36 months old have been removed.
+
+```{r, fig.show="hold", out.width="50%", fig.cap="Fewer replacement males intervention (left) vs. baseline (right)", message=FALSE}
+
+uncertainty::varscatter(in_var = lessReplace_res$y$pc_milk_yield,
+                        out_var = lessReplace_res$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = lessReplace_res$y$pc_milk_yield,
+                        out_var = lessReplace_res$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernelslicerange(in_var = lessReplace_res$y$pc_milk_yield,
+                                 out_var = lessReplace_res$y$pc_on_farm, 
+                                 min_in_var = round(min(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)),
+                                 max_in_var = round(max(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)), 
+                                 xlab_vars = "Total emissions given milk yield range")
+
+uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
+                                 xlab_vars = "Total emissions given milk yield range")
+
+```
+
 This document was generated using R Markdown [@R-rmarkdown].

### 2021-12-20 GHG_Emissions_Function.R

@@ -19,7 +19,7 @@ make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
 for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
 
 # run the function on the input_table
-make_variables(estimate_read_csv(input_table))
+# make_variables(estimate_read_csv(input_table))
 
 # Generate the GHG emissions function ####
 ghg_emissions<-function(x, varnames){
@@ -273,11 +273,23 @@ ghg_emissions<-function(x, varnames){
 
   on_farm<-sum(enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq)+feed_CO2
   
+  total_animals<-sum(sapply(1:12,function(x) eval(parse(text=paste0("N_animal_type_",x)))))
+
   return(list(enteric_CH4=enteric_CH4,
               milk_yield=sum(total_milk),
               enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
               mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
               mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
               feed_CO2=feed_CO2,
-              on_farm=on_farm))
+              on_farm=on_farm,
+              pc_enteric_CH4=enteric_CH4/total_animals,
+              pc_milk_yield=sum(total_milk)/total_animals,
+              pc_enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq)/total_animals,
+              pc_mm_CH4_CO2eq=sum(mm_CH4_CO2eq)/total_animals,
+              pc_mm_N2O_CO2eq=sum(mm_N2O_CO2eq)/total_animals,
+              pc_feed_CO2=feed_CO2/total_animals,
+              pc_on_farm=on_farm/total_animals
+              
+              
+              ))
 }

### 2021-12-20 Scenarios_report.Rmd

@@
 
+Here's the resulting file:
+
+```{r}
+knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE))
+
+```
+
+Here we run the model with the `scenario_mc` function.
+All variables that aren't specified in the file called in the `scenarios` option remain at their original values, given in the usual input table, called with `base_estimate`.
+
 ```{r monte_carlo_simulation, warning=FALSE, eval=FALSE}

### 2021-12-21 Scenarios_report.Rmd
adding some evaluations to the scenarios

@@
 ```{r}
-knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE))
+knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[,1:5])
 
@@
 
-uncertainty::varkernelslicerange(in_var = breed_res$y$pc_milk_yield,
-                                 out_var = breed_res$y$pc_on_farm, 
-                                 min_in_var = round(min(breed_res$y$pc_milk_yield)),
-                                 max_in_var = round(max(breed_res$y$pc_milk_yield)), 
-                                 xlab_vars = "Total emissions given milk yield range")
-
-uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-                                 xlab_vars = "Total emissions given milk yield range")
+# uncertainty::varkernelslicerange(in_var = breed_res$y$pc_milk_yield,
+#                                  out_var = breed_res$y$pc_on_farm, 
+#                                  min_in_var = round(min(breed_res$y$pc_milk_yield)),
+#                                  max_in_var = round(max(breed_res$y$pc_milk_yield)), 
+#                                  xlab_vars = "Total emissions given milk yield range")
+# 
+# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
+#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
+#                                  xlab_vars = "Total emissions given milk yield range")
 
@@
 
-uncertainty::varkernelslicerange(in_var = nobull_res$y$pc_milk_yield,
-                                 out_var = nobull_res$y$pc_on_farm, 
-                                 min_in_var = round(min(nobull_res$y$pc_milk_yield,na.rm=TRUE)),
-                                 max_in_var = round(max(nobull_res$y$pc_milk_yield,na.rm=TRUE)), 
-                                 xlab_vars = "Total emissions given milk yield range")
-
-uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-                                 xlab_vars = "Total emissions given milk yield range")
+# uncertainty::varkernelslicerange(in_var = nobull_res$y$pc_milk_yield,
+#                                  out_var = nobull_res$y$pc_on_farm, 
+#                                  min_in_var = round(min(nobull_res$y$pc_milk_yield,na.rm=TRUE)),
+#                                  max_in_var = round(max(nobull_res$y$pc_milk_yield,na.rm=TRUE)), 
+#                                  xlab_vars = "Total emissions given milk yield range")
+# 
+# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
+#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
+#                                  xlab_vars = "Total emissions given milk yield range")
 
@@
 
-uncertainty::varkernelslicerange(in_var = lessReplace_res$y$pc_milk_yield,
-                                 out_var = lessReplace_res$y$pc_on_farm, 
-                                 min_in_var = round(min(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)),
-                                 max_in_var = round(max(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)), 
-                                 xlab_vars = "Total emissions given milk yield range")
+# uncertainty::varkernelslicerange(in_var = lessReplace_res$y$pc_milk_yield,
+#                                  out_var = lessReplace_res$y$pc_on_farm, 
+#                                  min_in_var = round(min(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)),
+#                                  max_in_var = round(max(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)), 
+#                                  xlab_vars = "Total emissions given milk yield range")
+# 
+# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
+#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
+#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
+#                                  xlab_vars = "Total emissions given milk yield range")
 
-uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-                                 xlab_vars = "Total emissions given milk yield range")
+```
+
+# Comparing the scenarios
+
+Below we compare the density distributions for the scenarios.
+This required loading some helper functions that contain modified versions of some functions from the `uncertainty` package (`multi_data_varkernelslicerange` and `data_varkernelslicerange`).
+
+```{r uncertainty_evaluation_functions, include=FALSE}
+# Generate the GHG emissions function 
+source("uncertainty_evaluation.R", local = knitr::knit_global())
+```
+
+The plot below shows that in terms of per-capita emissions, the difference between the scenarios is not impressive.
+
+```{r}
+per_capita_scenarios<-list(list(scenario="Baseline",
+                     in_var=GHG_simulation_scenarios$y$pc_milk_yield,
+                     out_var=GHG_simulation_scenarios$y$pc_on_farm),
+                list(scenario="Less male replacement",
+                     in_var=lessReplace_res$y$pc_milk_yield,
+                     out_var=lessReplace_res$y$pc_on_farm),
+                list(scenario="No bulls",
+                     in_var=nobull_res$y$pc_milk_yield,
+                     out_var=nobull_res$y$pc_on_farm),
+                list(scenario="Breed upgrade",
+                     in_var=breed_res$y$pc_milk_yield,
+                     out_var=breed_res$y$pc_on_farm))
+
+per_capita_emissions<-multi_data_varkernelslicerange(per_capita_scenarios)
+
+library(ggplot2)
+
+ggplot(data = per_capita_emissions$slice_values, aes(x = y, y = z, color = scenario)) +
+  geom_line() + theme_bw(base_size = 12) +
+  xlab(expression(Per~capita~emissions~(kg~CO[2]-eq~year^-1))) +
+  ylab("Density") +
+  ggtitle("Per capita emissions") +
+  scale_color_discrete(name = "Scenario")
+
+```
+
+We can also look at the distribution of emissions per kg of milk yield.
+The plot below shows the cumulative probability distribution for this metric.
+The differences are a bit more visible here, but still fairly minor.
+
+```{r}
+
+ghg_per_milk <-
+  rbind(
+    data.frame(
+      scenario = "Baseline",
+      milk_yield = GHG_simulation_scenarios$y$pc_milk_yield,
+      GHG = GHG_simulation_scenarios$y$pc_on_farm
+    ),data.frame(
+      scenario = "Less male replacement",
+      milk_yield = lessReplace_res$y$pc_milk_yield,
+      GHG = lessReplace_res$y$pc_on_farm
+    ),data.frame(
+      scenario = "No bulls",
+      milk_yield = nobull_res$y$pc_milk_yield,
+      GHG = nobull_res$y$pc_on_farm
+    ),data.frame(
+      scenario = "Breed upgrade",
+      milk_yield = breed_res$y$pc_milk_yield,
+      GHG = breed_res$y$pc_on_farm
+    ))
+ghg_per_milk[,"ghg_per_milk"]<-ghg_per_milk$GHG/ghg_per_milk$milk_yield
+
+ggplot(data = ghg_per_milk, aes(x = ghg_per_milk, color = scenario)) +
+  stat_ecdf() + xlim(0, 10) + theme_bw(base_size = 12) + ggtitle("Emissions per kg of milk (truncated at 10 kg)") + xlab(expression(Emissions~per~kg~milk~(kg~CO[2]-eq~year^-1))) + ylab("Cumulative probability")
 

### 2021-12-21 uncertainty_evaluation.R

@@ -1,0 +1,102 @@
+# uncertainty package data processing functions
+
+data_varkernelslicerange<-
+function (in_var, out_var, min_in_var, max_in_var, scenario="Scenario1",
+          xlab_vars = "Outcome variable dist. given influence variable",
+          show_plot = FALSE) 
+{
+  if (!requireNamespace("ggplot2", quietly = TRUE)) {
+    stop("Package \"ggplot2\" needed for this function to work. Please install it.", 
+         call. = FALSE)
+  }
+  if (!requireNamespace("stats", quietly = TRUE)) {
+    stop("Package \"stats\" needed for this function to work. Please install it.", 
+         call. = FALSE)
+  }
+  in_outdata <- in_out <- NULL
+  assertthat::validate_that(length(in_var) == length(out_var), 
+                            msg = "\"in_var\" and \"out_var\" are not equal lengths.")
+  assertthat::validate_that(is.numeric(in_var), msg = "\"in_var\" is not numeric.")
+  assertthat::validate_that(is.numeric(min_in_var), msg = "\"min_in_var\" is not numeric.")
+  assertthat::validate_that(is.numeric(max_in_var), msg = "\"max_in_var\" is not numeric.")
+  assertthat::validate_that(is.numeric(out_var), msg = "\"out_var\" is not numeric.")
+  in_out <- as.data.frame(cbind(in_var, out_var))
+  in_outdata <- in_out[stats::complete.cases(in_out), ]
+  assertthat::see_if(length(in_out) == length(in_outdata), 
+                     msg = "Rows with NA were removed.")
+  in_outkernel <- MASS::kde2d(in_outdata$in_var, in_outdata$out_var, 
+                              n = 100)
+  lbound <- which(in_outkernel$x == min(in_outkernel$x[which(in_outkernel$x > 
+                                                               min_in_var)]))
+  rbound <- which(in_outkernel$x == max(in_outkernel$x[which(in_outkernel$x <= 
+                                                               max_in_var)]))
+  if (show_plot) {graphics::plot(in_outkernel$y, rowMeans(in_outkernel$z[, 
+                                                         lbound:rbound]), type = "l", col = "seagreen", 
+                 lwd = 2, xlab = paste(xlab_vars, as.character(min_in_var), 
[…]

### 2022-05-10 Scenarios_report.Rmd
trying to get varslice_resample to work

@@
                                  xlab_vars = "Total emissions given milk yield range")
+```
+
+# Probability of an increase in milk yield being associated with an increase in GHG emission intensity
+
+```{r}
+
+GHG_simulation_scenarios$y[,"emission_intensity"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
+
+GHG_sim_scen_y<-GHG_simulation_scenarios$y
+
+GHG_sim_scen_y<-GHG_sim_scen_y[which(!is.infinite(GHG_sim_scen_y$emission_intensity)),]
+
+uncertainty::varkernelslicerange(in_var = GHG_sim_scen_y$pc_milk_yield,
+                                 out_var = GHG_sim_scen_y$emission_intensity, 
+                                 min_in_var = round(min(GHG_sim_scen_y$pc_milk_yield)),
+                                 max_in_var = round(max(GHG_sim_scen_y$pc_milk_yield)), 
+                                 xlab_vars = "Emission intensity (XX/L) given milk yield range")
+
+  in_var<-GHG_sim_scen_y$pc_milk_yield
+  out_var<-GHG_sim_scen_y$emission_intensity
+baseline_milk_yield <- min(in_var)
+sliceb <-
+  varslice_resample(in_var, out_var, expectedin_var = baseline_milk_yield, n_runs = 100000)
+more_milk <- baseline_milk_yield
+results <- data.frame(milk_yield = more_milk, probability = 0.5)
+
+# the following isn't working yet
+# I suspect this is because of an error in the varslice_resample function (and also the varkernelslice functions).
+# I suspect the problem is in   Relative_probability <- in_outkernel$z[, expectedin_var]
+# There may not be an expectedin_var'th column - this should be the column that corresponds to the x value expectedin_var
+
+while (more_milk < max(in_var))
+{
+  more_milk <- more_milk + 10
+  slice_more_milk <-
+    varslice_resample(in_var,
+                      out_var,
+                      expectedin_var = more_milk,
+                      n_runs = 100000)
+  results <-
+    rbind(results, c(more_milk, sum(sliceb > slice_more_milk) / length(sliceb)))
+}
+
 ```

### 2022-05-11 Scenarios_report.Rmd
scenarios - added first version of probability of emissions reduction for milk yield increases

@@
 library(chillR)
+library(raster)
 
@@
 
-# the following isn't working yet
-# I suspect this is because of an error in the varslice_resample function (and also the varkernelslice functions).
-# I suspect the problem is in   Relative_probability <- in_outkernel$z[, expectedin_var]
-# There may not be an expectedin_var'th column - this should be the column that corresponds to the x value expectedin_var
+#devtools::install_github("CWWhitney/uncertainty")
+#require(uncertainty)
 
-while (more_milk < max(in_var))
-{
-  more_milk <- more_milk + 10
-  slice_more_milk <-
+?varslice_resample
+
+milk_yields_to_sample<-seq(100,10000,100)
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(in_var))]
+slices<-list()
+
+# slice through the distribution at every 100 L milk yield per cow
+
+for(i in 1:length(milk_yields_to_sample))
+{slices[[i]]<-
   varslice_resample(in_var,
                       out_var,
-                      expectedin_var = more_milk,
-                      n_runs = 100000)
-  results <-
-    rbind(results, c(more_milk, sum(sliceb > slice_more_milk) / length(sliceb)))
+                      expectedin_var = milk_yields_to_sample[i],
+                      n_runs = 100000,
+  out_var_sampling = (max(out_var) - min(out_var))/10000)$resample
  }
+
+#decreased_emissions_probability <-
+#  data.frame(Base_yield = milk_yields_to_sample)#
+#
+#for (i in 1:(length(milk_yields_to_sample) - 1))
+#{
+#  for (j in (i + 1):length(milk_yields_to_sample))
+#    decreased_emissions_probability[i, j] <-
+#      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
+#  
+#}
+#colnames(decreased_emissions_probability)[2:ncol(decreased_emissions_probability)] <-
+#  paste0("new", milk_yields_to_sample[2:length(milk_yields_to_sample)])
+
+decreased_emissions_probability2 <-
+  data.frame(Base_yield = NA,new_yield=NA,probability=NA)
+
+for (i in 1:(length(milk_yields_to_sample) - 1))
+{
+  for (j in (i + 1):length(milk_yields_to_sample))
+    {decreased_emissions_probability2[nrow(decreased_emissions_probability2)+1,"Base_yield"]<-milk_yields_to_sample[i]
+     decreased_emissions_probability2$new_yield[nrow(decreased_emissions_probability2)]<-milk_yields_to_sample[j]
+     decreased_emissions_probability2$probability[nrow(decreased_emissions_probability2)] <-
+      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
+
+}
+}
+
+decreased_emissions_probability2<-decreased_emissions_probability2[which(!is.na(decreased_emissions_probability2$Base_yield)),]
+
+decreased_emissions_probability2[,"yield_increase"]<-decreased_emissions_probability2$new_yield-decreased_emissions_probability2$Base_yield
+
+ggplot(data = decreased_emissions_probability2, aes(Base_yield, yield_increase, z =
+                                                      probability)) + geom_contour_filled() +
+  ggtitle("Probability that emission reductions occurred") + 
+  xlab("Baseline milk yield per head and year (kg)") +
+  ylab("Milk yield increase per head and year (kg)") +
+  stat_contour(breaks = c(0.9),col="black")
 

### 2022-05-11 Scenarios_report.Rmd
scenario_report update

--- 2021-12-21 Scenarios_report.Rmd
@@
                                  xlab_vars = "Total emissions given milk yield range")
+```
+
+# Probability of an increase in milk yield being associated with an increase in GHG emission intensity
+
+```{r}
+
+GHG_simulation_scenarios$y[,"emission_intensity"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
+
+GHG_sim_scen_y<-GHG_simulation_scenarios$y
+
+GHG_sim_scen_y<-GHG_sim_scen_y[which(!is.infinite(GHG_sim_scen_y$emission_intensity)),]
+
+uncertainty::varkernelslicerange(in_var = GHG_sim_scen_y$pc_milk_yield,
+                                 out_var = GHG_sim_scen_y$emission_intensity, 
+                                 min_in_var = round(min(GHG_sim_scen_y$pc_milk_yield)),
+                                 max_in_var = round(max(GHG_sim_scen_y$pc_milk_yield)), 
+                                 xlab_vars = "Emission intensity (XX/L) given milk yield range")
+
+in_var = GHG_sim_scen_y$pc_milk_yield
+out_var = GHG_sim_scen_y$emission_intensity
+
+milk_yields_to_sample<-seq(100,10000,100)
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(in_var))]
+slices<-list()
+
+# slice through the distribution at every 100 L milk yield per cow
+
+for(i in 1:length(milk_yields_to_sample))
+{slices[[i]]<-
+  varslice_resample(in_var,
+                      out_var,
+                      expectedin_var = milk_yields_to_sample[i],
+                      n_samples = 100000,
+  out_var_sampling = (max(out_var) - min(out_var))/10000)$resample
+ }
+
+decreased_emissions_probability <-
+  data.frame(Base_yield = NA,new_yield=NA,probability=NA)
+
+for (i in 1:(length(milk_yields_to_sample) - 1))
+{
+  for (j in (i + 1):length(milk_yields_to_sample))
+    {decreased_emissions_probability[nrow(decreased_emissions_probability)+1,"Base_yield"]<-milk_yields_to_sample[i]
+     decreased_emissions_probability$new_yield[nrow(decreased_emissions_probability)]<-milk_yields_to_sample[j]
+     decreased_emissions_probability$probability[nrow(decreased_emissions_probability)] <-
+      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
+
+}
+}
+
+decreased_emissions_probability<-decreased_emissions_probability[which(!is.na(decreased_emissions_probability$Base_yield)),]
+
+decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase, z =
+                                                      probability)) + geom_contour_filled() +
+  ggtitle("Probability that emission intensity reductions occurred") + 
+  xlab("Baseline milk yield per head and year (kg)") +
+  ylab("Milk yield increase per head and year (kg)") +
+  stat_contour(breaks = c(0.9),col="black")
+
 ```

### 2022-05-12 Manuscript_analyses.Rmd
started a new rmd for manuscript. some things not working yet.

--- 2022-05-10 Scenarios_report.Rmd
@@
 ---
-title: "Greenhouse gas emissions model - making scenarios"
-author: "Cory Whitney and Eike Luedeling"
-bibliography:
- - references/packages.bib
- - references/papers.bib
+title: "Manuscript_analyses"
+author: "Eike Luedeling, Cory Whitney, Todd Rosenstock"
+date: "5/12/2022"
 output: html_document
@@
 
-We're defining the function as in the ['Greenhouse gas emissions model'](https://github.com/ToddRosenstock/ghg-uncertainty) document.
-Please look there for more information about the original model design.
+# Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya
 
-```{r setup, include=FALSE}
+## Set up
+
+Install and load required packages
+
+```{r setup, warning=FALSE, include=FALSE}
+
 knitr::opts_chunk$set(echo = TRUE)
-```
 
-```{r, warning=FALSE, include=FALSE}
 #install and load packages
+# devtools::install_github("CWWhitney/uncertainty")
 library(DiagrammeR)
-# devtools::install_github("eikeluedeling/decisionSupport")
 library(decisionSupport)
@@
 library(rmarkdown)
-# devtools::install_github("CWWhitney/uncertainty")
 library(uncertainty)
 library(chillR)
+library(raster)
 
@@
 
-Here we load the library and the data.
-The data contains ranges of possible values for the simulation.
-Each variable is represented with a confidence interval range and a specification of the distribution shape.
+```{r, warning=FALSE, eval=FALSE}
+devtools::install_github("CWWhitney/uncertainty")
+library(DiagrammeR)
+library(decisionSupport)
+library(ggplot2)
+library(kableExtra)
+library(magrittr)
+library(plyr)
+library(rmarkdown)
+library(uncertainty)
+library(chillR)
+library(raster)
+```
 
-```{r load_data, include=FALSE}
+# Load emission calculation function
+
+The calculations are based on the IPCC xxx equation (describe and cite here)
+
+```{r ghg_emissions_function}
+# Generate the GHG emissions function 
+source("GHG_emissions_function.R", local = knitr::knit_global())
+```
+
+This equation takes a wide range of inputs.
+Many of these are usually not known precisely, yet the classic IPCC equation does not account for potential errors and uncertainties.
+Here, we express this uncertainty by defining distributions for each variable.
+The following table shows all the variables, as well as examples of distributions that may describe their quantities.
+Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals.
+
+```{r load_data}
 
@@
 
-```{r ghg_emissions_function, include=FALSE}
-# Generate the GHG emissions function 
-source("GHG_emissions_function.R", local = knitr::knit_global())
+## Monte Carlo simulation
+
+With the model equation and the input table defined above, we can run a Monte Carlo simulation that returns a probability distribution for the greenhouse gas emissions resulting from the particular herd that is specified in the input table.
+
+```{r, warning=FALSE,eval=FALSE}
+GHG_emissions<-mcSimulation(estimate=estimate_read_csv(input_table),
+                            model_function=ghg_emissions,
+                            numberOfModelRuns=1000,
+                            functionSyntax="plainNames")
 ```
 
-## Running the model with scenarios
+## Running a Monte Carlo simulation with scenarios
 
-The normal `mcSimulation` function takes random draws for each of the input variables from the distributions specified in the input table.
+The `mcSimulation` function takes random draws for each of the input variables from the distributions specified in the input table.
 The distributions that are used are the same for every run of the Monte Carlo simulation.
-Sometimes it's desirable to specify particular scenarios *a priori*, e.g.
-to simulate outcomes for particular farm types, climate scenarios etc.
+This may be desirable in homogeneous populations where all combinations of random values are plausible, but it often misses the mark where populations are composed of particular member types with specific characteristics.
+To account for populations that are heterogeneous, which is the case for livestock-keeping households in Kenya, we use the `scenario_mc` function, which allows specifying particular subgroups.
+Defining such scenarios can be desirable when it's necessary to simulate outcomes for particular farm types, climate scenarios etc.
 To do this, we would normally have to define separate input tables and run separate simulations.
 This is quite inconvenient.
-Here, we usep a function that can do this job automatically.
+Here, we use a function that can do this job automatically.
 
@@
 This is restricted to the parameters that should be *changed* for this scenario.
+
 Here's an example `scenarios` file:
@@
 If "both" is selected, both the lower and upper bounds are modified (this should only be done for constants).
+
+We define subgroups based on the populations of household that was described by a household survey in Kenya.
+Each of the XXX encountered in this survey is converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.
+
+## Household-based scenarios
+
+The information for the household is extracted from a spreadsheet that contains all the survey responses.
+Extracting all relevant information is a slightly complex task, which is accomplished by the following code.
 
@@
 
-Here's the resulting file:
+Here's the resulting file (only the first 3 farms are shown, and only the first 20 rows):
 
 ```{r}
-knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[,1:5])
+knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5])
 
@@
 
-Here we run the model with the `scenario_mc` function.
-All variables that aren't specified in the file called in the `scenarios` option remain at their original values, given in the usual input table, called with `base_estimate`.
+The first few rows in this table adjust the herd composition parameters, so that the scenario herd corresponds to the herd of the specific household.
+Further parameters that are adjusted include the milk yield per cow and various other factors.
+
+Now we can run the Monte Carlo simulation with these scenarios.
 
@@
 
-Here we plot the distributions of the various outputs from the model.
+Can can now plot distributions of the various outputs from the model, using the `plot_distributions` function.
 
@@
                    old_names = "enteric_CH4",
-                   new_names = "Total Methane Emissions from Livestock by Enteric Fermentation") + 
+                   new_names = "CH4 Emissions from Enteric Fermentation per Farm") + 
   theme(axis.title.x=element_blank(),
@@
 plot_layout(guides = "collect", design = layout)
+
+ggsave("Fig_1_GHG_simulation_outputs.png")
 ```
 
-# Uncertainty for per-head data
+# THIS STILL NEEDS IMPROVEMENT - e.g. units are missing, and plot titles don't fit (use concise symbols with subscripts etc.)
+
+## Possible errors resulting from input uncertainty
+
+To illustrate the possible influence of input uncertainty on estimates of total farm emissions, we show the example of farm #5 from the household survey.
+This farm has 6 head of cattle, including 3 lactating cows.
+We simulate 1000 versions of this farm.
+
+```{r, warning=FALSE}
+
+scenarios<-read.csv("data/farm_scenarios.csv", fileEncoding = "UTF-8-BOM")
+farm_5<-scenarios[,c(1,2,7)]
+
+farm5_emissions<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                             scenarios = farm_5,
+                             model_function = ghg_emissions,
+                             numberOfModelRuns = 1000,
+                             functionSyntax = "plainNames")
+
+plot_distributions(mcSimulation_object = farm5_emissions,
+                   vars = "on_farm",
+                   method = "boxplot_density",
+                   old_names = "on_farm",
+                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
+  theme(axis.title.y=element_blank())
+
+ggsave("Fig_2_uncertainty_illustration.png")
+
+```
+
+# NEED TO ADD THE USUAL PRECISE VALUE HERE FOR COMPARISON (also need to add units and possible fix up other things)
+
+This
+
+```{r}
+farm5_emissions$y[,"CO2eq_per_milk"]<-farm5_emissions$y$pc_on_farm/farm5_emissions$y$pc_milk_yield
+
+ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") +
+    xlab(expression(Emission~density~of~milk~production~(XX~CO[2]-eq~kg^-1)))
+
+ggsave("Fig_3_emission_density_farm5.png")
+
+```
+
+# UNIT NEEDS TO BE ADDED HERE (and checked for milk, though this is probably right)
+# ALSO need to add the usual precise value estimate
+
+# Population-wide emission density of milk production
+
+Since the emission density of each household thus involves considerable uncertainty, such uncertainty should also be expected at the population level.
+
+We can illustrate this uncertainty using a scatter plot. The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.
+
+```{r}
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+```
+
+# NICE PLOT, but would be nicer still if this was also a ggplot figure. Kinda inconsistent if it isn't. Also can't save it with ggsave, so will need another way.
 
@@
 
-```{r uncertainty_package}
+```{r}
 uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
@@
 ```
+
+An indicator that is often of interest is the ratio between these two variables: total emissions divided by the milk yield.
+We'll calculate this next.
+We then have to remove all households that don't produce any milk, because this ratio would be infinity for them.
+
+```{r}
+GHG_simulation_scenarios$y[,"CO2eq_per_milk"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
+GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[is.finite(GHG_simulation_scenarios$y$CO2eq_per_milk),]
+
+```
+
+Now we can plot the emissions intensity vs. the milk yield.
+
+```{r}
+
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
+```
+
+This can also be illustrated by a density surface.
+
 ```{r varkernel}
 uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
 ```
+
+# THIS LOOKS BAD, BECAUSE OF Low-Milk-high-emissions households. Would be good to break the axis - any ideas?
+
+The milk yield per capita is sometimes proposed as an indicator of greenhouse gas emissions.
+We can now examine the merits of this indicator.
+Essentially, its usefulness hinges on its relationship with greenhouse gas emission intensity.
+We can investigate this by taking slices through the density surface.
 
@@
 ```{r varkernelslicerange}
-uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                                 out_var = GHG_simulation_scenarios$y$pc_on_farm, 
-                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-                                 xlab_vars = "Total emissions given milk yield range")
-```
+slice<-uncertainty::varslice_resample(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                                 out_var = GHG_simulation_scenarios$y$CO2eq_per_milk, 
+                                 expectedin_var=2000,
+                           n = 1000,
+                           n_samples = 1000,
+                           out_var_sampling = (max(out_var)-min(out_var))/1000)
 
-# Probability of an increase in milk yield being associated with an increase in GHG emission intensity
-
-```{r}
-
-GHG_simulation_scenarios$y[,"emission_intensity"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
-
-GHG_sim_scen_y<-GHG_simulation_scenarios$y
-
-GHG_sim_scen_y<-GHG_sim_scen_y[which(!is.infinite(GHG_sim_scen_y$emission_intensity)),]
-
-uncertainty::varkernelslicerange(in_var = GHG_sim_scen_y$pc_milk_yield,
-                                 out_var = GHG_sim_scen_y$emission_intensity, 
-                                 min_in_var = round(min(GHG_sim_scen_y$pc_milk_yield)),
-                                 max_in_var = round(max(GHG_sim_scen_y$pc_milk_yield)), 
-                                 xlab_vars = "Emission intensity (XX/L) given milk yield range")
-
-  in_var<-GHG_sim_scen_y$pc_milk_yield
-  out_var<-GHG_sim_scen_y$emission_intensity
-baseline_milk_yield <- min(in_var)
-sliceb <-
-  varslice_resample(in_var, out_var, expectedin_var = baseline_milk_yield, n_runs = 100000)
-more_milk <- baseline_milk_yield
-results <- data.frame(milk_yield = more_milk, probability = 0.5)
-
-# the following isn't working yet
-# I suspect this is because of an error in the varslice_resample function (and also the varkernelslice functions).
-# I suspect the problem is in   Relative_probability <- in_outkernel$z[, expectedin_var]
-# There may not be an expectedin_var'th column - this should be the column that corresponds to the x value expectedin_var
-
-while (more_milk < max(in_var))
-{
-  more_milk <- more_milk + 10
-  slice_more_milk <-
-    varslice_resample(in_var,
-                      out_var,
-                      expectedin_var = more_milk,
-                      n_runs = 100000)
-  results <-
-    rbind(results, c(more_milk, sum(sliceb > slice_more_milk) / length(sliceb)))
-}
+graphics::plot(x = slice$slice$Output_values, 
+                 y = slice$slice$Relative_probability, 
+                 type = "l", 
+                 ylab = "Relative probability", 
+                 xlab = "Emissions intensity", 
+                 col = "seagreen", 
+                 lwd = 2)
 
@@
 
-# Interventions
-
-To clearly characterize the impact of interventions, we should compare their impact with a counterfactual, i.e.
-with the same farms without the intervention.
-While this is usually impossible in practice, it is easy with a model.
-We just have to run the simulation again with identical inputs except for those that are affected by the intervention.
-
-We already simulated emissions for a large population of farms, and we can now apply our interventions to this population by modifying certain parameters of the input tables.
-What we don't want now, however, is to randomize all values again, since this would add additional noise to the results, which would mask the signal caused by the interventions.
-
-We now first need a function that can apply our greenhouse gas emissions model to a pre-defined set of input values.
-We can find syntax that can achieve this in the mcSimulation function.
-Here it is (with slight modifications):
-
-```{r}
-run_model<-function(x_data,model)
-  {
-  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
-  model_function_ <- function(x) {
-
-    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
-    eval(expr = body(model), envir = e)
-  }
-  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
-                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
-  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
-  returnObject$call <- match.call()
-  class(returnObject) <- cbind("mcSimulation", class(returnObject))
-  return(returnObject)
-}
-
-```
-
-The following code can precisely reproduce the results of the previous analysis:
-
-```{r, eval=FALSE}
-res<-run_model(GHG_simulation_scenarios$x,ghg_emissions)
-```
-
-To simulate interventions, we can now feed this function with modified x_data data.frames.
-
-## Breed intervention
-
-Introduction of higher-performance breeds may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.
-
-To retain the farm-specific performance levels obtained from the farm survey, we simulate these impacts in a relative manner.
-For this, we first need to determine the performance and characteristics of each of the breeds.
-We'll base this on the farm survey data (`Kenyaherds` dataset from above).
-
-```{r}
-cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
-
-ggplot(cows, aes(factor(breed),milk_yield)) +
-  geom_violin(draw_quantiles=c(0.5),fill="light green") +
-  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
-
-milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
-milk_summary<-milk_summary[,c(1,4,2)]
-colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
-milk_summary
-
-```
-
-This evaluation shows that the difference in median milk yields is relatively small, but that the milk yield distributions show a lot of variation that is apparently not explained by the breed.
-It is also apparent that the number of animals these estimates are based on varies considerably.
-Two breeds make up `r round(sum(milk_summary$n[1:2])/sum(milk_summary$n)*100,1)` % of the entire population, with others having very few animals, which means that the medians may be somewhat unreliable (especially for the single animal of breed 7).
-
-Let's define the intervention as follows:
-
-All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences.
-This means that milk yields are multiplied by the following factors:
-
-```{r}
-milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))
-
-milk_summary             
-```
-
-A change of the breed also causes changes in weight, mature weight and weight gain.
-
-```{r}
-live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
-colnames(live_weights)<-c("animal_type","breed","median_live_weight")
-live_weights[,"Intervention_factor"]<-NA
-for(i in 1:nrow(live_weights))
-  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
-     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]
-
-# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.
-
-mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
-colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
-mature_weights[,"Intervention_factor"]<-NA
-for(i in 1:nrow(mature_weights))
-  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
-     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]
-
-# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1
-
-weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
-colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
-weight_gains[,"Intervention_factor"]<-NA
-for(i in 1:nrow(weight_gains))
-  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
-     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
-weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1
-
-```
-
-Now we can modify the input table accordingly.
-
-go through all the farms. Determine the 'improvement' coefficients (based on herd structure)
-
-```{r}
-farms<-unique(Kenyaherds$QQNO)
-breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)
-
-Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
-                                        "system",
-                                        "animal_type",
-                                        "breed",
-                                        "live_weight_LW",
-                                        "mature_weight_MW",
-                                        "weight_gain_WG",
-                                        "milk_yield",
-                                        "dig_energy",
-                                        "CP.")]
-
-Kenyaherds_interventions[,"milk_change"]<-0
-cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
-Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )
-
-milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)
-
-breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% milk_change[,1])]<-milk_change[,2]
-breed_intervention$milk_yield_5<-breed_intervention$milk_yield_4
-breed_intervention$milk_yield_6<-breed_intervention$milk_yield_4
-
-for(i in 1:nrow(Kenyaherds_interventions))
-  {Kenyaherds_interventions[i,"live_weight_change"]<-
-    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
-                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
-#   Kenyaherds_interventions[i,"mature_weight_change"]<-
-#      1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
-#                                                 mature_weights$breed==Kenyaherds_interventions$breed[i])]
-   Kenyaherds_interventions[i,"weight_gain_change"]<-
-     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
-                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
-}
-
-live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
-                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
-#mature_weight_agg<-aggregate(Kenyaherds_interventions$mature_weight_change,
-#by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
-weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
-                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
-
-colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
-#colnames(mature_weight_agg)<-c("farm","animal_type","conv_factor")
-colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")
-
-for(i in 1:nrow(live_weight_agg))
-  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
-                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
-    live_weight_agg$conv_factor[i]
-
-#for(i in 1:nrow(mature_weight_agg))
-#  breed_intervention[which(breed_intervention$Farm==mature_weight_agg$farm[i]),
-#                     paste0("mature_weight_LW_",mature_weight_agg$animal_type[i])]<-
-#    mature_weight_agg$conv_factor[i]
-
-for(i in 1:nrow(weight_gain_agg))
-  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
-                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
-    weight_gain_agg$conv_factor[i]
-
-breed_intervention[is.na(breed_intervention)]<-1
-
-```
-
-go through the x file
-for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors
-
-```{r, eval=FALSE}
-x_breed_intervention <- GHG_simulation_scenarios$x
-
-adoption_rate <- 0.6
-
-for (i in 1:nrow(x_breed_intervention))
-{
-  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
-    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
-      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
-      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
-                                 x_breed_intervention$Scenario[i]),
-                         2:ncol(breed_intervention)]
-}
-# run model for modified x file
-breed_res <- run_model(x_breed_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-breed_res$call<-"multi-scenario call"
-save_temperature_scenarios(breed_res,"results","Breed_intervention")
-```
-
-```{r, include=FALSE}
-adoption_rate <- 0.6
-breed_res<-load_temperature_scenarios("results","Breed_intervention")
-class(breed_res)<-"mcSimulation"
-```
-
-### Evaluate impact of breed intervention
-
-Assumed adoption rate, based on all farms (even the ones that already have high-performance farms and therefore don't get an 'upgrade'): `r adoption_rate*100` %
-
-```{r, fig.show="hold", out.width="50%", fig.cap="Breed intervention (left) vs. baseline (right)", message=FALSE}
-
-uncertainty::varscatter(in_var = breed_res$y$pc_milk_yield,
-                        out_var = breed_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = breed_res$y$pc_milk_yield,
-                        out_var = breed_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-# uncertainty::varkernelslicerange(in_var = breed_res$y$pc_milk_yield,
-#                                  out_var = breed_res$y$pc_on_farm, 
-#                                  min_in_var = round(min(breed_res$y$pc_milk_yield)),
-#                                  max_in_var = round(max(breed_res$y$pc_milk_yield)), 
-#                                  xlab_vars = "Total emissions given milk yield range")
-# 
-# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
-#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-#                                  xlab_vars = "Total emissions given milk yield range")
-
-```
-
-```{r}
-breed_change_in_CO2_eq_per_animal<-breed_res$y$pc_on_farm-
-  GHG_simulation_scenarios$y$pc_on_farm
-
-breed_change_in_milk_yield_per_animal<-breed_res$y$pc_milk_yield-
-  GHG_simulation_scenarios$y$pc_milk_yield
-```
-
-## Feed
-
-## Herd structure
-
-### Retiring unproductive bulls
-
-To simulate this scenario, we simply have to remove all animals of type 1 from the herds.
-
-```{r , warning=FALSE, eval=FALSE}
-x_nobull_intervention <- GHG_simulation_scenarios$x
-
-nobull_adoption_rate <- 1.0
-
-for (i in 1:nrow(x_nobull_intervention))
-{
-  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
-    x_nobull_intervention$N_animal_type_1[i]<-0
-}
-# run model for modified x file
-nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-nobull_res$call<-"multi-scenario call"
-save_temperature_scenarios(nobull_res,"results","noBull_intervention")
-```
-
-```{r, include=FALSE}
-nobull_res<-load_temperature_scenarios("results","noBull_intervention")
-class(nobull_res)<-"mcSimulation"
-```
-
-#### Evaluate impact of "Retiring unproductive bulls" intervention
-
-We're assuming that all bulls >36 months old have been removed.
-
-```{r, fig.show="hold", out.width="50%", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE}
-
-uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
-                        out_var = nobull_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
-                        out_var = nobull_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-# uncertainty::varkernelslicerange(in_var = nobull_res$y$pc_milk_yield,
-#                                  out_var = nobull_res$y$pc_on_farm, 
-#                                  min_in_var = round(min(nobull_res$y$pc_milk_yield,na.rm=TRUE)),
-#                                  max_in_var = round(max(nobull_res$y$pc_milk_yield,na.rm=TRUE)), 
-#                                  xlab_vars = "Total emissions given milk yield range")
-# 
-# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
-#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-#                                  xlab_vars = "Total emissions given milk yield range")
-
-```
-
-### Reducing unproductive replacement males
-
-The idea is to raise the off-take rate of replacement males by 23% (not sure why this number).
-We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate.
-The result should be quite similar.
-
-```{r , warning=FALSE, eval=FALSE}
-x_lessReplace_intervention <- GHG_simulation_scenarios$x
-
-lessReplace_adoption_rate <- 0.23
-
-for (i in 1:nrow(x_lessReplace_intervention))
-{
-  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
-    x_lessReplace_intervention$N_animal_type_3[i]<-0
-}
-# run model for modified x file
-lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-lessReplace_res$call<-"multi-scenario call"
-save_temperature_scenarios(lessReplace_res,"results","lessReplace_intervention")
-```
-
-```{r, include=FALSE}
-lessReplace_res<-load_temperature_scenarios("results","lessReplace_intervention")
-class(lessReplace_res)<-"mcSimulation"
-```
-
-#### Evaluate impact of "Retiring unproductive bulls" intervention
-
-We're assuming that all bulls >36 months old have been removed.
-
-```{r, fig.show="hold", out.width="50%", fig.cap="Fewer replacement males intervention (left) vs. baseline (right)", message=FALSE}
-
-uncertainty::varscatter(in_var = lessReplace_res$y$pc_milk_yield,
-                        out_var = lessReplace_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = lessReplace_res$y$pc_milk_yield,
-                        out_var = lessReplace_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-# uncertainty::varkernelslicerange(in_var = lessReplace_res$y$pc_milk_yield,
-#                                  out_var = lessReplace_res$y$pc_on_farm, 
-#                                  min_in_var = round(min(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)),
-#                                  max_in_var = round(max(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)), 
-#                                  xlab_vars = "Total emissions given milk yield range")
-# 
-# uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-#                         out_var = GHG_simulation_scenarios$y$pc_on_farm,
-#                                  min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
-#                                  max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
-#                                  xlab_vars = "Total emissions given milk yield range")
-
-```
-
-# Comparing the scenarios
-
-Below we compare the density distributions for the scenarios.
-This required loading some helper functions that contain modified versions of some functions from the `uncertainty` package (`multi_data_varkernelslicerange` and `data_varkernelslicerange`).
-
-```{r uncertainty_evaluation_functions, include=FALSE}
-# Generate the GHG emissions function 
-source("uncertainty_evaluation.R", local = knitr::knit_global())
-```
-
-The plot below shows that in terms of per-capita emissions, the difference between the scenarios is not impressive.
-
-```{r}
-per_capita_scenarios<-list(list(scenario="Baseline",
-                     in_var=GHG_simulation_scenarios$y$pc_milk_yield,
-                     out_var=GHG_simulation_scenarios$y$pc_on_farm),
-                list(scenario="Less male replacement",
-                     in_var=lessReplace_res$y$pc_milk_yield,
-                     out_var=lessReplace_res$y$pc_on_farm),
-                list(scenario="No bulls",
-                     in_var=nobull_res$y$pc_milk_yield,
-                     out_var=nobull_res$y$pc_on_farm),
-                list(scenario="Breed upgrade",
-                     in_var=breed_res$y$pc_milk_yield,
-                     out_var=breed_res$y$pc_on_farm))
-
-per_capita_emissions<-multi_data_varkernelslicerange(per_capita_scenarios)
-
-library(ggplot2)
-
-ggplot(data = per_capita_emissions$slice_values, aes(x = y, y = z, color = scenario)) +
-  geom_line() + theme_bw(base_size = 12) +
-  xlab(expression(Per~capita~emissions~(kg~CO[2]-eq~year^-1))) +
-  ylab("Density") +
-  ggtitle("Per capita emissions") +
-  scale_color_discrete(name = "Scenario")
-
-```
-
-We can also look at the distribution of emissions per kg of milk yield.
-The plot below shows the cumulative probability distribution for this metric.
-The differences are a bit more visible here, but still fairly minor.
-
-```{r}
-
-ghg_per_milk <-
-  rbind(
-    data.frame(
-      scenario = "Baseline",
-      milk_yield = GHG_simulation_scenarios$y$pc_milk_yield,
-      GHG = GHG_simulation_scenarios$y$pc_on_farm
-    ),data.frame(
-      scenario = "Less male replacement",
-      milk_yield = lessReplace_res$y$pc_milk_yield,
-      GHG = lessReplace_res$y$pc_on_farm
-    ),data.frame(
-      scenario = "No bulls",
-      milk_yield = nobull_res$y$pc_milk_yield,
-      GHG = nobull_res$y$pc_on_farm
-    ),data.frame(
-      scenario = "Breed upgrade",
-      milk_yield = breed_res$y$pc_milk_yield,
-      GHG = breed_res$y$pc_on_farm
-    ))
-ghg_per_milk[,"ghg_per_milk"]<-ghg_per_milk$GHG/ghg_per_milk$milk_yield
-
-ggplot(data = ghg_per_milk, aes(x = ghg_per_milk, color = scenario)) +
-  stat_ecdf() + xlim(0, 10) + theme_bw(base_size = 12) + ggtitle("Emissions per kg of milk (truncated at 10 kg)") + xlab(expression(Emissions~per~kg~milk~(kg~CO[2]-eq~year^-1))) + ylab("Cumulative probability")
-
-```
-
-This document was generated using R Markdown [@R-rmarkdown].
-
-# References
+# THIS WON'T REALLY WORK... peak is too high, everything else becomes invisible

### 2022-05-13 Manuscript_analyses.Rmd
incremental progress, still not working

@@
 We then have to remove all households that don't produce any milk, because this ratio would be infinity for them.
+In fact, we'll remove all households that reported very low milk yields of <500 kg per year per capita, since these are unlikely to represent regularly operating dairy operations.
+For instance, they may be mostly focused on beef cattle production, or they may only have a small number of dairy cows that happened to be unproductive at the time of the survey.
 
@@
 GHG_simulation_scenarios$y[,"CO2eq_per_milk"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
-GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[is.finite(GHG_simulation_scenarios$y$CO2eq_per_milk),]
 
+GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+
+#GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[is.finite(GHG_simulation_scenarios$y$CO2eq_per_milk),]
+
+#GHG_simulation_scenarios$y[GHG_simulation_scenarios$y[,"CO2eq_per_milk"]>10,]
+
+#GHG_simulation_scenarios$y
 ```

### 2022-05-13 Manuscript_analyses.Rmd
update

@@
 
-# THIS LOOKS BAD, BECAUSE OF Low-Milk-high-emissions households. Would be good to break the axis - any ideas?
+# THIS LOOKS BAD, BECAUSE OF Low-Milk-high-emissions households. Would be good to break the axis - any ideas? (I removed the extreme peak now by excluding underperforming dairy households)
 
@@
                            n_samples = 1000,
-                           out_var_sampling = (max(out_var)-min(out_var))/1000)
+                           out_var_sampling = 1000)
 
@@
 
+ggplot(data=slice$slice, aes(CO2eq_per_milk)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") +
+    xlab(expression(Emission~density~of~milk~production~(XX~CO[2]-eq~kg^-1)))
+
+ggsave("Fig_3_emission_density_farm5.png")
+
 ```
 
-# THIS WON'T REALLY WORK... peak is too high, everything else becomes invisible
+# THIS WON'T REALLY WORK... peak is too high, everything else becomes invisible (should work now, afte excluding underperforming households)

### 2022-05-15 Manuscript_analyses.Rmd
added scenarios

--- 2022-05-13 Manuscript_analyses.Rmd
@@
 ---
-title: "Manuscript_analyses"
+title: "Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya"
 author: "Eike Luedeling, Cory Whitney, Todd Rosenstock"
@@
 
-# Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya
-
-## Set up
+# Set up
 
@@
 
-## Monte Carlo simulation
+# Monte Carlo simulation
 
@@
 
-## Running a Monte Carlo simulation with scenarios
+# Running a Monte Carlo simulation with scenarios
 
@@
 
-## Household-based scenarios
+# Household-based scenarios
 
@@
 
-## Possible errors resulting from input uncertainty
+# Possible errors resulting from input uncertainty
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~density~of~milk~production~(XX~CO[2]-eq~kg^-1)))
+    xlab(expression(Emission~density~of~milk~production~(kg~CO[2]-eq~kg^-1)))
 
@@
 
-# UNIT NEEDS TO BE ADDED HERE (and checked for milk, though this is probably right)
-# ALSO need to add the usual precise value estimate
+# STILL NEED TO add the usual precise value estimate
 
@@
 We can illustrate this uncertainty using a scatter plot. The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.
+
+As with the GHG Model, we apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
+These come from the model results (`r nrow(GHG_simulation_scenarios$y)`) of the `GHG_simulation_scenarios` model results to the `in_var` (x) and an outcome variable `out_var` (y).
+In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
 
@@
 
-As with the GHG Model, we apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
-These come from the model results (`r nrow(GHG_simulation_scenarios$y)`) of the `GHG_simulation_scenarios` model results to the `in_var` (x) and an outcome variable `out_var` (y).
-In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
-
-## Milk_yield (x-axis) effect on on-farm GHG emissions (y-axis)
-
-```{r}
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-```
+# Emissions intensity of milk production
 
@@
 
-GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)
 
-#GHG_simulation_scenarios$y<-GHG_simulation_scenarios$y[is.finite(GHG_simulation_scenarios$y$CO2eq_per_milk),]
+dairy_households<-GHG_simulation_scenarios
+dairy_households$x<-dairy_households$x[normal_milk_yield,]
+dairy_households$y<-dairy_households$y[normal_milk_yield,]
 
-#GHG_simulation_scenarios$y[GHG_simulation_scenarios$y[,"CO2eq_per_milk"]>10,]
-
-#GHG_simulation_scenarios$y
 ```
@@
 
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
+uncertainty::varscatter(in_var = dairy_households$y$pc_milk_yield,
+                        out_var = dairy_households$y$CO2eq_per_milk,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 ```{r varkernel}
-uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
+uncertainty::varkernel(in_var = dairy_households$y$pc_milk_yield,
+                        out_var = dairy_households$y$CO2eq_per_milk,
                         xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
@@
 ```
-
-# THIS LOOKS BAD, BECAUSE OF Low-Milk-high-emissions households. Would be good to break the axis - any ideas?
 
@@
 
-## Probability of GHG emissions for a given range of milk yield
+# Probability of GHG emissions for a given milk yield
+
+The following plot shows the emissions intensity that corresponds to a milk yield of 2000 kg per capita per year, based on the population of households covered in the survey.
 
 ```{r varkernelslicerange}
-slice<-uncertainty::varslice_resample(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                                 out_var = GHG_simulation_scenarios$y$CO2eq_per_milk, 
+slice<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
+                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                  expectedin_var=2000,
@@
 
+ggplot(data=data.frame(CO2eq_per_milk=slice$resampled), aes(CO2eq_per_milk)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") +
+    xlab(expression(Emission~density~of~milk~production~at~2000~kg~milk~per~capita~(XX~CO[2]-eq~kg^-1)))
+
+ggsave("Fig_X_emission_density_all_farms.png")
+
 ```
 
-# THIS WON'T REALLY WORK... peak is too high, everything else becomes invisible
+We see considerable uncertainty about the emissions at this particular level of production, and the previous figure shows that emissions of all other levels is similarly variable.
+This raises considerable doubts about the usefulness of milk yield as an indicator of emissions intensity - at least when this indicator is taken at face value, without considering this uncertainty.
+Yet, some information is contained even in such an imperfect indicator, since, as also shown in the previous figure, there is clearly a correlation between milk yield and emissions intensity.
+The following proposes a mechanism for quantifying the level of certainty, with which a change in milk yield can reasonably be associated with a change in emissions intensity.
+
+# Probability of an increase in milk yield being associated with an increase in GHG emission intensity
+
+To derive the level of certainty, with which a change in milk yield is associated with an increase or decrease in emissions intensity, we have to investigate the probability distributions (as shown in Fig.
+X) of emissions intensity at two levels of milk production.
+This can be achieved using the `varslice_resample` function for both production levels.
+
+Let's look at this for production of 2000 and 3000 kg milk per capita per year.
+We'll take 1000 random draws of the emissions intensities that associated with both levels and compare values among these samples.
+
+```{r}
+slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
+                                 out_var = dairy_households$y$CO2eq_per_milk, 
+                                 expectedin_var=2000,
+                           n = 1000,
+                           n_samples = 1000,
+                           out_var_sampling = 1000)
+slice_3000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
+                                 out_var = dairy_households$y$CO2eq_per_milk, 
+                                 expectedin_var=3000,
+                           n = 1000,
+                           n_samples = 1000,
+                           out_var_sampling = 1000)
+
+sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)
+
+```
+The comparison in the last line of code above reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.
+
+We can now generalize this result by doing similar calculations for multiple production levels and illustrating them for the whole population.
+
+```{r}
+
+in_var = dairy_households$y$pc_milk_yield
+out_var = dairy_households$y$CO2eq_per_milk
+
+milk_yields_to_sample<-seq(100,10000,100)
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(in_var))]
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample>min(in_var))]
+slices<-list()
+
+# slice through the distribution at every 100 L milk yield per cow
+
+for(i in 1:length(milk_yields_to_sample))
+{slices[[i]]<-
+  varslice_resample(in_var,
+                      out_var,
+                      expectedin_var = milk_yields_to_sample[i],
+                      n_samples = 100000,
+  out_var_sampling = 1000)$resample
+ }
+
+decreased_emissions_probability <-
+  data.frame(Base_yield = NA,new_yield=NA,probability=NA)
+
+for (i in 1:(length(milk_yields_to_sample) - 1))
+{
+  for (j in (i + 1):length(milk_yields_to_sample))
+    {decreased_emissions_probability[nrow(decreased_emissions_probability)+1,"Base_yield"]<-milk_yields_to_sample[i]
+     decreased_emissions_probability$new_yield[nrow(decreased_emissions_probability)]<-milk_yields_to_sample[j]
+     decreased_emissions_probability$probability[nrow(decreased_emissions_probability)] <-
+      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
+
+}
+}
+
+decreased_emissions_probability<-decreased_emissions_probability[which(!is.na(decreased_emissions_probability$Base_yield)),]
+
+decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase, z =
+                                                      probability)) + geom_contour_filled() +
+  ggtitle("Probability that emission intensity reductions occurred") + 
+  xlab("Baseline milk yield per head and year (kg)") +
+  ylab("Milk yield increase per head and year (kg)") +
+  stat_contour(breaks = c(0.9),col="black")
+
+ggsave("Fig_X1_probability_of_intensity_reduction.png")
+
+```
+
+The 90% confidence contour is shown in the plot.
+The results indicate that whenever milk yields increase, the probability that emissions intensity decreased is usually greater than 50% (slightly lower values are due to randomness in the resampling procedure), but that large increases are needed to be sure of this.
+
+# Testing interventions - strategy
+
+Farms may try to reduce their emissions intensity by making changes to their herd structure, feeding strategy or management practices.
+Programs aiming to effect emissions reductions in dairy production may want to reward such changes, and they may be looking to the milk production per capita as an indicator for the effectiveness of such measures.
+We now want explore whether this is a reasonable strategy.
+
+## Strategy
+
+Develop scenarios that implement innovations
+
+* breed adjustment
+* herd structure
+
+Quantify improvements on farms by comparing with the baseline.
+Plot improvements vs. the baseline on the above figure.
+Plot improvements vs. baseline in the emissions intensity figure.
+
+For each slice, and each level of confidence, calculate the percentage of farms that made changes that would be correctly rewarded vs.
+the ones that did not change (the baseline population).
+Make a figure showing the Type I and Type II errors resulting from each confidence level at each baseline yield.
+
+# Testing interventions
+
+## Single-run function
+
+To clearly characterize the impact of interventions, we should compare their impact with a counterfactual, i.e.
+with the same farms without the intervention.
+While this is usually impossible in practice, it is easy with a model.
+We just have to run the simulation again with identical inputs except for those that are affected by the intervention.
+
+We already simulated emissions for a large population of farms, and we can now apply our interventions to this population by modifying certain parameters of the input tables.
+What we don't want now, however, is to randomize all values again, since this would add additional noise to the results, which would mask the signal caused by the interventions.
+
+We now first need a function that can apply our greenhouse gas emissions model to a pre-defined set of input values.
+We can find syntax that can achieve this in the mcSimulation function.
+Here it is (with slight modifications):
+
+```{r}
+run_model<-function(x_data,model)
+  {
+  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
+  model_function_ <- function(x) {
+
+    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
+    eval(expr = body(model), envir = e)
+  }
+  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
+                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
+  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
+  returnObject$call <- match.call()
+  class(returnObject) <- cbind("mcSimulation", class(returnObject))
+  return(returnObject)
+}
+
+```
+
+The following code can precisely reproduce the results of the previous analysis:
+
+```{r, eval=FALSE}
+res<-run_model(GHG_simulation_scenarios$x,ghg_emissions)
+```
+
+To simulate interventions, we can now feed this function with modified x_data data.frames.
+
+## Breed intervention
+
+Introduction of higher-performance breeds may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.
+
+To retain the farm-specific performance levels obtained from the farm survey, we simulate these impacts in a relative manner.
+For this, we first need to determine the performance and characteristics of each of the breeds.
+We'll base this on the farm survey data (`Kenyaherds` dataset from above).
+
+```{r}
+cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
+
+ggplot(cows, aes(factor(breed),milk_yield)) +
+  geom_violin(draw_quantiles=c(0.5),fill="light green") +
+  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
+
+milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
+milk_summary<-milk_summary[,c(1,4,2)]
+colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
+milk_summary
+
+```
+
+This evaluation shows that the difference in median milk yields is relatively small, but that the milk yield distributions show a lot of variation that is apparently not explained by the breed.
+It is also apparent that the number of animals these estimates are based on varies considerably.
+Two breeds make up `r round(sum(milk_summary$n[1:2])/sum(milk_summary$n)*100,1)` % of the entire population, with others having very few animals, which means that the medians may be somewhat unreliable (especially for the single animal of breed 7).
+
+Let's define the intervention as follows:
+
+All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences.
+This means that milk yields are multiplied by the following factors:
+
+```{r}
+milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))
+
+milk_summary             
+```
+
+A change of the breed also causes changes in weight, mature weight and weight gain.
+
+```{r}
+live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(live_weights)<-c("animal_type","breed","median_live_weight")
+live_weights[,"Intervention_factor"]<-NA
+for(i in 1:nrow(live_weights))
+  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
+     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]
+
+# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.
+
+mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
+mature_weights[,"Intervention_factor"]<-NA
+for(i in 1:nrow(mature_weights))
+  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
+     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]
+
+# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1
+
+weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
+weight_gains[,"Intervention_factor"]<-NA
+for(i in 1:nrow(weight_gains))
+  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
+     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
+weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1
+
+```
+
+Now we can modify the input table accordingly.
+
+go through all the farms. Determine the 'improvement' coefficients (based on herd structure)
+
+```{r}
+farms<-unique(Kenyaherds$QQNO)
+breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)
+
+Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
+                                        "system",
+                                        "animal_type",
+                                        "breed",
+                                        "live_weight_LW",
+                                        "mature_weight_MW",
+                                        "weight_gain_WG",
+                                        "milk_yield",
+                                        "dig_energy",
+                                        "CP.")]
+
+Kenyaherds_interventions[,"milk_change"]<-0
+cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
+Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )
+
+milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)
+
+breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% milk_change[,1])]<-milk_change[,2]
+breed_intervention$milk_yield_5<-breed_intervention$milk_yield_4
+breed_intervention$milk_yield_6<-breed_intervention$milk_yield_4
+
+for(i in 1:nrow(Kenyaherds_interventions))
+  {Kenyaherds_interventions[i,"live_weight_change"]<-
+    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
+                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
+#   Kenyaherds_interventions[i,"mature_weight_change"]<-
+#      1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
+#                                                 mature_weights$breed==Kenyaherds_interventions$breed[i])]
+   Kenyaherds_interventions[i,"weight_gain_change"]<-
+     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
+                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
+}
+
+live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
+                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+#mature_weight_agg<-aggregate(Kenyaherds_interventions$mature_weight_change,
+#by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
+                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+
+colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
+#colnames(mature_weight_agg)<-c("farm","animal_type","conv_factor")
+colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")
+
+for(i in 1:nrow(live_weight_agg))
+  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
+                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
+    live_weight_agg$conv_factor[i]
+
+#for(i in 1:nrow(mature_weight_agg))
+#  breed_intervention[which(breed_intervention$Farm==mature_weight_agg$farm[i]),
+#                     paste0("mature_weight_LW_",mature_weight_agg$animal_type[i])]<-
+#    mature_weight_agg$conv_factor[i]
+
+for(i in 1:nrow(weight_gain_agg))
+  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
+                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
+    weight_gain_agg$conv_factor[i]
+
+breed_intervention[is.na(breed_intervention)]<-1
+
+```
+
+go through the x file
+for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors. Actually, in this case we want to simulate full adoption, so we'll set this to 1.
+
+```{r, eval=FALSE}
+x_breed_intervention <- GHG_simulation_scenarios$x
+
+adoption_rate <- 1
+
+for (i in 1:nrow(x_breed_intervention))
+{
+  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
+    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
+      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
+      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
+                                 x_breed_intervention$Scenario[i]),
+                         2:ncol(breed_intervention)]
+}
+# run model for modified x file
+breed_res <- run_model(x_breed_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+breed_res$call<-"multi-scenario call"
+save_temperature_scenarios(breed_res,"results","Breed_intervention")
+```
+
+```{r, include=FALSE}
+adoption_rate <- 1
+breed_res<-load_temperature_scenarios("results","Breed_intervention")
+class(breed_res)<-"mcSimulation"
+```
+
+# Compare outcomes of breed intervention with baseline
+
+```{r}
+breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
+
+breed_res$y<-breed_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+breed_res$x<-breed_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+
+impact_of_breed<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=breed_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=breed_res$y$CO2eq_per_milk)
+impact_of_breed[,"milk_change"]<-round(impact_of_breed$breed_milk-impact_of_breed$baseline,2)
+impact_of_breed[,"emissions_intensity_change"]<-round(impact_of_breed$CO2_intensity_breed-impact_of_breed$CO2_intensity_baseline ,3)
+
+ggplot(data=impact_of_breed,aes(x=milk_change,y=emissions_intensity_change)) + geom_point() + theme_bw() + ylab("Change in emissions intensity (kg CO2-eq kg-1)") + xlab("Change in per capita milk yield (kg)")
+
+ggsave("Fig_XX_breed_reductions.png")
+
+```
+
+```{r}
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase)) + geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=milk_change)) +
+  ggtitle("Probability that emission intensity reductions occurred") + 
+  xlab("Baseline milk yield per head and year (kg)") +
+  ylab("Milk yield increase per head and year (kg)") 
+ggsave("Fig_XX1_breed_probability_of_intensity_reduction.png")
+
+```
+
+This plot shows a lot of linear patterns that result from the application of fixed factors to convert the milk yield between cows of different breeds. Not so pretty, but maybe ok.
+
+## Herd structure
+
+### Retiring unproductive bulls
+
+To simulate this scenario, we simply have to remove all animals of type 1 from the herds.
+
+```{r , warning=FALSE, eval=FALSE}
+x_nobull_intervention <- GHG_simulation_scenarios$x
+
+nobull_adoption_rate <- 1.0
+
+for (i in 1:nrow(x_nobull_intervention))
+{
+  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
+    x_nobull_intervention$N_animal_type_1[i]<-0
+}
+# run model for modified x file
+nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+nobull_res$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_res,"results","noBull_intervention")
+```
+
+```{r, include=FALSE}
+nobull_res<-load_temperature_scenarios("results","noBull_intervention")
+class(nobull_res)<-"mcSimulation"
+```
+
+#### Evaluate impact of "Retiring unproductive bulls" intervention
+
+We're assuming that all bulls >36 months old have been removed.
+
+```{r, fig.show="hold", out.width="50%", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE}
+
+uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
+                        out_var = nobull_res$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
+                        out_var = nobull_res$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+```
+
+### Reducing unproductive replacement males
+
+The idea is to raise the off-take rate of replacement males by 23% (not sure why this number).
+We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate.
+The result should be quite similar.
+
+```{r , warning=FALSE, eval=FALSE}
+x_lessReplace_intervention <- GHG_simulation_scenarios$x
+
+lessReplace_adoption_rate <- 0.23
+
+for (i in 1:nrow(x_lessReplace_intervention))
+{
+  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
+    x_lessReplace_intervention$N_animal_type_3[i]<-0
+}
+# run model for modified x file
+lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+lessReplace_res$call<-"multi-scenario call"
+save_temperature_scenarios(lessReplace_res,"results","lessReplace_intervention")
+```
+
+```{r, include=FALSE}
+lessReplace_res<-load_temperature_scenarios("results","lessReplace_intervention")
+class(lessReplace_res)<-"mcSimulation"
+```
+
+# Compare outcomes of herd intervention with baseline
+
+## No bulls scenario
+
+```{r}
+nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
+
+nobull_res$y<-nobull_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+nobull_res$x<-nobull_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+
+impact_of_nobull<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=nobull_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=nobull_res$y$CO2eq_per_milk)
+impact_of_nobull[,"milk_change"]<-round(impact_of_nobull$breed_milk-impact_of_nobull$baseline,2)
+impact_of_nobull[,"emissions_intensity_change"]<-round(impact_of_nobull$CO2_intensity_breed-impact_of_nobull$CO2_intensity_baseline ,3)
+
+ggplot(data=impact_of_nobull,aes(x=milk_change,y=emissions_intensity_change)) + geom_point() + theme_bw() + ylab("Change in emissions intensity (kg CO2-eq kg-1)") + xlab("Change in per capita milk yield (kg)")
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase)) + geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],aes(x=baseline,y=milk_change)) +
+  ggtitle("Probability that emission intensity reductions occurred") + 
+  xlab("Baseline milk yield per head and year (kg)") +
+  ylab("Milk yield increase per head and year (kg)") 
+ggsave("Fig_XX1_breed_probability_of_intensity_reduction.png")
+
+ggsave("Fig_XX_impact_of_nobull_reductions.png")
+
+```
+
+## Fewer replacement males scenario
+
+```{r}
+
+lessReplace_res$y[,"CO2eq_per_milk"]<-lessReplace_res$y$pc_on_farm/lessReplace_res$y$pc_milk_yield
+
+lessReplace_res$y<-lessReplace_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+lessReplace_res$x<-lessReplace_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+
+impact_of_lessReplace<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=lessReplace_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=lessReplace_res$y$CO2eq_per_milk)
+impact_of_lessReplace[,"milk_change"]<-round(impact_of_lessReplace$breed_milk-impact_of_lessReplace$baseline,2)
+impact_of_lessReplace[,"emissions_intensity_change"]<-round(impact_of_lessReplace$CO2_intensity_breed-impact_of_lessReplace$CO2_intensity_baseline ,3)
+
+ggplot(data=impact_of_lessReplace,aes(x=milk_change,y=emissions_intensity_change)) + geom_point() + theme_bw() + ylab("Change in emissions intensity (kg CO2-eq kg-1)") + xlab("Change in per capita milk yield (kg)")
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase)) + geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],aes(x=baseline,y=milk_change)) +
+  ggtitle("Probability that emission intensity reductions occurred") + 
+  xlab("Baseline milk yield per head and year (kg)") +
+  ylab("Milk yield increase per head and year (kg)") 
+ggsave("Fig_XX1_breed_probability_of_intensity_reduction.png")
+
+ggsave("Fig_XX_lessReplace_reductions.png")

### 2022-05-30 Manuscript_analyses.Rmd
Add citations - still working on the actual formulas from teh IPCC (@eikeluedeling can we just cite these and reference the equation?)

@@
 author: "Eike Luedeling, Cory Whitney, Todd Rosenstock"
-date: "5/12/2022"
+bibliography:
+ - references/packages.bib
+ - references/papers.bib
 output: html_document
@@
                    'decisionSupport',
+                   'ggplot2',
+                   'kableExtra',
+                   'magrittr',
+                   'plyr',
+                   'rmarkdown',
+                   'uncertainty',
+                   'chillR',
+                   'raster',
                    'MASS'), 'references/packages.bib')
@@
 
-The calculations are based on the IPCC xxx equation (describe and cite here)
+The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
+The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
+The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021], also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
+Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].
+We run the model with the decisionSupport package [@R-decisionSupport].
 
@@
 With the model equation and the input table defined above, we can run a Monte Carlo simulation that returns a probability distribution for the greenhouse gas emissions resulting from the particular herd that is specified in the input table.
+We use the functions provided by the decisionSupport package [@R-decisionSupport] in the R programming language [@R-base].
 
@@
 
-The `scenario_mc` function in the `decisionSupport` package (@R-decisionSupport) consists of a wrapper around the `mcSimulation` function.
+The `scenario_mc` function in the `decisionSupport` package [@R-decisionSupport] consists of a wrapper around the `mcSimulation` function.
 The function is able to modify the input table according to a scenario file we provide.
@@
 
-The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for for it.
+The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for it.
 Each column beyond the first two columns in the file contain information for a particular scenario.
-This is restricted to the parameters that should be *changed* for this scenario.
+This is restricted to the parameters that *should be changed* for this scenario.
 
@@
 The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both".
-If "both" is selected, both the lower and upper bounds are modified (this should only be done for constants).
-
-We define subgroups based on the populations of household that was described by a household survey in Kenya.
-Each of the XXX encountered in this survey is converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.
+If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables witha constant distribution, since the rplacement will be for both the upper and lower bound).
+
+We define subgroups based on the populations of households described by a household survey in Kenya [@wilkes_variation_2020; @wilkes_methods_2019].
+Each of the farms encountered in this survey were converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.
 
@@
 
-The information for the household is extracted from a spreadsheet that contains all the survey responses.
-Extracting all relevant information is a slightly complex task, which is accomplished by the following code.
-
-```{r}
+The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019].
+Extracting all relevant information from these spreadsheets is a slightly complex task, which is accomplished by the following code.
+
+```{r}
+# import the data from the survey and ensure that the .csv encoding can be read by R
+
 Kenyaherds<-read.table("data/Kenya_baseline_individual_cow_wNotes.csv", sep = ",",
@@
                            c(1-feed_prod_error,1+feed_prod_error)
+
   # update feed transport CO2 (with error)
@@
                            c(1-feed_trans_error,1+feed_trans_error)
+
    # update milk yield for each cow type (animal type 4-6) (with error)
@@
 The first few rows in this table adjust the herd composition parameters, so that the scenario herd corresponds to the herd of the specific household.
-Further parameters that are adjusted include the milk yield per cow and various other factors.
+Other parameters including the milk yield per cow are also adjusted.
 
@@
 
-Can can now plot distributions of the various outputs from the model, using the `plot_distributions` function.
+Given these results we can now plot distributions of the various outputs from the model, using the `plot_distributions` function.
 
@@
 
-# THIS STILL NEEDS IMPROVEMENT - e.g. units are missing, and plot titles don't fit (use concise symbols with subscripts etc.)
+*# THIS STILL NEEDS IMPROVEMENT - e.g. units are missing, and plot titles don't fit (use concise symbols with subscripts etc.)*
 
@@
 
-# NEED TO ADD THE USUAL PRECISE VALUE HERE FOR COMPARISON (also need to add units and possible fix up other things)
-
-This
+*# NEED TO ADD THE USUAL PRECISE VALUE HERE FOR COMPARISON (also need to add units and possibly fix up other things)*
+
+*This*
 
@@
 
-# STILL NEED TO add the usual precise value estimate
+*# STILL NEED TO add the usual precise value estimate*
 
@@
 
-# NICE PLOT, but would be nicer still if this was also a ggplot figure. Kinda inconsistent if it isn't. Also can't save it with ggsave, so will need another way.
+*# NICE PLOT, but would be nicer still if this was also a ggplot figure. Kinda inconsistent if it isn't. Also can't save it with ggsave, so will need another way.*
 
@@
 ```
+
 The comparison in the last line of code above reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.
@@
 
-go through all the farms. Determine the 'improvement' coefficients (based on herd structure)
+We go through all the farms and determine the 'improvement' coefficients (based on herd structure).
 
@@
 
-go through the x file
-for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors. Actually, in this case we want to simulate full adoption, so we'll set this to 1.
+Go through the x file, for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors.
+Actually, in this case we want to simulate full adoption, so we'll set this to 1.
 
@@
 ggsave("Fig_XX_lessReplace_reductions.png")
+
+```
+
+# References

### 2022-05-31 Manuscript_analyses.Rmd
worked on figures, mainly correcting labels and producing precise estimates (in the traditional sense) to add to Figs. 2 and 3.

@@
 ```{r}
-knitr::kable(read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE))
+read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)   %>%
+    kableExtra::kbl(digits = 2)  %>%
+    kableExtra::kable_classic_2(full_width = F)
 
@@
 ```{r}
-knitr::kable(read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5])
+read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5]  %>%
+    kableExtra::kbl(digits = 2)  %>%
+    kableExtra::kable_classic_2(full_width = F)
+
+```
+
+We'll also make precise scenarios that contain only the exact values that were observed in the survey.
+
+```{r}
+# count the number of animals for each type
+
+animals<-table(Kenyaherds[,c("QQNO","animal_type")])
+
+variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
+                             paste0("Cp_rate_",4:6),
+                             "feed_prod_CO2",
+                             "feed_trans_CO2",
+                             paste0("live_weight_LW_",1:12),
+                             paste0("weight_gain_WG_",1:12),
+                             "DE_perc",
+                             "perc_crude_protein_diet")
+
+precise_estimates<-c("Ca_pasture","Ca_large_grazing","C_intact_males","C_castrates","C_females","maintenance_coefficient_Cfi_other",
+                     "maintenance_coefficient_Cfi_lact_cow","maintenance_coefficient_Cfi_bulls","mature_weight_MW_female","mature_weight_MW_male",
+                     "milk_fat","Cp_rate_4","Cp_rate_5","Cp_rate_6","Y_m","Bo_animaltype_1","Bo_animaltype_2","Bo_animaltype_3","Bo_animaltype_4",
+                     "Bo_animaltype_5","Bo_animaltype_6","Bo_animaltype_7","Bo_animaltype_8","Bo_animaltype_9","Bo_animaltype_10","Bo_animaltype_11",
+                     "Bo_animaltype_12","MCFpasture","MCFspread","MCFdrylot","MCFsolidstore","MCFcomposted","MCFliquid","MCFbiogas","MCFburn","MCFsold",
+                     "mndec_pasture","mndec_spread","mndec_drylot","mndec_solidstore","mndec_composted","mndec_liquid","mnvc_pasture","mnvc_spread",
+                     "mnvc_drylot","mnvc_solidstore","mnvc_composted","mnvc_liquid","frac_leach_pasture","frac_leach_spread","frac_leach_drylot",
+                     "frac_leach_solidstore","frac_leach_composted","frac_leach_liquid")
+
+precise_scenarios<-data.frame(Variable=c(variables_to_update_const,variables_to_update_posnorm,precise_estimates),param="both")
+
+for(i in 1:nrow(animals))
+  {coln<-paste0("Farm_",row.names(animals)[i])
+   precise_scenarios[,coln]<-NA
+     precise_scenarios[which(precise_scenarios$Variable %in% paste0("N_animal_type_",1:12)),coln]<-animals[i,]
+}
+
+# farm-scale variables
+
+Farms<-unique(Kenyaherds$QQNO)
+
+for (f in Farms)
+  {coln<-paste0("Farm_",f)
+  # update feeding system 
+  precise_scenarios[which(precise_scenarios$Variable == "feeding_system"),coln]<-
+     median(Kenyaherds[which(Kenyaherds$QQNO==f),"system"])
+
+  # update digestible energy (mean for all animals for now) (with error)
+  precise_scenarios[which(precise_scenarios$Variable == "DE_perc"),coln]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"dig_energy"])
+
+  # update crude protein (mean for all animals for now) (with error)
+  precise_scenarios[which(precise_scenarios$Variable == "perc_crude_protein_diet"),coln]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"CP."])
+
+  # update feed production CO2 (with error)
+   precise_scenarios[which(precise_scenarios$Variable == "feed_prod_CO2"),coln]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])
+
+  # update feed transport CO2 (with error)
+   precise_scenarios[which(precise_scenarios$Variable == "feed_trans_CO2"),coln]<-
+     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])
+
+   # update milk yield for each cow type (animal type 4-6) (with error)
+   for(typ in 4:6)
+     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
+       {precise_scenarios[which(precise_scenarios$Variable==paste0("milk_yield_",typ)),coln]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "milk_yield"])
+       precise_scenarios[which(precise_scenarios$Variable==paste0("CP_rate_",typ)),coln]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "Cp"]/0.1)
+
+     }
+  # update live weight and weight gain for each animal type (with error)
+   for(typ in 1:12)
+     if(sum(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ)>0)
+       {precise_scenarios[which(precise_scenarios$Variable==paste0("live_weight_LW_",typ)),coln]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "live_weight_LW"])
+       precise_scenarios[which(precise_scenarios$Variable==paste0("weight_gain_WG_",typ)),coln]<-
+         mean(Kenyaherds[which(Kenyaherds$QQNO==f&Kenyaherds$animal_type==typ),
+                         "weight_gain_WG"])
+     }
+
+   # now add the best estimates from the IPCC guidelines to the precise scenarios
+
+   precise_scenarios[which(precise_scenarios$Variable=="Ca_pasture"),coln]<-0.17
+   precise_scenarios[which(precise_scenarios$Variable=="Ca_large_grazing"),coln]<-0.36
+   precise_scenarios[which(precise_scenarios$Variable=="C_intact_males"),coln]<-1.2   
+   precise_scenarios[which(precise_scenarios$Variable=="C_castrates"),coln]<-1.0   
+   precise_scenarios[which(precise_scenarios$Variable=="C_females"),coln]<-0.8   
+   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_other"),coln]<-0.322   
+   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_lact_cow"),coln]<-0.386
+   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_bulls"),coln]<-0.370
+   precise_scenarios[which(precise_scenarios$Variable=="mature_weight_MW_female"),coln]<-349 
+   precise_scenarios[which(precise_scenarios$Variable=="mature_weight_MW_male"),coln]<-483   
+   precise_scenarios[which(precise_scenarios$Variable=="milk_fat"),coln]<-4 
+   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_4"),coln]<-0.1  
+   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_5"),coln]<-0.1   
+   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_6"),coln]<-0.1
+   precise_scenarios[which(precise_scenarios$Variable=="Y_m"),coln]<-0.06
+
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_1"),coln]<-0.13
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_2"),coln]<-0.13
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_3"),coln]<-0.13
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_4"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_5"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_6"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_7"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_8"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_9"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_10"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_11"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_12"),coln]<-0.13   
+   precise_scenarios[which(precise_scenarios$Variable=="MCFpasture"),coln]<-0.0047   
+   precise_scenarios[which(precise_scenarios$Variable=="MCFspread"),coln]<-0.01
+   precise_scenarios[which(precise_scenarios$Variable=="MCFdrylot"),coln]<-0.02     
+   precise_scenarios[which(precise_scenarios$Variable=="MCFsolidstore"),coln]<-0.05
+   precise_scenarios[which(precise_scenarios$Variable=="MCFcomposted"),coln]<-0.025
+   precise_scenarios[which(precise_scenarios$Variable=="MCFliquid"),coln]<-0.8   
+   precise_scenarios[which(precise_scenarios$Variable=="MCFbiogas"),coln]<-0.1085   
+   precise_scenarios[which(precise_scenarios$Variable=="MCFburn"),coln]<-0.1   
+   precise_scenarios[which(precise_scenarios$Variable=="MCFsold"),coln]<-0   
+   precise_scenarios[which(precise_scenarios$Variable=="mndec_pasture"),coln]<-0.02   
+   precise_scenarios[which(precise_scenarios$Variable=="mndec_spread"),coln]<-0   
+   precise_scenarios[which(precise_scenarios$Variable=="mndec_drylot"),coln]<-0.02   
+   precise_scenarios[which(precise_scenarios$Variable=="mndec_solidstore"),coln]<-0.01   
+   precise_scenarios[which(precise_scenarios$Variable=="mndec_composted"),coln]<-0.01   
+   precise_scenarios[which(precise_scenarios$Variable=="mndec_liquid"),coln]<-0.01   
+   precise_scenarios[which(precise_scenarios$Variable=="mnvc_pasture"),coln]<-0.02
+   precise_scenarios[which(precise_scenarios$Variable=="mnvc_spread"),coln]<-0.07
+   precise_scenarios[which(precise_scenarios$Variable=="mnvc_drylot"),coln]<-0.3
+   precise_scenarios[which(precise_scenarios$Variable=="mnvc_solidstore"),coln]<-0.45   
+   precise_scenarios[which(precise_scenarios$Variable=="mnvc_composted"),coln]<-0.65   
+   precise_scenarios[which(precise_scenarios$Variable=="mnvc_liquid"),coln]<-0.3   
+   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_pasture"),coln]<-0.24   
+   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_spread"),coln]<-0.24   
+   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_drylot"),coln]<-0.24   
+   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_solidstore"),coln]<-0.24   
+   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_composted"),coln]<-0.24  
+   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_liquid"),coln]<-0.24
+
+   }
+
+write.csv(precise_scenarios,"data/farm_scenarios_precise.csv",row.names = FALSE)
 
@@
                    old_names = "enteric_CH4",
-                   new_names = "CH4 Emissions from Enteric Fermentation per Farm") + 
+                   new_names = "CH[4]~Emissions~from~Enteric~Fermentation~(kg~CH[4])") + 
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())
+
+enteric_CH4_dist_plot <- enteric_CH4_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
                    old_names = "milk_yield",
-                   new_names = "Total Milk Yield") + 
+                   new_names = "Total~Milk~Yield~(kg)") + 
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())
+
+milk_yield_dist_plot <- milk_yield_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
                    old_names = "enteric_CH4_CO2eq",
-                   new_names = "Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock") + 
+                   new_names = "Enteric~CH[4]~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
   theme(axis.title.x=element_blank())
+
+enteric_CH4_CO2eq_dist_plot <- enteric_CH4_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
                    old_names = "mm_CH4_CO2eq",
-                   new_names = "Carbon Dioxide Equivalent Methane Emissions from Livestock Manure") + 
+                   new_names = "CH[4]~Emissions~from~Livestock~Manure~(kg~CO[2]*'-eq')") + 
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())
+
+mm_CH4_CO2eq_dist_plot <- mm_CH4_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
                    old_names = "mm_N2O_CO2eq",
-                   new_names = "Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure") + 
+                   new_names = "N[2]*O~Emissions~from~Livestock~Manure~(kg~CO[2]*'-eq')") + 
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())
+
+mm_N2O_CO2eq_dist_plot <- mm_N2O_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
                    old_names = "feed_CO2",
-                   new_names = "Carbon Dioxide Emissions from Livestock Feed Production and Transport") + 
+                   new_names = "CO[2]~Emissions~from~Livestock~Feed~Production~and~Transport~(kg)") + 
   theme(axis.title.x=element_blank(),
         axis.title.y=element_blank())
+feed_CO2_dist_plot <- feed_CO2_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
                    old_names = "on_farm",
-                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
+                   new_names = "Total~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
   theme(axis.title.y=element_blank())
+
+on_farm_dist_plot <- on_farm_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
 
@@
 
-ggsave("Fig_1_GHG_simulation_outputs.png")
-```
-
-*# THIS STILL NEEDS IMPROVEMENT - e.g. units are missing, and plot titles don't fit (use concise symbols with subscripts etc.)*
+ggsave("Fig_1_GHG_simulation_outputs.png",width=9,height=9)
+```
+
+*# THIS STILL NEEDS CHECKING - are all the units correct?*
 
@@
 
-plot_distributions(mcSimulation_object = farm5_emissions,
+precise_scenarios<-read.csv("data/farm_scenarios_precise.csv", fileEncoding = "UTF-8-BOM")
+
+farm5_precise<-precise_scenarios[,c(1,2,7)]
+
+farm5_emissions_precise<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                             scenarios = farm5_precise,
+                             model_function = ghg_emissions,
+                             numberOfModelRuns = 10,
+                             functionSyntax = "plainNames")
+
+pd<-plot_distributions(mcSimulation_object = farm5_emissions,
                    vars = "on_farm",
@@
                    old_names = "on_farm",
-                   new_names = "Total Carbon Dioxide Equivalent Emissions from Livestock") + 
+                   new_names = "Total~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
   theme(axis.title.y=element_blank())
 
-ggsave("Fig_2_uncertainty_illustration.png")
-
-```
-
-*# NEED TO ADD THE USUAL PRECISE VALUE HERE FOR COMPARISON (also need to add units and possibly fix up other things)*
-
-*This*
+pd + facet_wrap(. ~ name, labeller = label_parsed) + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=on_farm),color="blue", lwd=3)
+
+ggsave("Fig_2_uncertainty_illustration.png",width=5,height=3)
+
+```
+
+The blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.
 
@@
 
-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
+histo<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~density~of~milk~production~(kg~CO[2]-eq~kg^-1)))
-
-ggsave("Fig_3_emission_density_farm5.png")
-
-```
-
-*# STILL NEED TO add the usual precise value estimate*
+    xlab(expression(Emission~density~of~milk~production~(~kg~CO[2]*'-eq'~kg^-1)))
+
+farm5_emissions_precise$y[,"CO2eq_per_milk"]<-farm5_emissions_precise$y$pc_on_farm/farm5_emissions_precise$y$pc_milk_yield
+
+histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3)
+
+ggsave("Fig_3_emission_density_farm5.png",width=5,height=3)
+
+```
+
+Also here, the blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.
 

### 2022-06-01 Manuscript_analyses.Rmd
most figures ggized

@@
 
-```{r}
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-```
-
-*# NICE PLOT, but would be nicer still if this was also a ggplot figure. Kinda inconsistent if it isn't. Also can't save it with ggsave, so will need another way.*
+```{r, warning=FALSE, message=FALSE}
+#uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+#                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+plotdata<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
+                     out_var=GHG_simulation_scenarios$y$pc_on_farm,
+                     Group="just one")
+
+plot1 <- ggplot(plotdata, aes(x = in_var, y = out_var)) + 
+  geom_point(shape = 20, color = "blue", size = 2) + 
+  stat_smooth(method = "loessm", fullrange = TRUE) +
+  geom_density_2d() +
+  geom_rug(color="lightskyblue3") + 
+  labs(y=expression(Total~emissions~(kg~CO[2]*'-eq'~head^{-1}~year^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() +
+  theme(legend.position = c(0.15, 0.9))
+
+dens1 <- ggplot(plotdata, aes(x = in_var)) + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, fill = "white") +
+  geom_density(fill="lightskyblue3") + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, alpha=0) + 
+  theme_void() + 
+  theme(legend.position = "none")
+
+dens2 <- ggplot(plotdata, aes(x = out_var)) + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, fill = "white") +
+  geom_density(fill="lightskyblue3") + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, alpha=0) + 
+  theme_void() + 
+  theme(legend.position = "none") + 
+  coord_flip()
+
+dens1 + plot_spacer() + plot1 + dens2 + 
+  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
+
+ggsave("Fig_4_scatter_per_head_milk_emissions.png",width=5,height=3)
+
+```
 
@@
 
-```{r}
-
-uncertainty::varscatter(in_var = dairy_households$y$pc_milk_yield,
-                        out_var = dairy_households$y$CO2eq_per_milk,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
+```{r, warning=FALSE, message=FALSE}
+
+#uncertainty::varscatter(in_var = dairy_households$y$pc_milk_yield,
+#                        out_var = dairy_households$y$CO2eq_per_milk,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
+
+plotdata<-data.frame(in_var=dairy_households$y$pc_milk_yield,
+                     out_var=dairy_households$y$CO2eq_per_milk,
+                     Group="just one")
+
+plot1 <- ggplot(plotdata, aes(x = in_var, y = out_var)) + 
+  geom_point(shape = 20, color = "blue", size = 2) + 
+  stat_smooth(method = "loess", fullrange = TRUE) +
+  geom_density_2d() +
+  geom_rug(color="lightskyblue3") + 
+  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() +
+  theme(legend.position = c(0.15, 0.9))
+
+dens1 <- ggplot(plotdata, aes(x = in_var)) + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, fill = "white") +
+  geom_density(fill="lightskyblue3") + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, alpha=0) + 
+  theme_void() + 
+  theme(legend.position = "none")
+
+dens2 <- ggplot(plotdata, aes(x = out_var)) + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, fill = "white") +
+  geom_density(fill="lightskyblue3") + 
+  geom_histogram(aes(y = ..density..),
+                 colour = 1, alpha=0) + 
+  theme_void() + 
+  theme(legend.position = "none") + 
+  coord_flip()
+
+dens1 + plot_spacer() + plot1 + dens2 + 
+  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
+
+ggsave("Fig_5_scatter_per_head_milk_emissions_intensity.png",width=7,height=7)
+
 ```
@@
 ```{r varkernel}
-uncertainty::varkernel(in_var = dairy_households$y$pc_milk_yield,
-                        out_var = dairy_households$y$CO2eq_per_milk,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
+#uncertainty::varkernel(in_var = dairy_households$y$pc_milk_yield,
+#                        out_var = dairy_households$y$CO2eq_per_milk,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
+
+ggplot(plotdata, aes(x = in_var, y = out_var)) +
+  stat_density_2d(
+    geom = "raster",
+    aes(fill = after_stat(density)),
+    contour = FALSE
+    ) +
+  scale_fill_viridis_c() + 
+  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw()
+
+ggsave("Fig_6_surface_per_head_milk_emissions_intensity.png",width=7,height=7)
+
 ```
@@
 
-graphics::plot(x = slice$slice$Output_values, 
-                 y = slice$slice$Relative_probability, 
-                 type = "l", 
-                 ylab = "Relative probability", 
-                 xlab = "Emissions intensity", 
-                 col = "seagreen", 
-                 lwd = 2)
+#graphics::plot(x = slice$slice$Output_values, 
+#                 y = slice$slice$Relative_probability, 
+#                 type = "l", 
+#                 ylab = "Relative probability", 
+#                 xlab = "Emissions intensity", 
+#                 col = "seagreen", 
+#                 lwd = 2)
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~density~of~milk~production~at~2000~kg~milk~per~capita~(XX~CO[2]-eq~kg^-1)))
-
-ggsave("Fig_X_emission_density_all_farms.png")
+    xlab(expression(Emission~intensity~of~milk~production~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1)))
+
+ggsave("Fig_7_emission_density_all_farms.png")
 
@@
 
-ggsave("Fig_X1_probability_of_intensity_reduction.png")
+ggsave("Fig_8_probability_of_intensity_reduction.png", width=7,height=5)
 
@@
   xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
+
+ggsave("Fig_9_breeds.png", width=7,height=5)
 
@@
 
-ggsave("Fig_XX_breed_reductions.png")
+ggsave("Fig_10_breed_reductions.png", width=7,height=5)
 
@@
   ylab("Milk yield increase per head and year (kg)") 
-ggsave("Fig_XX1_breed_probability_of_intensity_reduction.png")
+ggsave("Fig_11_breed_probability_of_intensity_reduction.png", width=7,height=5)
 

### 2022-06-01 Manuscript_analyses.Rmd
finished ggizing

@@
 options(knitr.kable.NA = '', encoding = "Windows-1252")
- read.csv(input_table)[,c(1:2,4:6)]   %>%
+ read.csv(input_table)[,c(1:2,4,7,5:6)]   %>%
     kableExtra::kbl(digits = 2)  %>%
@@
 
-precise_estimates<-c("Ca_pasture","Ca_large_grazing","C_intact_males","C_castrates","C_females","maintenance_coefficient_Cfi_other",
-                     "maintenance_coefficient_Cfi_lact_cow","maintenance_coefficient_Cfi_bulls","mature_weight_MW_female","mature_weight_MW_male",
-                     "milk_fat","Cp_rate_4","Cp_rate_5","Cp_rate_6","Y_m","Bo_animaltype_1","Bo_animaltype_2","Bo_animaltype_3","Bo_animaltype_4",
-                     "Bo_animaltype_5","Bo_animaltype_6","Bo_animaltype_7","Bo_animaltype_8","Bo_animaltype_9","Bo_animaltype_10","Bo_animaltype_11",
-                     "Bo_animaltype_12","MCFpasture","MCFspread","MCFdrylot","MCFsolidstore","MCFcomposted","MCFliquid","MCFbiogas","MCFburn","MCFsold",
-                     "mndec_pasture","mndec_spread","mndec_drylot","mndec_solidstore","mndec_composted","mndec_liquid","mnvc_pasture","mnvc_spread",
-                     "mnvc_drylot","mnvc_solidstore","mnvc_composted","mnvc_liquid","frac_leach_pasture","frac_leach_spread","frac_leach_drylot",
-                     "frac_leach_solidstore","frac_leach_composted","frac_leach_liquid")
-
-precise_scenarios<-data.frame(Variable=c(variables_to_update_const,variables_to_update_posnorm,precise_estimates),param="both")
+precise_inputs<-read.csv(input_table)[,c(1:7)] 
+precise_vars<-precise_inputs[which(!is.na(precise_inputs$IPCC_default)),]
+
+precise_scenarios<-data.frame(Variable=c(variables_to_update_const,variables_to_update_posnorm,precise_vars$variable),param="both")
+
+                                      #   precise_estimates),param="both")
 
@@
 
-   precise_scenarios[which(precise_scenarios$Variable=="Ca_pasture"),coln]<-0.17
-   precise_scenarios[which(precise_scenarios$Variable=="Ca_large_grazing"),coln]<-0.36
-   precise_scenarios[which(precise_scenarios$Variable=="C_intact_males"),coln]<-1.2   
-   precise_scenarios[which(precise_scenarios$Variable=="C_castrates"),coln]<-1.0   
-   precise_scenarios[which(precise_scenarios$Variable=="C_females"),coln]<-0.8   
-   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_other"),coln]<-0.322   
-   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_lact_cow"),coln]<-0.386
-   precise_scenarios[which(precise_scenarios$Variable=="maintenance_coefficient_Cfi_bulls"),coln]<-0.370
-   precise_scenarios[which(precise_scenarios$Variable=="mature_weight_MW_female"),coln]<-349 
-   precise_scenarios[which(precise_scenarios$Variable=="mature_weight_MW_male"),coln]<-483   
-   precise_scenarios[which(precise_scenarios$Variable=="milk_fat"),coln]<-4 
-   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_4"),coln]<-0.1  
-   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_5"),coln]<-0.1   
-   precise_scenarios[which(precise_scenarios$Variable=="Cp_rate_6"),coln]<-0.1
-   precise_scenarios[which(precise_scenarios$Variable=="Y_m"),coln]<-0.06
-
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_1"),coln]<-0.13
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_2"),coln]<-0.13
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_3"),coln]<-0.13
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_4"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_5"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_6"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_7"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_8"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_9"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_10"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_11"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="Bo_animaltype_12"),coln]<-0.13   
-   precise_scenarios[which(precise_scenarios$Variable=="MCFpasture"),coln]<-0.0047   
-   precise_scenarios[which(precise_scenarios$Variable=="MCFspread"),coln]<-0.01
-   precise_scenarios[which(precise_scenarios$Variable=="MCFdrylot"),coln]<-0.02     
-   precise_scenarios[which(precise_scenarios$Variable=="MCFsolidstore"),coln]<-0.05
-   precise_scenarios[which(precise_scenarios$Variable=="MCFcomposted"),coln]<-0.025
-   precise_scenarios[which(precise_scenarios$Variable=="MCFliquid"),coln]<-0.8   
-   precise_scenarios[which(precise_scenarios$Variable=="MCFbiogas"),coln]<-0.1085   
-   precise_scenarios[which(precise_scenarios$Variable=="MCFburn"),coln]<-0.1   
-   precise_scenarios[which(precise_scenarios$Variable=="MCFsold"),coln]<-0   
-   precise_scenarios[which(precise_scenarios$Variable=="mndec_pasture"),coln]<-0.02   
-   precise_scenarios[which(precise_scenarios$Variable=="mndec_spread"),coln]<-0   
-   precise_scenarios[which(precise_scenarios$Variable=="mndec_drylot"),coln]<-0.02   
-   precise_scenarios[which(precise_scenarios$Variable=="mndec_solidstore"),coln]<-0.01   
-   precise_scenarios[which(precise_scenarios$Variable=="mndec_composted"),coln]<-0.01   
-   precise_scenarios[which(precise_scenarios$Variable=="mndec_liquid"),coln]<-0.01   
-   precise_scenarios[which(precise_scenarios$Variable=="mnvc_pasture"),coln]<-0.02
-   precise_scenarios[which(precise_scenarios$Variable=="mnvc_spread"),coln]<-0.07
-   precise_scenarios[which(precise_scenarios$Variable=="mnvc_drylot"),coln]<-0.3
-   precise_scenarios[which(precise_scenarios$Variable=="mnvc_solidstore"),coln]<-0.45   
-   precise_scenarios[which(precise_scenarios$Variable=="mnvc_composted"),coln]<-0.65   
-   precise_scenarios[which(precise_scenarios$Variable=="mnvc_liquid"),coln]<-0.3   
-   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_pasture"),coln]<-0.24   
-   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_spread"),coln]<-0.24   
-   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_drylot"),coln]<-0.24   
-   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_solidstore"),coln]<-0.24   
-   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_composted"),coln]<-0.24  
-   precise_scenarios[which(precise_scenarios$Variable=="frac_leach_liquid"),coln]<-0.24
+   for(vv in 1:nrow(precise_vars))
+     precise_scenarios[which(precise_scenarios$Variable==precise_vars$variable[vv]),coln]<-
+     precise_vars$IPCC_default[vv]
 
@@
 
-```{r, fig.show="hold", out.width="50%", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE}
-
-uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
-                        out_var = nobull_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
-                        out_var = nobull_res$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+```{r, fig.show="hold", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE, warning=FALSE}
+
+nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
+
+#uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
+#                        out_var = nobull_res$y$CO2eq_per_milk,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+#uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+#                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+#uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
+#                        out_var = nobull_res$y$pc_on_farm,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+#uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
+#                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
+#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
+#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
+
+plotdata_nobull<-data.frame(in_var=nobull_res$y$pc_milk_yield,
+                     out_var=nobull_res$y$CO2eq_per_milk)
+
+plotdata_baseline<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
+                     out_var=GHG_simulation_scenarios$y$CO2eq_per_milk)
+
+plot_nobull <- ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) + 
+  geom_point(shape = 20, color = "blue", size = 2) + 
+  stat_smooth(method = "loess", fullrange = TRUE) +
+  geom_density_2d() +
+  geom_rug(color="lightskyblue3") + 
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() +
+  theme(legend.position = c(0.15, 0.9)) + ggtitle("'No bulls' scenario")
+
+plot_baseline <- ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) + 
+  geom_point(shape = 20, color = "blue", size = 2) + 
+  stat_smooth(method = "loess", fullrange = TRUE) +
+  geom_density_2d() +
+  scale_y_continuous(position = "right") +
+  geom_rug(color="lightskyblue3", sides="bl") + 
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() +
+  theme(legend.position = c(0.15, 0.9)) + ggtitle("Baseline scenario")
+
+dens_nobull<-ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) +
+  stat_density_2d(
+    geom = "raster",
+    aes(fill = after_stat(density)),
+    contour = FALSE
+    ) +
+  scale_fill_viridis_c() + 
+  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw()
+
+dens_baseline<-ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) +
+  stat_density_2d(
+    geom = "raster",
+    aes(fill = after_stat(density)),
+    contour = FALSE
+    ) +
+    scale_y_continuous(position = "right") +
+  scale_fill_viridis_c() + 
+  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw()
+
+plot_nobull + plot_baseline + dens_nobull + dens_baseline +
+  plot_layout(ncol = 2, nrow = 2) + guides(fill="none")
+
+ggsave("Fig_12_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
 
@@
   ylab("Milk yield increase per head and year (kg)")
-ggsave("Fig_XX1_breed_probability_of_intensity_reduction.png")
-
-ggsave("Fig_XX_impact_of_nobull_reductions.png")
+
+ggsave("Fig_13_impact_of_nobull_reductions.png")
 
@@
   ylab("Milk yield increase per head and year (kg)")
-ggsave("Fig_XX1_breed_probability_of_intensity_reduction.png")
-
-ggsave("Fig_XX_lessReplace_reductions.png")
+
+ggsave("Fig_14_lessReplace_reductions.png")
 

### 2022-06-01 Manuscript_analyses.Rmd
some polishing

@@
  - references/papers.bib
-output: html_document
+output:
+  html_document:
+    theme: united
+    toc: true
+    toc_float: true
+    number_sections: true
 ---
@@
 
-# Load emission calculation function
-
-The formulas come from Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
+## Load emission calculation function
+
+Greenhouse gas emissions from livestock are calculated according to Chapter 10 of the *2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@dong_ipcc_2006] (Volume 4 - Agriculture, Forestry and Other Land Use).
 The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
@@
 Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].
-We run the model with the decisionSupport package [@R-decisionSupport].
+We also draw on recommendations provided in Chapter 10 of Volume 4 of the *2019 Refinement to the 2006 IPCC Guidelines for National Greenhouse Gas Inventories* [@ipcc_2019], which is also accessible via [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2019rf/vol4.html).
 
@@
 Here, we express this uncertainty by defining distributions for each variable.
-The following table shows all the variables, as well as examples of distributions that may describe their quantities.
+The following table (**data/livestock_ghg_input_table.csv**) shows all the variables, as well as examples of distributions that may describe their quantities.
 Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals.
 
-```{r load_data}
-
-### # This project will be stored in relative file paths (i.e. same or lower folders)
+```{r load_data, include=FALSE}
 
@@
     kableExtra::kbl(digits = 2)  %>%
-    kableExtra::kable_classic_2(full_width = F)
-
-```
-
-# Monte Carlo simulation
+    kableExtra::kable_classic_2(full_width = F) %>% 
+ scroll_box(height = "300px")
+
+```
+
+## Monte Carlo simulation
 
@@
 
-# Running a Monte Carlo simulation with scenarios
+## Running a Monte Carlo simulation with scenarios
 
@@
 The `scenario_mc` function in the `decisionSupport` package [@R-decisionSupport] consists of a wrapper around the `mcSimulation` function.
-The function is able to modify the input table according to a scenario file we provide.
+The function can modify the input table according to a scenario file we provide.
 This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
@@
 The .csv file for the `scenarios` option specifies variables that should be modified for each of the scenarios, as well as the number of model runs for it.
-Each column beyond the first two columns in the file contain information for a particular scenario.
+Each column beyond the first two columns in the file contains information for a particular scenario.
 This is restricted to the parameters that *should be changed* for this scenario.
@@
 
-# Household-based scenarios
+## Household-based scenarios
 
@@
 
+## Uncertainty at the farm scale
+
 To illustrate the possible influence of input uncertainty on estimates of total farm emissions, we show the example of farm #5 from the household survey.
@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 farm5_emissions$y[,"CO2eq_per_milk"]<-farm5_emissions$y$pc_on_farm/farm5_emissions$y$pc_milk_yield
@@
 
-# Population-wide emission density of milk production
+## Population-wide emission density of milk production
 
@@
 ```{r, warning=FALSE, message=FALSE}
-#uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-#                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
 plotdata<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
@@
 
-# Emissions intensity of milk production
+## Emissions intensity of milk production
 
@@
 ```{r, warning=FALSE, message=FALSE}
-
-#uncertainty::varscatter(in_var = dairy_households$y$pc_milk_yield,
-#                        out_var = dairy_households$y$CO2eq_per_milk,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
 
@@
 
-# Probability of GHG emissions for a given milk yield
+# Milk yield as an indicator of GHG emissions intensity
+
+## Probability of GHG emissions for a given milk yield
 
@@
 
-#graphics::plot(x = slice$slice$Output_values, 
-#                 y = slice$slice$Relative_probability, 
-#                 type = "l", 
-#                 ylab = "Relative probability", 
-#                 xlab = "Emissions intensity", 
-#                 col = "seagreen", 
-#                 lwd = 2)
-
 ggplot(data=data.frame(CO2eq_per_milk=slice$resampled), aes(CO2eq_per_milk)) +
@@
 
-# Probability of an increase in milk yield being associated with an increase in GHG emission intensity
+## Probability of an increase in milk yield being associated with an increase in GHG emission intensity
 
@@
 
-# Testing interventions - strategy
+# Methodology for testing interventions
 
@@
 Make a figure showing the Type I and Type II errors resulting from each confidence level at each baseline yield.
-
-# Testing interventions
 
@@
 
-## Breed intervention
+# Emission reductions through higher-performance breeds
 
@@
 
-# Compare outcomes of breed intervention with baseline
+## Emissions intensity response to introduction of high-performance breeds
 
@@
 
-ggplot(data=impact_of_breed,aes(x=milk_change,y=emissions_intensity_change)) + geom_point() + theme_bw() + ylab("Change in emissions intensity (kg CO2-eq kg-1)") + xlab("Change in per capita milk yield (kg)")
-
-ggsave("Fig_10_breed_reductions.png", width=7,height=5)
-
-```
-
-```{r}
-
-ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase)) + geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+breed_scatter<-ggplot(data=impact_of_breed,
+                      aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Milk yield and emissions response") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+breed_probs<-ggplot(data = decreased_emissions_probability,
+                    aes(Base_yield, yield_increase)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),
+               aes(z=probability),col="black") +
   geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=milk_change)) +
   ggtitle("Probability that emission intensity reductions occurred") + 
-  xlab("Baseline milk yield per head and year (kg)") +
-  ylab("Milk yield increase per head and year (kg)") 
-ggsave("Fig_11_breed_probability_of_intensity_reduction.png", width=7,height=5)
+  labs(y=expression(Milk~yield~increase~(kg~head^{-1}~year^{-1})),
+       x=expression(Baseline~milk~yield~(kg~head^{-1}~year^{-1})),
+       fill="Probability")
+
+breed_scatter + breed_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_10_breed_reductions.png", width=9,height=9)
 
@@
 
-## Herd structure
-
-### Retiring unproductive bulls
+# Herd structure adjustments
+
+## Retiring unproductive bulls
 
@@
 
-#### Evaluate impact of "Retiring unproductive bulls" intervention
+## Emissions intensity response to retiring of unproductive bulls
 
@@
 
-```{r, fig.show="hold", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE}
 
 nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
-
-#uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
-#                        out_var = nobull_res$y$CO2eq_per_milk,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-#uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-#                        out_var = GHG_simulation_scenarios$y$CO2eq_per_milk,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-#uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
-#                        out_var = nobull_res$y$pc_on_farm,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
-
-#uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
-#                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
 
@@
   scale_fill_viridis_c() + 
-  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
        x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw()
+  theme_bw() + guides(fill="none")
 
@@
   scale_fill_viridis_c() + 
-  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
        x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw()
+  theme_bw() + guides(fill="none")
 
 plot_nobull + plot_baseline + dens_nobull + dens_baseline +
-  plot_layout(ncol = 2, nrow = 2) + guides(fill="none")
-
-ggsave("Fig_12_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
-
-```
-
-### Reducing unproductive replacement males
+  plot_layout(ncol = 2, nrow = 2)
+
+ggsave("Fig_11_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
+
+```
+
+## Reducing unproductive replacement males
 
@@
 
-# Compare outcomes of herd intervention with baseline
-
-## No bulls scenario
+## Herd interventions vs. baseline scenario
+
+### No bulls scenario
 
@@
 
-impact_of_nobull<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=nobull_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=nobull_res$y$CO2eq_per_milk)
-impact_of_nobull[,"milk_change"]<-round(impact_of_nobull$breed_milk-impact_of_nobull$baseline,2)
-impact_of_nobull[,"emissions_intensity_change"]<-round(impact_of_nobull$CO2_intensity_breed-impact_of_nobull$CO2_intensity_baseline ,3)
-
-ggplot(data=impact_of_nobull,aes(x=milk_change,y=emissions_intensity_change)) + geom_point() + theme_bw() + ylab("Change in emissions intensity (kg CO2-eq kg-1)") + xlab("Change in per capita milk yield (kg)")
-
-ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase)) + geom_contour_filled(aes(z=probability)) +
+impact_of_nobull<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                             breed_milk=nobull_res$y$pc_milk_yield,
+                             CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                             CO2_intensity_breed=nobull_res$y$CO2eq_per_milk)
+impact_of_nobull[,"milk_change"]<-
+  round(impact_of_nobull$breed_milk-impact_of_nobull$baseline,2)
+impact_of_nobull[,"emissions_intensity_change"]<-
+  round(impact_of_nobull$CO2_intensity_breed-impact_of_nobull$CO2_intensity_baseline ,3)
+
+nobull_scatter<-ggplot(data=impact_of_nobull,
+                       aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Milk yield and emissions response") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+nobull_probs<-ggplot(data = decreased_emissions_probability,
+                     aes(Base_yield, yield_increase)) +
+  geom_contour_filled(aes(z=probability)) +
   stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],aes(x=baseline,y=milk_change)) +
+  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],
+             aes(x=baseline,y=milk_change)) +
   ggtitle("Probability that emission intensity reductions occurred") + 
-  xlab("Baseline milk yield per head and year (kg)") +
-  ylab("Milk yield increase per head and year (kg)")
-
-ggsave("Fig_13_impact_of_nobull_reductions.png")
-
-```
-
-## Fewer replacement males scenario
+  labs(y=expression(Milk~yield~increase~(kg~head^{-1}~year^{-1})),
+       x=expression(Baseline~milk~yield~(kg~head^{-1}~year^{-1})),
+       fill="Probability")
+
+nobull_scatter + nobull_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_12_impact_of_nobull_reductions.png", width=9,height=5)
+
+```
+
+### Fewer replacement males scenario
 
@@
 
-impact_of_lessReplace<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=lessReplace_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=lessReplace_res$y$CO2eq_per_milk)
-impact_of_lessReplace[,"milk_change"]<-round(impact_of_lessReplace$breed_milk-impact_of_lessReplace$baseline,2)
-impact_of_lessReplace[,"emissions_intensity_change"]<-round(impact_of_lessReplace$CO2_intensity_breed-impact_of_lessReplace$CO2_intensity_baseline ,3)
-
-ggplot(data=impact_of_lessReplace,aes(x=milk_change,y=emissions_intensity_change)) + geom_point() + theme_bw() + ylab("Change in emissions intensity (kg CO2-eq kg-1)") + xlab("Change in per capita milk yield (kg)")
-
-ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase)) + geom_contour_filled(aes(z=probability)) +
+impact_of_lessReplace<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  breed_milk=lessReplace_res$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_breed=lessReplace_res$y$CO2eq_per_milk)
+impact_of_lessReplace[,"milk_change"]<-
+  round(impact_of_lessReplace$breed_milk-impact_of_lessReplace$baseline,2)
+impact_of_lessReplace[,"emissions_intensity_change"]<-
+  round(impact_of_lessReplace$CO2_intensity_breed-impact_of_lessReplace$CO2_intensity_baseline ,3)
+
+lessRep_scatter<-ggplot(data=impact_of_lessReplace,
+                        aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+lessRep_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, yield_increase)) +
+  geom_contour_filled(aes(z=probability)) +
   stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],aes(x=baseline,y=milk_change)) +
+  geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],
+             aes(x=baseline,y=milk_change)) +
   ggtitle("Probability that emission intensity reductions occurred") + 
-  xlab("Baseline milk yield per head and year (kg)") +
-  ylab("Milk yield increase per head and year (kg)")
-
-ggsave("Fig_14_lessReplace_reductions.png")
+  labs(y=expression(Milk~yield~increase~(kg~head^{-1}~year^{-1})),
+       x=expression(Baseline~milk~yield~(kg~head^{-1}~year^{-1})),
+       fill="Probability")
+
+lessRep_scatter + lessRep_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_13_lessReplace_reductions.png", width=9,height=9)
 

### 2022-06-02 Manuscript_analyses.Rmd
cleaned up and reformated html

@@
     number_sections: true
+    code_folding: hide
 ---
@@
 
-```{r plot_distributions, fig.width=10, fig.height=10}
+```{r plot_distributions, fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
 # enteric_CH4=enteric_CH4,
@@
 
-```{r varkernelslicerange}
+```{r varkernelslicerange, message=FALSE, warning=FALSE}
 slice<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
@@
 Let's look at this for production of 2000 and 3000 kg milk per capita per year.
-We'll take 1000 random draws of the emissions intensities that associated with both levels and compare values among these samples.
+We'll take 1000 random draws of the emissions intensities that are associated with both levels and compare values among these samples.
 
@@
 
-sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)
+# sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)
 
@@
 
-```{r}
+```{r, warning=FALSE}
 cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
@@
 
-```{r}
+```{r, fig.dim = c(8, 8)}
 breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
@@
 
-```{r, message=FALSE, warning=FALSE}
+```{r, message=FALSE, warning=FALSE, fig.dim = c(8, 8)}
 
@@
 
-```{r}
+```{r, fig.dim = c(8, 8)}
 nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
@@
 
-```{r}
+```{r, fig.dim = c(8, 8)}
 

### 2022-06-03 Manuscript_analyses.Rmd
cleaned up code - for comments

@@
 
-# Set up
-
-Install and load required packages
+# Setting up
+
+Note that the code for all operations is hidden, so that it doesn't clutter the document. You can make it appear by clicking the code button on the right.
+
+We first need to install and load all required packages.
 
@@
 
+The equations provided in these documents, in particular in the 2006 guideline document [@dong_ipcc_2006] were converted into an R function called `ghg_emissions`, which is defined here (hit the `Code` button on the right).
+
 ```{r ghg_emissions_function}
 # Generate the GHG emissions function 
-source("GHG_emissions_function.R", local = knitr::knit_global())
+ghg_emissions<-function(x, varnames){
+
+  # Enteric fermentation emissions ($CH_4$)
+
+  # The main source of all this is Chapter 10 of the *2006 IPCC Guidelines 
+  # for National Greenhouse Gas Inventories* 
+  # (Volume 4 - Agriculture, Forestry and Other Land Use). 
+  # The document can be found on the IPCC web 
+  # https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf
+
+  # We start with the net energy for maintenance (NE_m; in MJ/day)
+  # This includes a maintenance coefficient, 
+  # for which different values are recommended depending on animal type
+
+  # The following animal types were used in the example:
+
+  # animal type 1 - 'intact' grown-up males >=36 months
+  # animal type 2 - castrated grown-up males >=36 months
+  # animal type 3 - males 12-36 months
+  # animal type 4 - lactating cows
+  # animal type 5 - lactating cows
+  # animal type 6 - lactating cows
+  # animal type 7 - weaned young females <= 12 months
+  # animal type 8 - weaned young males <= 12 months
+  # animal type 9 - females 12-36 months
+  # animal type 10 - pre-weaned female calves (<12 months)
+  # animal type 11 - pre-weaned male calves (<12 months)
+  # animal type 12 - first-time pregnant cows
+
+  # animal types 4, 5 and 6 are all lactating cows - 
+  # important to note here that the original data table 
+  # appears to have worked with the actual number of months 
+  # that cows were pregnant. This resulted in a weighted mean 
+  # of the various maintenance coefficients. 
+  # We're not sure where we could get such data. 
+  # In the original dataset, types 4, 5 and 6 
+  # referred to different pregnancy/lactation states, 
+  # but it doesn't seem like this scheme was rigorously followed.
+
+  # We make placeholder vectors to collect emissions data 
+  # for all animal types
+  mm_N2O_CO2eq<-c()
+  mm_CH4_CO2eq<-c()
+  enteric_CH4_CO2eq<-c()
+  total_milk<-c()
+
+  # Now we start a long loop, 
+  # where emissions are calculated for each animal type
+
+  for (animal_type in 1:12)
+  {
+    # define C and Cfi based on sex (Page 10.17, below Equation 10.6)
+    if(animal_type %in% c(1,3,8,11)) growth_coefficient_C<-C_intact_males # males (1.2)
+    if(animal_type %in% c(2)) growth_coefficient_C<-C_castrates # castrates (1.0)
+    if(animal_type %in% c(4,5,6,7,9,10,12)) growth_coefficient_C<-C_females # females (0.8)
+
+    if(animal_type %in% c(4,5,6)) Cfi<-maintenance_coefficient_Cfi_lact_cow # lactating cows (0.386)
+    if(animal_type %in% c(1)) Cfi<-maintenance_coefficient_Cfi_bulls # bulls (0.37)
+    if(animal_type %in% c(2,3,7,8,9,10,11,12)) Cfi<-maintenance_coefficient_Cfi_other # others (0.322)
+
+    # define mature body weight (assuming this needs to be disaggregated by sex, for Eqation 10.6)
+    if(animal_type %in% c(1,2,3,8,11)) mature_weight_MW<-mature_weight_MW_male
+    if(animal_type %in% c(4,5,6,7,9,10,12)) mature_weight_MW<-mature_weight_MW_female
+
+    # define milk yield
+    if(animal_type %in% c(1,2,3,7,8,9,10,11,12)) milk_yield<-0
+    if(animal_type %in% c(4,5,6)) milk_yield<-eval(parse(text=paste0("milk_yield_",animal_type)))
+
+    #define pregnancy coefficient (full Cp for first-time pregnant, lower for later in most cases)
+    if(animal_type==12) Cp<-Cp_0
+    if(animal_type==4) Cp<-Cp_0*Cp_rate_4
+    if(animal_type==5) Cp<-Cp_0*Cp_rate_5
+    if(animal_type==6) Cp<-Cp_0*Cp_rate_6
+    if(animal_type %in% c(1,2,3,7,8,9,10,11)) Cp<-0
+
+    N_animal_type<-eval(parse(text=paste0("N_animal_type_",animal_type)))
+
+    # define activity coefficient Ca (Table 10.5)
+    # Stall 0.00 - Animals are confined to a small area 
+    #     (i.e., tethered, pen, barn) with the result that they expend 
+    #     very little or no energy to acquire feed.
+    # Pasture 0.17 - Animals are confined in areas with sufficient forage 
+    #     requiring modest energy expense to acquire feed.
+    # Grazing large areas 0.36 - Animals graze in open range land 
+    #     or hilly terrain and expend significant energy to acquire feed.
+
+    if(feeding_system==1) C_activity<-Ca_stall # Stall (0.0)
+    if(feeding_system==2) C_activity<-Ca_pasture # Stall (0.17)
+    if(feeding_system==3) C_activity<-Ca_large_grazing # Stall (0.36)
+
+    # Mean live weight (kg) of animals surely varies by animal type, 
+    # so we need separate estimates for them
+    live_weight_LW<-eval(parse(text=paste0("live_weight_LW_",animal_type)))
+
+    # Mean weight gain of animal type (kg/day)
+    weight_gain_WG<-eval(parse(text=paste0("weight_gain_WG_",animal_type)))
+
+    # Calculate the Net Energy for Maintenance (NE_m) (MJ/day) (Equation 10.3)
+    NE_m<-Cfi*(live_weight_LW^0.75)
+
+    # Calculate the Net Energy for Activity (NE_activity) (MJ/day) (Equation 10.4)
+    NE_activity<-C_activity*NE_m
+
+    #Net energy for growth (NE_growth) (MJ/day) (Equation 10.6)
+    NE_growth<-22.02*
+      ((live_weight_LW/(growth_coefficient_C*mature_weight_MW))^0.75)*
+      (weight_gain_WG^1.097)
+
+    # Net energy for lactation
+    NE_lactation<-milk_yield*(1.47+0.40*milk_fat)
+
+    #Energy need during pregnancy (Equation 10.13)
+    NE_pregnancy<-Cp*NE_m
+
+    # Ratio of net energy available in a diet for maintenance 
+    # to digestible energy consumed (Equation 10.14)
+    REM<-1.123-(4.092/1000*DE_perc)+(1.126/100000*DE_perc^2)-(25.4/DE_perc)
+
+    # Ratio of net energy available for growth in a diet 
+    # to digestible energy consumed (Equation 10.15)
+    REG<-1.164-(5.160/1000*DE_perc)+(1.308/100000*DE_perc^2)-(37.4/DE_perc)
+
+    # Gross Energy (GE) MJ/day
+    NE_work<-0 # work animals aren't covered here, so we set this to zero
+    GE<-(((NE_m+NE_activity+NE_lactation+NE_work+NE_pregnancy)/REM)+
+           (NE_growth/REG))/(DE_perc/100)
+
+    # CH4 emission factors for enteric fermentation from a livestock category 
+    # (Equation 10.21)
+    enteric_CH4<-(GE*Y_m*365)/55.65 # kg CH4 per animal per year
+
+    # Compute the CO2-equivalent of methane emissions for enteric methane
+    # This converts methane emissions to their Global Warming Potential 
+    # (CO2-equivalent of CH4 is 25)
+    enteric_CH4_CO2eq[animal_type]<-N_animal_type*enteric_CH4*25
+
+    # Manure management methane ($CH_4$) and nitrous oxide ($N_2O$)
+    # daily volatile solid excreted (Equation 10.24)
+
+    volsol<-(GE*
+               (1-(DE_perc/100))+
+               (urinary_energy*GE))*
+      ((1-ash_content)/18.45)
+
+    # The maximum methane emission from manure (Bo) depends on the animal type. 
+    # In the input table, we treat these as constants for each animal type. 
+    # The following line assigns the corresponding Bo value
+
+    Bo<-eval(parse(text=paste0("Bo_animaltype_",animal_type)))
+
+    # CH4 emission factor from manure management (Equation 10.23)
+    mm_ch4<-(volsol*365)*
+      (Bo*0.67*
+         ((mms_pasture*MCFpasture)+
+            (mms_spread*MCFspread)+
+            (mms_drylot*MCFdrylot)+
+            (mms_solidstore*MCFsolidstore)+
+            (mms_composted*MCFcomposted)+
+            (mms_liquid*MCFliquid)+
+            (mms_biogas*MCFbiogas)+
+            (mms_burn*MCFburn)+
+          # the original code was lacking the 'sold' fraction - 
+          # we added this for completeness
+          (mms_sold*MCFsold)))
+
+    # Estimate CO2eq for manure methane
+    mm_CH4_CO2eq[animal_type]<-N_animal_type*mm_ch4*25
+
+    # Calculate Nitrous oxide emissions from manure and urine on pasture
+
+    # estimate N intake (kg/animal/day) (Equation 10.32)
+    N_intake<-(GE/18.45)*((perc_crude_protein_diet/100)/6.25)
+
+    # estimate N retention (kg/animal/day) (Equation 10.33)
+
+    if(weight_gain_WG==0)
+      N_retention <- ((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)
+
+    if(!weight_gain_WG==0)
+      N_retention<-((milk_yield*(1.9+0.4*milk_fat)/100)/6.38)+
+      ((weight_gain_WG*(268-(7.03*NE_growth/weight_gain_WG)))/(1000*6.25))
+
+    # estimate N excretion 
+    # (kg N/animal/yr) (Equation 10.31)
+    N_excretion<-N_intake*(1-N_retention)*365
+
+    # Direct nitrous oxide emissions 
+    # (kg N2O/animal/yr) (Eq 10.25 from IPCC guidelines)
+    mm_direct_n2o<-N_excretion*((mms_pasture*mndec_pasture)+
+                                  (mms_spread*mndec_spread)+
+                                  (mms_drylot*mndec_drylot)+
+                                  (mms_solidstore*mndec_solidstore)+
+                                  (mms_composted*mndec_composted)+
+                                  (mms_liquid*mndec_liquid)+
+                                  (mms_biogas*mndec_biogas)+
+                                  (mms_burn*mndec_burn)+
+                                  (mms_sold*mndec_sold))*(44/28)
+
+    # N losses from volatilization 
+    # (kg N/animal/yr) (Eq 10.26 from IPCC guidelines)
+    mm_vol_n<-N_excretion*((mms_pasture*mnvc_pasture)+
+                             (mms_spread*mnvc_spread)+
+                             (mms_drylot*mnvc_drylot)+
+                             (mms_solidstore*mnvc_solidstore)+
+                             (mms_composted*mnvc_composted)+
+                             (mms_liquid*mnvc_liquid)+
+                             (mms_biogas*mnvc_biogas)+
+                             (mms_burn*mnvc_burn)+
+                             (mms_sold*mnvc_sold))
+
+    # indirect N2O emissions due to volatilization of N from manure management 
+    # (kg N2O/animal/year) (Equation 10.27)
+    mm_vol_n2o<-mm_vol_n*EF4*(44/28)
+
+    # Nitrous oxide (N2O) from leaching (kg N/animal/year) (equation 10.28)
+    mm_leach_n<-N_excretion*((mms_pasture*frac_leach_pasture)+
+                               (mms_spread*frac_leach_spread)+
+                               (mms_drylot*frac_leach_drylot)+
+                               (mms_solidstore*frac_leach_solidstore)+
+                               (mms_composted*frac_leach_composted)+
+                               (mms_liquid*frac_leach_liquid)+
+                               (mms_biogas*frac_leach_biogas)+
+                               (mms_burn*frac_leach_burn)+
+                               (mms_sold*frac_leach_sold))
+
+    # indirect N2O emissions due to leaching of N from manure management 
+    # (kg N2O/animal/year) (Equation 10.29)
+    mm_leach_n2o<-mm_leach_n*EF5*(44/28)
+
+    # CW EL Comment ####
+    # # comment: we removed this, it was probably wrong, 
+    # since leached nitrogen doesn't emit (much) nitrous oxide
+    #mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
+    #              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)
+
+    # Estimate CO2eq for manure nitrous
+    mm_N2O_CO2eq[animal_type]<-N_animal_type*(mm_direct_n2o+mm_vol_n2o+mm_leach_n2o)*298 # (298 is the CO2-equivalent of N2O)
+
+    total_milk[animal_type]<-N_animal_type*milk_yield*365
+
+  }
+
+  # Emissions from feed production and transport.
+
+  feed_CO2<-feed_prod_CO2+feed_trans_CO2
+
+  # Total emissions
+
+  on_farm<-sum(enteric_CH4_CO2eq+mm_CH4_CO2eq+mm_N2O_CO2eq)+feed_CO2
+
+  total_animals<-sum(sapply(1:12,function(x) eval(parse(text=paste0("N_animal_type_",x)))))
+
+  return(list(enteric_CH4=enteric_CH4,
+              milk_yield=sum(total_milk),
+              enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
+              mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
+              mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
+              feed_CO2=feed_CO2,
+              on_farm=on_farm,
+              pc_enteric_CH4=enteric_CH4/total_animals,
+              pc_milk_yield=sum(total_milk)/total_animals,
+              pc_enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq)/total_animals,
+              pc_mm_CH4_CO2eq=sum(mm_CH4_CO2eq)/total_animals,
+              pc_mm_N2O_CO2eq=sum(mm_N2O_CO2eq)/total_animals,
+              pc_feed_CO2=feed_CO2/total_animals,
+              pc_on_farm=on_farm/total_animals
+
+              ))
+}
+
 ```
@@
 Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals.
+We use the following distribution shapes:
+
+* const - constants (upper and lower bounds are the same)
+* posnorm - positive normal distribution; this is a normal distribution that is truncated at 0.
+
+The decisionSupport package [@R-decisionSupport], which we use to run the simulations, also accommodates other distribution types, but we did not consider these appropriate here.
+The positive normal distribution (and other distributions such as normal or lognormal distributions) are defined based on the 5% and 95% quantiles.
+Note that it isn't always possible to fit a positive normal distribution based on such specifications, especially if the 5% quantile is set so close to zero that the remaining 5% of the data do not **fit** in the space between zero and the lower bound.
+In such cases, decisionSupport [@R-decisionSupport] returns a warning message, but it still does its best to generate the requested distribution.
+It is worth being on the lookout for the respective warning messages and try to adjust the bounds of the distribution to **possible** values.
 
@@
 This file contains information on which variables should be modified for each scenario and how many model runs should be included in the simulation.
-The function is essentially the same as the `mcSimulation` function, but with an additional `scenarios` option which imports the contents of a .csv file.
+The function call is essentially the same as for the `mcSimulation` function, but with an additional `scenarios` option which imports the contents of a .csv file.
 
@@
 Each column beyond the first two columns in the file contains information for a particular scenario.
-This is restricted to the parameters that *should be changed* for this scenario.
+This is restricted to the parameters that *should be changed* for this scenario (all others don't have to be included).
 
@@
 The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both".
-If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables witha constant distribution, since the rplacement will be for both the upper and lower bound).
+If "both" is selected, both the lower and upper bounds are modified (this should only be done for variables with a constant distribution, since the replacement will be for both the upper and lower bound).
 
 We define subgroups based on the populations of households described by a household survey in Kenya [@wilkes_variation_2020; @wilkes_methods_2019].
-Each of the farms encountered in this survey were converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.
+Each of the farms encountered in this survey was converted into one scenario, for which Monte Carlo simulation is then used to produce multiple replicates.
 
@@
 The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019].
-Extracting all relevant information from these spreadsheets is a slightly complex task, which is accomplished by the following code.
+
+### Probabilistic scenarios
+
+Extracting all relevant information from the survey results spreadsheet is a slightly complex task, which is accomplished by the following code.
+The code includes specification of lower and upper bounds of probability distributions for most of the variables.
+We keep the herd composition constant, but all variables that in all likelihood weren't determined with high precision and accuracy are equipped with such uncertainty estimates.
+Estimates of the uncertainty for model input variables were extracted from the IPCC documents, or, if no such indications were available, based on our expert judgement.
 
@@
 
-We'll also make precise scenarios that contain only the exact values that were observed in the survey.
+The first few rows in this table adjust the herd composition parameters, so that the scenario herd corresponds to the herd of the specific household.
+Other parameters including the milk yield per cow are also adjusted.
+
+### Precise scenarios
+
+We also make precise scenarios that contain only the exact values that were observed in the survey.
+This is needed for illustrating the impact of accounting for uncertainty later.
+For these scenarios, we follow the commonly used practice of applying default values recommended by the various IPCC documents mentioned above [@dong_ipcc_2006;@ipcc_2019;@ipcc_good_2021].
 
@@
 
-The first few rows in this table adjust the herd composition parameters, so that the scenario herd corresponds to the herd of the specific household.
-Other parameters including the milk yield per cow are also adjusted.
+## Running the simulation
 
@@
 We simulate 1000 versions of this farm.
+We also do a precise calculation based on the exact numbers returned by the household survey and the default IPCC values.
+
+### Total emissions of the farm
+
+The following code and the resulting distribution estimate the total emissions of farm #5.
 
@@
 
+The light blue distribution illustrates the estimated emissions from farm #5, considering uncertainty about all input variables.
 The blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.
+
+### Emissions intensity
+
+Naturally, the total emissions of the farm depend strongly on the size of the farm.
+A better indicator of the carbon footprint of milk production is the emissions intensity, i.e.
+the emissions per quantity of milk produced.
 
@@
 
-We can illustrate this uncertainty using a scatter plot. The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.
-
-As with the GHG Model, we apply the uncertainty functions on the `milk_yield` and `on_farm` variables.
-These come from the model results (`r nrow(GHG_simulation_scenarios$y)`) of the `GHG_simulation_scenarios` model results to the `in_var` (x) and an outcome variable `out_var` (y).
-In the model we defined `on_farm` as the sum of Carbon Dioxide Equivalent Enteric Methane Emissions from Livestock (`enteric_CH4_CO2eq`), Carbon Dioxide Equivalent Methane Emissions from Livestock Manure (`mm_CH4_CO2eq`), Carbon Dioxide Equivalent Nitrous Oxide Emissions from Livestock Manure (`mm_N2O_CO2eq`) and Carbon Dioxide Emissions from Livestock Feed Production and Transport (`feed_CO2`).
+We can illustrate this uncertainty using a scatter plot.
+The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.
+Total emissions are the sum of methane emissions from enteric fermentation and livestock manure, nitrous oxide emissions from manure and carbon dioxide emissions from livestock feed production and transport.
 
@@
 plotdata<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
-                     out_var=GHG_simulation_scenarios$y$pc_on_farm,
-                     Group="just one")
+                     out_var=GHG_simulation_scenarios$y$pc_on_farm)
 
@@
 
-An indicator that is often of interest is the ratio between these two variables: total emissions divided by the milk yield.
+As mentioned above, an indicator that is often of interest is the ratio between total emissions and the milk yield.
 We'll calculate this next.
@@
 ```{r varkernel}
-#uncertainty::varkernel(in_var = dairy_households$y$pc_milk_yield,
-#                        out_var = dairy_households$y$CO2eq_per_milk,
-#                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
-#                        ylab = expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]-eq~kg^{-1})))
 
@@
 We can now examine the merits of this indicator.
-Essentially, its usefulness hinges on its relationship with greenhouse gas emission intensity.
+Essentially, its usefulness hinges on its relationship with greenhouse gas emissions intensity.
 We can investigate this by taking slices through the density surface.
@@
 
-We see considerable uncertainty about the emissions at this particular level of production, and the previous figure shows that emissions of all other levels is similarly variable.
+We see considerable uncertainty about the emissions at this particular level of production, and the previous figure shows that emissions of all other levels are similarly variable.
 This raises considerable doubts about the usefulness of milk yield as an indicator of emissions intensity - at least when this indicator is taken at face value, without considering this uncertainty.
@@
 
-To derive the level of certainty, with which a change in milk yield is associated with an increase or decrease in emissions intensity, we have to investigate the probability distributions (as shown in Fig.
-X) of emissions intensity at two levels of milk production.
-This can be achieved using the `varslice_resample` function for both production levels.
+To derive the level of certainty, with which a change in milk yield is associated with an increase or decrease in emissions intensity, we have to investigate the probability distributions (as shown in the previous figure) of emissions intensity at two levels of milk production.
+This can be achieved by using the `varslice_resample` function for both production levels.
 
@@
 
+slices<-rbind(data.frame(Values=slice_2000$resampled, Level="2000 kg"), data.frame(Values=slice_3000$resampled, Level="3000 kg"))
+
+ggplot(data=slices, aes(Values, fill=Level)) +
+ geom_density(alpha=.2) +
+  ylab("Relative probability") +
+    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level')
+
+ggsave("Fig_8_emission_density_all_farms.png", width=7,height=5)
+
 # sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)
@@
 
-The comparison in the last line of code above reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.
+Now we want to determine the probability of emissions intensity at a production level of 3000 kg milk per capita and year being lower than emissions intensity at 2000 kg per capita and year.
+We can do this empirically by comparing the two sets of 1000 random values that are assembled in the probability distribution slices representing per capita milk production levels of 3000 and 2000 kg.
+We test for each index number (1-1000) if the respective element of the 3000-kg slice is smaller than the respective element of the 2000-kg slice.
+By adding up the results of all these comparisons (with TRUE coded as 1 and FALSE as 0) and dividing by 1000 (and optionally multiplying by 100% for a percentage), we obtain the probability that emissions intensity is really lower at the higher milk production level.
+
+In this case, this comparison reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.
 
 We can now generalize this result by doing similar calculations for multiple production levels and illustrating them for the whole population.
+We'll base this on slices representing production levels between 100 and 10,000 kg per capita and year, at 100-kg intervals.
 
@@
 
-decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield
-
-ggplot(data = decreased_emissions_probability, aes(Base_yield, yield_increase, z =
+#decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
                                                       probability)) + geom_contour_filled() +
-  ggtitle("Probability that emission intensity reductions occurred") + 
-  xlab("Baseline milk yield per head and year (kg)") +
-  ylab("Milk yield increase per head and year (kg)") +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") +
   stat_contour(breaks = c(0.9),col="black")
 
-ggsave("Fig_8_probability_of_intensity_reduction.png", width=7,height=5)
-
-```
-
-The 90% confidence contour is shown in the plot.
-The results indicate that whenever milk yields increase, the probability that emissions intensity decreased is usually greater than 50% (slightly lower values are due to randomness in the resampling procedure), but that large increases are needed to be sure of this.
+ggsave("Fig_9_probability_of_lower_emissions_intensity.png", width=7,height=5)
+
+```
+
+The results indicate that whenever milk yields are higher on a particular farm than on another one, the probability that emissions intensity is lower is usually greater than 50% (slightly lower values are due to randomness in the resampling procedure), but that large differences are needed to be sure of this.
+The 90% confidence contour (shown in the plot as a black line) indicates the minimum level of milk yield that would be needed to state with 90% certainty that the emissions intensity is lower on the higher-productivity farm than on the respective lower-productivity one.
 
@@
 Programs aiming to effect emissions reductions in dairy production may want to reward such changes, and they may be looking to the milk production per capita as an indicator for the effectiveness of such measures.
-We now want explore whether this is a reasonable strategy.
-
-## Strategy
-
-Develop scenarios that implement innovations
-
-* breed adjustment
-* herd structure
-
-Quantify improvements on farms by comparing with the baseline.
-Plot improvements vs. the baseline on the above figure.
-Plot improvements vs. baseline in the emissions intensity figure.
-
-For each slice, and each level of confidence, calculate the percentage of farms that made changes that would be correctly rewarded vs.
-the ones that did not change (the baseline population).
-Make a figure showing the Type I and Type II errors resulting from each confidence level at each baseline yield.
-
-## Single-run function
-
-To clearly characterize the impact of interventions, we should compare their impact with a counterfactual, i.e.
-with the same farms without the intervention.
-While this is usually impossible in practice, it is easy with a model.
-We just have to run the simulation again with identical inputs except for those that are affected by the intervention.
-
-We already simulated emissions for a large population of farms, and we can now apply our interventions to this population by modifying certain parameters of the input tables.
-What we don't want now, however, is to randomize all values again, since this would add additional noise to the results, which would mask the signal caused by the interventions.
-
-We now first need a function that can apply our greenhouse gas emissions model to a pre-defined set of input values.
-We can find syntax that can achieve this in the mcSimulation function.
-Here it is (with slight modifications):
+We now want to explore whether this is a reasonable strategy.
+
+## Interventions
+
+We will explore the impacts of three interventions that farmers may consider. We'll restrict this assessment to interventions we can easily implement in our modeling framework.
+
+### Breed adjustment
+
+We evaluate milk production for all the breeds found in the survey dataset. All cows of low-productivity breeds are then replaced by high-productivity cows.
+
+### Herd structure intervention 1 - retiring unproductive bulls
+
+All bulls that are beyond their productive life span are removed from the herd. We implement this by removing all animals of animal type 1, which denotes bulls >36 months of age.
+
+### Herd structure intervention 2 - fewer unproductive replacement males
+
+Many farms have more male calves and young bulls than are needed to meet the objectives of dairy production.
+Based on a real-life scenario, we target a reduction in this number by 23%.
+For this, we randomly select 23% of the farms and set the number of replacement males (animal type 3) to 0.
+
+## Analysis strategy
+
+In a model, we can simulate the implications of specific interventions by leaving all variables unchanged, except those that are modified by the intervention.
+To implement this using the decisionSupport package, we have to disable randomness for the variables we don't want to modify.
+For this, we can create a function (`run_model`) that simply applies our model to a set of values we specify.
+We can use this function to exactly reproduce the results of a model run.
+Note that this only works, because our model is perfectly deterministic.
+All variation in model outputs results from variation in inputs; the model itself doesn't include additional random processes.
+
+We can now generate the outputs of a Monte Carlo model for our baseline situation (without applying interventions).
+The output object of the `mcSimulation` function doesn't only contain the output variables, however, but also a complete list of all random inputs for all model runs.
+As mentioned above, we could now apply the `run_model` function to these inputs (the `x` element of the `mcSimulation` result) to reproduce the outputs of the simulation.
+More importantly, we can now modify certain variables included in this `x` element to simulate interventions, and then run the model on these modified inputs to simulate the impacts of these interventions.
+
+The following code contains the `run_model` function:
 
@@
 
-The following code can precisely reproduce the results of the previous analysis:
+The following code can therefore precisely reproduce the results of the previous analysis:
 
@@
 
-# Emission reductions through higher-performance breeds
-
-Introduction of higher-performance breeds may reduce greenhouse gas emissions. We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.
-
-To retain the farm-specific performance levels obtained from the farm survey, we simulate these impacts in a relative manner.
-For this, we first need to determine the performance and characteristics of each of the breeds.
-We'll base this on the farm survey data (`Kenyaherds` dataset from above).
+# Upgrading to higher-performance breeds
+
+Replacing cows of low-productivity breeds with better-performing ones may reduce greenhouse gas emissions.
+We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.
+
+We first need to determine the performance and characteristics of each of the breeds that are present in the farm survey dataset (`Kenyaherds` dataset from above).
 
@@
 
-ggsave("Fig_9_breeds.png", width=7,height=5)
+ggsave("Fig_10_breeds.png", width=7,height=5)
 
@@
 colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
-milk_summary
+
+milk_summary  %>%
+    kableExtra::kbl(digits = 2)  %>%
+    kableExtra::kable_classic_2(full_width = F)
 
@@
 
-Let's define the intervention as follows:
+We define the intervention as follows:
 
 All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences.
-This means that milk yields are multiplied by the following factors:
+This means that milk yields are divided by the following factors:
 
@@
 
-milk_summary             
+milk_summary   %>%
+    kableExtra::kbl(digits = 2)  %>%
+    kableExtra::kable_classic_2(full_width = F)            
 ```
@@
 
-milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows]), FUN=mean)
-
-breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% milk_change[,1])]<-milk_change[,2]
-breed_intervention$milk_yield_5<-breed_intervention$milk_yield_4
-breed_intervention$milk_yield_6<-breed_intervention$milk_yield_4
+milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows], Kenyaherds_interventions$animal_type[cows]), FUN=mean)
+
+Milk_changes_4<-milk_change[which(milk_change$Group.2==4),]
+breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% Milk_changes_4[,1])]<-Milk_changes_4[,3]
+Milk_changes_5<-milk_change[which(milk_change$Group.2==5),]
+breed_intervention$milk_yield_5[which(breed_intervention$Farm %in% Milk_changes_5[,1])]<-Milk_changes_5[,3]
+Milk_changes_6<-milk_change[which(milk_change$Group.2==6),]
+breed_intervention$milk_yield_6[which(breed_intervention$Farm %in% Milk_changes_6[,1])]<-Milk_changes_6[,3]
 
@@
 
-Go through the x file, for each row take a random draw based on adoption rate (does farm adopt or not), then (possibly) apply factors.
-Actually, in this case we want to simulate full adoption, so we'll set this to 1.
+We can now go through the `x` file, and for each row take a random draw based on an adoption rate we can specify (does the farm adopt the intervention or not), then (possibly) apply factors.
+In this case we want to simulate full adoption, so we'll set this to 1.
 
@@
 
-## Emissions intensity response to introduction of high-performance breeds
-
-```{r, fig.dim = c(8, 8)}
+## Emissions intensity response to upgrading breeds
+
+Exchanging cows of unproductive breeds to more productive individuals should have an effect on emissions intensity, yet this effect is difficult to see at the population level, both in the scatter plot and in the surface plot:
+
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 8)}
 breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
 
-breed_res$y<-breed_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-breed_res$x<-breed_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-
-impact_of_breed<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=breed_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=breed_res$y$CO2eq_per_milk)
-impact_of_breed[,"milk_change"]<-round(impact_of_breed$breed_milk-impact_of_breed$baseline,2)
-impact_of_breed[,"emissions_intensity_change"]<-round(impact_of_breed$CO2_intensity_breed-impact_of_breed$CO2_intensity_baseline ,3)
-
-breed_scatter<-ggplot(data=impact_of_breed,
-                      aes(x=milk_change,y=emissions_intensity_change)) +
-  geom_point() +
-  theme_bw() +
-  ggtitle("Milk yield and emissions response") + 
-  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)")
-
-breed_probs<-ggplot(data = decreased_emissions_probability,
-                    aes(Base_yield, yield_increase)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),
-               aes(z=probability),col="black") +
-  geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=milk_change)) +
-  ggtitle("Probability that emission intensity reductions occurred") + 
-  labs(y=expression(Milk~yield~increase~(kg~head^{-1}~year^{-1})),
-       x=expression(Baseline~milk~yield~(kg~head^{-1}~year^{-1})),
-       fill="Probability")
-
-breed_scatter + breed_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_10_breed_reductions.png", width=9,height=9)
-
-```
-
-This plot shows a lot of linear patterns that result from the application of fixed factors to convert the milk yield between cows of different breeds. Not so pretty, but maybe ok.
-
-# Herd structure adjustments
-
-## Retiring unproductive bulls
-
-To simulate this scenario, we simply have to remove all animals of type 1 from the herds.
-
-```{r , warning=FALSE, eval=FALSE}
-x_nobull_intervention <- GHG_simulation_scenarios$x
-
-nobull_adoption_rate <- 1.0
-
-for (i in 1:nrow(x_nobull_intervention))
-{
-  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
-    x_nobull_intervention$N_animal_type_1[i]<-0
-}
-# run model for modified x file
-nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-nobull_res$call<-"multi-scenario call"
-save_temperature_scenarios(nobull_res,"results","noBull_intervention")
-```
-
-```{r, include=FALSE}
-nobull_res<-load_temperature_scenarios("results","noBull_intervention")
-class(nobull_res)<-"mcSimulation"
-```
-
-## Emissions intensity response to retiring of unproductive bulls
-
-We're assuming that all bulls >36 months old have been removed.
-
-```{r, message=FALSE, warning=FALSE, fig.dim = c(8, 8)}
-
-nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
-
-plotdata_nobull<-data.frame(in_var=nobull_res$y$pc_milk_yield,
-                     out_var=nobull_res$y$CO2eq_per_milk)
+plotdata_breed<-data.frame(in_var=breed_res$y$pc_milk_yield,
+                     out_var=breed_res$y$CO2eq_per_milk)
 
@@
 
-plot_nobull <- ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) + 
+plot_nobull <- ggplot(plotdata_breed, aes(x = in_var, y = out_var)) + 
   geom_point(shape = 20, color = "blue", size = 2) + 
@@
   theme_bw() +
-  theme(legend.position = c(0.15, 0.9)) + ggtitle("'No bulls' scenario")
+  theme(legend.position = c(0.15, 0.9)) + ggtitle("'Upgrading breeds' scenario")
 
@@
 
-dens_nobull<-ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) +
+dens_nobull<-ggplot(plotdata_breed, aes(x = in_var, y = out_var)) +
   stat_density_2d(
@@
 
-ggsave("Fig_11_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
-
-```
-
-## Reducing unproductive replacement males
-
-The idea is to raise the off-take rate of replacement males by 23% (not sure why this number).
+ggsave("Fig_11_breed_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
+
+```
+
+We may even feel the need to confirm that changes did in fact take place. The scatter plot below shows changes in emissions intensity and per capita milk yield in response to the intervention.
+
+To put these results into perspective, we can plot them on top of the probability-of-difference plot we generated above (bottom plot below).
+The data points indicate the milk yield level before implementing the intervention (x-axis) and the milk yield level after upgrading breeds.
+We can see that the differences that are caused by the intervention are usually associated with relatively low probabilities.
+This means that the magnitude of changes that occurred in response to the intervention may also have arisen by chance.
+We know that this is not the case here (since we made these changes happen).
+Nevertheless, these results imply that an evaluator using milk yield as an indicator and requiring a high level of certainty that emissions intensity is in fact low may easily fail to reward the farmers' efforts.
+
+```{r, fig.dim = c(8, 8)}
+breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
+
+breed_res$y<-breed_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+breed_res$x<-breed_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+
+impact_of_breed<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=breed_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=breed_res$y$CO2eq_per_milk)
+impact_of_breed[,"milk_change"]<-round(impact_of_breed$breed_milk-impact_of_breed$baseline,2)
+impact_of_breed[,"emissions_intensity_change"]<-round(impact_of_breed$CO2_intensity_breed-impact_of_breed$CO2_intensity_baseline ,3)
+
+breed_scatter<-ggplot(data=impact_of_breed,
+                      aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Milk yield and emissions intensity response") + 
+  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+breed_probs<-ggplot(data = decreased_emissions_probability,
+                    aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),
+               aes(z=probability),col="black") +
+  geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=breed_milk)) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity")
+
+breed_scatter + breed_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_12_breed_reductions.png", width=9,height=9)
+
+```
+
+Note that the linear patterns (of ten points each) in this plot result from the 10 replicates that were considered for each farm in the original Monte Carlo simulation.
+The originally random milk yields increased by the same factor in all these replicates, leading to linear relationships.
+
+# Herd structure intervention 1 - retiring unproductive bulls
+
+To simulate this scenario, we simply have to remove all animals of type 1 from the herds. Again, we assume an adoption rate of 100%.
+
+```{r , warning=FALSE, eval=FALSE}
+x_nobull_intervention <- GHG_simulation_scenarios$x
+
+nobull_adoption_rate <- 1.0
+
+for (i in 1:nrow(x_nobull_intervention))
+{
+  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
+    x_nobull_intervention$N_animal_type_1[i]<-0
+}
+# run model for modified x file
+nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+nobull_res$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_res,"results","noBull_intervention")
+```
+
+```{r, include=FALSE}
+nobull_res<-load_temperature_scenarios("results","noBull_intervention")
+class(nobull_res)<-"mcSimulation"
+```
+
+## Emissions intensity response to retiring unproductive bulls
+
+Removing all bulls >36 months of age clearly leads to a reduction in the overall emissions intensity of milk production, yet this effect is difficult to see when looking at the overall distribution of milk yield and emissions intensity in the region.
+
+```{r, message=FALSE, warning=FALSE, fig.dim = c(8, 8)}
+
+nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
+
+plotdata_nobull<-data.frame(in_var=nobull_res$y$pc_milk_yield,
+                     out_var=nobull_res$y$CO2eq_per_milk)
+
+plotdata_baseline<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
+                     out_var=GHG_simulation_scenarios$y$CO2eq_per_milk)
+
+plot_nobull <- ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) + 
+  geom_point(shape = 20, color = "blue", size = 2) + 
+  stat_smooth(method = "loess", fullrange = TRUE) +
+  geom_density_2d() +
+  geom_rug(color="lightskyblue3") + 
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() +
+  theme(legend.position = c(0.15, 0.9)) + ggtitle("'No bulls' scenario")
+
+plot_baseline <- ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) + 
+  geom_point(shape = 20, color = "blue", size = 2) + 
+  stat_smooth(method = "loess", fullrange = TRUE) +
+  geom_density_2d() +
+  scale_y_continuous(position = "right") +
+  geom_rug(color="lightskyblue3", sides="bl") + 
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() +
+  theme(legend.position = c(0.15, 0.9)) + ggtitle("Baseline scenario")
+
+dens_nobull<-ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) +
+  stat_density_2d(
+    geom = "raster",
+    aes(fill = after_stat(density)),
+    contour = FALSE
+    ) +
+  scale_fill_viridis_c() + 
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() + guides(fill="none")
+
+dens_baseline<-ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) +
+  stat_density_2d(
+    geom = "raster",
+    aes(fill = after_stat(density)),
+    contour = FALSE
+    ) +
+    scale_y_continuous(position = "right") +
+  scale_fill_viridis_c() + 
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  theme_bw() + guides(fill="none")
+
+plot_nobull + plot_baseline + dens_nobull + dens_baseline +
+  plot_layout(ncol = 2, nrow = 2)
+
+ggsave("Fig_13_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
+
+```
+
+Also for this intervention, we can verify that emissions intensity has actually decreased in response to the intervention (top panel below).
+Also here, however, we can see that most of the farms that made such changes are again located in the "low-confidence" region of the plot, where the impacts of the intervention are difficult to distinguish from the general variability within the population.
+For some farms, the effect is strong enough to possibly lead to a milk-yield-based reward, yet such high-confidence cases are rare also in this dataset.
+
+```{r, fig.dim = c(8, 8)}
+nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
+
+nobull_res$y<-nobull_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+nobull_res$x<-nobull_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+
+impact_of_nobull<-data.frame(baseline=dairy_households$y$pc_milk_yield,nobull_milk=nobull_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_nobull=nobull_res$y$CO2eq_per_milk)
+impact_of_nobull[,"milk_change"]<-round(impact_of_nobull$nobull_milk-impact_of_nobull$baseline,2)
+impact_of_nobull[,"emissions_intensity_change"]<-round(impact_of_nobull$CO2_intensity_nobull-impact_of_nobull$CO2_intensity_baseline ,3)
+
+nobull_scatter<-ggplot(data=impact_of_nobull,
+                      aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Milk yield and emissions intensity response") + 
+  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+nobull_probs<-ggplot(data = decreased_emissions_probability,
+                    aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),
+               aes(z=probability),col="black") +
+  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],aes(x=baseline,y=nobull_milk)) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity")
+
+nobull_scatter + nobull_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_14_nobull_reductions.png", width=9,height=9)
+
+```
+
+# Herd structure intervention 2 - fewer unproductive replacement males
+
+Another strategy in reducing emissions intensity is to raise the off-take rate of replacement males.
 We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate.
-The result should be quite similar.
 
@@
 
-## Herd interventions vs. baseline scenario
-
-### No bulls scenario
-
-```{r, fig.dim = c(8, 8)}
-nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
-
-nobull_res$y<-nobull_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-nobull_res$x<-nobull_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-
-impact_of_nobull<-data.frame(baseline=dairy_households$y$pc_milk_yield,
-                             breed_milk=nobull_res$y$pc_milk_yield,
-                             CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
-                             CO2_intensity_breed=nobull_res$y$CO2eq_per_milk)
-impact_of_nobull[,"milk_change"]<-
-  round(impact_of_nobull$breed_milk-impact_of_nobull$baseline,2)
-impact_of_nobull[,"emissions_intensity_change"]<-
-  round(impact_of_nobull$CO2_intensity_breed-impact_of_nobull$CO2_intensity_baseline ,3)
-
-nobull_scatter<-ggplot(data=impact_of_nobull,
-                       aes(x=milk_change,y=emissions_intensity_change)) +
-  geom_point() +
-  theme_bw() +
-  ggtitle("Milk yield and emissions response") + 
-  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)")
-
-nobull_probs<-ggplot(data = decreased_emissions_probability,
-                     aes(Base_yield, yield_increase)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],
-             aes(x=baseline,y=milk_change)) +
-  ggtitle("Probability that emission intensity reductions occurred") + 
-  labs(y=expression(Milk~yield~increase~(kg~head^{-1}~year^{-1})),
-       x=expression(Baseline~milk~yield~(kg~head^{-1}~year^{-1})),
-       fill="Probability")
-
-nobull_scatter + nobull_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_12_impact_of_nobull_reductions.png", width=9,height=5)
-
-```
-
-### Fewer replacement males scenario
+## Emissions intensity response to fewer unproductive replacement males
+
+As for the other interventions, the impact of reducing the number of unproductive replacement males is clearly reflected in a drop in emissions intensity (see scatter plot below).
+However, also here few data points were located in the high-confidence region of the probability-of-difference plot, so that a milk yield-based reward scheme may easily fail to recognize the efforts of the respective farms.
 
@@
 impact_of_lessReplace<-data.frame(baseline=dairy_households$y$pc_milk_yield,
-                                  breed_milk=lessReplace_res$y$pc_milk_yield,
+                                  lessReplace_milk=lessReplace_res$y$pc_milk_yield,
                                   CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
@@
 impact_of_lessReplace[,"milk_change"]<-
-  round(impact_of_lessReplace$breed_milk-impact_of_lessReplace$baseline,2)
+  round(impact_of_lessReplace$lessReplace_milk-impact_of_lessReplace$baseline,2)
 impact_of_lessReplace[,"emissions_intensity_change"]<-
@@
 lessRep_probs<-ggplot(data = decreased_emissions_probability,
-                      aes(Base_yield, yield_increase)) +
+                      aes(Base_yield, new_yield)) +
   geom_contour_filled(aes(z=probability)) +
@@
   geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],
-             aes(x=baseline,y=milk_change)) +
-  ggtitle("Probability that emission intensity reductions occurred") + 
-  labs(y=expression(Milk~yield~increase~(kg~head^{-1}~year^{-1})),
-       x=expression(Baseline~milk~yield~(kg~head^{-1}~year^{-1})),
-       fill="Probability")
+             aes(x=baseline,y=lessReplace_milk)) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity")
 
@@
 
-ggsave("Fig_13_lessReplace_reductions.png", width=9,height=9)
-
-```
+ggsave("Fig_15_lessReplace_reductions.png", width=9,height=9)
+
+```
+
+# Discussion
+
+* considerable uncertainty in inputs to IPCC equations
+* cause major uncertainty in emissions intensity estimates
+* this makes it difficult to evaluate the emissions intensity of dairy farms in Kenya based on their milk yield alone - large differences are needed to be certain that one farm is more emissions-efficient than another one
+* interventions clearly impact emissions intensity
+* emissions intensity savings are substantial for all interventions, but still relatively small in comparison to the overall uncertainty associated with limited knowledge about values of input variables.
+* it seems likely that a milk yield-based emissions reduction scheme might miss measures taken by farmers, at least if the reward scheme wants to be really sure that emissions reductions have in fact occurred.
+* it seems the information content in milk yields isn't sufficient to base mitigation reward schemes on them
+* might be a better idea after all to observe farmers' practices - did they implement any measures? Then they should be rewarded
+* maybe rewarding measures taken is a better idea than evaluating emissions intensity based on milk yield alone.
+
+# some remaining notes
+
+For each slice, and each level of confidence, calculate the percentage of farms that made changes that would be correctly rewarded vs.
+the ones that did not change (the baseline population).
+Make a figure showing the Type I and Type II errors resulting from each confidence level at each baseline yield.
 

### 2022-06-06 Manuscript_analyses.Rmd
diagnostic plots for breed and herd structure interventions

@@
 library(raster)
+library(reshape2)
 
@@
 library(raster)
+library(reshape2)
 ```
@@
 
-histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3)
+histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3) + theme_bw()
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~intensity~of~milk~production~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1)))
+    xlab(expression(Emission~intensity~of~milk~production~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw()
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level')
+    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level') + theme_bw()
 
@@
        fill="Probability of \nlower emissions \nintensity") +
-  stat_contour(breaks = c(0.9),col="black")
+  stat_contour(breaks = c(0.9),col="black") + theme_bw()
 
@@
 
-plot_nobull <- ggplot(plotdata_breed, aes(x = in_var, y = out_var)) + 
+plot_breed <- ggplot(plotdata_breed, aes(x = in_var, y = out_var)) + 
   geom_point(shape = 20, color = "blue", size = 2) + 
@@
 
-dens_nobull<-ggplot(plotdata_breed, aes(x = in_var, y = out_var)) +
+dens_breed<-ggplot(plotdata_breed, aes(x = in_var, y = out_var)) +
   stat_density_2d(
@@
 
-plot_nobull + plot_baseline + dens_nobull + dens_baseline +
+plot_breed + plot_baseline + dens_breed + dens_baseline +
   plot_layout(ncol = 2, nrow = 2)
@@
 
-breed_res$y<-breed_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-breed_res$x<-breed_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+breed_res$y<-breed_res$y[normal_milk_yield,]
+breed_res$x<-breed_res$x[normal_milk_yield,]
 
@@
        x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity")
+       fill="Probability of \nlower emissions \nintensity") + theme_bw()
 
@@
 The originally random milk yields increased by the same factor in all these replicates, leading to linear relationships.
+
+### Characteristics of sensitive farms
+
+The figure above shows the probability of detecting changes in greenhouse gas emissions intensity by comparing milk yields, for farms that have implemented the breed upgrade intervention.
+We can now extract the most likely probability of such detection for each of the farms (based on the nearest data point on the probability surface).
+This allows us to detect farms that are associated with particularly high probabilities (e.g.
+90%).
+We can then look in detail at the characteristics of these farms to identify farms that are particularly likely to benefit (in terms of reducing emissions intensities) from the breed upgrade intervention.
+
+```{r}
+
+breed_diagnosis<-breed_res
+breed_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(breed_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-breed_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-breed_diagnosis$y$pc_milk_yield[i])
+  breed_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+breed_90_farms<-breed_diagnosis$x$Scenario[which(breed_diagnosis$y$confidence_level>0.9)]
+
+```
+
+The only farm characteristics that we modified in this intervention is the breed composition, so the success of the intervention can only depend on the array of dairy cow breeds present on the farms.
+We can plot the breed composition of the farms against the probability that higher per-capita milk production is associated with lower emissions intensity.
+
+```{r}
+
+farm_structures<-data.frame(Farm=unique(Kenyaherds$QQNO))
+
+for(f in farm_structures$Farm)
+  {samp<-Kenyaherds[which(Kenyaherds$QQNO==f),]
+   cows<-samp[which(samp$animal_type %in% 4:6),]
+   farm_structures[which(farm_structures$Farm==f),"breed_1"]<-length(which(cows$breed==1))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_2"]<-length(which(cows$breed==2))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_3"]<-length(which(cows$breed==3))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_4"]<-length(which(cows$breed==4))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_5"]<-length(which(cows$breed==5))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_6"]<-length(which(cows$breed==6))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_7"]<-length(which(cows$breed==7))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"breed_8"]<-length(which(cows$breed==8))/nrow(cows)
+   farm_structures[which(farm_structures$Farm==f),"unproductive_bulls"]<-
+     length(which(samp$animal_type==1))/nrow(samp)
+   farm_structures[which(farm_structures$Farm==f),"replacement_males"]<-
+     length(which(samp$animal_type==3))/nrow(samp)
+
+}
+
+breed_diagnosis$x["Farm"]<-sapply(breed_diagnosis$x$Scenario,
+                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
+
+breed_structures<-merge(breed_diagnosis$x,farm_structures)
+breed_structures<-list(x=breed_structures, y=breed_diagnosis$y)
+
+to_plot<-cbind(Farm=breed_structures$x$Farm,Confidence_level=breed_structures$y$confidence_level,
+               breed_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
+
+melted<-melt(to_plot, id.vars=c("Farm","Confidence_level","unproductive_bulls","replacement_males"))
+melted<-melted[which(!is.na(melted$value)),]
+
+melted$Confidence_level_rounded<-round(melted$Confidence_level,2)*100
+
+breedcol<-palette.colors(n = 8, palette = "Okabe-Ito", alpha=1, recycle = FALSE)
+
+ggplot(melted, aes(group=variable, fill=variable, y=value, x=Confidence_level_rounded)) + 
+    geom_bar(position="fill", stat="identity") + coord_flip() +
+  labs(y="Cumulative share of breed in herd",
+       x="Probability of lower emissions intensity (%)",
+       fill="Breed of \ndairy cows") + scale_fill_manual(values= c("breed_1" = breedcol[1],
+                                                                   "breed_2" = breedcol[2],
+                                                                   "breed_3" = breedcol[3],
+                                                                   "breed_4" = breedcol[4],
+                                                                   "breed_5" = breedcol[5],
+                                                                   "breed_6" = breedcol[6],
+                                                                   "breed_7" = breedcol[7],
+                                                                   "breed_8" = breedcol[8]),
+                                               labels = c("Holstein Friesian (pure/improved)",
+                                                          "Ayrshire (pure/improved)",
+                                                          "Jersey (pure/improved)",
+                                                          "Guernsey (pure/improved)",
+                                                          "Cross-bred unknown",
+                                                          "Sahiwal",
+                                                          "Boran",
+                                                          "Local zebu")) +
+  theme_bw()
+
+ggsave("Fig_13_breed_diagnostics.png",width=5,height=5)
+
+```
 
@@
 
-ggsave("Fig_13_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
+ggsave("Fig_14_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
 
@@
 
-nobull_res$y<-nobull_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-nobull_res$x<-nobull_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+nobull_res$y<-nobull_res$y[normal_milk_yield,]
+nobull_res$x<-nobull_res$x[normal_milk_yield,]
 
@@
 
-ggsave("Fig_14_nobull_reductions.png", width=9,height=9)
+ggsave("Fig_15_nobull_reductions.png", width=9,height=9)
+
+```
+
+The confidence level depends on the farms' herd structure, specifically on the share of unproductive males with the farms' herds.
+
+```{r}
+
+nobull_diagnosis<-nobull_res
+nobull_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(nobull_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-nobull_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-nobull_diagnosis$y$pc_milk_yield[i])
+  nobull_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+nobull_90_farms<-nobull_diagnosis$x$Scenario[which(nobull_diagnosis$y$confidence_level>0.9)]
+
+nobull_diagnosis$x["Farm"]<-sapply(nobull_diagnosis$x$Scenario,
+                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
+
+nobull_structures<-merge(nobull_diagnosis$x,farm_structures)
+nobull_structures<-list(x=nobull_structures, y=nobull_diagnosis$y)
+
+to_plot<-cbind(Farm=nobull_structures$x$Farm,Confidence_level=nobull_structures$y$confidence_level,
+               nobull_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
+to_plot$other_animals<-1-to_plot$unproductive_bulls
+
+melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","unproductive_bulls")], id.vars=c("Farm","Confidence_level"))
+melted<-melted[which(!is.na(melted$value)),]
+
+melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100
+
+ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
+    geom_bar(position="fill", stat="identity") + coord_flip() +
+  labs(y="Share of animal type in the herd",
+       x="Probability of lower emissions intensity (%)",
+       fill="Animal type") + scale_fill_manual(values= c("unproductive_bulls" = "burlywood",
+                                                         "other_animals" = "darkslategrey"),
+                                               labels = c("Unproductive bulls","Other animals")) +
+  theme_bw()
+
+ggsave("Fig_16_nobull_diagnostics.png",width=5,height=5)
 
@@
 
-lessReplace_res$y<-lessReplace_res$y[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
-lessReplace_res$x<-lessReplace_res$x[which(GHG_simulation_scenarios$y$pc_milk_yield>500),]
+lessReplace_res$y<-lessReplace_res$y[normal_milk_yield,]
+lessReplace_res$x<-lessReplace_res$x[normal_milk_yield,]
 
@@
 
-ggsave("Fig_15_lessReplace_reductions.png", width=9,height=9)
+ggsave("Fig_17_lessReplace_reductions.png", width=9,height=9)
+
+```
+
+Also with this intervention, the level of confidence we can have in higher per-capita milk yields being associated with lower emissions intensity is related to the farms' herd structures.
+In this case, the share of replacement males within the herds is a key determinant.
+
+```{r}
+
+lessReplace_diagnosis<-lessReplace_res
+lessReplace_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(lessReplace_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-lessReplace_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-lessReplace_diagnosis$y$pc_milk_yield[i])
+  lessReplace_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+lessReplace_90_farms<-lessReplace_diagnosis$x$Scenario[which(lessReplace_diagnosis$y$confidence_level>0.9)]
+
+lessReplace_diagnosis$x["Farm"]<-sapply(lessReplace_diagnosis$x$Scenario,
+                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
+
+lessReplace_structures<-merge(lessReplace_diagnosis$x,farm_structures)
+lessReplace_structures<-list(x=lessReplace_structures, y=lessReplace_diagnosis$y)
+
+to_plot<-cbind(Farm=lessReplace_structures$x$Farm,Confidence_level=lessReplace_structures$y$confidence_level,
+               lessReplace_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
+to_plot$other_animals<-1-to_plot$replacement_males
+
+melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","replacement_males")], id.vars=c("Farm","Confidence_level"))
+melted<-melted[which(!is.na(melted$value)),]
+
+melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100
+
+ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
+    geom_bar(position="fill", stat="identity") + coord_flip() +
+  labs(y="Share of animal type in the herd",
+       x="Probability of lower emissions intensity (%)",
+       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
+                                                         "other_animals" = "darkslategrey"),
+                                               labels = c("Replacement males","Other animals")) +
+  theme_bw()
+
+ggsave("Fig_18_lessReplace_diagnostics.png",width=5,height=5)
 

### 2022-06-06 Manuscript_analyses.Rmd
Second MC analysis and Type I/II errors

@@
 
-```{r, eval=FALSE}
+```{r}
 x_breed_intervention <- GHG_simulation_scenarios$x
@@
 
-```{r , warning=FALSE, eval=FALSE}
+```{r , warning=FALSE}
 x_nobull_intervention <- GHG_simulation_scenarios$x
@@
 
-```{r, include=FALSE, eval=FALSE}
+```{r, include=FALSE}
 nobull_res$call<-"multi-scenario call"
@@
 
-```{r , warning=FALSE, eval=FALSE}
+```{r , warning=FALSE}
 x_lessReplace_intervention <- GHG_simulation_scenarios$x
@@
 
-```{r, include=FALSE, eval=FALSE}
+```{r, include=FALSE}
 lessReplace_res$call<-"multi-scenario call"
@@
 
+# Reliability of effort detection via milk yield per capity
+
+- run another Monte Carlo
+- transfer interventions to the results
+- identify for each intervention all the households that didn't adopt the intervention
+- use this to
+
+In practice, rewards for emission reductions are likely to be awarded for changes in changes in per-capita milk yield compared to a baseline.
+These changes may result from farmers adopting innovative practices or from random processes that are not reflective of conscious adaptation action.
+Development programs aiming to reduce emissions should aim to reward farmers who adopt low-emissions practices, rather than providing undeserved rewards for farmers who retain traditional production patterns but simply experience rising milk yield caused by random processes.
+
+To determine the suitability of per-capita milk yield as an indicator of adaptation action, we now simulate ,milk production and emissions at a second point in time.
+The farms that were selected for the three innovations implement the intended measures, but the production and emissions of their farms - like all other farms - are again subjected to random processes.
+This is implemented through a second Monte Carlo simulation, with the interventions applied to the results in the same way as before.
+We can now quantify changes in per-capita milk yields between the two simulated times.
+We can also determine the reliability of identifying intervention adopters through this metric.
+We assume that evaluators will require a certain level of confidence that emissions reductions have occurred, which they derive from the probabilistic analysis of the survey data (see above).
+For each level of required confidence, we can calculate the type 1 error - the percentage of innovators that remain undetected, as well as the type 2 error - the percentage of non-innovators that are falsely rewarded for changes that arose by chance.
+
+```{r monte_carlo_simulation2, warning=FALSE, eval=FALSE}
+
+GHG_simulation_scenarios_time2<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                                            scenarios = read.csv("data/farm_scenarios.csv",
+                                                                 fileEncoding = "UTF-8-BOM"),
+                                            model_function = ghg_emissions,
+                                            numberOfModelRuns = 10,
+                                            functionSyntax = "plainNames")
+
+```
+
+```{r, eval=FALSE, include=FALSE}
+dir.create("results")
+save_temperature_scenarios(GHG_simulation_scenarios_time2,"results","Endline")
+```
+
+```{r, include=FALSE}
+GHG_simulation_scenarios_time2<-load_temperature_scenarios("results","Endline")
+class(GHG_simulation_scenarios)<-"mcSimulation"
+```
+
+## Breed intervention
+
+We apply the same intervention to the dataset as described above.
+
+```{r}
+xb<-x_breed_intervention
+basel<-GHG_simulation_scenarios$x
+
+# breed intervention
+breed_adopters<-which(!xb$milk_yield_4*xb$N_animal_type_4==basel$milk_yield_4*basel$N_animal_type_4 |
+                        !xb$milk_yield_5*xb$N_animal_type_5==basel$milk_yield_5*basel$N_animal_type_5 |
+                        !xb$milk_yield_6*xb$N_animal_type_6==basel$milk_yield_6*basel$N_animal_type_6)
+
+x_breed_intervention_endline <- GHG_simulation_scenarios_time2$x
+
+adoption_rate <- 1
+
+for (i in 1:nrow(x_breed_intervention_endline))
+{
+  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
+    x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
+      x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
+      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
+                                 x_breed_intervention_endline$Scenario[i]),
+                         2:ncol(breed_intervention)]
+}
+
+# run model for modified x file
+breed_res_endline <- run_model(x_breed_intervention_endline, ghg_emissions)
+
+```
+
+```{r, include=FALSE}
+breed_res_endline$call<-"multi-scenario call"
+save_temperature_scenarios(breed_res_endline,"results","Breed_intervention_endline")
+```
+
+```{r, include=FALSE}
+adoption_rate <- 1
+breed_res_endline<-load_temperature_scenarios("results","Breed_intervention_endline")
+class(breed_res_endline)<-"mcSimulation"
+```
+
+We can make scatter and probability plots as above.
+The results are richer now, because the second Monte Carlo simulation introduced additional variation.
+We can now also visually distinguish between changes that occurred on adopting farms and those occurring on non-adopting farms.
+
+```{r}
+breed_res_endline$y$breed_adopter[breed_adopters]<-"Adopter"
+breed_res_endline$y$breed_adopter[is.na(breed_res_endline$y$breed_adopter)]<-"Non-adopter"
+
+breed_res_endline$y[,"CO2eq_per_milk"]<-breed_res_endline$y$pc_on_farm/breed_res_endline$y$pc_milk_yield
+
+breed_res_endline$y<-breed_res_endline$y[normal_milk_yield,]
+breed_res_endline$x<-breed_res_endline$x[normal_milk_yield,]
+
+impact_of_breed_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  breed_endline_milk=breed_res_endline$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_breed=breed_res_endline$y$CO2eq_per_milk,
+                                  Adopter=breed_res_endline$y$breed_adopter)
+impact_of_breed_endline[,"milk_change"]<-
+  round(impact_of_breed_endline$breed_endline_milk-impact_of_breed_endline$baseline,2)
+impact_of_breed_endline[,"emissions_intensity_change"]<-
+  round(impact_of_breed_endline$CO2_intensity_breed-impact_of_breed_endline$CO2_intensity_baseline ,3)
+
+breed_endline_scatter<-ggplot(data=impact_of_breed_endline,
+                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
+  geom_point(size=0.5) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+breed_endline_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_breed_endline[which(impact_of_breed_endline$milk_change>0),],
+             aes(x=baseline,y=breed_endline_milk,color=Adopter), size=0.5) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+breed_endline_scatter + breed_endline_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_19_breed_endline.png", width=9,height=9)
+
+```
+
+Finally, we can calculate the percentatge of adopting farms that would be overlooked at a particular level of required confidence (Type I error).
+We can also compute the percentage of farms that would be identified as adopters, even though they made no changes in their production practices (Type II error).
+
+```{r}
+breed_endline_diagnosis<-breed_res_endline
+breed_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(breed_endline_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-breed_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-breed_endline_diagnosis$y$pc_milk_yield[i])
+  breed_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+confidence_level_analysis_breed<-data.frame(confidence=seq(0.5,1,0.01))
+n_adopt<-length(which(breed_endline_diagnosis$y$breed_adopter=="Adopter"))
+n_nonadopt<-length(which(breed_endline_diagnosis$y$breed_adopter=="Non-adopter"))
+for(i in 1:nrow(confidence_level_analysis_breed))
+{confidence_level_analysis_breed$Type1error[i]<-length(which(breed_endline_diagnosis$y$breed_adopter=="Adopter" &
+                                                 breed_endline_diagnosis$y$confidence_level<confidence_level_analysis_breed$confidence[i])) / n_adopt * 100
+ confidence_level_analysis_breed$Type2error[i]<-length(which(breed_endline_diagnosis$y$breed_adopter=="Non-adopter" &
+                                                 breed_endline_diagnosis$y$confidence_level>confidence_level_analysis_breed$confidence[i])) / n_nonadopt * 100
+
+}
+confidence_level_analysis_breed$confidence<-confidence_level_analysis_breed$confidence*100
+
+toplot<-melt(confidence_level_analysis_breed,id.vars = "confidence")
+toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
+                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
+
+ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
+
+ggsave("Fig_20_breed_endline_errors.png", width=5,height=5)
+
+```
+
+## Herd structure intervention 1 - retiring unproductive bulls
+
+```{r}
+
+# nobull intervention
+nobull_adopters<-which(!x_nobull_intervention$N_animal_type_1==GHG_simulation_scenarios$x$N_animal_type_1)
+
+x_nobull_intervention_endline <- GHG_simulation_scenarios_time2$x
+
+nobull_adoption_rate <- 1.0
+
+for (i in 1:nrow(x_nobull_intervention_endline))
+{
+  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
+    x_nobull_intervention_endline$N_animal_type_1[i]<-0
+}
+# run model for modified x file
+nobull_res_endline <- run_model(x_nobull_intervention_endline, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+nobull_res_endline$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_res_endline,"results","Nobull_intervention_endline")
+```
+
+```{r, include=FALSE}
+adoption_rate <- 1
+nobull_res_endline<-load_temperature_scenarios("results","Nobull_intervention_endline")
+class(nobull_res_endline)<-"mcSimulation"
+```
+
+```{r}
+nobull_res_endline$y$nobull_adopters<-NA
+nobull_res_endline$y$nobull_adopters[nobull_adopters]<-"Adopter"
+nobull_res_endline$y$nobull_adopters[is.na(nobull_res_endline$y$nobull_adopters)]<-"Non-adopter"
+
+nobull_res_endline$y[,"CO2eq_per_milk"]<-nobull_res_endline$y$pc_on_farm/nobull_res_endline$y$pc_milk_yield
+
+nobull_res_endline$y<-nobull_res_endline$y[normal_milk_yield,]
+nobull_res_endline$x<-nobull_res_endline$x[normal_milk_yield,]
+
+impact_of_nobull_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  nobull_endline_milk=nobull_res_endline$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_nobull=nobull_res_endline$y$CO2eq_per_milk,
+                                  Adopter=nobull_res_endline$y$nobull_adopters)
+impact_of_nobull_endline[,"milk_change"]<-
+  round(impact_of_nobull_endline$nobull_endline_milk-impact_of_nobull_endline$baseline,2)
+impact_of_nobull_endline[,"emissions_intensity_change"]<-
+  round(impact_of_nobull_endline$CO2_intensity_nobull-impact_of_nobull_endline$CO2_intensity_baseline ,3)
+
+nobull_endline_scatter<-ggplot(data=impact_of_nobull_endline,
+                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
+  geom_point(size=0.5) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+nobull_endline_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_nobull_endline[which(impact_of_nobull_endline$milk_change>0),],
+             aes(x=baseline,y=nobull_endline_milk,color=Adopter), size=0.5) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+nobull_endline_scatter + nobull_endline_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_21_nobull_endline.png", width=9,height=9)
+
+```
+
+```{r}
+nobull_endline_diagnosis<-nobull_res_endline
+nobull_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(nobull_endline_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-nobull_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-nobull_endline_diagnosis$y$pc_milk_yield[i])
+  nobull_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+confidence_level_analysis_nobull<-data.frame(confidence=seq(0.5,1,0.01))
+n_adopt<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Adopter"))
+n_nonadopt<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Non-adopter"))
+for(i in 1:nrow(confidence_level_analysis_nobull))
+{confidence_level_analysis_nobull$Type1error[i]<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Adopter" &
+                                                 nobull_endline_diagnosis$y$confidence_level<confidence_level_analysis_nobull$confidence[i])) / n_adopt * 100
+ confidence_level_analysis_nobull$Type2error[i]<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Non-adopter" &
+                                                 nobull_endline_diagnosis$y$confidence_level>confidence_level_analysis_nobull$confidence[i])) / n_nonadopt * 100
+
+}
+confidence_level_analysis_nobull$confidence<-confidence_level_analysis_nobull$confidence*100
+
+toplot<-melt(confidence_level_analysis_nobull,id.vars = "confidence")
+toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
+                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
+
+ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
+
+ggsave("Fig_22_nobull_endline_errors.png", width=5,height=5)
+
+```
+
+## Herd structure intervention 2 - fewer unproductive replacement males
+
+We can do the same analysis for the reduction in the number of unproductive replacement males. First we implement the intervention as in the first Monte Carlo analysis.
+
+```{r}
+
+# lessReplace intervention
+lessReplace_adopters<-which(!x_lessReplace_intervention$N_animal_type_3==GHG_simulation_scenarios$x$N_animal_type_3)
+
+x_lessReplace_intervention_endline <- GHG_simulation_scenarios_time2$x
+
+x_lessReplace_intervention_endline$N_animal_type_3[lessReplace_adopters]<-0
+
+# run model for modified x file
+lessReplace_res_endline <- run_model(x_lessReplace_intervention_endline, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+lessReplace_res_endline$call<-"multi-scenario call"
+save_temperature_scenarios(lessReplace_res_endline,"results","lessReplace_intervention_endline")
+```
+
+```{r, include=FALSE}
+lessReplace_res_endline<-load_temperature_scenarios("results","lessReplace_intervention_endline")
+class(lessReplace_res_endline)<-"mcSimulation"
+```
+
+Now we can compare the performance of adopters and non-adopters and plot this against the confidence level for recognizing that emissions reductions have occurred.
+
+```{r}
+lessReplace_res_endline$y$lessReplace_adopters<-NA
+lessReplace_res_endline$y$lessReplace_adopters[lessReplace_adopters]<-"Adopter"
+lessReplace_res_endline$y$lessReplace_adopters[is.na(lessReplace_res_endline$y$lessReplace_adopters)]<-"Non-adopter"
+
+lessReplace_res_endline$y[,"CO2eq_per_milk"]<-lessReplace_res_endline$y$pc_on_farm/lessReplace_res_endline$y$pc_milk_yield
+
+lessReplace_res_endline$y<-lessReplace_res_endline$y[normal_milk_yield,]
+lessReplace_res_endline$x<-lessReplace_res_endline$x[normal_milk_yield,]
+
+impact_of_lessReplace_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  lessReplace_endline_milk=lessReplace_res_endline$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_lessReplace=lessReplace_res_endline$y$CO2eq_per_milk,
+                                  Adopter=lessReplace_res_endline$y$lessReplace_adopters)
+impact_of_lessReplace_endline[,"milk_change"]<-
+  round(impact_of_lessReplace_endline$lessReplace_endline_milk-impact_of_lessReplace_endline$baseline,2)
+impact_of_lessReplace_endline[,"emissions_intensity_change"]<-
+  round(impact_of_lessReplace_endline$CO2_intensity_lessReplace-impact_of_lessReplace_endline$CO2_intensity_baseline ,3)
+
+lessReplace_endline_scatter<-ggplot(data=impact_of_lessReplace_endline,
+                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
+  geom_point(size=0.5) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+lessReplace_endline_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_lessReplace_endline[which(impact_of_lessReplace_endline$milk_change>0),],
+             aes(x=baseline,y=lessReplace_endline_milk,color=Adopter), size=0.5) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+lessReplace_endline_scatter + lessReplace_endline_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_23_lessReplace_endline.png", width=9,height=9)
+
+```
+
+Finally, we can evaluate the Type 1 and Type 2 errors of a reward scheme based on per-capita milk yield.
+
+```{r}
+lessReplace_endline_diagnosis<-lessReplace_res_endline
+lessReplace_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(lessReplace_endline_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-lessReplace_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-lessReplace_endline_diagnosis$y$pc_milk_yield[i])
+  lessReplace_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+confidence_level_analysis_lessReplace<-data.frame(confidence=seq(0.5,1,0.01))
+n_adopt<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Adopter"))
+n_nonadopt<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Non-adopter"))
+for(i in 1:nrow(confidence_level_analysis_lessReplace))
+{confidence_level_analysis_lessReplace$Type1error[i]<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Adopter" &
+                                                 lessReplace_endline_diagnosis$y$confidence_level<confidence_level_analysis_lessReplace$confidence[i])) / n_adopt * 100
+ confidence_level_analysis_lessReplace$Type2error[i]<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Non-adopter" &
+                                                 lessReplace_endline_diagnosis$y$confidence_level>confidence_level_analysis_lessReplace$confidence[i])) / n_nonadopt * 100
+
+}
+confidence_level_analysis_lessReplace$confidence<-confidence_level_analysis_lessReplace$confidence*100
+
+toplot<-melt(confidence_level_analysis_lessReplace,id.vars = "confidence")
+toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
+                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
+
+ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
+
+ggsave("Fig_24_lessReplace_endline_errors.png", width=5,height=5)
+
+```
+
 # Discussion

### 2022-06-07 Manuscript_analyses.Rmd
figure format fix

@@
 
-```{r}
+```{r, fig.dim = c(8, 8)}
 breed_res_endline$y$breed_adopter[breed_adopters]<-"Adopter"
@@
 
-```{r}
+```{r, fig.dim = c(8, 8)}
 nobull_res_endline$y$nobull_adopters<-NA
@@
 
-```{r}
+```{r, fig.dim = c(8, 8)}
 lessReplace_res_endline$y$lessReplace_adopters<-NA

### 2022-06-07 Manuscript_analyses.Rmd
sensitivity analysis

@@
   ylab("Relative probability") +
-    xlab(expression(Emission~density~of~milk~production~(~kg~CO[2]*'-eq'~kg^-1)))
+    xlab(expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^-1)))
 
@@
 
-ggsave("Fig_3_emission_density_farm5.png",width=5,height=3)
+ggsave("Fig_3_emissions_intensity_farm5.png",width=5,height=3)
 
@@
 
-## Population-wide emission density of milk production
-
-Since the emission density of each household thus involves considerable uncertainty, such uncertainty should also be expected at the population level.
+## Population-wide emissions intensity of milk production
+
+Since the emissions intensity of each household thus involves considerable uncertainty, such uncertainty should also be expected at the population level.
 
@@
 
-ggsave("Fig_7_emission_density_all_farms.png")
+ggsave("Fig_7_emissions_intensity_all_farms.png")
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level') + theme_bw()
-
-ggsave("Fig_8_emission_density_all_farms.png", width=7,height=5)
+    xlab(expression(Emissions~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level') + theme_bw()
+
+ggsave("Fig_8_emissions_intensity_all_farms.png", width=7,height=5)
 
@@
 
+# Sensitivity analysis - which uncertainties matter?
+
+To identify influential variables in the population that are particularly strongly related to emissions intensity, we can use the statistical procedure of Projection-to-Latent-Structures (PLS) regression.
+Based on the Variable-Importance-in-the-Projection (VIP) score, we can see, for which uncertain variables knowledge gains would be particularly fruitful.
+
+```{r}
+
+pls_GHG<-dairy_households
+pls_GHG$x<-pls_GHG$x[,which(!colnames(pls_GHG$x)=="Scenario")]
+
+pls_out<-plsr.mcSimulation(object=pls_GHG, resultName = "CO2eq_per_milk")
+
+legend_table<-read.csv("data/legend.csv", fileEncoding = "UTF-8-BOM")
+
+plot_pls(pls_out,input_table=legend_table)
+ggsave("Fig_25_PLS_dairy_households.png", width=8,height=5)
+```
+
 # Discussion
@@
 * might be a better idea after all to observe farmers' practices - did they implement any measures? Then they should be rewarded
-* maybe rewarding measures taken is a better idea than evaluating emissions intensity based on milk yield alone.
-
-# some remaining notes
-
-For each slice, and each level of confidence, calculate the percentage of farms that made changes that would be correctly rewarded vs.
-the ones that did not change (the baseline population).
-Make a figure showing the Type I and Type II errors resulting from each confidence level at each baseline yield.
+* maybe rewarding measures taken is a better idea than evaluating emissions intensity based on milk yield alone
+* sensitivity analysis highlights the need to gain more clarity on milk yields - we specified a 90% confidence interval of mean +- 27.5%
+* Improving reliability of milk yield measurements might make this indicator more viable
+
+# Remaining jobs
+
+- Feed intervention (how exactly?)
+- Select what to present in the manuscript and the supplement
 

### 2022-06-09 Detailed_analyses_for_paper.Rmd
renamed rmd file (since these aren't really the figures for the manuscript)

--- 2022-06-07 Manuscript_analyses.Rmd
@@
    scenarios[which(scenarios$Variable == "feed_prod_CO2"),coln][1:2]<-
-     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])*
+     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])*
                            c(1-feed_prod_error,1+feed_prod_error)
@@
    scenarios[which(scenarios$Variable == "feed_trans_CO2"),coln][1:2]<-
-     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])*
+     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])*
                            c(1-feed_trans_error,1+feed_trans_error)
@@
 
-precise_inputs<-read.csv(input_table)[,c(1:7)] 
+precise_inputs<-read.csv(input_table)[,c(1:8)] 
 precise_vars<-precise_inputs[which(!is.na(precise_inputs$IPCC_default)),]
@@
    precise_scenarios[which(precise_scenarios$Variable == "feed_prod_CO2"),coln]<-
-     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])
+     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed_kgCO2"])
 
@@
    precise_scenarios[which(precise_scenarios$Variable == "feed_trans_CO2"),coln]<-
-     mean(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])
+     sum(Kenyaherds[which(Kenyaherds$QQNO==f),"feed.transport_kgCO2"])
 
@@
 ```
-
-*# THIS STILL NEEDS CHECKING - are all the units correct?*
 
@@
 
-ggsave("Fig_14_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
+ggsave("Fig_14_nobull_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
 
@@
 
+# Feed intervention 1 - Supplementary legume fodder
+
+Animal feeds are supplemented with cuttings from the legumious tree Calliandra. This expected to have the following impacts:
+
+- milk yield of non-dry cows (i.e. animal types 5 and 6) rises by 0.156 kg per day
+- the digestible energy (DE) of the diet rises by 0.14 percentage points
+- the crude protein content of the diet (CP) drops by 0.07 percentage points
+- feed emissions drop by 0.726 kg CO2 per animal and year
+
+We assume that 30% of households adopt this intervention.
+
+```{r}
+x_Calliandra_intervention <- GHG_simulation_scenarios$x
+
+Calliandra_adoption_rate <- 0.3
+
+for (i in 1:nrow(x_Calliandra_intervention))
+{
+  if (rbinom(1, 1, Calliandra_adoption_rate)) # if farmers adopt, make changes in x file
+    {x_Calliandra_intervention$milk_yield_5[i]<-x_Calliandra_intervention$milk_yield_5[i]+0.156
+     x_Calliandra_intervention$milk_yield_6[i]<-x_Calliandra_intervention$milk_yield_6[i]+0.156
+     x_Calliandra_intervention$DE_perc[i]<-x_Calliandra_intervention$DE_perc[i]+0.14
+     x_Calliandra_intervention$perc_crude_protein_diet[i]<-x_Calliandra_intervention$perc_crude_protein_diet[i]-0.07
+     x_Calliandra_intervention$feed_prod_CO2[i]<-x_Calliandra_intervention$feed_prod_CO2[i]- (0.726 *
+                                        sum(x_Calliandra_intervention[i,paste0("N_animal_type_",1:12)]))
+
+  }
+}
+# run model for modified x file
+Calliandra_res <- run_model(x_Calliandra_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE}
+Calliandra_res$call<-"multi-scenario call"
+save_temperature_scenarios(Calliandra_res,"results","Calliandra_intervention")
+```
+
+```{r, include=FALSE}
+Calliandra_res<-load_temperature_scenarios("results","Calliandra_intervention")
+class(Calliandra_res)<-"mcSimulation"
+```
+
+## Emissions intensity response to introducing legume fodder
+
+As for the other interventions, the impact of reducing the number of unproductive replacement males is clearly reflected in a drop in emissions intensity (see scatter plot below).
+However, also here few data points were located in the high-confidence region of the probability-of-difference plot, so that a milk yield-based reward scheme may easily fail to recognize the efforts of the respective farms.
+
+```{r, fig.dim = c(8, 8)}
+
+Calliandra_res$y[,"CO2eq_per_milk"]<-Calliandra_res$y$pc_on_farm/Calliandra_res$y$pc_milk_yield
+
+Calliandra_res$y<-Calliandra_res$y[normal_milk_yield,]
+Calliandra_res$x<-Calliandra_res$x[normal_milk_yield,]
+
+impact_of_Calliandra<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  Calliandra_milk=Calliandra_res$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_Calliandra=Calliandra_res$y$CO2eq_per_milk)
+impact_of_Calliandra[,"milk_change"]<-
+  round(impact_of_Calliandra$Calliandra_milk-impact_of_Calliandra$baseline,2)
+impact_of_Calliandra[,"emissions_intensity_change"]<-
+  round(impact_of_Calliandra$CO2_intensity_Calliandra-impact_of_Calliandra$CO2_intensity_baseline ,3)
+
+lessRep_scatter<-ggplot(data=impact_of_Calliandra,
+                        aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+lessRep_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_Calliandra[which(impact_of_Calliandra$milk_change>0),],
+             aes(x=baseline,y=Calliandra_milk)) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw()
+
+lessRep_scatter + lessRep_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_19_Calliandra_reductions.png", width=9,height=9)
+
+```
+
+Still need to define something adequate for the next section.
+
+```{r}
+
+Calliandra_diagnosis<-Calliandra_res
+Calliandra_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(Calliandra_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-Calliandra_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-Calliandra_diagnosis$y$pc_milk_yield[i])
+  Calliandra_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+Calliandra_90_farms<-Calliandra_diagnosis$x$Scenario[which(Calliandra_diagnosis$y$confidence_level>0.9)]
+
+Calliandra_diagnosis$x["Farm"]<-sapply(Calliandra_diagnosis$x$Scenario,
+                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
+
+Calliandra_structures<-merge(Calliandra_diagnosis$x,farm_structures)
+Calliandra_structures<-list(x=Calliandra_structures, y=Calliandra_diagnosis$y)
+
+to_plot<-cbind(Farm=Calliandra_structures$x$Farm,Confidence_level=Calliandra_structures$y$confidence_level,
+               Calliandra_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
+to_plot$other_animals<-1-to_plot$replacement_males
+
+melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","replacement_males")], id.vars=c("Farm","Confidence_level"))
+melted<-melted[which(!is.na(melted$value)),]
+
+melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100
+
+ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
+    geom_bar(position="fill", stat="identity") + coord_flip() +
+  labs(y="Share of animal type in the herd",
+       x="Probability of lower emissions intensity (%)",
+       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
+                                                         "other_animals" = "darkslategrey"),
+                                               labels = c("Replacement males","Other animals")) +
+  theme_bw()
+
+ggsave("Fig_20_Calliandra_diagnostics.png",width=5,height=5)
+
+```
+
+**This needs change - sensitive factors not analyzed so far**
+
+# Feed intervention 2 - Balanced diets
+
+- milk yield per cow rises by 0.75 kg per day
+- feed emissions drop by 328.5 kg CO2 per animal and year
+
+We assume that 14% of farms adopt this intervention.
+
+```{r}
+x_balance_intervention <- GHG_simulation_scenarios$x
+
+balance_adoption_rate <- 0.14
+
+for (i in 1:nrow(x_balance_intervention))
+{
+  if (rbinom(1, 1, balance_adoption_rate)) # if farmers adopt, make changes in x file
+    {x_balance_intervention$milk_yield_5[i]<-x_balance_intervention$milk_yield_5[i]+0.75
+     x_balance_intervention$milk_yield_6[i]<-x_balance_intervention$milk_yield_6[i]+0.75
+     x_balance_intervention$feed_prod_CO2[i]<-x_balance_intervention$feed_prod_CO2[i]-(328.5 * 
+                    sum(x_Calliandra_intervention[i,paste0("N_animal_type_",1:12)]))
+  }
+}
+# run model for modified x file
+balance_res <- run_model(x_balance_intervention, ghg_emissions)
+
+```
+
+```{r, include=FALSE}
+balance_res$call<-"multi-scenario call"
+save_temperature_scenarios(balance_res,"results","balance_intervention")
+```
+
+```{r, include=FALSE}
+balance_res<-load_temperature_scenarios("results","balance_intervention")
+class(balance_res)<-"mcSimulation"
+```
+
+## Emissions intensity response to introduction of balanced diets
+
+As for the other interventions, the impact of reducing the number of unproductive replacement males is clearly reflected in a drop in emissions intensity (see scatter plot below).
+However, also here few data points were located in the high-confidence region of the probability-of-difference plot, so that a milk yield-based reward scheme may easily fail to recognize the efforts of the respective farms.
+
+```{r, fig.dim = c(8, 8)}
+
+balance_res$y[,"CO2eq_per_milk"]<-balance_res$y$pc_on_farm/balance_res$y$pc_milk_yield
+
+balance_res$y<-balance_res$y[normal_milk_yield,]
+balance_res$x<-balance_res$x[normal_milk_yield,]
+
+impact_of_balance<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  balance_milk=balance_res$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_balance=balance_res$y$CO2eq_per_milk)
+impact_of_balance[,"milk_change"]<-
+  round(impact_of_balance$balance_milk-impact_of_balance$baseline,2)
+impact_of_balance[,"emissions_intensity_change"]<-
+  round(impact_of_balance$CO2_intensity_balance-impact_of_balance$CO2_intensity_baseline ,3)
+
+balance_scatter<-ggplot(data=impact_of_balance,
+                        aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point() +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)")
+
+balance_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_balance[which(impact_of_balance$milk_change>0),],
+             aes(x=baseline,y=balance_milk)) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw()
+
+balance_scatter + balance_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_21_balance_reductions.png", width=9,height=9)
+
+```
+
+Still need to define something adequate for the next section.
+
+```{r}
+
+balance_diagnosis<-balance_res
+balance_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(balance_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-balance_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-balance_diagnosis$y$pc_milk_yield[i])
+  balance_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+balance_90_farms<-balance_diagnosis$x$Scenario[which(balance_diagnosis$y$confidence_level>0.9)]
+
+balance_diagnosis$x["Farm"]<-sapply(balance_diagnosis$x$Scenario,
+                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
+
+balance_structures<-merge(balance_diagnosis$x,farm_structures)
+balance_structures<-list(x=balance_structures, y=balance_diagnosis$y)
+
+to_plot<-cbind(Farm=balance_structures$x$Farm,Confidence_level=balance_structures$y$confidence_level,
+               balance_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
+to_plot$other_animals<-1-to_plot$replacement_males
+
+melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","replacement_males")], id.vars=c("Farm","Confidence_level"))
+melted<-melted[which(!is.na(melted$value)),]
+
+melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100
+
+ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
+    geom_bar(position="fill", stat="identity") + coord_flip() +
+  labs(y="Share of animal type in the herd",
+       x="Probability of lower emissions intensity (%)",
+       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
+                                                         "other_animals" = "darkslategrey"),
+                                               labels = c("Replacement males","Other animals")) +
+  theme_bw()
+
+ggsave("Fig_22_balance_diagnostics.png",width=5,height=5)
+
+```
+
+**This needs change - sensitive factors not analyzed so far**
+
 # Reliability of effort detection via milk yield per capity
@@
 breed_res_endline$call<-"multi-scenario call"
-save_temperature_scenarios(breed_res_endline,"results","Breed_intervention_endline")
+save_temperature_scenarios(breed_res_endline,"results","End_Breed_intervention")
 ```
@@
 adoption_rate <- 1
-breed_res_endline<-load_temperature_scenarios("results","Breed_intervention_endline")
+breed_res_endline<-load_temperature_scenarios("results","End_Breed_intervention")
 class(breed_res_endline)<-"mcSimulation"
@@
 
-ggsave("Fig_19_breed_endline.png", width=9,height=9)
+ggsave("Fig_23_breed_endline.png", width=9,height=9)
 
@@
 
-ggsave("Fig_20_breed_endline_errors.png", width=5,height=5)
+ggsave("Fig_24_breed_endline_errors.png", width=5,height=5)
 
@@
 nobull_res_endline$call<-"multi-scenario call"
-save_temperature_scenarios(nobull_res_endline,"results","Nobull_intervention_endline")
+save_temperature_scenarios(nobull_res_endline,"results","End_nobull_intervention")
 ```
@@
 adoption_rate <- 1
-nobull_res_endline<-load_temperature_scenarios("results","Nobull_intervention_endline")
+nobull_res_endline<-load_temperature_scenarios("results","End_nobull_intervention")
 class(nobull_res_endline)<-"mcSimulation"
@@
 
-ggsave("Fig_21_nobull_endline.png", width=9,height=9)
+ggsave("Fig_25_nobull_endline.png", width=9,height=9)
 
@@
 
-ggsave("Fig_22_nobull_endline_errors.png", width=5,height=5)
+ggsave("Fig_26_nobull_endline_errors.png", width=5,height=5)
 
@@
 lessReplace_res_endline$call<-"multi-scenario call"
-save_temperature_scenarios(lessReplace_res_endline,"results","lessReplace_intervention_endline")
+save_temperature_scenarios(lessReplace_res_endline,"results","End_lessReplace_intervention")
 ```
@@
 ```{r, include=FALSE}
-lessReplace_res_endline<-load_temperature_scenarios("results","lessReplace_intervention_endline")
+lessReplace_res_endline<-load_temperature_scenarios("results","End_lessReplace_intervention")
 class(lessReplace_res_endline)<-"mcSimulation"
@@
 
-ggsave("Fig_23_lessReplace_endline.png", width=9,height=9)
+ggsave("Fig_27_lessReplace_endline.png", width=9,height=9)
 
@@
 
-ggsave("Fig_24_lessReplace_endline_errors.png", width=5,height=5)
+ggsave("Fig_28_lessReplace_endline_errors.png", width=5,height=5)
+
+```
+
+## Feed intervention 1 - Supplementary legume fodder
+
+We can do the same analysis for supplementing cow nutrition with Calliandra.
+
+```{r}
+
+# Calliandra intervention
+Calliandra_adopters<-which(!x_Calliandra_intervention$DE_perc==GHG_simulation_scenarios$x$DE_perc)
+
+x_Calliandra_intervention_endline <- GHG_simulation_scenarios_time2$x
+
+x_Calliandra_intervention_endline$N_animal_type_3[Calliandra_adopters]<-0
+
+x_Calliandra_intervention$milk_yield_5[Calliandra_adopters]<-x_Calliandra_intervention$milk_yield_5[Calliandra_adopters]+0.156
+x_Calliandra_intervention$milk_yield_6[Calliandra_adopters]<-x_Calliandra_intervention$milk_yield_6[Calliandra_adopters]+0.156
+x_Calliandra_intervention$DE_perc[Calliandra_adopters]<-x_Calliandra_intervention$DE_perc[Calliandra_adopters]+0.14
+x_Calliandra_intervention$perc_crude_protein_diet[Calliandra_adopters]<-x_Calliandra_intervention$perc_crude_protein_diet[Calliandra_adopters]-0.07
+x_Calliandra_intervention$feed_prod_CO2[Calliandra_adopters]<-x_Calliandra_intervention$feed_prod_CO2[Calliandra_adopters]- (0.726 *
+                                  sum(x_Calliandra_intervention[Calliandra_adopters,paste0("N_animal_type_",1:12)]))
+
+# run model for modified x file
+Calliandra_res_endline <- run_model(x_Calliandra_intervention_endline, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+Calliandra_res_endline$call<-"multi-scenario call"
+save_temperature_scenarios(Calliandra_res_endline,"results","End_Calliandra_intervention")
+```
+
+```{r, include=FALSE}
+Calliandra_res_endline<-load_temperature_scenarios("results","End_Calliandra_intervention")
+class(Calliandra_res_endline)<-"mcSimulation"
+```
+
+Now we can compare the performance of adopters and non-adopters and plot this against the confidence level for recognizing that emissions reductions have occurred.
+
+```{r, fig.dim = c(8, 8)}
+Calliandra_res_endline$y$Calliandra_adopters<-NA
+Calliandra_res_endline$y$Calliandra_adopters[Calliandra_adopters]<-"Adopter"
+Calliandra_res_endline$y$Calliandra_adopters[is.na(Calliandra_res_endline$y$Calliandra_adopters)]<-"Non-adopter"
+
+Calliandra_res_endline$y[,"CO2eq_per_milk"]<-Calliandra_res_endline$y$pc_on_farm/Calliandra_res_endline$y$pc_milk_yield
+
+Calliandra_res_endline$y<-Calliandra_res_endline$y[normal_milk_yield,]
+Calliandra_res_endline$x<-Calliandra_res_endline$x[normal_milk_yield,]
+
+impact_of_Calliandra_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  Calliandra_endline_milk=Calliandra_res_endline$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_Calliandra=Calliandra_res_endline$y$CO2eq_per_milk,
+                                  Adopter=Calliandra_res_endline$y$Calliandra_adopters)
+impact_of_Calliandra_endline[,"milk_change"]<-
+  round(impact_of_Calliandra_endline$Calliandra_endline_milk-impact_of_Calliandra_endline$baseline,2)
+impact_of_Calliandra_endline[,"emissions_intensity_change"]<-
+  round(impact_of_Calliandra_endline$CO2_intensity_Calliandra-impact_of_Calliandra_endline$CO2_intensity_baseline ,3)
+
+Calliandra_endline_scatter<-ggplot(data=impact_of_Calliandra_endline,
+                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
+  geom_point(size=0.5) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+Calliandra_endline_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_Calliandra_endline[which(impact_of_Calliandra_endline$milk_change>0),],
+             aes(x=baseline,y=Calliandra_endline_milk,color=Adopter), size=0.5) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+Calliandra_endline_scatter + Calliandra_endline_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_29_Calliandra_endline.png", width=9,height=9)
+
+```
+
+Finally, we can evaluate the Type 1 and Type 2 errors of a reward scheme based on per-capita milk yield.
+
+```{r}
+Calliandra_endline_diagnosis<-Calliandra_res_endline
+Calliandra_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(Calliandra_endline_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-Calliandra_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-Calliandra_endline_diagnosis$y$pc_milk_yield[i])
+  Calliandra_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+confidence_level_analysis_Calliandra<-data.frame(confidence=seq(0.5,1,0.01))
+n_adopt<-length(which(Calliandra_endline_diagnosis$y$Calliandra_adopter=="Adopter"))
+n_nonadopt<-length(which(Calliandra_endline_diagnosis$y$Calliandra_adopter=="Non-adopter"))
+for(i in 1:nrow(confidence_level_analysis_Calliandra))
+{confidence_level_analysis_Calliandra$Type1error[i]<-length(which(Calliandra_endline_diagnosis$y$Calliandra_adopter=="Adopter" &
+                                                 Calliandra_endline_diagnosis$y$confidence_level<confidence_level_analysis_Calliandra$confidence[i])) / n_adopt * 100
+ confidence_level_analysis_Calliandra$Type2error[i]<-length(which(Calliandra_endline_diagnosis$y$Calliandra_adopter=="Non-adopter" &
+                                                 Calliandra_endline_diagnosis$y$confidence_level>confidence_level_analysis_Calliandra$confidence[i])) / n_nonadopt * 100
+
+}
+confidence_level_analysis_Calliandra$confidence<-confidence_level_analysis_Calliandra$confidence*100
+
+toplot<-melt(confidence_level_analysis_Calliandra,id.vars = "confidence")
+toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
+                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
+
+ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
+
+ggsave("Fig_30_Calliandra_endline_errors.png", width=5,height=5)
+
+```
+
+## Feed intervention 2 - Balanced diets
+
+We can do the same analysis for a balanced diet.
+
+```{r}
+
+# balance intervention
+balance_adopters<-which(!x_balance_intervention$feed_prod_CO2==GHG_simulation_scenarios$x$feed_prod_CO2)
+
+x_balance_intervention_endline <- GHG_simulation_scenarios_time2$x
+
+x_balance_intervention$milk_yield_5[balance_adopters]<-x_balance_intervention$milk_yield_5[balance_adopters]+0.75
+x_balance_intervention$milk_yield_6[balance_adopters]<-x_balance_intervention$milk_yield_6[balance_adopters]+0.75
+x_balance_intervention$feed_prod_CO2[balance_adopters]<-x_balance_intervention$feed_prod_CO2[balance_adopters]-(328.5 * 
+                    sum(x_Calliandra_intervention[balance_adopters,paste0("N_animal_type_",1:12)]))
+
+# run model for modified x file
+balance_res_endline <- run_model(x_balance_intervention_endline, ghg_emissions)
+
+```
+
+```{r, include=FALSE, eval=FALSE}
+balance_res_endline$call<-"multi-scenario call"
+save_temperature_scenarios(balance_res_endline,"results","End_balance_intervention")
+```
+
+```{r, include=FALSE}
+balance_res_endline<-load_temperature_scenarios("results","End_balance_intervention")
+class(balance_res_endline)<-"mcSimulation"
+```
+
+Now we can compare the performance of adopters and non-adopters and plot this against the confidence level for recognizing that emissions reductions have occurred.
+
+```{r, fig.dim = c(8, 8)}
+balance_res_endline$y$balance_adopters<-NA
+balance_res_endline$y$balance_adopters[balance_adopters]<-"Adopter"
+balance_res_endline$y$balance_adopters[is.na(balance_res_endline$y$balance_adopters)]<-"Non-adopter"
+
+balance_res_endline$y[,"CO2eq_per_milk"]<-balance_res_endline$y$pc_on_farm/balance_res_endline$y$pc_milk_yield
+
+balance_res_endline$y<-balance_res_endline$y[normal_milk_yield,]
+balance_res_endline$x<-balance_res_endline$x[normal_milk_yield,]
+
+impact_of_balance_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
+                                  balance_endline_milk=balance_res_endline$y$pc_milk_yield,
+                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
+                                  CO2_intensity_balance=balance_res_endline$y$CO2eq_per_milk,
+                                  Adopter=balance_res_endline$y$balance_adopters)
+impact_of_balance_endline[,"milk_change"]<-
+  round(impact_of_balance_endline$balance_endline_milk-impact_of_balance_endline$baseline,2)
+impact_of_balance_endline[,"emissions_intensity_change"]<-
+  round(impact_of_balance_endline$CO2_intensity_balance-impact_of_balance_endline$CO2_intensity_baseline ,3)
+
+balance_endline_scatter<-ggplot(data=impact_of_balance_endline,
+                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
+  geom_point(size=0.5) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+balance_endline_probs<-ggplot(data = decreased_emissions_probability,
+                      aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
+  geom_point(data=impact_of_balance_endline[which(impact_of_balance_endline$milk_change>0),],
+             aes(x=baseline,y=balance_endline_milk,color=Adopter), size=0.5) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
+
+balance_endline_scatter + balance_endline_probs +
+  plot_layout(ncol = 1, nrow = 2)
+
+ggsave("Fig_31_balance_endline.png", width=9,height=9)
+
+```
+
+Finally, we can evaluate the Type 1 and Type 2 errors of a reward scheme based on per-capita milk yield.
+
+```{r}
+balance_endline_diagnosis<-balance_res_endline
+balance_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
+
+# get closest probability value
+
+for(i in 1:nrow(balance_endline_diagnosis$x))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-balance_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
+    abs(decreased_emissions_probability$new_yield-balance_endline_diagnosis$y$pc_milk_yield[i])
+  balance_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+confidence_level_analysis_balance<-data.frame(confidence=seq(0.5,1,0.01))
+n_adopt<-length(which(balance_endline_diagnosis$y$balance_adopter=="Adopter"))
+n_nonadopt<-length(which(balance_endline_diagnosis$y$balance_adopter=="Non-adopter"))
+for(i in 1:nrow(confidence_level_analysis_balance))
+{confidence_level_analysis_balance$Type1error[i]<-length(which(balance_endline_diagnosis$y$balance_adopter=="Adopter" &
+                                                 balance_endline_diagnosis$y$confidence_level<confidence_level_analysis_balance$confidence[i])) / n_adopt * 100
+ confidence_level_analysis_balance$Type2error[i]<-length(which(balance_endline_diagnosis$y$balance_adopter=="Non-adopter" &
+                                                 balance_endline_diagnosis$y$confidence_level>confidence_level_analysis_balance$confidence[i])) / n_nonadopt * 100
+
+}
+confidence_level_analysis_balance$confidence<-confidence_level_analysis_balance$confidence*100
+
+toplot<-melt(confidence_level_analysis_balance,id.vars = "confidence")
+toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
+                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
+
+ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
+
+ggsave("Fig_32_balance_endline_errors.png", width=5,height=5)
 
@@
 plot_pls(pls_out,input_table=legend_table)
-ggsave("Fig_25_PLS_dairy_households.png", width=8,height=5)
+ggsave("Fig_33_PLS_dairy_households.png", width=8,height=5)
 ```
@@
 
-- Feed intervention (how exactly?)
 - Select what to present in the manuscript and the supplement

### 2022-06-09 Manuscript_figures.Rmd
new rmd for the manuscript

--- 2022-06-07 Manuscript_analyses.Rmd
@@
 ---
-title: "Accounting for uncertainty in simulation greenhouse gas emissions from dairy cattle in Kenya"
-author: "Eike Luedeling, Cory Whitney, Todd Rosenstock"
+title: "Accounting for uncertainty in simulating greenhouse gas emissions from dairy cattle in Kenya"
+author: "Eike Luedeling, Cory Whitney, Andreas Wilkes, Todd Rosenstock"
 bibliography:
@@
 library(reshape2)
-
-#Automatically write R package citation entries to a .bib file
+library(patchwork)
+
 knitr::write_bib(c(.packages(), 
@@
                    'raster',
-                   'MASS'), 'references/packages.bib')
+                   'reshape2',
+                   'patchwork'), 'references/packages.bib')
 #resolves an issue with the scientific style of numbers,answer from @Paul Hiemstra: https://stackoverflow.com/a/25947542/4061993 
@@
 library(reshape2)
+library(patchwork)
 ```
@@
 ```{r, warning=FALSE,eval=FALSE}
+input_table <- "data/livestock_ghg_input_table.csv"
 GHG_emissions<-mcSimulation(estimate=estimate_read_csv(input_table),
@@
 
-```
-
-Here's the resulting file (only the first 3 farms are shown, and only the first 20 rows):
-
-```{r}
 read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5]  %>%
@@
 ```
+
+This is the resulting file (only the first 3 farms are shown, and only the first 20 rows):
 
@@
                                  functionSyntax = "plainNames")
-
-```
-
-```{r, eval=FALSE, include=FALSE}
+GHG_simulation_scenarios$y[,"CO2eq_per_milk"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
 dir.create("results")
 save_temperature_scenarios(GHG_simulation_scenarios,"results","Baseline")
+
 ```
@@
 
-Given these results we can now plot distributions of the various outputs from the model, using the `plot_distributions` function.
-
-```{r plot_distributions, fig.width=10, fig.height=10, message=FALSE, warning=FALSE}
-# enteric_CH4=enteric_CH4,
-enteric_CH4_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "enteric_CH4",
-                   method = "boxplot_density",
-                   old_names = "enteric_CH4",
-                   new_names = "CH[4]~Emissions~from~Enteric~Fermentation~(kg~CH[4])") + 
-  theme(axis.title.x=element_blank(),
-        axis.title.y=element_blank())
-
-enteric_CH4_dist_plot <- enteric_CH4_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-# milk_yield=sum(total_milk),
-milk_yield_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "milk_yield",
-                   method = "boxplot_density",
-                   old_names = "milk_yield",
-                   new_names = "Total~Milk~Yield~(kg)") + 
-  theme(axis.title.x=element_blank(),
-        axis.title.y=element_blank())
-
-milk_yield_dist_plot <- milk_yield_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-# enteric_CH4_CO2eq=sum(enteric_CH4_CO2eq),
-enteric_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "enteric_CH4_CO2eq",
-                   method = "boxplot_density",
-                   old_names = "enteric_CH4_CO2eq",
-                   new_names = "Enteric~CH[4]~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
-  theme(axis.title.x=element_blank())
-
-enteric_CH4_CO2eq_dist_plot <- enteric_CH4_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-# mm_CH4_CO2eq=sum(mm_CH4_CO2eq),
-mm_CH4_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "mm_CH4_CO2eq",
-                   method = "boxplot_density",
-                   old_names = "mm_CH4_CO2eq",
-                   new_names = "CH[4]~Emissions~from~Livestock~Manure~(kg~CO[2]*'-eq')") + 
-  theme(axis.title.x=element_blank(),
-        axis.title.y=element_blank())
-
-mm_CH4_CO2eq_dist_plot <- mm_CH4_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-# mm_N2O_CO2eq=sum(mm_N2O_CO2eq),
-mm_N2O_CO2eq_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "mm_N2O_CO2eq",
-                   method = "boxplot_density",
-                   old_names = "mm_N2O_CO2eq",
-                   new_names = "N[2]*O~Emissions~from~Livestock~Manure~(kg~CO[2]*'-eq')") + 
-  theme(axis.title.x=element_blank(),
-        axis.title.y=element_blank())
-
-mm_N2O_CO2eq_dist_plot <- mm_N2O_CO2eq_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-# feed_CO2=feed_CO2,
-feed_CO2_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "feed_CO2",
-                   method = "boxplot_density",
-                   old_names = "feed_CO2",
-                   new_names = "CO[2]~Emissions~from~Livestock~Feed~Production~and~Transport~(kg)") + 
-  theme(axis.title.x=element_blank(),
-        axis.title.y=element_blank())
-feed_CO2_dist_plot <- feed_CO2_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-# on_farm=sum(on_farm)
-on_farm_dist_plot <- plot_distributions(mcSimulation_object = GHG_simulation_scenarios,
-                   vars = "on_farm",
-                   method = "boxplot_density",
-                   old_names = "on_farm",
-                   new_names = "Total~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
-  theme(axis.title.y=element_blank())
-
-on_farm_dist_plot <- on_farm_dist_plot + facet_wrap(. ~ name, labeller = label_parsed)
-
-library(patchwork)
-# remove the axis marks from the two lefthand-most plots and the x-axis title from the left and right plots
-
-# plot with patchwork
-layout <- "
-           AAABBB
-           CCCDDD
-           EEEFFF
-           GGGGGG
-          "
-
-enteric_CH4_dist_plot + milk_yield_dist_plot + 
-  enteric_CH4_CO2eq_dist_plot + mm_CH4_CO2eq_dist_plot + 
-  mm_N2O_CO2eq_dist_plot + feed_CO2_dist_plot + 
-  on_farm_dist_plot + 
-plot_layout(guides = "collect", design = layout)
-
-ggsave("Fig_1_GHG_simulation_outputs.png",width=9,height=9)
-```
-
-*# THIS STILL NEEDS CHECKING - are all the units correct?*
-
-# Possible errors resulting from input uncertainty
-
-## Uncertainty at the farm scale
-
-To illustrate the possible influence of input uncertainty on estimates of total farm emissions, we show the example of farm #5 from the household survey.
-This farm has 6 head of cattle, including 3 lactating cows.
-We simulate 1000 versions of this farm.
-We also do a precise calculation based on the exact numbers returned by the household survey and the default IPCC values.
-
-### Total emissions of the farm
-
-The following code and the resulting distribution estimate the total emissions of farm #5.
-
-```{r, warning=FALSE}
+Before we explore the results for the full population of farms, let's first look at one example.
+The following figure shows the greenhouse gas emissions, per farm and per head of cattle, for farm #5, which holds 12 head of cattle.
+
+```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4)}
 
@@
 
-pd<-plot_distributions(mcSimulation_object = farm5_emissions,
-                   vars = "on_farm",
-                   method = "boxplot_density",
-                   old_names = "on_farm",
-                   new_names = "Total~Emissions~from~Livestock~(kg~CO[2]*'-eq')") + 
-  theme(axis.title.y=element_blank())
-
-pd + facet_wrap(. ~ name, labeller = label_parsed) + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=on_farm),color="blue", lwd=3)
-
-ggsave("Fig_2_uncertainty_illustration.png",width=5,height=3)
-
-```
-
-The light blue distribution illustrates the estimated emissions from farm #5, considering uncertainty about all input variables.
-The blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.
-
-### Emissions intensity
-
-Naturally, the total emissions of the farm depend strongly on the size of the farm.
-A better indicator of the carbon footprint of milk production is the emissions intensity, i.e.
-the emissions per quantity of milk produced.
-
-```{r, warning=FALSE, message=FALSE}
-farm5_emissions$y[,"CO2eq_per_milk"]<-farm5_emissions$y$pc_on_farm/farm5_emissions$y$pc_milk_yield
-
-histo<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
+farm5a<-ggplot(data=farm5_emissions$y, aes(on_farm)) +
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
@@
   ylab("Relative probability") +
-    xlab(expression(Emission~density~of~milk~production~(~kg~CO[2]*'-eq'~kg^-1)))
+    xlab(expression(Total~emissions~from~livestock~(kg~CO[2]*'-eq'~year^-1))) + theme_bw() + ggtitle("a) Total emissions")
+
+farm5_total <- farm5a + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=on_farm),color="blue", lwd=3)
+
+farm5_emissions$y[,"CO2eq_per_milk"]<-farm5_emissions$y$pc_on_farm/farm5_emissions$y$pc_milk_yield
+
+farm5b<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") +
+    xlab(expression(Emissions~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emissions intensity")
 
@@
 
-histo + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3) + theme_bw()
-
-ggsave("Fig_3_emission_density_farm5.png",width=5,height=3)
-
-```
-
-Also here, the blue line represents a classic best-bet estimate based on default values in the IPCC guidelines and some context-specific reasonable guesses.
-
-## Population-wide emission density of milk production
-
-Since the emission density of each household thus involves considerable uncertainty, such uncertainty should also be expected at the population level.
-
-We can illustrate this uncertainty using a scatter plot.
-The following figure shows all combinations of total milk yield and total greenhouse gas emissions across all households.
-Total emissions are the sum of methane emissions from enteric fermentation and livestock manure, nitrous oxide emissions from manure and carbon dioxide emissions from livestock feed production and transport.
-
-```{r, warning=FALSE, message=FALSE}
-plotdata<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
-                     out_var=GHG_simulation_scenarios$y$pc_on_farm)
-
-plot1 <- ggplot(plotdata, aes(x = in_var, y = out_var)) + 
+farm5_intensity<-farm5b + geom_vline(data=farm5_emissions_precise$y, aes(xintercept=CO2eq_per_milk),color="blue", lwd=3) + theme_bw()
+
+farm5_total+farm5_intensity
+
+ggsave("figures/Fig_1_emissions_from_farm_5.png",width=9,height=5)
+
+```
+
+Now we look at the overall population.
+
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4)}
+
+all_farms_total<-ggplot(data=GHG_simulation_scenarios$y, aes(on_farm)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") +
+    xlab(expression(Total~emissions~from~livestock~(kg~CO[2]*'-eq'~year^-1))) + theme_bw() + ggtitle("a) Total emissions")
+
+all_farms_intensity<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") + theme_bw() +
+    xlab(expression(Emissions~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emissions intensity")
+
+all_farms_total+all_farms_intensity
+
+ggsave("figures/Fig_2_emissions_from_all_farms.png",width=9,height=5)
+```
+
+We're interested in the suitability of total per capita milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.
+
+Here's the relationship between these two variables.
+We exclude farms with a per capita milk yield of <500 kg per year, since such farms can't be considered dairy farms and they would create outliers in many analysis.
+
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6)}
+normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)
+
+dairy_households<-GHG_simulation_scenarios
+dairy_households$x<-dairy_households$x[normal_milk_yield,]
+dairy_households$y<-dairy_households$y[normal_milk_yield,]
+
+plot1 <- ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) + 
   geom_point(shape = 20, color = "blue", size = 2) + 
-  stat_smooth(method = "loessm", fullrange = TRUE) +
+  stat_smooth(method = "loess", fullrange = TRUE) +
   geom_density_2d() +
   geom_rug(color="lightskyblue3") + 
-  labs(y=expression(Total~emissions~(kg~CO[2]*'-eq'~head^{-1}~year^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})))+
   theme_bw() +
@@
 
-dens1 <- ggplot(plotdata, aes(x = in_var)) + 
+dens1 <- ggplot(dairy_households$y, aes(x = pc_milk_yield)) + 
   geom_histogram(aes(y = ..density..),
@@
 
-dens2 <- ggplot(plotdata, aes(x = out_var)) + 
+dens2 <- ggplot(dairy_households$y, aes(x = CO2eq_per_milk)) + 
   geom_histogram(aes(y = ..density..),
@@
 
-dens1 + plot_spacer() + plot1 + dens2 + 
+scatt<-dens1 + plot_spacer() + plot1 + dens2 + 
   plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
 
-ggsave("Fig_4_scatter_per_head_milk_emissions.png",width=5,height=3)
-
-```
-
-## Emissions intensity of milk production
-
-As mentioned above, an indicator that is often of interest is the ratio between total emissions and the milk yield.
-We'll calculate this next.
-We then have to remove all households that don't produce any milk, because this ratio would be infinity for them.
-In fact, we'll remove all households that reported very low milk yields of <500 kg per year per capita, since these are unlikely to represent regularly operating dairy operations.
-For instance, they may be mostly focused on beef cattle production, or they may only have a small number of dairy cows that happened to be unproductive at the time of the survey.
-
-```{r}
-GHG_simulation_scenarios$y[,"CO2eq_per_milk"]<-GHG_simulation_scenarios$y$pc_on_farm/GHG_simulation_scenarios$y$pc_milk_yield
-
-normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)
-
-dairy_households<-GHG_simulation_scenarios
-dairy_households$x<-dairy_households$x[normal_milk_yield,]
-dairy_households$y<-dairy_households$y[normal_milk_yield,]
-
-```
-
-Now we can plot the emissions intensity vs. the milk yield.
-
-```{r, warning=FALSE, message=FALSE}
-
-plotdata<-data.frame(in_var=dairy_households$y$pc_milk_yield,
-                     out_var=dairy_households$y$CO2eq_per_milk,
-                     Group="just one")
-
-plot1 <- ggplot(plotdata, aes(x = in_var, y = out_var)) + 
-  geom_point(shape = 20, color = "blue", size = 2) + 
-  stat_smooth(method = "loess", fullrange = TRUE) +
-  geom_density_2d() +
-  geom_rug(color="lightskyblue3") + 
-  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() +
-  theme(legend.position = c(0.15, 0.9))
-
-dens1 <- ggplot(plotdata, aes(x = in_var)) + 
-  geom_histogram(aes(y = ..density..),
-                 colour = 1, fill = "white") +
-  geom_density(fill="lightskyblue3") + 
-  geom_histogram(aes(y = ..density..),
-                 colour = 1, alpha=0) + 
-  theme_void() + 
-  theme(legend.position = "none")
-
-dens2 <- ggplot(plotdata, aes(x = out_var)) + 
-  geom_histogram(aes(y = ..density..),
-                 colour = 1, fill = "white") +
-  geom_density(fill="lightskyblue3") + 
-  geom_histogram(aes(y = ..density..),
-                 colour = 1, alpha=0) + 
-  theme_void() + 
-  theme(legend.position = "none") + 
-  coord_flip()
-
-dens1 + plot_spacer() + plot1 + dens2 + 
-  plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
-
-ggsave("Fig_5_scatter_per_head_milk_emissions_intensity.png",width=7,height=7)
-
-```
-
-This can also be illustrated by a density surface.
-
-```{r varkernel}
-
-ggplot(plotdata, aes(x = in_var, y = out_var)) +
+densi<-ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) +
   stat_density_2d(
@@
   scale_fill_viridis_c() + 
-  labs(y=expression(Total~emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
+  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1}))) +
   theme_bw()
 
-ggsave("Fig_6_surface_per_head_milk_emissions_intensity.png",width=7,height=7)
-
-```
-
-The milk yield per capita is sometimes proposed as an indicator of greenhouse gas emissions.
-We can now examine the merits of this indicator.
-Essentially, its usefulness hinges on its relationship with greenhouse gas emissions intensity.
-We can investigate this by taking slices through the density surface.
-
-# Milk yield as an indicator of GHG emissions intensity
-
-## Probability of GHG emissions for a given milk yield
-
-The following plot shows the emissions intensity that corresponds to a milk yield of 2000 kg per capita per year, based on the population of households covered in the survey.
-
-```{r varkernelslicerange, message=FALSE, warning=FALSE}
-slice<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
-                                 out_var = dairy_households$y$CO2eq_per_milk, 
-                                 expectedin_var=2000,
-                           n = 1000,
-                           n_samples = 1000,
-                           out_var_sampling = 1000)
-
-ggplot(data=data.frame(CO2eq_per_milk=slice$resampled), aes(CO2eq_per_milk)) +
-geom_histogram(aes(y=..density..), colour="black", fill="white")+
- geom_density(alpha=.2, fill="#FF6666") +
-  ylab("Relative probability") +
-    xlab(expression(Emission~intensity~of~milk~production~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw()
-
-ggsave("Fig_7_emission_density_all_farms.png")
-
-```
-
-We see considerable uncertainty about the emissions at this particular level of production, and the previous figure shows that emissions of all other levels are similarly variable.
-This raises considerable doubts about the usefulness of milk yield as an indicator of emissions intensity - at least when this indicator is taken at face value, without considering this uncertainty.
-Yet, some information is contained even in such an imperfect indicator, since, as also shown in the previous figure, there is clearly a correlation between milk yield and emissions intensity.
-The following proposes a mechanism for quantifying the level of certainty, with which a change in milk yield can reasonably be associated with a change in emissions intensity.
-
-## Probability of an increase in milk yield being associated with an increase in GHG emission intensity
-
-To derive the level of certainty, with which a change in milk yield is associated with an increase or decrease in emissions intensity, we have to investigate the probability distributions (as shown in the previous figure) of emissions intensity at two levels of milk production.
-This can be achieved by using the `varslice_resample` function for both production levels.
-
-Let's look at this for production of 2000 and 3000 kg milk per capita per year.
-We'll take 1000 random draws of the emissions intensities that are associated with both levels and compare values among these samples.
-
-```{r}
+dens1 + plot_spacer() + plot_spacer() + plot1 + dens2 + densi + 
+  plot_layout(ncol = 3, nrow = 2, widths = c(4, 1), heights = c(1, 4))
+
+ggsave("figures/Fig_3_scatter_and_surface_per_head_milk_emissions_intensity.png",width=10,height=6)
+```
+
+According to these distributions, the emissions intensity at a particular per capita milk yield can be described by a slice through the probability surface.
+
+```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4)}
 slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
@@
 
+one_slice<-ggplot(data=data.frame(CO2eq_per_milk=slice_2000$resampled), aes(CO2eq_per_milk)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") +
+    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")
+
 slices<-rbind(data.frame(Values=slice_2000$resampled, Level="2000 kg"), data.frame(Values=slice_3000$resampled, Level="3000 kg"))
 
-ggplot(data=slices, aes(Values, fill=Level)) +
+two_slices<-ggplot(data=slices, aes(Values, fill=Level)) +
  geom_density(alpha=.2) +
   ylab("Relative probability") +
-    xlab(expression(Emission~intensity~of~milk~production~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Production level') + theme_bw()
-
-ggsave("Fig_8_emission_density_all_farms.png", width=7,height=5)
-
-# sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)
-
-```
-
-Now we want to determine the probability of emissions intensity at a production level of 3000 kg milk per capita and year being lower than emissions intensity at 2000 kg per capita and year.
-We can do this empirically by comparing the two sets of 1000 random values that are assembled in the probability distribution slices representing per capita milk production levels of 3000 and 2000 kg.
-We test for each index number (1-1000) if the respective element of the 3000-kg slice is smaller than the respective element of the 2000-kg slice.
-By adding up the results of all these comparisons (with TRUE coded as 1 and FALSE as 0) and dividing by 1000 (and optionally multiplying by 100% for a percentage), we obtain the probability that emissions intensity is really lower at the higher milk production level.
-
-In this case, this comparison reveals that in `r sum(slice_3000$resampled < slice_2000$resampled) / length(slice_2000$resampled)*100` % of cases, emissions intensity at the higher milk production level is lower than at the lower level.
-
-We can now generalize this result by doing similar calculations for multiple production levels and illustrating them for the whole population.
-We'll base this on slices representing production levels between 100 and 10,000 kg per capita and year, at 100-kg intervals.
-
-```{r}
-
-in_var = dairy_households$y$pc_milk_yield
-out_var = dairy_households$y$CO2eq_per_milk
+    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")
+
+one_slice + two_slices
+
+ggsave("figures/Fig_4_slices.png",width=13,height=6)
+
+```
+
+By resampling from these distributions, we can determine the likelihood that a value of one of these distributions is greater than one of the other.
+We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per capita.
+
+```{r, fig.dim = c(8, 4)}
 
 milk_yields_to_sample<-seq(100,10000,100)
-milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(in_var))]
-milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample>min(in_var))]
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(dairy_households$y$pc_milk_yield))]
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample>min(dairy_households$y$pc_milk_yield))]
 slices<-list()
@@
 {slices[[i]]<-
-  varslice_resample(in_var,
-                      out_var,
+  varslice_resample(dairy_households$y$pc_milk_yield,
+                      dairy_households$y$CO2eq_per_milk,
                       expectedin_var = milk_yields_to_sample[i],
@@
 
-#decreased_emissions_probability[,"yield_increase"]<-decreased_emissions_probability$new_yield-decreased_emissions_probability$Base_yield
-
 ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
@@
 
-ggsave("Fig_9_probability_of_lower_emissions_intensity.png", width=7,height=5)
-
-```
-
-The results indicate that whenever milk yields are higher on a particular farm than on another one, the probability that emissions intensity is lower is usually greater than 50% (slightly lower values are due to randomness in the resampling procedure), but that large differences are needed to be sure of this.
-The 90% confidence contour (shown in the plot as a black line) indicates the minimum level of milk yield that would be needed to state with 90% certainty that the emissions intensity is lower on the higher-productivity farm than on the respective lower-productivity one.
-
-# Methodology for testing interventions
-
-Farms may try to reduce their emissions intensity by making changes to their herd structure, feeding strategy or management practices.
-Programs aiming to effect emissions reductions in dairy production may want to reward such changes, and they may be looking to the milk production per capita as an indicator for the effectiveness of such measures.
-We now want to explore whether this is a reasonable strategy.
-
-## Interventions
-
-We will explore the impacts of three interventions that farmers may consider. We'll restrict this assessment to interventions we can easily implement in our modeling framework.
-
-### Breed adjustment
-
-We evaluate milk production for all the breeds found in the survey dataset. All cows of low-productivity breeds are then replaced by high-productivity cows.
-
-### Herd structure intervention 1 - retiring unproductive bulls
-
-All bulls that are beyond their productive life span are removed from the herd. We implement this by removing all animals of animal type 1, which denotes bulls >36 months of age.
-
-### Herd structure intervention 2 - fewer unproductive replacement males
-
-Many farms have more male calves and young bulls than are needed to meet the objectives of dairy production.
-Based on a real-life scenario, we target a reduction in this number by 23%.
-For this, we randomly select 23% of the farms and set the number of replacement males (animal type 3) to 0.
-
-## Analysis strategy
-
-In a model, we can simulate the implications of specific interventions by leaving all variables unchanged, except those that are modified by the intervention.
-To implement this using the decisionSupport package, we have to disable randomness for the variables we don't want to modify.
-For this, we can create a function (`run_model`) that simply applies our model to a set of values we specify.
-We can use this function to exactly reproduce the results of a model run.
-Note that this only works, because our model is perfectly deterministic.
-All variation in model outputs results from variation in inputs; the model itself doesn't include additional random processes.
-
-We can now generate the outputs of a Monte Carlo model for our baseline situation (without applying interventions).
-The output object of the `mcSimulation` function doesn't only contain the output variables, however, but also a complete list of all random inputs for all model runs.
-As mentioned above, we could now apply the `run_model` function to these inputs (the `x` element of the `mcSimulation` result) to reproduce the outputs of the simulation.
-More importantly, we can now modify certain variables included in this `x` element to simulate interventions, and then run the model on these modified inputs to simulate the impacts of these interventions.
-
-The following code contains the `run_model` function:
-
-```{r}
-run_model<-function(x_data,model)
-  {
-  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
-  model_function_ <- function(x) {
-
-    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
-    eval(expr = body(model), envir = e)
-  }
-  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
-                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
-  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
-  returnObject$call <- match.call()
-  class(returnObject) <- cbind("mcSimulation", class(returnObject))
-  return(returnObject)
-}
-
-```
-
-The following code can therefore precisely reproduce the results of the previous analysis:
-
-```{r, eval=FALSE}
-res<-run_model(GHG_simulation_scenarios$x,ghg_emissions)
-```
-
-To simulate interventions, we can now feed this function with modified x_data data.frames.
-
-# Upgrading to higher-performance breeds
-
-Replacing cows of low-productivity breeds with better-performing ones may reduce greenhouse gas emissions.
-We simulate an intervention that replaces animals of low-performance breeds with animals of high-performance breeds.
-
-We first need to determine the performance and characteristics of each of the breeds that are present in the farm survey dataset (`Kenyaherds` dataset from above).
-
-```{r, warning=FALSE}
-cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
-
-ggplot(cows, aes(factor(breed),milk_yield)) +
-  geom_violin(draw_quantiles=c(0.5),fill="light green") +
-  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
-
-ggsave("Fig_10_breeds.png", width=7,height=5)
-
-milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
-milk_summary<-milk_summary[,c(1,4,2)]
-colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
-
-milk_summary  %>%
-    kableExtra::kbl(digits = 2)  %>%
-    kableExtra::kable_classic_2(full_width = F)
-
-```
-
-This evaluation shows that the difference in median milk yields is relatively small, but that the milk yield distributions show a lot of variation that is apparently not explained by the breed.
-It is also apparent that the number of animals these estimates are based on varies considerably.
-Two breeds make up `r round(sum(milk_summary$n[1:2])/sum(milk_summary$n)*100,1)` % of the entire population, with others having very few animals, which means that the medians may be somewhat unreliable (especially for the single animal of breed 7).
-
-We define the intervention as follows:
-
-All animals are replaced by high-performance animals of breed 1, which increases milk yields proportionally to the median milk yield differences.
-This means that milk yields are divided by the following factors:
-
-```{r}
-milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))
-
-milk_summary   %>%
-    kableExtra::kbl(digits = 2)  %>%
-    kableExtra::kable_classic_2(full_width = F)            
-```
-
-A change of the breed also causes changes in weight, mature weight and weight gain.
-
-```{r}
-live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
-colnames(live_weights)<-c("animal_type","breed","median_live_weight")
-live_weights[,"Intervention_factor"]<-NA
-for(i in 1:nrow(live_weights))
-  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
-     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]
-
-# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.
-
-mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
-colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
-mature_weights[,"Intervention_factor"]<-NA
-for(i in 1:nrow(mature_weights))
-  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
-     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]
-
-# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1
-
-weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
-colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
-weight_gains[,"Intervention_factor"]<-NA
-for(i in 1:nrow(weight_gains))
-  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
-     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
-weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1
-
-```
-
-Now we can modify the input table accordingly.
-
-We go through all the farms and determine the 'improvement' coefficients (based on herd structure).
-
-```{r}
-farms<-unique(Kenyaherds$QQNO)
-breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)
-
-Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
-                                        "system",
-                                        "animal_type",
-                                        "breed",
-                                        "live_weight_LW",
-                                        "mature_weight_MW",
-                                        "weight_gain_WG",
-                                        "milk_yield",
-                                        "dig_energy",
-                                        "CP.")]
-
-Kenyaherds_interventions[,"milk_change"]<-0
-cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
-Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )
-
-milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows], Kenyaherds_interventions$animal_type[cows]), FUN=mean)
-
-Milk_changes_4<-milk_change[which(milk_change$Group.2==4),]
-breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% Milk_changes_4[,1])]<-Milk_changes_4[,3]
-Milk_changes_5<-milk_change[which(milk_change$Group.2==5),]
-breed_intervention$milk_yield_5[which(breed_intervention$Farm %in% Milk_changes_5[,1])]<-Milk_changes_5[,3]
-Milk_changes_6<-milk_change[which(milk_change$Group.2==6),]
-breed_intervention$milk_yield_6[which(breed_intervention$Farm %in% Milk_changes_6[,1])]<-Milk_changes_6[,3]
-
-for(i in 1:nrow(Kenyaherds_interventions))
-  {Kenyaherds_interventions[i,"live_weight_change"]<-
-    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
-                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
-#   Kenyaherds_interventions[i,"mature_weight_change"]<-
-#      1/mature_weights$Intervention_factor[which(mature_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
-#                                                 mature_weights$breed==Kenyaherds_interventions$breed[i])]
-   Kenyaherds_interventions[i,"weight_gain_change"]<-
-     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
-                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
-}
-
-live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
-                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
-#mature_weight_agg<-aggregate(Kenyaherds_interventions$mature_weight_change,
-#by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
-weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
-                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
-
-colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
-#colnames(mature_weight_agg)<-c("farm","animal_type","conv_factor")
-colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")
-
-for(i in 1:nrow(live_weight_agg))
-  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
-                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
-    live_weight_agg$conv_factor[i]
-
-#for(i in 1:nrow(mature_weight_agg))
-#  breed_intervention[which(breed_intervention$Farm==mature_weight_agg$farm[i]),
-#                     paste0("mature_weight_LW_",mature_weight_agg$animal_type[i])]<-
-#    mature_weight_agg$conv_factor[i]
-
-for(i in 1:nrow(weight_gain_agg))
-  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
-                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
-    weight_gain_agg$conv_factor[i]
-
-breed_intervention[is.na(breed_intervention)]<-1
-
-```
-
-We can now go through the `x` file, and for each row take a random draw based on an adoption rate we can specify (does the farm adopt the intervention or not), then (possibly) apply factors.
-In this case we want to simulate full adoption, so we'll set this to 1.
-
-```{r}
-x_breed_intervention <- GHG_simulation_scenarios$x
-
-adoption_rate <- 1
-
-for (i in 1:nrow(x_breed_intervention))
-{
-  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
-    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
-      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
-      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
-                                 x_breed_intervention$Scenario[i]),
-                         2:ncol(breed_intervention)]
-}
-# run model for modified x file
-breed_res <- run_model(x_breed_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-breed_res$call<-"multi-scenario call"
-save_temperature_scenarios(breed_res,"results","Breed_intervention")
-```
-
-```{r, include=FALSE}
-adoption_rate <- 1
-breed_res<-load_temperature_scenarios("results","Breed_intervention")
-class(breed_res)<-"mcSimulation"
-```
-
-## Emissions intensity response to upgrading breeds
-
-Exchanging cows of unproductive breeds to more productive individuals should have an effect on emissions intensity, yet this effect is difficult to see at the population level, both in the scatter plot and in the surface plot:
-
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 8)}
-breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
-
-plotdata_breed<-data.frame(in_var=breed_res$y$pc_milk_yield,
-                     out_var=breed_res$y$CO2eq_per_milk)
-
-plotdata_baseline<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
-                     out_var=GHG_simulation_scenarios$y$CO2eq_per_milk)
-
-plot_breed <- ggplot(plotdata_breed, aes(x = in_var, y = out_var)) + 
-  geom_point(shape = 20, color = "blue", size = 2) + 
-  stat_smooth(method = "loess", fullrange = TRUE) +
-  geom_density_2d() +
-  geom_rug(color="lightskyblue3") + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() +
-  theme(legend.position = c(0.15, 0.9)) + ggtitle("'Upgrading breeds' scenario")
-
-plot_baseline <- ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) + 
-  geom_point(shape = 20, color = "blue", size = 2) + 
-  stat_smooth(method = "loess", fullrange = TRUE) +
-  geom_density_2d() +
-  scale_y_continuous(position = "right") +
-  geom_rug(color="lightskyblue3", sides="bl") + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() +
-  theme(legend.position = c(0.15, 0.9)) + ggtitle("Baseline scenario")
-
-dens_breed<-ggplot(plotdata_breed, aes(x = in_var, y = out_var)) +
-  stat_density_2d(
-    geom = "raster",
-    aes(fill = after_stat(density)),
-    contour = FALSE
-    ) +
-  scale_fill_viridis_c() + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() + guides(fill="none")
-
-dens_baseline<-ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) +
-  stat_density_2d(
-    geom = "raster",
-    aes(fill = after_stat(density)),
-    contour = FALSE
-    ) +
-    scale_y_continuous(position = "right") +
-  scale_fill_viridis_c() + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() + guides(fill="none")
-
-plot_breed + plot_baseline + dens_breed + dens_baseline +
-  plot_layout(ncol = 2, nrow = 2)
-
-ggsave("Fig_11_breed_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
-
-```
-
-We may even feel the need to confirm that changes did in fact take place. The scatter plot below shows changes in emissions intensity and per capita milk yield in response to the intervention.
-
-To put these results into perspective, we can plot them on top of the probability-of-difference plot we generated above (bottom plot below).
-The data points indicate the milk yield level before implementing the intervention (x-axis) and the milk yield level after upgrading breeds.
-We can see that the differences that are caused by the intervention are usually associated with relatively low probabilities.
-This means that the magnitude of changes that occurred in response to the intervention may also have arisen by chance.
-We know that this is not the case here (since we made these changes happen).
-Nevertheless, these results imply that an evaluator using milk yield as an indicator and requiring a high level of certainty that emissions intensity is in fact low may easily fail to reward the farmers' efforts.
-
-```{r, fig.dim = c(8, 8)}
-breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
-
-breed_res$y<-breed_res$y[normal_milk_yield,]
-breed_res$x<-breed_res$x[normal_milk_yield,]
-
-impact_of_breed<-data.frame(baseline=dairy_households$y$pc_milk_yield,breed_milk=breed_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_breed=breed_res$y$CO2eq_per_milk)
-impact_of_breed[,"milk_change"]<-round(impact_of_breed$breed_milk-impact_of_breed$baseline,2)
-impact_of_breed[,"emissions_intensity_change"]<-round(impact_of_breed$CO2_intensity_breed-impact_of_breed$CO2_intensity_baseline ,3)
-
-breed_scatter<-ggplot(data=impact_of_breed,
-                      aes(x=milk_change,y=emissions_intensity_change)) +
-  geom_point() +
-  theme_bw() +
-  ggtitle("Milk yield and emissions intensity response") + 
-  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)")
-
-breed_probs<-ggplot(data = decreased_emissions_probability,
-                    aes(Base_yield, new_yield)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),
-               aes(z=probability),col="black") +
-  geom_point(data=impact_of_breed[which(impact_of_breed$milk_change>0),],aes(x=baseline,y=breed_milk)) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") + theme_bw()
-
-breed_scatter + breed_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_12_breed_reductions.png", width=9,height=9)
-
-```
-
-Note that the linear patterns (of ten points each) in this plot result from the 10 replicates that were considered for each farm in the original Monte Carlo simulation.
-The originally random milk yields increased by the same factor in all these replicates, leading to linear relationships.
-
-### Characteristics of sensitive farms
-
-The figure above shows the probability of detecting changes in greenhouse gas emissions intensity by comparing milk yields, for farms that have implemented the breed upgrade intervention.
-We can now extract the most likely probability of such detection for each of the farms (based on the nearest data point on the probability surface).
-This allows us to detect farms that are associated with particularly high probabilities (e.g.
-90%).
-We can then look in detail at the characteristics of these farms to identify farms that are particularly likely to benefit (in terms of reducing emissions intensities) from the breed upgrade intervention.
-
-```{r}
-
-breed_diagnosis<-breed_res
-breed_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
-
-# get closest probability value
-
-for(i in 1:nrow(breed_diagnosis$x))
-  {decreased_emissions_probability[,"diff"]<-
-    abs(decreased_emissions_probability$Base_yield-breed_diagnosis$y$baseline_pc_milk_yield[i]) +
-    abs(decreased_emissions_probability$new_yield-breed_diagnosis$y$pc_milk_yield[i])
-  breed_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
-}
-
-breed_90_farms<-breed_diagnosis$x$Scenario[which(breed_diagnosis$y$confidence_level>0.9)]
-
-```
-
-The only farm characteristics that we modified in this intervention is the breed composition, so the success of the intervention can only depend on the array of dairy cow breeds present on the farms.
-We can plot the breed composition of the farms against the probability that higher per-capita milk production is associated with lower emissions intensity.
-
-```{r}
-
-farm_structures<-data.frame(Farm=unique(Kenyaherds$QQNO))
-
-for(f in farm_structures$Farm)
-  {samp<-Kenyaherds[which(Kenyaherds$QQNO==f),]
-   cows<-samp[which(samp$animal_type %in% 4:6),]
-   farm_structures[which(farm_structures$Farm==f),"breed_1"]<-length(which(cows$breed==1))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_2"]<-length(which(cows$breed==2))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_3"]<-length(which(cows$breed==3))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_4"]<-length(which(cows$breed==4))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_5"]<-length(which(cows$breed==5))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_6"]<-length(which(cows$breed==6))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_7"]<-length(which(cows$breed==7))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"breed_8"]<-length(which(cows$breed==8))/nrow(cows)
-   farm_structures[which(farm_structures$Farm==f),"unproductive_bulls"]<-
-     length(which(samp$animal_type==1))/nrow(samp)
-   farm_structures[which(farm_structures$Farm==f),"replacement_males"]<-
-     length(which(samp$animal_type==3))/nrow(samp)
-
-}
-
-breed_diagnosis$x["Farm"]<-sapply(breed_diagnosis$x$Scenario,
-                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
-
-breed_structures<-merge(breed_diagnosis$x,farm_structures)
-breed_structures<-list(x=breed_structures, y=breed_diagnosis$y)
-
-to_plot<-cbind(Farm=breed_structures$x$Farm,Confidence_level=breed_structures$y$confidence_level,
-               breed_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
-
-melted<-melt(to_plot, id.vars=c("Farm","Confidence_level","unproductive_bulls","replacement_males"))
-melted<-melted[which(!is.na(melted$value)),]
-
-melted$Confidence_level_rounded<-round(melted$Confidence_level,2)*100
-
-breedcol<-palette.colors(n = 8, palette = "Okabe-Ito", alpha=1, recycle = FALSE)
-
-ggplot(melted, aes(group=variable, fill=variable, y=value, x=Confidence_level_rounded)) + 
-    geom_bar(position="fill", stat="identity") + coord_flip() +
-  labs(y="Cumulative share of breed in herd",
-       x="Probability of lower emissions intensity (%)",
-       fill="Breed of \ndairy cows") + scale_fill_manual(values= c("breed_1" = breedcol[1],
-                                                                   "breed_2" = breedcol[2],
-                                                                   "breed_3" = breedcol[3],
-                                                                   "breed_4" = breedcol[4],
-                                                                   "breed_5" = breedcol[5],
-                                                                   "breed_6" = breedcol[6],
-                                                                   "breed_7" = breedcol[7],
-                                                                   "breed_8" = breedcol[8]),
-                                               labels = c("Holstein Friesian (pure/improved)",
-                                                          "Ayrshire (pure/improved)",
-                                                          "Jersey (pure/improved)",
-                                                          "Guernsey (pure/improved)",
-                                                          "Cross-bred unknown",
-                                                          "Sahiwal",
-                                                          "Boran",
-                                                          "Local zebu")) +
-  theme_bw()
-
-ggsave("Fig_13_breed_diagnostics.png",width=5,height=5)
-
-```
-
-# Herd structure intervention 1 - retiring unproductive bulls
-
-To simulate this scenario, we simply have to remove all animals of type 1 from the herds. Again, we assume an adoption rate of 100%.
-
-```{r , warning=FALSE}
-x_nobull_intervention <- GHG_simulation_scenarios$x
-
-nobull_adoption_rate <- 1.0
-
-for (i in 1:nrow(x_nobull_intervention))
-{
-  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
-    x_nobull_intervention$N_animal_type_1[i]<-0
-}
-# run model for modified x file
-nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE}
-nobull_res$call<-"multi-scenario call"
-save_temperature_scenarios(nobull_res,"results","noBull_intervention")
-```
-
-```{r, include=FALSE}
-nobull_res<-load_temperature_scenarios("results","noBull_intervention")
-class(nobull_res)<-"mcSimulation"
-```
-
-## Emissions intensity response to retiring unproductive bulls
-
-Removing all bulls >36 months of age clearly leads to a reduction in the overall emissions intensity of milk production, yet this effect is difficult to see when looking at the overall distribution of milk yield and emissions intensity in the region.
-
-```{r, message=FALSE, warning=FALSE, fig.dim = c(8, 8)}
-
-nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
-
-plotdata_nobull<-data.frame(in_var=nobull_res$y$pc_milk_yield,
-                     out_var=nobull_res$y$CO2eq_per_milk)
-
-plotdata_baseline<-data.frame(in_var=GHG_simulation_scenarios$y$pc_milk_yield,
-                     out_var=GHG_simulation_scenarios$y$CO2eq_per_milk)
-
-plot_nobull <- ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) + 
-  geom_point(shape = 20, color = "blue", size = 2) + 
-  stat_smooth(method = "loess", fullrange = TRUE) +
-  geom_density_2d() +
-  geom_rug(color="lightskyblue3") + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() +
-  theme(legend.position = c(0.15, 0.9)) + ggtitle("'No bulls' scenario")
-
-plot_baseline <- ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) + 
-  geom_point(shape = 20, color = "blue", size = 2) + 
-  stat_smooth(method = "loess", fullrange = TRUE) +
-  geom_density_2d() +
-  scale_y_continuous(position = "right") +
-  geom_rug(color="lightskyblue3", sides="bl") + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() +
-  theme(legend.position = c(0.15, 0.9)) + ggtitle("Baseline scenario")
-
-dens_nobull<-ggplot(plotdata_nobull, aes(x = in_var, y = out_var)) +
-  stat_density_2d(
-    geom = "raster",
-    aes(fill = after_stat(density)),
-    contour = FALSE
-    ) +
-  scale_fill_viridis_c() + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() + guides(fill="none")
-
-dens_baseline<-ggplot(plotdata_baseline, aes(x = in_var, y = out_var)) +
-  stat_density_2d(
-    geom = "raster",
-    aes(fill = after_stat(density)),
-    contour = FALSE
-    ) +
-    scale_y_continuous(position = "right") +
-  scale_fill_viridis_c() + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Total~milk~yield~(kg~head^{-1}~year^{-1}))) +
-  theme_bw() + guides(fill="none")
-
-plot_nobull + plot_baseline + dens_nobull + dens_baseline +
-  plot_layout(ncol = 2, nrow = 2)
-
-ggsave("Fig_14_no_bulls_scatter_per_head_milk_emissions_intensity.png",width=9,height=9)
-
-```
-
-Also for this intervention, we can verify that emissions intensity has actually decreased in response to the intervention (top panel below).
-Also here, however, we can see that most of the farms that made such changes are again located in the "low-confidence" region of the plot, where the impacts of the intervention are difficult to distinguish from the general variability within the population.
-For some farms, the effect is strong enough to possibly lead to a milk-yield-based reward, yet such high-confidence cases are rare also in this dataset.
-
-```{r, fig.dim = c(8, 8)}
-nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
-
-nobull_res$y<-nobull_res$y[normal_milk_yield,]
-nobull_res$x<-nobull_res$x[normal_milk_yield,]
-
-impact_of_nobull<-data.frame(baseline=dairy_households$y$pc_milk_yield,nobull_milk=nobull_res$y$pc_milk_yield,CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,CO2_intensity_nobull=nobull_res$y$CO2eq_per_milk)
-impact_of_nobull[,"milk_change"]<-round(impact_of_nobull$nobull_milk-impact_of_nobull$baseline,2)
-impact_of_nobull[,"emissions_intensity_change"]<-round(impact_of_nobull$CO2_intensity_nobull-impact_of_nobull$CO2_intensity_baseline ,3)
-
-nobull_scatter<-ggplot(data=impact_of_nobull,
-                      aes(x=milk_change,y=emissions_intensity_change)) +
-  geom_point() +
-  theme_bw() +
-  ggtitle("Milk yield and emissions intensity response") + 
-  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)")
-
-nobull_probs<-ggplot(data = decreased_emissions_probability,
-                    aes(Base_yield, new_yield)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),
-               aes(z=probability),col="black") +
-  geom_point(data=impact_of_nobull[which(impact_of_nobull$milk_change>0),],aes(x=baseline,y=nobull_milk)) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") + theme_bw()
-
-nobull_scatter + nobull_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_15_nobull_reductions.png", width=9,height=9)
-
-```
-
-The confidence level depends on the farms' herd structure, specifically on the share of unproductive males with the farms' herds.
-
-```{r}
-
-nobull_diagnosis<-nobull_res
-nobull_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
-
-# get closest probability value
-
-for(i in 1:nrow(nobull_diagnosis$x))
-  {decreased_emissions_probability[,"diff"]<-
-    abs(decreased_emissions_probability$Base_yield-nobull_diagnosis$y$baseline_pc_milk_yield[i]) +
-    abs(decreased_emissions_probability$new_yield-nobull_diagnosis$y$pc_milk_yield[i])
-  nobull_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
-}
-
-nobull_90_farms<-nobull_diagnosis$x$Scenario[which(nobull_diagnosis$y$confidence_level>0.9)]
-
-nobull_diagnosis$x["Farm"]<-sapply(nobull_diagnosis$x$Scenario,
-                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
-
-nobull_structures<-merge(nobull_diagnosis$x,farm_structures)
-nobull_structures<-list(x=nobull_structures, y=nobull_diagnosis$y)
-
-to_plot<-cbind(Farm=nobull_structures$x$Farm,Confidence_level=nobull_structures$y$confidence_level,
-               nobull_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
-to_plot$other_animals<-1-to_plot$unproductive_bulls
-
-melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","unproductive_bulls")], id.vars=c("Farm","Confidence_level"))
-melted<-melted[which(!is.na(melted$value)),]
-
-melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100
-
-ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
-    geom_bar(position="fill", stat="identity") + coord_flip() +
-  labs(y="Share of animal type in the herd",
-       x="Probability of lower emissions intensity (%)",
-       fill="Animal type") + scale_fill_manual(values= c("unproductive_bulls" = "burlywood",
-                                                         "other_animals" = "darkslategrey"),
-                                               labels = c("Unproductive bulls","Other animals")) +
-  theme_bw()
-
-ggsave("Fig_16_nobull_diagnostics.png",width=5,height=5)
-
-```
-
-# Herd structure intervention 2 - fewer unproductive replacement males
-
-Another strategy in reducing emissions intensity is to raise the off-take rate of replacement males.
-We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate.
-
-```{r , warning=FALSE}
-x_lessReplace_intervention <- GHG_simulation_scenarios$x
-
-lessReplace_adoption_rate <- 0.23
-
-for (i in 1:nrow(x_lessReplace_intervention))
-{
-  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
-    x_lessReplace_intervention$N_animal_type_3[i]<-0
-}
-# run model for modified x file
-lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)
-
-```
-
-```{r, include=FALSE}
-lessReplace_res$call<-"multi-scenario call"
-save_temperature_scenarios(lessReplace_res,"results","lessReplace_intervention")
-```
-
-```{r, include=FALSE}
-lessReplace_res<-load_temperature_scenarios("results","lessReplace_intervention")
-class(lessReplace_res)<-"mcSimulation"
-```
-
-## Emissions intensity response to fewer unproductive replacement males
-
-As for the other interventions, the impact of reducing the number of unproductive replacement males is clearly reflected in a drop in emissions intensity (see scatter plot below).
-However, also here few data points were located in the high-confidence region of the probability-of-difference plot, so that a milk yield-based reward scheme may easily fail to recognize the efforts of the respective farms.
-
-```{r, fig.dim = c(8, 8)}
-
-lessReplace_res$y[,"CO2eq_per_milk"]<-lessReplace_res$y$pc_on_farm/lessReplace_res$y$pc_milk_yield
-
-lessReplace_res$y<-lessReplace_res$y[normal_milk_yield,]
-lessReplace_res$x<-lessReplace_res$x[normal_milk_yield,]
-
-impact_of_lessReplace<-data.frame(baseline=dairy_households$y$pc_milk_yield,
-                                  lessReplace_milk=lessReplace_res$y$pc_milk_yield,
-                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
-                                  CO2_intensity_breed=lessReplace_res$y$CO2eq_per_milk)
-impact_of_lessReplace[,"milk_change"]<-
-  round(impact_of_lessReplace$lessReplace_milk-impact_of_lessReplace$baseline,2)
-impact_of_lessReplace[,"emissions_intensity_change"]<-
-  round(impact_of_lessReplace$CO2_intensity_breed-impact_of_lessReplace$CO2_intensity_baseline ,3)
-
-lessRep_scatter<-ggplot(data=impact_of_lessReplace,
-                        aes(x=milk_change,y=emissions_intensity_change)) +
-  geom_point() +
-  theme_bw() +
-  ggtitle("Intervention impact on milk yield and emissions") + 
-  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)")
-
-lessRep_probs<-ggplot(data = decreased_emissions_probability,
-                      aes(Base_yield, new_yield)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_lessReplace[which(impact_of_lessReplace$milk_change>0),],
-             aes(x=baseline,y=lessReplace_milk)) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") + theme_bw()
-
-lessRep_scatter + lessRep_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_17_lessReplace_reductions.png", width=9,height=9)
-
-```
-
-Also with this intervention, the level of confidence we can have in higher per-capita milk yields being associated with lower emissions intensity is related to the farms' herd structures.
-In this case, the share of replacement males within the herds is a key determinant.
-
-```{r}
-
-lessReplace_diagnosis<-lessReplace_res
-lessReplace_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
-
-# get closest probability value
-
-for(i in 1:nrow(lessReplace_diagnosis$x))
-  {decreased_emissions_probability[,"diff"]<-
-    abs(decreased_emissions_probability$Base_yield-lessReplace_diagnosis$y$baseline_pc_milk_yield[i]) +
-    abs(decreased_emissions_probability$new_yield-lessReplace_diagnosis$y$pc_milk_yield[i])
-  lessReplace_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
-}
-
-lessReplace_90_farms<-lessReplace_diagnosis$x$Scenario[which(lessReplace_diagnosis$y$confidence_level>0.9)]
-
-lessReplace_diagnosis$x["Farm"]<-sapply(lessReplace_diagnosis$x$Scenario,
-                                  function(x) as.numeric(strsplit(x,"_")[[1]][2]))
-
-lessReplace_structures<-merge(lessReplace_diagnosis$x,farm_structures)
-lessReplace_structures<-list(x=lessReplace_structures, y=lessReplace_diagnosis$y)
-
-to_plot<-cbind(Farm=lessReplace_structures$x$Farm,Confidence_level=lessReplace_structures$y$confidence_level,
-               lessReplace_structures$x[,c(paste0("breed_",1:8),"unproductive_bulls","replacement_males")])
-to_plot$other_animals<-1-to_plot$replacement_males
-
-melted<-melt(to_plot[,c("Farm","Confidence_level","other_animals","replacement_males")], id.vars=c("Farm","Confidence_level"))
-melted<-melted[which(!is.na(melted$value)),]
-
-melted$Confidence_level_rounded<-round(to_plot$Confidence_level,2)*100
-
-ggplot(melted, aes(group=variable, fill=variable, y=value , x=Confidence_level_rounded)) + 
-    geom_bar(position="fill", stat="identity") + coord_flip() +
-  labs(y="Share of animal type in the herd",
-       x="Probability of lower emissions intensity (%)",
-       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
-                                                         "other_animals" = "darkslategrey"),
-                                               labels = c("Replacement males","Other animals")) +
-  theme_bw()
-
-ggsave("Fig_18_lessReplace_diagnostics.png",width=5,height=5)
-
-```
-
-# Reliability of effort detection via milk yield per capity
-
-- run another Monte Carlo
-- transfer interventions to the results
-- identify for each intervention all the households that didn't adopt the intervention
-- use this to
-
-In practice, rewards for emission reductions are likely to be awarded for changes in changes in per-capita milk yield compared to a baseline.
-These changes may result from farmers adopting innovative practices or from random processes that are not reflective of conscious adaptation action.
-Development programs aiming to reduce emissions should aim to reward farmers who adopt low-emissions practices, rather than providing undeserved rewards for farmers who retain traditional production patterns but simply experience rising milk yield caused by random processes.
-
-To determine the suitability of per-capita milk yield as an indicator of adaptation action, we now simulate ,milk production and emissions at a second point in time.
-The farms that were selected for the three innovations implement the intended measures, but the production and emissions of their farms - like all other farms - are again subjected to random processes.
-This is implemented through a second Monte Carlo simulation, with the interventions applied to the results in the same way as before.
-We can now quantify changes in per-capita milk yields between the two simulated times.
-We can also determine the reliability of identifying intervention adopters through this metric.
-We assume that evaluators will require a certain level of confidence that emissions reductions have occurred, which they derive from the probabilistic analysis of the survey data (see above).
-For each level of required confidence, we can calculate the type 1 error - the percentage of innovators that remain undetected, as well as the type 2 error - the percentage of non-innovators that are falsely rewarded for changes that arose by chance.
-
-```{r monte_carlo_simulation2, warning=FALSE, eval=FALSE}
-
-GHG_simulation_scenarios_time2<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
-                                            scenarios = read.csv("data/farm_scenarios.csv",
-                                                                 fileEncoding = "UTF-8-BOM"),
-                                            model_function = ghg_emissions,
-                                            numberOfModelRuns = 10,
-                                            functionSyntax = "plainNames")
-
-```
-
-```{r, eval=FALSE, include=FALSE}
-dir.create("results")
-save_temperature_scenarios(GHG_simulation_scenarios_time2,"results","Endline")
-```
-
-```{r, include=FALSE}
-GHG_simulation_scenarios_time2<-load_temperature_scenarios("results","Endline")
-class(GHG_simulation_scenarios)<-"mcSimulation"
-```
-
-## Breed intervention
-
-We apply the same intervention to the dataset as described above.
-
-```{r}
-xb<-x_breed_intervention
-basel<-GHG_simulation_scenarios$x
-
-# breed intervention
-breed_adopters<-which(!xb$milk_yield_4*xb$N_animal_type_4==basel$milk_yield_4*basel$N_animal_type_4 |
-                        !xb$milk_yield_5*xb$N_animal_type_5==basel$milk_yield_5*basel$N_animal_type_5 |
-                        !xb$milk_yield_6*xb$N_animal_type_6==basel$milk_yield_6*basel$N_animal_type_6)
-
-x_breed_intervention_endline <- GHG_simulation_scenarios_time2$x
-
-adoption_rate <- 1
-
-for (i in 1:nrow(x_breed_intervention_endline))
-{
-  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
-    x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
-      x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
-      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
-                                 x_breed_intervention_endline$Scenario[i]),
-                         2:ncol(breed_intervention)]
-}
-
-# run model for modified x file
-breed_res_endline <- run_model(x_breed_intervention_endline, ghg_emissions)
-
-```
-
-```{r, include=FALSE}
-breed_res_endline$call<-"multi-scenario call"
-save_temperature_scenarios(breed_res_endline,"results","Breed_intervention_endline")
-```
-
-```{r, include=FALSE}
-adoption_rate <- 1
-breed_res_endline<-load_temperature_scenarios("results","Breed_intervention_endline")
-class(breed_res_endline)<-"mcSimulation"
-```
-
-We can make scatter and probability plots as above.
-The results are richer now, because the second Monte Carlo simulation introduced additional variation.
-We can now also visually distinguish between changes that occurred on adopting farms and those occurring on non-adopting farms.
-
-```{r, fig.dim = c(8, 8)}
-breed_res_endline$y$breed_adopter[breed_adopters]<-"Adopter"
-breed_res_endline$y$breed_adopter[is.na(breed_res_endline$y$breed_adopter)]<-"Non-adopter"
-
-breed_res_endline$y[,"CO2eq_per_milk"]<-breed_res_endline$y$pc_on_farm/breed_res_endline$y$pc_milk_yield
-
-breed_res_endline$y<-breed_res_endline$y[normal_milk_yield,]
-breed_res_endline$x<-breed_res_endline$x[normal_milk_yield,]
-
-impact_of_breed_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
-                                  breed_endline_milk=breed_res_endline$y$pc_milk_yield,
-                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
-                                  CO2_intensity_breed=breed_res_endline$y$CO2eq_per_milk,
-                                  Adopter=breed_res_endline$y$breed_adopter)
-impact_of_breed_endline[,"milk_change"]<-
-  round(impact_of_breed_endline$breed_endline_milk-impact_of_breed_endline$baseline,2)
-impact_of_breed_endline[,"emissions_intensity_change"]<-
-  round(impact_of_breed_endline$CO2_intensity_breed-impact_of_breed_endline$CO2_intensity_baseline ,3)
-
-breed_endline_scatter<-ggplot(data=impact_of_breed_endline,
-                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
-  geom_point(size=0.5) +
-  theme_bw() +
-  ggtitle("Intervention impact on milk yield and emissions") + 
-  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
-
-breed_endline_probs<-ggplot(data = decreased_emissions_probability,
-                      aes(Base_yield, new_yield)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_breed_endline[which(impact_of_breed_endline$milk_change>0),],
-             aes(x=baseline,y=breed_endline_milk,color=Adopter), size=0.5) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
-
-breed_endline_scatter + breed_endline_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_19_breed_endline.png", width=9,height=9)
-
-```
-
-Finally, we can calculate the percentatge of adopting farms that would be overlooked at a particular level of required confidence (Type I error).
-We can also compute the percentage of farms that would be identified as adopters, even though they made no changes in their production practices (Type II error).
-
-```{r}
-breed_endline_diagnosis<-breed_res_endline
-breed_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
-
-# get closest probability value
-
-for(i in 1:nrow(breed_endline_diagnosis$x))
-  {decreased_emissions_probability[,"diff"]<-
-    abs(decreased_emissions_probability$Base_yield-breed_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
-    abs(decreased_emissions_probability$new_yield-breed_endline_diagnosis$y$pc_milk_yield[i])
-  breed_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
-}
-
-confidence_level_analysis_breed<-data.frame(confidence=seq(0.5,1,0.01))
-n_adopt<-length(which(breed_endline_diagnosis$y$breed_adopter=="Adopter"))
-n_nonadopt<-length(which(breed_endline_diagnosis$y$breed_adopter=="Non-adopter"))
-for(i in 1:nrow(confidence_level_analysis_breed))
-{confidence_level_analysis_breed$Type1error[i]<-length(which(breed_endline_diagnosis$y$breed_adopter=="Adopter" &
-                                                 breed_endline_diagnosis$y$confidence_level<confidence_level_analysis_breed$confidence[i])) / n_adopt * 100
- confidence_level_analysis_breed$Type2error[i]<-length(which(breed_endline_diagnosis$y$breed_adopter=="Non-adopter" &
-                                                 breed_endline_diagnosis$y$confidence_level>confidence_level_analysis_breed$confidence[i])) / n_nonadopt * 100
-
-}
-confidence_level_analysis_breed$confidence<-confidence_level_analysis_breed$confidence*100
-
-toplot<-melt(confidence_level_analysis_breed,id.vars = "confidence")
-toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
-                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
-
-ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
-  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
-
-ggsave("Fig_20_breed_endline_errors.png", width=5,height=5)
-
-```
-
-## Herd structure intervention 1 - retiring unproductive bulls
-
-```{r}
-
-# nobull intervention
-nobull_adopters<-which(!x_nobull_intervention$N_animal_type_1==GHG_simulation_scenarios$x$N_animal_type_1)
-
-x_nobull_intervention_endline <- GHG_simulation_scenarios_time2$x
-
-nobull_adoption_rate <- 1.0
-
-for (i in 1:nrow(x_nobull_intervention_endline))
-{
-  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
-    x_nobull_intervention_endline$N_animal_type_1[i]<-0
-}
-# run model for modified x file
-nobull_res_endline <- run_model(x_nobull_intervention_endline, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-nobull_res_endline$call<-"multi-scenario call"
-save_temperature_scenarios(nobull_res_endline,"results","Nobull_intervention_endline")
-```
-
-```{r, include=FALSE}
-adoption_rate <- 1
-nobull_res_endline<-load_temperature_scenarios("results","Nobull_intervention_endline")
-class(nobull_res_endline)<-"mcSimulation"
-```
-
-```{r, fig.dim = c(8, 8)}
-nobull_res_endline$y$nobull_adopters<-NA
-nobull_res_endline$y$nobull_adopters[nobull_adopters]<-"Adopter"
-nobull_res_endline$y$nobull_adopters[is.na(nobull_res_endline$y$nobull_adopters)]<-"Non-adopter"
-
-nobull_res_endline$y[,"CO2eq_per_milk"]<-nobull_res_endline$y$pc_on_farm/nobull_res_endline$y$pc_milk_yield
-
-nobull_res_endline$y<-nobull_res_endline$y[normal_milk_yield,]
-nobull_res_endline$x<-nobull_res_endline$x[normal_milk_yield,]
-
-impact_of_nobull_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
-                                  nobull_endline_milk=nobull_res_endline$y$pc_milk_yield,
-                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
-                                  CO2_intensity_nobull=nobull_res_endline$y$CO2eq_per_milk,
-                                  Adopter=nobull_res_endline$y$nobull_adopters)
-impact_of_nobull_endline[,"milk_change"]<-
-  round(impact_of_nobull_endline$nobull_endline_milk-impact_of_nobull_endline$baseline,2)
-impact_of_nobull_endline[,"emissions_intensity_change"]<-
-  round(impact_of_nobull_endline$CO2_intensity_nobull-impact_of_nobull_endline$CO2_intensity_baseline ,3)
-
-nobull_endline_scatter<-ggplot(data=impact_of_nobull_endline,
-                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
-  geom_point(size=0.5) +
-  theme_bw() +
-  ggtitle("Intervention impact on milk yield and emissions") + 
-  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
-
-nobull_endline_probs<-ggplot(data = decreased_emissions_probability,
-                      aes(Base_yield, new_yield)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_nobull_endline[which(impact_of_nobull_endline$milk_change>0),],
-             aes(x=baseline,y=nobull_endline_milk,color=Adopter), size=0.5) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
-
-nobull_endline_scatter + nobull_endline_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_21_nobull_endline.png", width=9,height=9)
-
-```
-
-```{r}
-nobull_endline_diagnosis<-nobull_res_endline
-nobull_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
-
-# get closest probability value
-
-for(i in 1:nrow(nobull_endline_diagnosis$x))
-  {decreased_emissions_probability[,"diff"]<-
-    abs(decreased_emissions_probability$Base_yield-nobull_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
-    abs(decreased_emissions_probability$new_yield-nobull_endline_diagnosis$y$pc_milk_yield[i])
-  nobull_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
-}
-
-confidence_level_analysis_nobull<-data.frame(confidence=seq(0.5,1,0.01))
-n_adopt<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Adopter"))
-n_nonadopt<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Non-adopter"))
-for(i in 1:nrow(confidence_level_analysis_nobull))
-{confidence_level_analysis_nobull$Type1error[i]<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Adopter" &
-                                                 nobull_endline_diagnosis$y$confidence_level<confidence_level_analysis_nobull$confidence[i])) / n_adopt * 100
- confidence_level_analysis_nobull$Type2error[i]<-length(which(nobull_endline_diagnosis$y$nobull_adopter=="Non-adopter" &
-                                                 nobull_endline_diagnosis$y$confidence_level>confidence_level_analysis_nobull$confidence[i])) / n_nonadopt * 100
-
-}
-confidence_level_analysis_nobull$confidence<-confidence_level_analysis_nobull$confidence*100
-
-toplot<-melt(confidence_level_analysis_nobull,id.vars = "confidence")
-toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
-                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
-
-ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
-  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
-
-ggsave("Fig_22_nobull_endline_errors.png", width=5,height=5)
-
-```
-
-## Herd structure intervention 2 - fewer unproductive replacement males
-
-We can do the same analysis for the reduction in the number of unproductive replacement males. First we implement the intervention as in the first Monte Carlo analysis.
-
-```{r}
-
-# lessReplace intervention
-lessReplace_adopters<-which(!x_lessReplace_intervention$N_animal_type_3==GHG_simulation_scenarios$x$N_animal_type_3)
-
-x_lessReplace_intervention_endline <- GHG_simulation_scenarios_time2$x
-
-x_lessReplace_intervention_endline$N_animal_type_3[lessReplace_adopters]<-0
-
-# run model for modified x file
-lessReplace_res_endline <- run_model(x_lessReplace_intervention_endline, ghg_emissions)
-
-```
-
-```{r, include=FALSE, eval=FALSE}
-lessReplace_res_endline$call<-"multi-scenario call"
-save_temperature_scenarios(lessReplace_res_endline,"results","lessReplace_intervention_endline")
-```
-
-```{r, include=FALSE}
-lessReplace_res_endline<-load_temperature_scenarios("results","lessReplace_intervention_endline")
-class(lessReplace_res_endline)<-"mcSimulation"
-```
-
-Now we can compare the performance of adopters and non-adopters and plot this against the confidence level for recognizing that emissions reductions have occurred.
-
-```{r, fig.dim = c(8, 8)}
-lessReplace_res_endline$y$lessReplace_adopters<-NA
-lessReplace_res_endline$y$lessReplace_adopters[lessReplace_adopters]<-"Adopter"
-lessReplace_res_endline$y$lessReplace_adopters[is.na(lessReplace_res_endline$y$lessReplace_adopters)]<-"Non-adopter"
-
-lessReplace_res_endline$y[,"CO2eq_per_milk"]<-lessReplace_res_endline$y$pc_on_farm/lessReplace_res_endline$y$pc_milk_yield
-
-lessReplace_res_endline$y<-lessReplace_res_endline$y[normal_milk_yield,]
-lessReplace_res_endline$x<-lessReplace_res_endline$x[normal_milk_yield,]
-
-impact_of_lessReplace_endline<-data.frame(baseline=dairy_households$y$pc_milk_yield,
-                                  lessReplace_endline_milk=lessReplace_res_endline$y$pc_milk_yield,
-                                  CO2_intensity_baseline= dairy_households$y$CO2eq_per_milk,
-                                  CO2_intensity_lessReplace=lessReplace_res_endline$y$CO2eq_per_milk,
-                                  Adopter=lessReplace_res_endline$y$lessReplace_adopters)
-impact_of_lessReplace_endline[,"milk_change"]<-
-  round(impact_of_lessReplace_endline$lessReplace_endline_milk-impact_of_lessReplace_endline$baseline,2)
-impact_of_lessReplace_endline[,"emissions_intensity_change"]<-
-  round(impact_of_lessReplace_endline$CO2_intensity_lessReplace-impact_of_lessReplace_endline$CO2_intensity_baseline ,3)
-
-lessReplace_endline_scatter<-ggplot(data=impact_of_lessReplace_endline,
-                        aes(x=milk_change,y=emissions_intensity_change,color=Adopter)) +
-  geom_point(size=0.5) +
-  theme_bw() +
-  ggtitle("Intervention impact on milk yield and emissions") + 
-  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)") + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
-
-lessReplace_endline_probs<-ggplot(data = decreased_emissions_probability,
-                      aes(Base_yield, new_yield)) +
-  geom_contour_filled(aes(z=probability)) +
-  stat_contour(breaks = c(0.9),aes(z=probability),col="black") +
-  geom_point(data=impact_of_lessReplace_endline[which(impact_of_lessReplace_endline$milk_change>0),],
-             aes(x=baseline,y=lessReplace_endline_milk,color=Adopter), size=0.5) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") + theme_bw() + scale_color_manual(values=c("Adopter" = "darkslategrey", "Non-adopter" = "burlywood"))
-
-lessReplace_endline_scatter + lessReplace_endline_probs +
-  plot_layout(ncol = 1, nrow = 2)
-
-ggsave("Fig_23_lessReplace_endline.png", width=9,height=9)
-
-```
-
-Finally, we can evaluate the Type 1 and Type 2 errors of a reward scheme based on per-capita milk yield.
-
-```{r}
-lessReplace_endline_diagnosis<-lessReplace_res_endline
-lessReplace_endline_diagnosis$y[,"baseline_pc_milk_yield"]<-dairy_households$y$pc_milk_yield
-
-# get closest probability value
-
-for(i in 1:nrow(lessReplace_endline_diagnosis$x))
-  {decreased_emissions_probability[,"diff"]<-
-    abs(decreased_emissions_probability$Base_yield-lessReplace_endline_diagnosis$y$baseline_pc_milk_yield[i]) +
-    abs(decreased_emissions_probability$new_yield-lessReplace_endline_diagnosis$y$pc_milk_yield[i])
-  lessReplace_endline_diagnosis$y[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
-}
-
-confidence_level_analysis_lessReplace<-data.frame(confidence=seq(0.5,1,0.01))
-n_adopt<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Adopter"))
-n_nonadopt<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Non-adopter"))
-for(i in 1:nrow(confidence_level_analysis_lessReplace))
-{confidence_level_analysis_lessReplace$Type1error[i]<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Adopter" &
-                                                 lessReplace_endline_diagnosis$y$confidence_level<confidence_level_analysis_lessReplace$confidence[i])) / n_adopt * 100
- confidence_level_analysis_lessReplace$Type2error[i]<-length(which(lessReplace_endline_diagnosis$y$lessReplace_adopter=="Non-adopter" &
-                                                 lessReplace_endline_diagnosis$y$confidence_level>confidence_level_analysis_lessReplace$confidence[i])) / n_nonadopt * 100
-
-}
-confidence_level_analysis_lessReplace$confidence<-confidence_level_analysis_lessReplace$confidence*100
-
-toplot<-melt(confidence_level_analysis_lessReplace,id.vars = "confidence")
-toplot$variable <- factor(toplot$variable, levels = c("Type1error", "Type2error"),
-                  labels = c("Type I error - failure to reward adopters", "Type II error - rewarding non-adopters"))
-
-ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=2) + theme_bw() +
-  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
-
-ggsave("Fig_24_lessReplace_endline_errors.png", width=5,height=5)
-
-```
-
-# Discussion
-
-* considerable uncertainty in inputs to IPCC equations
-* cause major uncertainty in emissions intensity estimates
-* this makes it difficult to evaluate the emissions intensity of dairy farms in Kenya based on their milk yield alone - large differences are needed to be certain that one farm is more emissions-efficient than another one
-* interventions clearly impact emissions intensity
-* emissions intensity savings are substantial for all interventions, but still relatively small in comparison to the overall uncertainty associated with limited knowledge about values of input variables.
-* it seems likely that a milk yield-based emissions reduction scheme might miss measures taken by farmers, at least if the reward scheme wants to be really sure that emissions reductions have in fact occurred.
-* it seems the information content in milk yields isn't sufficient to base mitigation reward schemes on them
-* might be a better idea after all to observe farmers' practices - did they implement any measures? Then they should be rewarded
-* maybe rewarding measures taken is a better idea than evaluating emissions intensity based on milk yield alone.
-
-# some remaining notes
-
-For each slice, and each level of confidence, calculate the percentage of farms that made changes that would be correctly rewarded vs.
-the ones that did not change (the baseline population).
-Make a figure showing the Type I and Type II errors resulting from each confidence level at each baseline yield.
-
-# References
+ggsave("figures/Fig_5_probability_of_lower_emissions_intensity.png", width=7,height=5)
+
+```
+
+Now a description of the 5 interventions
+
+Further figures to include:
+
+- Fig. 6: Scatter plot that shows the impact of all interventions on per capita milk yield and emissions intensity
+- Fig. 7: Scatter plot of all interventions against the backdrop of all samples drawn in second MC simulation (endline)
+- Fig. 8: Intervention households shown on probability surface (Fig. 5)
+- Fig. 9: Structural analysis - which household types have a chance of being detected
+- Fig. 10: Errors (Type I and Type 2) - would be good to combine all interventions in one figure (might get crowded)

### 2022-06-13 Detailed_analyses_for_paper.Rmd
corresponds to the first draft of methods

@@
      x_balance_intervention$feed_prod_CO2[i]<-x_balance_intervention$feed_prod_CO2[i]-(328.5 * 
-                    sum(x_Calliandra_intervention[i,paste0("N_animal_type_",1:12)]))
+                    sum(x_balance_intervention[i,paste0("N_animal_type_",1:12)]))
   }

### 2022-06-13 Manuscript_figures.Rmd
corresponds to the first draft of methods

@@
 variables_to_update_posnorm<-c(paste0("milk_yield_",4:6),
-                             paste0("Cp_rate_",4:6),
                              "feed_prod_CO2",
@@
 
-```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4)}
+```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "Total farm emissions and emissions intensity for farm number 5 from the farm survey, based on the outputs of a Monte Carlo simulation.
+The blue bar indicates a point estimate based on the default values recommended by the IPCC"}
 
@@
 
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4)}
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "Total greenhouse gas emissions per farm and emissions intensity across all farms in the farm survey, based on the outputs of a Monte Carlo simulation."}
 
@@
 
-all_farms_intensity<-ggplot(data=farm5_emissions$y, aes(CO2eq_per_milk)) +
+all_farms_intensity<-ggplot(data=GHG_simulation_scenarios$y, aes(CO2eq_per_milk)) +
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
@@
 
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6)}
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="Relationship between milk yield per capita and emissions intensity of milk production for farms in Kenya, based on a Monte Carlo simulation.
+Outputs are shown as a scatter plot (left) and a density surface (right)."}
 normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)
@@
   theme_void() + 
-  theme(legend.position = "none")
+  theme(legend.position = "none") + ggtitle("a) Scatter plot of farm-scale data")
 
@@
   labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1}))) +
-  theme_bw()
-
-dens1 + plot_spacer() + plot_spacer() + plot1 + dens2 + densi + 
+       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})),
+       fill="Probability \ndensity") +
+  theme_bw() + geom_vline(xintercept=c(2000,3000),color=c("#F8766D","#00BFC4"),alpha=.5, lwd=2)
+
+empty<-ggplot() +
+    theme_void() +
+    ggtitle("b) Interpolated probability density surface")
+
+dens1 + plot_spacer() + empty + plot1 + dens2 + densi + 
   plot_layout(ncol = 3, nrow = 2, widths = c(4, 1), heights = c(1, 4))
@@
 
-```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4)}
+```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="Probability distributions of greenhouse gas emissions intensity at specific levels of milk production per capita.
+Slices through the emissions intensity density surface are shown for a production level of 2000 kg milk per capita and year (left) and as a comparison between production levels of 2000 and 3000 kg milk per capita and year (right)."}
 slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
@@
 
-```{r, fig.dim = c(8, 4)}
+```{r, fig.dim = c(8, 4), fig.cap="Probability that farms that differ in per capita milk production also differ in greenhouse gas emissions intensity.
+The surface, which is based on the outputs of a Monte Carlo simulation, expresses the probability that a farm with higher milk productivity (y-axis) has a lower emissions intensity than a farm with lower milk productivity (x-axis)."}
 
@@
 
-for (i in 1:(length(milk_yields_to_sample) - 1))
+for (i in 2:(length(milk_yields_to_sample) - 1))
 {
-  for (j in (i + 1):length(milk_yields_to_sample))
+  for (j in 1:length(milk_yields_to_sample))
     {decreased_emissions_probability[nrow(decreased_emissions_probability)+1,"Base_yield"]<-milk_yields_to_sample[i]
@@
       sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
+     decreased_emissions_probability$probability[nrow(decreased_emissions_probability)] <-
+      sum(slices[[i]] > sample(slices[[j]], length(slices[[i]]))) / length(slices[[i]])
 
@@
                                                       probability)) + geom_contour_filled() +
+  ggtitle("Probability of lower emissions intensity on farm B") + 
+  labs(y=expression(Milk~yield~of~farm~B~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~farm~A~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity on farm B") +
+  stat_contour(breaks = c(0.9),col="black") + theme_bw()
+
+ggsave("figures/Fig_5_probability_of_lower_emissions_intensity.png", width=7,height=5)
+
+```
+
+Now a description of the 5 interventions
+
+```{r defining_interventions}
+# breed intervention
+
+cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
+
+ggplot(cows, aes(factor(breed),milk_yield)) +
+  geom_violin(draw_quantiles=c(0.5),fill="light green") +
+  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
+
+ggsave("figures/Fig_S1_breeds.png", width=7,height=5)
+
+milk_summary<-cbind(aggregate(cows$milk_yield, by=list(cows$breed), FUN=median),table(cows$breed))
+milk_summary<-milk_summary[,c(1,4,2)]
+colnames(milk_summary)<-c("Breed","n","Median_milk_yield")
+
+milk_summary[,"Intervention_factor"]<-sapply(milk_summary$Median_milk_yield, function(x) min(x/milk_summary$Median_milk_yield[1],1))
+
+live_weights<-aggregate(Kenyaherds$live_weight_LW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(live_weights)<-c("animal_type","breed","median_live_weight")
+live_weights[,"Intervention_factor"]<-NA
+for(i in 1:nrow(live_weights))
+  live_weights$Intervention_factor[i]<-live_weights$median_live_weight[i]/
+     live_weights$median_live_weight[which(live_weights$animal_type==live_weights$animal_type[i] &                                                                                                            live_weights$breed==1)]
+
+# mature weight should probably be changed, and the code below does this. However, the mature weight is currently the same for all the breeds, so this accomplished nothing.
+
+mature_weights<-aggregate(Kenyaherds$mature_weight_MW, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(mature_weights)<-c("animal_type","breed","median_mature_weight")
+mature_weights[,"Intervention_factor"]<-NA
+for(i in 1:nrow(mature_weights))
+  mature_weights$Intervention_factor[i]<-mature_weights$median_mature_weight[i]/
+     mature_weights$median_mature_weight[which(mature_weights$animal_type==mature_weights$animal_type[i] &                                                                                                            mature_weights$breed==1)]
+
+# same for the weight gain - doesn't depend on the breed at the moment, so all the factors are 1
+
+weight_gains<-aggregate(Kenyaherds$weight_gain_WG, by=list(Kenyaherds$animal_type,Kenyaherds$breed), FUN=median)
+colnames(weight_gains)<-c("animal_type","breed","median_weight_gain")
+weight_gains[,"Intervention_factor"]<-NA
+for(i in 1:nrow(weight_gains))
+  weight_gains$Intervention_factor[i]<-weight_gains$median_weight_gain[i]/
+     weight_gains$median_weight_gain[which(weight_gains$animal_type==weight_gains$animal_type[i] &                                                                                                            weight_gains$breed==1)]
+weight_gains$Intervention_factor[which(weight_gains$median_weight_gain==0)]<-1
+
+farms<-unique(Kenyaherds$QQNO)
+breed_intervention<-data.frame(Farm=farms, milk_yield_4=1, milk_yield_5=1, milk_yield_6=1)
+
+Kenyaherds_interventions<-Kenyaherds[,c("QQNO",
+                                        "system",
+                                        "animal_type",
+                                        "breed",
+                                        "live_weight_LW",
+                                        "mature_weight_MW",
+                                        "weight_gain_WG",
+                                        "milk_yield",
+                                        "dig_energy",
+                                        "CP.")]
+
+Kenyaherds_interventions[,"milk_change"]<-0
+cows<-which(Kenyaherds_interventions$animal_type %in% 4:6)
+Kenyaherds_interventions$milk_change[cows]<- sapply(cows,function(x) 1/milk_summary$Intervention_factor[milk_summary$Breed==Kenyaherds_interventions$breed[x]]  )
+
+milk_change<-aggregate(Kenyaherds_interventions$milk_change[cows], by=list(Kenyaherds_interventions$QQNO[cows], Kenyaherds_interventions$animal_type[cows]), FUN=mean)
+
+Milk_changes_4<-milk_change[which(milk_change$Group.2==4),]
+breed_intervention$milk_yield_4[which(breed_intervention$Farm %in% Milk_changes_4[,1])]<-Milk_changes_4[,3]
+Milk_changes_5<-milk_change[which(milk_change$Group.2==5),]
+breed_intervention$milk_yield_5[which(breed_intervention$Farm %in% Milk_changes_5[,1])]<-Milk_changes_5[,3]
+Milk_changes_6<-milk_change[which(milk_change$Group.2==6),]
+breed_intervention$milk_yield_6[which(breed_intervention$Farm %in% Milk_changes_6[,1])]<-Milk_changes_6[,3]
+
+for(i in 1:nrow(Kenyaherds_interventions))
+  {Kenyaherds_interventions[i,"live_weight_change"]<-
+    1/live_weights$Intervention_factor[which(live_weights$animal_type==Kenyaherds_interventions$animal_type[i]&
+                                             live_weights$breed==Kenyaherds_interventions$breed[i])]
+   Kenyaherds_interventions[i,"weight_gain_change"]<-
+     1/weight_gains$Intervention_factor[which(weight_gains$animal_type==Kenyaherds_interventions$animal_type[i]&
+                                              weight_gains$breed==Kenyaherds_interventions$breed[i])]
+}
+
+live_weight_agg<-aggregate(Kenyaherds_interventions$live_weight_change,
+                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+weight_gain_agg<-aggregate(Kenyaherds_interventions$weight_gain_change,
+                           by=list(Kenyaherds_interventions$QQNO,Kenyaherds_interventions$animal_type), FUN=mean)
+
+colnames(live_weight_agg)<-c("farm","animal_type","conv_factor")
+colnames(weight_gain_agg)<-c("farm","animal_type","conv_factor")
+
+for(i in 1:nrow(live_weight_agg))
+  breed_intervention[which(breed_intervention$Farm==live_weight_agg$farm[i]),
+                     paste0("live_weight_LW_",live_weight_agg$animal_type[i])]<-
+    live_weight_agg$conv_factor[i]
+
+for(i in 1:nrow(weight_gain_agg))
+  breed_intervention[which(breed_intervention$Farm==weight_gain_agg$farm[i]),
+                     paste0("weight_gain_WG_",weight_gain_agg$animal_type[i])]<-
+    weight_gain_agg$conv_factor[i]
+
+breed_intervention[is.na(breed_intervention)]<-1
+
+x_breed_intervention <- GHG_simulation_scenarios$x
+
+adoption_rate <- 1
+
+intervention_adopters<-which(rbinom(nrow(x_breed_intervention),1,adoption_rate)==1)
+
+for (i in 1:nrow(x_breed_intervention))
+{
+  if (i %in% intervention_adopters) # if farmers adopt, make changes in x file
+    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
+      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
+      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
+                                 x_breed_intervention$Scenario[i]),
+                         2:ncol(breed_intervention)]
+}
+
+breed_adopters<-which(!x_breed_intervention$milk_yield_4*x_breed_intervention$N_animal_type_4==GHG_simulation_scenarios$x$milk_yield_4*GHG_simulation_scenarios$x$N_animal_type_4 |
+                        !x_breed_intervention$milk_yield_5*x_breed_intervention$N_animal_type_5==GHG_simulation_scenarios$x$milk_yield_5*GHG_simulation_scenarios$x$N_animal_type_5 |
+                        !x_breed_intervention$milk_yield_6*x_breed_intervention$N_animal_type_6==GHG_simulation_scenarios$x$milk_yield_6*GHG_simulation_scenarios$x$N_animal_type_6)
+
+## Herd intervention 1 - nobulls
+
+x_nobull_intervention <- GHG_simulation_scenarios$x
+nobull_adoption_rate <- 1.0
+for (i in 1:nrow(x_nobull_intervention))
+{
+  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
+    x_nobull_intervention$N_animal_type_1[i]<-0
+}
+nobull_adopters<-which(!x_nobull_intervention$N_animal_type_1==GHG_simulation_scenarios$x$N_animal_type_1)
+
+## Herd intervention 2 - lessReplace
+
+x_lessReplace_intervention <- GHG_simulation_scenarios$x
+lessReplace_adoption_rate <- 0.23
+for (i in 1:nrow(x_lessReplace_intervention))
+{
+  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
+    x_lessReplace_intervention$N_animal_type_3[i]<-0
+}
+lessReplace_adopters<-which(!x_lessReplace_intervention$N_animal_type_3==GHG_simulation_scenarios$x$N_animal_type_3)
+
+## Feed intervention 1 - Supplementary legume fodder
+
+x_Calliandra_intervention <- GHG_simulation_scenarios$x
+Calliandra_adoption_rate <- 0.3
+for (i in 1:nrow(x_Calliandra_intervention))
+{
+  if (rbinom(1, 1, Calliandra_adoption_rate)) # if farmers adopt, make changes in x file
+    {x_Calliandra_intervention$milk_yield_5[i]<-x_Calliandra_intervention$milk_yield_5[i]+0.156
+     x_Calliandra_intervention$milk_yield_6[i]<-x_Calliandra_intervention$milk_yield_6[i]+0.156
+     x_Calliandra_intervention$DE_perc[i]<-x_Calliandra_intervention$DE_perc[i]+0.14
+     x_Calliandra_intervention$perc_crude_protein_diet[i]<-x_Calliandra_intervention$perc_crude_protein_diet[i]-0.07
+     x_Calliandra_intervention$feed_prod_CO2[i]<-x_Calliandra_intervention$feed_prod_CO2[i]- (0.726 *
+                                        sum(x_Calliandra_intervention[i,paste0("N_animal_type_",1:12)]))
+
+  }
+}
+Calliandra_adopters<-which(!x_Calliandra_intervention$DE_perc==GHG_simulation_scenarios$x$DE_perc)
+
+## Feed intervention 2 - Balanced diets
+
+x_balance_intervention <- GHG_simulation_scenarios$x
+balance_adoption_rate <- 0.14
+for (i in 1:nrow(x_balance_intervention))
+{
+  if (rbinom(1, 1, balance_adoption_rate)) # if farmers adopt, make changes in x file
+    {x_balance_intervention$milk_yield_5[i]<-x_balance_intervention$milk_yield_5[i]+0.75
+     x_balance_intervention$milk_yield_6[i]<-x_balance_intervention$milk_yield_6[i]+0.75
+     x_balance_intervention$feed_prod_CO2[i]<-x_balance_intervention$feed_prod_CO2[i]-(328.5 * 
+                    sum(x_balance_intervention[i,paste0("N_animal_type_",1:12)]))
+  }
+}
+balance_adopters<-which(!x_balance_intervention$feed_prod_CO2==GHG_simulation_scenarios$x$feed_prod_CO2)
+
+```
+
+To run the interventions, we need a function to run the model with a specific set of x values.
+
+```{r}
+# single model-run function
+run_model<-function(x_data,model)
+  {
+  varnames<-colnames(x_data)[which(!colnames(x_data)=="Scenario")]
+  model_function_ <- function(x) {
+
+    e <- as.list(sapply(X = varnames, FUN = function(i) as.numeric(unlist(x[i]))))
+    eval(expr = body(model), envir = e)
+  }
+  y <- do.call(what = rbind, args = lapply(X = apply(X = x_data, 
+                                                     MARGIN = 1, FUN = model_function_), FUN = unlist))
+  returnObject <- list(y = data.frame(y), x = data.frame(x_data))
+  returnObject$call <- match.call()
+  class(returnObject) <- cbind("mcSimulation", class(returnObject))
+  return(returnObject)
+}
+```
+
+```{r, eval=FALSE}
+# run model for modified x file
+breed_res <- run_model(x_breed_intervention, ghg_emissions)
+nobull_res <- run_model(x_nobull_intervention, ghg_emissions)
+lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)
+Calliandra_res <- run_model(x_Calliandra_intervention, ghg_emissions)
+balance_res <- run_model(x_balance_intervention, ghg_emissions)
+
+breed_res$y[,"CO2eq_per_milk"]<-breed_res$y$pc_on_farm/breed_res$y$pc_milk_yield
+nobull_res$y[,"CO2eq_per_milk"]<-nobull_res$y$pc_on_farm/nobull_res$y$pc_milk_yield
+lessReplace_res$y[,"CO2eq_per_milk"]<-lessReplace_res$y$pc_on_farm/lessReplace_res$y$pc_milk_yield
+Calliandra_res$y[,"CO2eq_per_milk"]<-Calliandra_res$y$pc_on_farm/Calliandra_res$y$pc_milk_yield
+balance_res$y[,"CO2eq_per_milk"]<-balance_res$y$pc_on_farm/balance_res$y$pc_milk_yield
+
+breed_res$call<-"multi-scenario call"
+save_temperature_scenarios(breed_res,"results","Breed_intervention")
+nobull_res$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_res,"results","Nobull_intervention")
+lessReplace_res$call<-"multi-scenario call"
+save_temperature_scenarios(lessReplace_res,"results","LessReplace_intervention")
+Calliandra_res$call<-"multi-scenario call"
+save_temperature_scenarios(Calliandra_res,"results","Calliandra_intervention")
+balance_res$call<-"multi-scenario call"
+save_temperature_scenarios(balance_res,"results","Balance_intervention")
+```
+
+```{r}
+breed_res<-load_temperature_scenarios("results","Breed_intervention")
+class(breed_res)<-"mcSimulation"
+nobull_res<-load_temperature_scenarios("results","Nobull_intervention")
+class(nobull_res)<-"mcSimulation"
+lessReplace_res<-load_temperature_scenarios("results","LessReplace_intervention")
+class(lessReplace_res)<-"mcSimulation"
+Calliandra_res<-load_temperature_scenarios("results","Calliandra_intervention")
+class(Calliandra_res)<-"mcSimulation"
+balance_res<-load_temperature_scenarios("results","Balance_intervention")
+class(balance_res)<-"mcSimulation"
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="Intervention impact on per capita milk production and greenhouse gas emissions intensity."}
+
+# make data.frame with all intervention results
+
+interventions<-rbind(cbind(breed_res$y[breed_adopters,],
+                           intervention="breed",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[breed_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[breed_adopters]),
+                     cbind(nobull_res$y[nobull_adopters,],
+                           intervention="nobull",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[nobull_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[nobull_adopters]),
+                     cbind(lessReplace_res$y[lessReplace_adopters,],
+                           intervention="lessReplace",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[lessReplace_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[lessReplace_adopters]),
+                     cbind(Calliandra_res$y[Calliandra_adopters,],
+                           intervention="Calliandra",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[Calliandra_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[Calliandra_adopters]),
+                     cbind(balance_res$y[balance_adopters,],
+                           intervention="balance",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[balance_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[balance_adopters]))
+
+interventions[,"milk_change"]<-interventions$pc_milk_yield-interventions$pc_milk_baseline
+interventions[,"emissions_intensity_change"]<-interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline
+interventions[,"milk_change_relative"]<-(interventions$pc_milk_yield-interventions$pc_milk_baseline)/interventions$pc_milk_baseline*100
+interventions[,"emissions_intensity_change_relative"]<-
+  (interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline)/interventions$CO2_intensity_baseline*100
+
+interventions_normal_milk<-interventions[which(interventions$pc_milk_yield>500),]
+
+intervention_change <- merge(data.frame(aggregate(interventions_normal_milk$emissions_intensity_change,
+                                                  by = list(interventions_normal_milk$intervention),
+                                                  FUN = mean)),
+                             data.frame(aggregate(interventions_normal_milk$milk_change,
+                                                  by = list(interventions_normal_milk$intervention),
+                                                  FUN = mean)),
+                             by = "Group.1")
+colnames(intervention_change) <-
+  c("intervention", "emissions_intensity_change", "milk_change")
+
+intervention_colors<-palette.colors(n = 6, palette = "Okabe-Ito", alpha=1, recycle = FALSE)
+
+interventions_scatter<-ggplot(data=interventions_normal_milk,
+                        aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point(aes(color=intervention),size=0.9) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)",
+       color="Intervention") +
+  scale_color_manual(values= c("breed" = intervention_colors[1],
+                              "nobull" = intervention_colors[2],
+                              "lessReplace" = intervention_colors[3],
+                              "Calliandra" = intervention_colors[4],
+                              "balance" = intervention_colors[6]),
+                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+  facet_wrap(~factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
+  theme(legend.position = "none")
+
+interventions_scatter
+
+ggsave("figures/Fig_6a_intervention_impacts.png", width=9,height=6)
+
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="Intervention impact on relative per capita milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not apply the respective intervention."}
+
+interventions_scatter_relative<-ggplot(data=interventions_normal_milk,
+                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
+  geom_point(aes(color=intervention),size=0.9) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y="Relative change in emissions intensity (%)",
+       x="Relative change in per capita milk yield (%)",
+       color="Intervention") +
+  scale_color_manual(values= c("breed" = intervention_colors[1],
+                              "nobull" = intervention_colors[2],
+                              "lessReplace" = intervention_colors[3],
+                              "Calliandra" = intervention_colors[4],
+                              "balance" = intervention_colors[6]),
+                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+  facet_wrap(~factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
+  theme(legend.position = "none")
+
+interventions_scatter_relative
+
+ggsave("figures/Fig_6_intervention_impacts_relative.png", width=9,height=6)
+
+```
+
+These are noticeable changes, but can they be detected against the random variation that all farms experience?
+We'll now take another Monte Carlo sample, apply the same interventions to the same farms and evaluate the results again.
+
+```{r, eval=FALSE}
+
+## Run endline Monte Carlo simulation
+GHG_simulation_scenarios_time2<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                                            scenarios = read.csv("data/farm_scenarios.csv",
+                                                                 fileEncoding = "UTF-8-BOM"),
+                                            model_function = ghg_emissions,
+                                            numberOfModelRuns = 10,
+                                            functionSyntax = "plainNames")
+
+GHG_simulation_scenarios_time2$call<-"multi-scenario call"
+save_temperature_scenarios(GHG_simulation_scenarios_time2,"results","Endline_no_intervention")
+
+```
+
+```{r}
+
+GHG_simulation_scenarios_time2<-load_temperature_scenarios("results","Endline_no_intervention")
+class(GHG_simulation_scenarios_time2)<-"mcSimulation"
+
+## Produce endline breed scenario
+x_breed_intervention_endline <- GHG_simulation_scenarios_time2$x
+
+for (i in breed_adopters)
+{x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
+ x_breed_intervention_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
+    breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
+                               x_breed_intervention_endline$Scenario[i]),
+                       2:ncol(breed_intervention)]
+}
+
+## Herd intervention 1 - nobulls
+
+x_nobull_intervention_endline <- GHG_simulation_scenarios_time2$x
+x_nobull_intervention_endline$N_animal_type_1[nobull_adopters]<-0
+
+## Herd intervention 2 - lessReplace
+
+x_lessReplace_intervention_endline <- GHG_simulation_scenarios_time2$x
+x_lessReplace_intervention_endline$N_animal_type_3[lessReplace_adopters]<-0
+
+## Feed intervention 1 - Supplementary legume fodder
+
+x_Calliandra_intervention_endline <- GHG_simulation_scenarios_time2$x
+for (i in Calliandra_adopters)
+{x_Calliandra_intervention_endline$milk_yield_5[i]<-x_Calliandra_intervention_endline$milk_yield_5[i]+0.156
+ x_Calliandra_intervention_endline$milk_yield_6[i]<-x_Calliandra_intervention_endline$milk_yield_6[i]+0.156
+ x_Calliandra_intervention_endline$DE_perc[i]<-x_Calliandra_intervention_endline$DE_perc[i]+0.14
+ x_Calliandra_intervention_endline$perc_crude_protein_diet[i]<-x_Calliandra_intervention_endline$perc_crude_protein_diet[i]-0.07
+ x_Calliandra_intervention_endline$feed_prod_CO2[i]<-x_Calliandra_intervention_endline$feed_prod_CO2[i]- (0.726 *
+                                        sum(x_Calliandra_intervention_endline[i,paste0("N_animal_type_",1:12)]))
+}
+
+## Feed intervention 2 - Balanced diets
+
+x_balance_intervention_endline <- GHG_simulation_scenarios_time2$x
+for (i in balance_adopters)
+{x_balance_intervention$milk_yield_5[i]<-x_balance_intervention$milk_yield_5[i]+0.75
+ x_balance_intervention$milk_yield_6[i]<-x_balance_intervention$milk_yield_6[i]+0.75
+ x_balance_intervention$feed_prod_CO2[i]<-x_balance_intervention$feed_prod_CO2[i]-(328.5 * 
+                    sum(x_balance_intervention[i,paste0("N_animal_type_",1:12)]))
+}
+
+```
+
+```{r, eval=FALSE}
+# run model for modified x file (endline)
+breed_end <- run_model(x_breed_intervention_endline, ghg_emissions)
+nobull_end <- run_model(x_nobull_intervention_endline, ghg_emissions)
+lessReplace_end <- run_model(x_lessReplace_intervention_endline, ghg_emissions)
+Calliandra_end <- run_model(x_Calliandra_intervention_endline, ghg_emissions)
+balance_end <- run_model(x_balance_intervention_endline, ghg_emissions)
+
+breed_end$y[,"CO2eq_per_milk"]<-breed_end$y$pc_on_farm/breed_end$y$pc_milk_yield
+nobull_end$y[,"CO2eq_per_milk"]<-nobull_end$y$pc_on_farm/nobull_end$y$pc_milk_yield
+lessReplace_end$y[,"CO2eq_per_milk"]<-lessReplace_end$y$pc_on_farm/lessReplace_end$y$pc_milk_yield
+Calliandra_end$y[,"CO2eq_per_milk"]<-Calliandra_end$y$pc_on_farm/Calliandra_end$y$pc_milk_yield
+balance_end$y[,"CO2eq_per_milk"]<-balance_end$y$pc_on_farm/balance_end$y$pc_milk_yield
+
+breed_end$call<-"multi-scenario call"
+save_temperature_scenarios(breed_end,"results","End_Breed_intervention")
+nobull_end$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_end,"results","End_Nobull_intervention")
+lessReplace_end$call<-"multi-scenario call"
+save_temperature_scenarios(lessReplace_end,"results","End_LessReplace_intervention")
+Calliandra_end$call<-"multi-scenario call"
+save_temperature_scenarios(Calliandra_end,"results","End_Calliandra_intervention")
+balance_end$call<-"multi-scenario call"
+save_temperature_scenarios(balance_end,"results","End_Balance_intervention")
+```
+
+```{r, include=FALSE}
+breed_end<-load_temperature_scenarios("results","End_Breed_intervention")
+class(breed_end)<-"mcSimulation"
+nobull_end<-load_temperature_scenarios("results","End_Nobull_intervention")
+class(nobull_end)<-"mcSimulation"
+lessReplace_end<-load_temperature_scenarios("results","End_LessReplace_intervention")
+class(lessReplace_end)<-"mcSimulation"
+Calliandra_end<-load_temperature_scenarios("results","End_Calliandra_intervention")
+class(Calliandra_end)<-"mcSimulation"
+balance_end<-load_temperature_scenarios("results","End_Balance_intervention")
+class(balance_end)<-"mcSimulation"
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="Intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions."}
+
+# make data.frame with all intervention results
+
+interventions_end<-rbind(cbind(breed_end$y[breed_adopters,],
+                           intervention="breed",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[breed_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[breed_adopters]),
+                     cbind(nobull_end$y[nobull_adopters,],
+                           intervention="nobull",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[nobull_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[nobull_adopters]),
+                     cbind(lessReplace_end$y[lessReplace_adopters,],
+                           intervention="lessReplace",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[lessReplace_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[lessReplace_adopters]),
+                     cbind(Calliandra_end$y[Calliandra_adopters,],
+                           intervention="Calliandra",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[Calliandra_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[Calliandra_adopters]),
+                     cbind(balance_end$y[balance_adopters,],
+                           intervention="balance",
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[balance_adopters],
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[balance_adopters]))
+
+normal_end<-cbind(GHG_simulation_scenarios_time2$y,
+                           pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield,
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk)
+normal_end[,"CO2eq_per_milk"]<-normal_end$pc_on_farm/normal_end$pc_milk_yield
+
+interventions_end[,"milk_change"]<-interventions_end$pc_milk_yield-interventions_end$pc_milk_baseline
+interventions_end[,"emissions_intensity_change"]<-interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline
+interventions_end[,"milk_change_relative"]<-(interventions_end$pc_milk_yield-interventions_end$pc_milk_baseline)/interventions_end$pc_milk_baseline*100
+interventions_end[,"emissions_intensity_change_relative"]<-
+  (interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline)/interventions_end$CO2_intensity_baseline*100
+
+normal_end[,"milk_change"]<-normal_end$pc_milk_yield-normal_end$pc_milk_baseline
+normal_end[,"emissions_intensity_change"]<-normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline
+normal_end[,"milk_change_relative"]<-(normal_end$pc_milk_yield-normal_end$pc_milk_baseline)/normal_end$pc_milk_baseline*100
+normal_end[,"emissions_intensity_change_relative"]<-
+  (normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline)/normal_end$CO2_intensity_baseline*100
+
+interventions_normal_milk_end<-interventions_end[which(interventions_end$pc_milk_baseline>500),]
+normal_normal_milk_end<-normal_end[which(normal_end$pc_milk_yield>500),]
+normal_plot<-rbind(cbind(normal_normal_milk_end,intervention="breed"),
+                   cbind(normal_normal_milk_end,intervention="nobull"),
+                   cbind(normal_normal_milk_end,intervention="lessReplace"),
+                   cbind(normal_normal_milk_end,intervention="Calliandra"),
+                   cbind(normal_normal_milk_end,intervention="balance"))
+
+intervention_colors<-palette.colors(n = 6, palette = "Okabe-Ito", alpha=1, recycle = FALSE)
+
+interventions_scatter_end<-ggplot(data=normal_plot,
+                        aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point(color="gray",size=1) +
+  geom_point(data=interventions_normal_milk_end,
+                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
+  scale_color_manual(values= c("breed" = intervention_colors[1],
+                              "nobull" = intervention_colors[2],
+                              "lessReplace" = intervention_colors[3],
+                              "Calliandra" = intervention_colors[4],
+                              "balance" = intervention_colors[6]),
+                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per capita milk yield (kg)",
+       color="Intervention")  +
+  facet_wrap(~factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
+  theme(legend.position = "none")
+
+interventions_scatter_end
+
+ggsave("figures/Fig_7a_intervention_impacts_endline.png", width=9,height=6)
+
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="Relative intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions."}
+
+interventions_scatter_relative_end<-ggplot(data=normal_plot,
+                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
+  geom_point(color="gray",size=1) +
+  geom_point(data=interventions_normal_milk_end,
+                        aes(x=milk_change_relative,y=emissions_intensity_change_relative, color=intervention),size=0.8) +
+  scale_color_manual(values= c("breed" = intervention_colors[1],
+                              "nobull" = intervention_colors[2],
+                              "lessReplace" = intervention_colors[3],
+                              "Calliandra" = intervention_colors[4],
+                              "balance" = intervention_colors[6]),
+                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y="Relative change in emissions intensity (%)",
+       x="Relative change in per capita milk yield (%)",
+       color="Intervention") +
+  facet_wrap(~factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
+  theme(legend.position = "none")
+
+interventions_scatter_relative_end
+
+ggsave("figures/Fig_7_intervention_impacts_endline_relative.png", width=9,height=6)
+
+```
+
+```{r, fig.dim = c(8, 4), fig.cap="Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
+the intial samping time.
+The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity."}
+
+interventions_plot<-interventions_normal_milk_end
+none_plot<-cbind(normal_normal_milk_end,intervention="none")
+int_none_plot<-rbind(interventions_plot,none_plot)
+
+interventions_probability<-ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
   ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
-  labs(y=expression(Milk~yield~of~'higher-productivity'~farm~(kg~head^{-1}~year^{-1})),
-       x=expression(Milk~yield~of~'lower-productivity'~farm~(kg~head^{-1}~year^{-1})),
+  labs(y=expression(Milk~yield~at~endline~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~at~baseline~(kg~head^{-1}~year^{-1})),
        fill="Probability of \nlower emissions \nintensity") +
-  stat_contour(breaks = c(0.9),col="black") + theme_bw()
-
-ggsave("figures/Fig_5_probability_of_lower_emissions_intensity.png", width=7,height=5)
-
-```
-
-Now a description of the 5 interventions
+  stat_contour(aes(z=probability),breaks = c(0.9),col="black") + theme_bw() +
+  geom_point(data=int_none_plot,aes(x=pc_milk_baseline,y=pc_milk_yield), size=0.8) +
+  ylim(c(0,NA)) + xlim(c(0,NA)) +
+  facet_wrap(~factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance","none"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet", "No intervention"))  )
+
+interventions_probability
+
+ggsave("figures/Fig_8_probability_of_lower_emissions_intensity_interventions.png", width=7,height=5)
+
+```
+
+```{r, fig.dim = c(4, 6), fig.cap="Probability of Type I and Type II error incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred.
+Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects."}
+
+# get closest probability value
+
+for(i in 1:nrow(int_none_plot))
+  {decreased_emissions_probability[,"diff"]<-
+    abs(decreased_emissions_probability$Base_yield-int_none_plot$pc_milk_baseline [i]) +
+    abs(decreased_emissions_probability$new_yield-int_none_plot$pc_milk_yield[i])
+  int_none_plot[i,"confidence_level"]<-decreased_emissions_probability$probability[which(decreased_emissions_probability$diff==min(decreased_emissions_probability$diff))][1]
+}
+
+confidence_level_analysis<-data.frame(confidence=seq(0.25,1,0.01))
+
+for(i in 1:nrow(confidence_level_analysis))
+{# compute Type 1 errors for all interventions
+  for(ints in c("breed", "nobull", "lessReplace", "Calliandra", "balance"))
+    confidence_level_analysis[i,ints]<-
+      length(which(int_none_plot$intervention==ints &
+                     int_none_plot$confidence_level <
+                     confidence_level_analysis$confidence[i])) /
+      length(which(int_none_plot$intervention==ints)) * 100
+
+  confidence_level_analysis$Type2error[i]<-
+    length(which(int_none_plot$intervention=="none" &
+                   int_none_plot$confidence_level>confidence_level_analysis$confidence[i])) /
+    length(which(int_none_plot$intervention=="none")) * 100
+
+}
+
+confidence_level_analysis$confidence<-confidence_level_analysis$confidence*100
+toplot<-melt(confidence_level_analysis,id.vars = "confidence")
+toplot$variable <- factor(toplot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
+                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
+                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
+                             "Type I error - failure to reward 'Fewer replacement males' adopters",
+                             "Type I error - failure to reward 'Calliandra feed' adopters",
+                             "Type I error - failure to reward 'Balanced diet' adopters",
+                             "Type II error - rewarding non-adopters"))
+
+ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
+
+ggsave("figures/Fig_9_erorr_types.png", width=5,height=6)
+
+```
+
+and now for the sensitivity analysis
+
+```{r, fig.dim = c(8, 5), fig.cap="Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
+"}
+
+pls_GHG<-dairy_households
+pls_GHG$x<-pls_GHG$x[,which(!colnames(pls_GHG$x)=="Scenario")]
+
+pls_out<-plsr.mcSimulation(object=pls_GHG, resultName = "CO2eq_per_milk")
+
+legend_table<-read.csv("data/legend.csv", fileEncoding = "UTF-8-BOM")
+
+plot_pls(pls_out,input_table=legend_table)
+ggsave("figures/Fig_10_PLS_dairy_households.png", width=8,height=5)
+```
 
@@
 
-- Fig. 6: Scatter plot that shows the impact of all interventions on per capita milk yield and emissions intensity
-- Fig. 7: Scatter plot of all interventions against the backdrop of all samples drawn in second MC simulation (endline)
-- Fig. 8: Intervention households shown on probability surface (Fig. 5)
-- Fig. 9: Structural analysis - which household types have a chance of being detected
-- Fig. 10: Errors (Type I and Type 2) - would be good to combine all interventions in one figure (might get crowded)
+- Fig. 11: Structural analysis - which household types have a chance of being detected
+maybe this is more of a supplementary figure
+- Fig. 10: Sensitivity analysis

### 2022-06-13 Manuscript_figures.Rmd
more progress

@@
 
-```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "Total farm emissions and emissions intensity for farm number 5 from the farm survey, based on the outputs of a Monte Carlo simulation.
-The blue bar indicates a point estimate based on the default values recommended by the IPCC"}
+```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "***Figure 1.*** *Estimated total emissions from livestock (a) and emissions intensity of milk production (b) for a smallholder dairy farm in Kenya.
+Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.
+The blue line indicates a precise estimate based on precise results from a farm survey and default parameters recommended by various guidance documents.*"}
 
@@
 
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "Total greenhouse gas emissions per farm and emissions intensity across all farms in the farm survey, based on the outputs of a Monte Carlo simulation."}
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livestock (a) and emissions intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms).
+Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
 
@@
 
-We're interested in the suitability of total per capita milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.
-
-Here's the relationship between these two variables.
-We exclude farms with a per capita milk yield of <500 kg per year, since such farms can't be considered dairy farms and they would create outliers in many analysis.
-
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="Relationship between milk yield per capita and emissions intensity of milk production for farms in Kenya, based on a Monte Carlo simulation.
-Outputs are shown as a scatter plot (left) and a density surface (right)."}
+The
+
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure S1.*** *Greenhouse gas emissions from livestock per head of cattle across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms).
+Plausible estimates are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
+all_farms_per_capita<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
+geom_histogram(aes(y=..density..), colour="black", fill="white")+
+ geom_density(alpha=.2, fill="#FF6666") +
+  ylab("Relative probability") + theme_bw() +
+    xlab(expression(Total~greenhouse~gas~emissions~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per capita")
+
+all_farms_per_capita
+
+ggsave("figures/Fig_S1_emissions_from_all_farms_per_capita.png",width=9,height=5)
+
+```
+
+We can now look at summary statistics of total and per capita emissions from cattle across livestock-keeping smallholder farms in Kenya.
+To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per capita milk yields of >500 kg of milk.
+
+```{r}
+
 normal_milk_yield<-which(GHG_simulation_scenarios$y$pc_milk_yield>500)
@@
 dairy_households$y<-dairy_households$y[normal_milk_yield,]
+
+emissions_summary<-rbind(data.frame("Emissions per farm"=round(quantile(GHG_simulation_scenarios$y$on_farm,
+                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
+                                    "Emissions per capita"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
+                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
+                                    "Emissions intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
+                                                                   c(0,0.05,0.5,0.95,1)),1)),
+                         data.frame("Emissions per farm"=round(quantile(dairy_households$y$on_farm,
+                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
+                                    "Emissions per capita"=round(quantile(dairy_households$y$pc_on_farm,
+                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
+                                    "Emissions intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
+                                                                   c(0,0.05,0.5,0.95,1)),1)))
+emissions_summary$Quantile<-c("Minimum","5% quantile", "Median", "95% quantile", "Maximum",
+                              "Minimum","5% quantile", "Median", "95% quantile", "Maximum")
+rownames(emissions_summary)<-NULL
+
+kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
+      col.names = c("Distribution attribute",
+                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
+                    paste0("CO","$_2$","-eq head","$^{-1}$"),
+                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
+      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
+  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
+column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
+  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per capita" = 1, "Emissions intensity" = 1))
+
+```
+
+\
+
+We're interested in the suitability of total per capita milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.
+
+Here's the relationship between these two variables, for farms with per-capita milk yields of >500 kg per year.
+
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per capita (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b).
+The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig.
+4b).*"}
 
@@
 
-```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="Probability distributions of greenhouse gas emissions intensity at specific levels of milk production per capita.
-Slices through the emissions intensity density surface are shown for a production level of 2000 kg milk per capita and year (left) and as a comparison between production levels of 2000 and 3000 kg milk per capita and year (right)."}
+```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-capita milk production among livestock-keeping smallholder farms in Kenya.
+Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
 slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="Probability that farms that differ in per capita milk production also differ in greenhouse gas emissions intensity.
-The surface, which is based on the outputs of a Monte Carlo simulation, expresses the probability that a farm with higher milk productivity (y-axis) has a lower emissions intensity than a farm with lower milk productivity (x-axis)."}
+```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-capita milk production levels.
+Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-capita milk production and emissions intensity.
+This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}
 
@@
 
-```{r defining_interventions}
+```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Observed distributions of milk yields for all cattle breeds encountered among dairy cows in a survey of livestock-keeping smallholder farms in Kenya.*"}
 # breed intervention
@@
 cows<-Kenyaherds[which(Kenyaherds$animal_type %in% 4:6),]
-
-ggplot(cows, aes(factor(breed),milk_yield)) +
+cowcount<-data.frame(table(cows$breed))
+
+breed_yield_plot<-ggplot(cows, aes(factor(breed),milk_yield)) +
   geom_violin(draw_quantiles=c(0.5),fill="light green") +
-  xlab("Breed") + ylab("Milk yield (kg per day)") + theme_bw()
-
-ggsave("figures/Fig_S1_breeds.png", width=7,height=5)
+  xlab("Breed") + ylab(expression(Milk~yield~(kg~cow^{-1}~day^{-1}))) +
+  geom_boxplot(width=0.1,fill="forestgreen") +
+  theme_bw() +
+  coord_flip() +
+  scale_x_discrete(labels=c("Holstein Friesian (pure/improved)",
+                            "Ayrshire (pure/improved)",
+                            "Jersey (pure/improved)",
+                            "Guernsey (pure/improved)",
+                            "Cross-bred unknown",
+                            "Boran",
+                            "Local zebu")) +
+  geom_text(data=cowcount ,aes(x = Var1, y = -2, label=paste("n =",Freq)),size = 4)
+
+breed_milk_summary<-data.frame(Breed=c("Holstein Friesian (pure/improved)",
+                            "Ayrshire (pure/improved)",
+                            "Jersey (pure/improved)",
+                            "Guernsey (pure/improved)",
+                            "Cross-bred unknown",
+                            "Boran",
+                            "Local zebu"),
+                            cow_number=cowcount$Freq,
+                            milk_yield=round(ggplot_build(breed_yield_plot)$data[[2]]$middle,1),
+                            intervention_factor= round(ggplot_build(breed_yield_plot)$data[[2]]$middle/
+                             ggplot_build(breed_yield_plot)$data[[2]]$middle[1],2))
+breed_milk_summary$intervention_factor[which(breed_milk_summary$Breed=="Boran")]<-1
+
+breed_yield_plot
+
+ggsave("figures/Fig_S2_breeds.png", width=9,height=4)
+
+```
+```{r}
+kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
+      caption = "<p style='color:black'><i><b>Table S2.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
+column_spec(2:4, width = "5cm")
+
+```
+
+\
+
+```{r}
 

### 2022-06-14 Manuscript_figures.Rmd
first complete draft of methods and results

@@
 
-```{r, fig.dim = c(8, 8), fig.cap="Intervention impact on per capita milk production and greenhouse gas emissions intensity."}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per capita milk production and greenhouse gas emissions intensity.*"}
 
@@
                            pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[breed_adopters],
-                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[breed_adopters]),
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[breed_adopters],
+                           Farm=breed_res$x$Scenario[breed_adopters]),
                      cbind(nobull_res$y[nobull_adopters,],
@@
                            pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[nobull_adopters],
-                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[nobull_adopters]),
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[nobull_adopters],
+                           Farm=nobull_res$x$Scenario[nobull_adopters]),
                      cbind(lessReplace_res$y[lessReplace_adopters,],
@@
                            pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[lessReplace_adopters],
-                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[lessReplace_adopters]),
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[lessReplace_adopters],
+                           Farm=lessReplace_res$x$Scenario[lessReplace_adopters]),
                      cbind(Calliandra_res$y[Calliandra_adopters,],
@@
                            pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[Calliandra_adopters],
-                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[Calliandra_adopters]),
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[Calliandra_adopters],
+                           Farm=Calliandra_res$x$Scenario[Calliandra_adopters]),
                      cbind(balance_res$y[balance_adopters,],
@@
                            pc_milk_baseline=GHG_simulation_scenarios$y$pc_milk_yield[balance_adopters],
-                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[balance_adopters]))
+                           CO2_intensity_baseline=GHG_simulation_scenarios$y$CO2eq_per_milk[balance_adopters],
+                           Farm=balance_res$x$Scenario[balance_adopters]))
 
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="Intervention impact on relative per capita milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not apply the respective intervention."}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per capita milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}
 
@@
 
+The herd structure interventions showed potential for causing major changes in milk yields and emissions intensity on particular farms.
+Plotting these impacts against the initial sets of relevant herd attributes serves to illustrate which farms are particularly responsive to these interventions (Fig.
+S3).
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S3.*** *Influence of the breed composition of farm-scale dairy cow populations on changes in milk yield and emissions intensity of milk production in response to a breed upgrade intervention*."}
+
+interventions_diagnostics<-interventions
+
+farm_structures<-data.frame(Farm_no=unique(Kenyaherds$QQNO))
+
+for(f in farm_structures$Farm_no)
+  {samp<-Kenyaherds[which(Kenyaherds$QQNO==f),]
+   cows<-samp[which(samp$animal_type %in% 4:6),]
+   farm_structures[which(farm_structures$Farm_no==f),"Farm"]<-paste0("Farm_",f)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_1"]<-length(which(cows$breed==1))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_2"]<-length(which(cows$breed==2))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_3"]<-length(which(cows$breed==3))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_4"]<-length(which(cows$breed==4))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_5"]<-length(which(cows$breed==5))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_6"]<-length(which(cows$breed==6))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_7"]<-length(which(cows$breed==7))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"breed_8"]<-length(which(cows$breed==8))/nrow(cows)
+   farm_structures[which(farm_structures$Farm_no==f),"unproductive_bulls"]<-
+     length(which(samp$animal_type==1))/nrow(samp)
+   farm_structures[which(farm_structures$Farm_no==f),"replacement_males"]<-
+     length(which(samp$animal_type==3))/nrow(samp)
+
+}
+
+intervention_structures<-merge(interventions_diagnostics,farm_structures)
+intervention_structures[,"not_unproductive_bulls"]<-1-intervention_structures$unproductive_bulls
+intervention_structures[,"not_replacement_males"]<-1-intervention_structures$replacement_males
+
+melted_breed<-melt(intervention_structures[,c("Farm",
+                                              "intervention",
+                                              "milk_change_relative",
+                                              "emissions_intensity_change_relative",
+                                              paste0("breed_",1:8),
+                                              "unproductive_bulls",
+                                              "replacement_males")],
+                   id.vars=c("Farm",
+                             "intervention",
+                             "milk_change_relative",
+                             "emissions_intensity_change_relative",
+                             "unproductive_bulls",
+                             "replacement_males"))
+melted_breed<-melted_breed[which(!is.na(melted_breed$value)),]
+
+#melted$Confidence_level_rounded<-round(melted$Confidence_level,2)*100
+
+colnames(melted_breed)[which(colnames(melted_breed)=="value")]<-"breed_share"
+colnames(melted_breed)[which(colnames(melted_breed)=="variable")]<-"breed"
+
+metric_melt_breed<-melt(melted_breed,id.vars=c("Farm","intervention",
+                                "unproductive_bulls","replacement_males","breed","breed_share"))
+metric_melt_breed$value<-round(metric_melt_breed$value)
+colnames(metric_melt_breed)[which(colnames(metric_melt_breed)=="value")]<-"change"
+colnames(metric_melt_breed)[which(colnames(metric_melt_breed)=="variable")]<-"metric"
+
+breedcol<-palette.colors(n = 8, palette = "Okabe-Ito", alpha=1, recycle = FALSE)
+
+metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
+names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")
+
+breed_diagnostic<-
+  ggplot(metric_melt_breed[which(metric_melt_breed$intervention=="breed"),], aes(group=breed, fill=breed, y=breed_share, x=change)) + 
+    geom_bar(position="fill", stat="identity") +
+  scale_y_continuous(labels = scales::percent) + coord_flip() +
+  labs(y="Cumulative share of breed in herd",
+       x="Relative change in outcome metric (%)",
+       fill="Breed of \ndairy cows") + scale_fill_manual(values= c("breed_1" = breedcol[1],
+                                                                   "breed_2" = breedcol[2],
+                                                                   "breed_3" = breedcol[3],
+                                                                   "breed_4" = breedcol[4],
+                                                                   "breed_5" = breedcol[5],
+                                                                   "breed_6" = breedcol[6],
+                                                                   "breed_7" = breedcol[7],
+                                                                   "breed_8" = breedcol[8]),
+                                               labels = c("Holstein Friesian (pure/improved)",
+                                                          "Ayrshire (pure/improved)",
+                                                          "Jersey (pure/improved)",
+                                                          "Guernsey (pure/improved)",
+                                                          "Cross-bred unknown",
+                                                          "Sahiwal",
+                                                          "Boran",
+                                                          "Local zebu")) +
+  theme_bw() + facet_wrap(~metric, 
+  labeller = labeller(metric = metrics))
+
+breed_diagnostic
+
+ggsave("figures/Fig_S3_breed_diagnostics.png",width=8,height=5)
+
+```
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S4.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}
+
+melted_nobull<-melt(intervention_structures[which(intervention_structures$intervention=="nobull"),
+                                            c("Farm",
+                                              "milk_change_relative",
+                                              "emissions_intensity_change_relative",
+                                              "unproductive_bulls",
+                                              "not_unproductive_bulls")],
+                   id.vars=c("Farm",
+                             "milk_change_relative",
+                             "emissions_intensity_change_relative"))
+
+colnames(melted_nobull)[which(colnames(melted_nobull)=="value")]<-"percentage"
+colnames(melted_nobull)[which(colnames(melted_nobull)=="variable")]<-"animal_type"
+
+metric_melt_nobull<-melt(melted_nobull,
+                         id.vars=c("Farm",
+                                   "animal_type",
+                                   "percentage"))
+metric_melt_nobull$value<-round(metric_melt_nobull$value)
+colnames(metric_melt_nobull)[which(colnames(metric_melt_nobull)=="value")]<-"change"
+colnames(metric_melt_nobull)[which(colnames(metric_melt_nobull)=="variable")]<-"metric"
+
+metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
+names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")
+
+metric_melt_nobull<-metric_melt_nobull[which(!is.na(metric_melt_nobull$change)),]
+
+metric_melt_nobull$percentage<-round(metric_melt_nobull$percentage,2)
+
+nobull_diagnostic<-
+  ggplot(metric_melt_nobull, aes(group=animal_type, fill=animal_type, y=percentage, x=change)) + 
+    geom_bar(position="fill", stat="identity") +
+  scale_y_continuous(labels = scales::percent) + coord_flip() +
+  labs(y="Cumulative share of breed in herd",
+       x="Relative change in outcome metric (%)",
+       fill="Animal type") + scale_fill_manual(values= c("unproductive_bulls" = "burlywood",
+                                                         "not_unproductive_bulls" = "darkslategrey"),
+                                               labels = c("Unproductive bulls",
+                                                          "Other animals")) +
+  theme_bw() + facet_wrap(~metric, 
+  labeller = labeller(metric = metrics))
+
+nobull_diagnostic
+
+ggsave("figures/Fig_S4_nobull_diagnostics.png",width=8,height=5)
+
+```
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S5.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'fewer replacement males' intervention*."}
+
+melted_lessReplace<-melt(intervention_structures[which(intervention_structures$intervention=="lessReplace"),
+                                            c("Farm",
+                                              "milk_change_relative",
+                                              "emissions_intensity_change_relative",
+                                              "replacement_males",
+                                              "not_replacement_males")],
+                   id.vars=c("Farm",
+                             "milk_change_relative",
+                             "emissions_intensity_change_relative"))
+
+colnames(melted_lessReplace)[which(colnames(melted_lessReplace)=="value")]<-"percentage"
+colnames(melted_lessReplace)[which(colnames(melted_lessReplace)=="variable")]<-"animal_type"
+
+metric_melt_lessReplace<-melt(melted_lessReplace,
+                         id.vars=c("Farm",
+                                   "animal_type",
+                                   "percentage"))
+metric_melt_lessReplace$value<-round(metric_melt_lessReplace$value)
+colnames(metric_melt_lessReplace)[which(colnames(metric_melt_lessReplace)=="value")]<-"change"
+colnames(metric_melt_lessReplace)[which(colnames(metric_melt_lessReplace)=="variable")]<-"metric"
+
+metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
+names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")
+
+metric_melt_lessReplace<-metric_melt_lessReplace[which(!is.na(metric_melt_lessReplace$change)),]
+
+metric_melt_lessReplace$percentage<-round(metric_melt_lessReplace$percentage,2)
+
+lessReplace_diagnostic<-
+  ggplot(metric_melt_lessReplace, aes(group=animal_type, fill=animal_type, y=percentage, x=change)) + 
+    geom_bar(position="fill", stat="identity") +
+  scale_y_continuous(labels = scales::percent) + coord_flip() +
+  labs(y="Cumulative share of breed in herd",
+       x="Relative change in outcome metric (%)",
+       fill="Animal type") + scale_fill_manual(values= c("replacement_males" = "burlywood",
+                                                         "not_replacement_males" = "darkslategrey"),
+                                               labels = c("Replacement males",
+                                                          "Other animals")) +
+  theme_bw() + facet_wrap(~metric, 
+  labeller = labeller(metric = metrics))
+
+lessReplace_diagnostic
+
+ggsave("figures/Fig_S5_lessReplace_diagnostics.png",width=8,height=5)
+
+```
+
+Summary statistics of interventions for the counterfactual simulation
+
+```{r}
+
+ints<-c("breed","nobull","lessReplace","Calliandra","balance")
+
+interventions_summary_milk<-data.frame(Intervention=c("Breed upgrade",
+                                                 "Retiring unproductive bulls",
+                                                 "Fewer replacement males",
+                                                 "Calliandra feed",
+                                                 "Balanced diet"))
+interventions_summary_emissions<-interventions_summary_milk
+
+for(i in 1:length(ints))
+  {interventions_summary_milk[i,c("Weakest effect",
+                                  "5% quantile",
+                                  "Median effect",
+                                  "95% quantile",
+                                  "Strongest effect")]<-
+    round(quantile(intervention_structures[which(intervention_structures$intervention==ints[i]),
+                                           "milk_change_relative"],c(0,0.05,0.5,0.95,1),
+                   na.rm=TRUE),1)
+  interventions_summary_emissions[i,c("Weakest effect",
+                                      "5% quantile",
+                                      "Median effect",
+                                      "95% quantile",
+                                      "Strongest effect")]<-
+    round(quantile(intervention_structures[which(intervention_structures$intervention==ints[i]),
+                                           "emissions_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
+                   na.rm=TRUE),1)}
+
+intervention_summary<-rbind(interventions_summary_milk,interventions_summary_emissions)
+
+kable(intervention_summary, booktabs = TRUE,digits=0,
+     # col.names = c("Intervention",
+       #             paste0("CO","$_2$","-eq farm","$^{-1}$"),
+        #            paste0("CO","$_2$","-eq head","$^{-1}$"),
+         #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
+      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
+Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.
+</i></p>") %>%
+  pack_rows(index = c("Impact on milk yield per capita (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
+column_spec(2:6, width = "3cm")
+
+```
+
+\
+
 These are noticeable changes, but can they be detected against the random variation that all farms experience?
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="Intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
-The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions."}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
 
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="Relative intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
-The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions."}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
 
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
-the intial samping time.
-The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity."}
+Summary statistics of interventions for the realistic-endline simulation
+
+```{r}
+
+ints<-c("breed","nobull","lessReplace","Calliandra","balance")
+
+interventions_end_summary_milk<-data.frame(Intervention=c("Breed upgrade",
+                                                 "Retiring unproductive bulls",
+                                                 "Fewer replacement males",
+                                                 "Calliandra feed",
+                                                 "Balanced diet",
+                                                 "No intervention"))
+interventions_end_summary_emissions<-interventions_end_summary_milk
+
+for(i in 1:length(ints))
+  {interventions_end_summary_milk[i,c("Weakest effect",
+                                  "5% quantile",
+                                  "Median effect",
+                                  "95% quantile",
+                                  "Strongest effect")]<-
+    round(quantile(interventions_normal_milk_end[which(interventions_normal_milk_end$intervention==ints[i]),
+                                           "milk_change_relative"],c(0,0.05,0.5,0.95,1),
+                                      na.rm=TRUE),1)
+interventions_end_summary_emissions[i,c("Weakest effect",
+                                      "5% quantile",
+                                      "Median effect",
+                                      "95% quantile",
+                                      "Strongest effect")]<-
+    round(quantile(interventions_normal_milk_end[which(interventions_normal_milk_end$intervention==ints[i]),
+                                           "emissions_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
+                   na.rm=TRUE),1)}
+
+interventions_end_summary_milk[6,c("Weakest effect",
+                                      "5% quantile",
+                                      "Median effect",
+                                      "95% quantile",
+                                      "Strongest effect")]<-
+  round(quantile(normal_normal_milk_end[,"milk_change_relative"],
+                 c(0,0.05,0.5,0.95,1),
+                 na.rm=TRUE),1)
+interventions_end_summary_emissions[6,c("Weakest effect",
+                                      "5% quantile",
+                                      "Median effect",
+                                      "95% quantile",
+                                      "Strongest effect")]<-
+  round(quantile(normal_normal_milk_end[,"emissions_intensity_change_relative"],
+                 c(1,0.95,0.5,0.05,0),
+                 na.rm=TRUE),1)
+
+intervention_end_summary<-rbind(interventions_end_summary_milk,interventions_end_summary_emissions)
+
+kable(intervention_end_summary, booktabs = TRUE,digits=0,
+      caption = "<p style='color:black'><i><b>Table S4.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
+Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures.
+At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
+  pack_rows(index = c("Impact on milk yield per capita (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
+column_spec(2:6, width = "3cm")
+
+```
+
+\
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure 8.
+*** *Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
+the initial sampling time.
+The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity.*"}
 
@@
   stat_contour(aes(z=probability),breaks = c(0.9),col="black") + theme_bw() +
-  geom_point(data=int_none_plot,aes(x=pc_milk_baseline,y=pc_milk_yield), size=0.8) +
+  geom_point(data=int_none_plot,aes(x=pc_milk_baseline,y=pc_milk_yield), size=0.8, color="coral1") +
   ylim(c(0,NA)) + xlim(c(0,NA)) +
@@
 
-```{r, fig.dim = c(4, 6), fig.cap="Probability of Type I and Type II error incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred.
-Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects."}
+```{r, fig.dim = c(4, 6), fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred.
+Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
 
@@
 
-```{r, fig.dim = c(8, 5), fig.cap="Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
-"}
+```{r, fig.dim = c(8, 5), fig.cap="***Figure 10.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.*"}
 
@@
 
-Further figures to include:
-
-- Fig. 11: Structural analysis - which household types have a chance of being detected
-maybe this is more of a supplementary figure
-- Fig. 10: Sensitivity analysis
+# References

### 2022-06-16 Manuscript_figures.Rmd
updated milk yield measurements

@@
 
-The
-
 ```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure S1.*** *Greenhouse gas emissions from livestock per head of cattle across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms).
 Plausible estimates are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
-all_farms_per_capita<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
+all_farms_per_head<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
@@
   ylab("Relative probability") + theme_bw() +
-    xlab(expression(Total~greenhouse~gas~emissions~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per capita")
-
-all_farms_per_capita
-
-ggsave("figures/Fig_S1_emissions_from_all_farms_per_capita.png",width=9,height=5)
-
-```
-
-We can now look at summary statistics of total and per capita emissions from cattle across livestock-keeping smallholder farms in Kenya.
-To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per capita milk yields of >500 kg of milk.
+    xlab(expression(Total~greenhouse~gas~emissions~per~head~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per head")
+
+all_farms_per_head
+
+ggsave("figures/Fig_S1_emissions_from_all_farms_per_head.png",width=9,height=5)
+
+```
+
+We can now look at summary statistics of total and per head emissions from cattle across livestock-keeping smallholder farms in Kenya.
+To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per head milk yields of >500 kg of milk.
 
@@
                                                                   c(0,0.05,0.5,0.95,1))/1000,1),
-                                    "Emissions per capita"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
+                                    "Emissions per head"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
                                                                     c(0,0.05,0.5,0.95,1))/1000,1),
@@
                                                                   c(0,0.05,0.5,0.95,1))/1000,1),
-                                    "Emissions per capita"=round(quantile(dairy_households$y$pc_on_farm,
+                                    "Emissions per head"=round(quantile(dairy_households$y$pc_on_farm,
                                                                     c(0,0.05,0.5,0.95,1))/1000,1),
@@
 column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
-  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per capita" = 1, "Emissions intensity" = 1))
+  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per head" = 1, "Emissions intensity" = 1))
 
@@
 
-We're interested in the suitability of total per capita milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.
-
-Here's the relationship between these two variables, for farms with per-capita milk yields of >500 kg per year.
-
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per capita (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b).
+We're interested in the suitability of total per head milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.
+
+Here's the relationship between these two variables, for farms with per-head milk yields of >500 kg per year.
+
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per head (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b).
 The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig.
@@
   labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})))+
+       x=expression(Milk~yield~per~head~(kg~head^{-1}~year^{-1})))+
   theme_bw() +
@@
   labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
-       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~per~head~(kg~head^{-1}~year^{-1})),
        fill="Probability \ndensity") +
@@
 
-According to these distributions, the emissions intensity at a particular per capita milk yield can be described by a slice through the probability surface.
-
-```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-capita milk production among livestock-keeping smallholder farms in Kenya.
+According to these distributions, the emissions intensity at a particular per head milk yield can be described by a slice through the probability surface.
+
+```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-head milk production among livestock-keeping smallholder farms in Kenya.
 Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
@@
   ylab("Relative probability") +
-    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")
+    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~head~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")
+    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~head~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")
 
@@
 By resampling from these distributions, we can determine the likelihood that a value of one of these distributions is greater than one of the other.
-We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per capita.
-
-```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-capita milk production levels.
-Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-capita milk production and emissions intensity.
+We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per head.
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels.
+Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emissions intensity.
 This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per capita milk production and greenhouse gas emissions intensity.*"}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per head milk production and greenhouse gas emissions intensity.*"}
 
@@
   labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)",
+       x="Change in per head milk yield (kg)",
        color="Intervention") +
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per capita milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per head milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}
 
@@
   labs(y="Relative change in emissions intensity (%)",
-       x="Relative change in per capita milk yield (%)",
+       x="Relative change in per head milk yield (%)",
        color="Intervention") +
@@
 
-metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
+metrics <- c("Milk yield per head", "Emissions intensity of milk production")
 names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")
@@
 </i></p>") %>%
-  pack_rows(index = c("Impact on milk yield per capita (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
+  pack_rows(index = c("Impact on milk yield per head (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
 column_spec(2:6, width = "3cm")
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
   labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
-       x="Change in per capita milk yield (kg)",
+       x="Change in per head milk yield (kg)",
        color="Intervention")  +
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
 At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
-  pack_rows(index = c("Impact on milk yield per capita (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
+  pack_rows(index = c("Impact on milk yield per head (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
 column_spec(2:6, width = "3cm")
@@
 confidence_level_analysis$confidence<-confidence_level_analysis$confidence*100
-toplot<-melt(confidence_level_analysis,id.vars = "confidence")
-toplot$variable <- factor(toplot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
+confidence_plot<-melt(confidence_level_analysis,id.vars = "confidence")
+confidence_plot$variable <- factor(confidence_plot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
                   labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
@@
 
-ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
+ggplot(confidence_plot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
   labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
@@
 
+# Narrowing key uncertainties
+
+The sensitivity analysis indicated that uncertainty about milk yields is a key driver of output variation, with unfavorable implications for using this metric as an indicator of emissions intensity.
+Published evidence suggests that, with some additional effort, this uncertainty can be reduced to a measurement error of approximately 15%.
+We can explore the impact of such measures on the suitability of milk yield per head as an indicator of emissions intensity.
+
+```{r, eval=FALSE}
+scen_upd<-scenarios
+
+new_error<-0.15
+
+mean_milk_4<-(as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="lower"),3:ncol(scen_upd)]) +
+  as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="upper"),3:ncol(scen_upd)])) / 2
+scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="lower"),3:ncol(scen_upd)]<-mean_milk_4*(1-new_error)
+scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="upper"),3:ncol(scen_upd)]<-mean_milk_4*(1+new_error)
+
+mean_milk_5<-(as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="lower"),3:ncol(scen_upd)]) +
+  as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="upper"),3:ncol(scen_upd)])) / 2
+scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="lower"),3:ncol(scen_upd)]<-mean_milk_5*(1-new_error)
+scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="upper"),3:ncol(scen_upd)]<-mean_milk_5*(1+new_error)
+
+mean_milk_6<-(as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="lower"),3:ncol(scen_upd)]) +
+  as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="upper"),3:ncol(scen_upd)])) / 2
+scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="lower"),3:ncol(scen_upd)]<-mean_milk_6*(1-new_error)
+scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="upper"),3:ncol(scen_upd)]<-mean_milk_6*(1+new_error)
+
+write.csv(scen_upd,"data/farm_scenarios_update.csv",row.names = FALSE)
+
+GHG_sim_update<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                                 scenarios = read.csv("data/farm_scenarios_update.csv",
+                                                      fileEncoding = "UTF-8-BOM"),
+                                 model_function = ghg_emissions,
+                                 numberOfModelRuns = 10,
+                                 functionSyntax = "plainNames")
+GHG_sim_update$y[,"CO2eq_per_milk"]<-GHG_sim_update$y$pc_on_farm/GHG_sim_update$y$pc_milk_yield
+dir.create("results")
+save_temperature_scenarios(GHG_sim_update,"results","Update_better_milk_yield_estimates")
+
+GHG_sim_update_endline<-scenario_mc(base_estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
+                                 scenarios = read.csv("data/farm_scenarios_update.csv",
+                                                      fileEncoding = "UTF-8-BOM"),
+                                 model_function = ghg_emissions,
+                                 numberOfModelRuns = 10,
+                                 functionSyntax = "plainNames")
+GHG_sim_update_endline$y[,"CO2eq_per_milk"]<-GHG_sim_update_endline$y$pc_on_farm/GHG_sim_update_endline$y$pc_milk_yield
+dir.create("results")
+save_temperature_scenarios(GHG_sim_update_endline,"results","Update__endline_better_milk_yield_estimates")
+
+```
+```{r}
+GHG_sim_update<-load_temperature_scenarios("results","Update_better_milk_yield_estimates")
+class(GHG_sim_update)<-"mcSimulation"
+GHG_sim_update_endline<-load_temperature_scenarios("results","Update__endline_better_milk_yield_estimates")
+class(GHG_sim_update_endline)<-"mcSimulation"
+
+```
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S7.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels, under the assumption that errors in milk yield estimates have been reduced to 15%.
+Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emissions intensity.
+This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}
+
+dairy_households_update<-GHG_sim_update
+dairy_households_update$x<-dairy_households_update$x[normal_milk_yield,]
+dairy_households_update$y<-dairy_households_update$y[normal_milk_yield,]
+
+milk_yields_to_sample<-seq(100,10000,100)
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample<max(dairy_households_update$y$pc_milk_yield))]
+milk_yields_to_sample<-milk_yields_to_sample[which(milk_yields_to_sample>min(dairy_households_update$y$pc_milk_yield))]
+slices<-list()
+
+# slice through the distribution at every 100 L milk yield per cow
+
+for(i in 1:length(milk_yields_to_sample))
+{slices[[i]]<-
+  varslice_resample(dairy_households_update$y$pc_milk_yield,
+                      dairy_households_update$y$CO2eq_per_milk,
+                      expectedin_var = milk_yields_to_sample[i],
+                      n_samples = 100000,
+  out_var_sampling = 1000)$resample
+ }
+
+decreased_emissions_probability_update <-
+  data.frame(Base_yield = NA,new_yield=NA,probability=NA)
+
+for (i in 2:(length(milk_yields_to_sample) - 1))
+{
+  for (j in 1:length(milk_yields_to_sample))
+    {decreased_emissions_probability_update[nrow(decreased_emissions_probability_update)+1,"Base_yield"]<-milk_yields_to_sample[i]
+     decreased_emissions_probability_update$new_yield[nrow(decreased_emissions_probability_update)]<-milk_yields_to_sample[j]
+     decreased_emissions_probability_update$probability[nrow(decreased_emissions_probability_update)] <-
+      sum(slices[[i]] > slices[[j]]) / length(slices[[i]])
+     decreased_emissions_probability_update$probability[nrow(decreased_emissions_probability_update)] <-
+      sum(slices[[i]] > sample(slices[[j]], length(slices[[i]]))) / length(slices[[i]])
+
+}
+}
+
+decreased_emissions_probability_update<-decreased_emissions_probability_update[which(!is.na(decreased_emissions_probability_update$Base_yield)),]
+
+decreased_emissions_probability_update[,"version"]<-"Precise milk yields"
+decreased_emissions_probability[,"version"]<-"Initial uncertainty"
+
+dec_emiss_update_effects<-rbind(decreased_emissions_probability[,c("Base_yield","new_yield","probability","version")],
+                                decreased_emissions_probability_update)
+
+dec_emis_prob_update<-ggplot(data = dec_emiss_update_effects, aes(Base_yield, new_yield, z =
+                                                      probability)) + geom_contour_filled() +
+  ggtitle("Probability of lower emissions intensity on farm B - impact on added precision for milk yield measurements") + 
+  labs(y=expression(Milk~yield~of~farm~B~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~farm~A~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity on farm B") +
+  stat_contour(breaks = c(0.9),col="black") + theme_bw() + facet_wrap(~version)
+
+dec_emis_prob_update
+
+ggsave("figures/Fig_S7_updated_probability_of_lower_emissions_intensity.png", width=10,height=5)
+
+```
+
+We need to implement the same interventions as for the initial simulation.
+
+```{r}
+## Produce endline breed scenario
+x_breed_intervention_update_endline <- GHG_sim_update_endline$x
+
+for (i in breed_adopters)
+{x_breed_intervention_update_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
+ x_breed_intervention_update_endline[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
+    breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
+                               x_breed_intervention_update_endline$Scenario[i]),
+                       2:ncol(breed_intervention)]
+}
+
+## Herd intervention 1 - nobulls
+
+x_nobull_intervention_update_endline <- GHG_sim_update_endline$x
+x_nobull_intervention_update_endline$N_animal_type_1[nobull_adopters]<-0
+
+## Herd intervention 2 - lessReplace
+
+x_lessReplace_intervention_update_endline <- GHG_sim_update_endline$x
+x_lessReplace_intervention_update_endline$N_animal_type_3[lessReplace_adopters]<-0
+
+## Feed intervention 1 - Supplementary legume fodder
+
+x_Calliandra_intervention_update_endline <- GHG_sim_update_endline$x
+for (i in Calliandra_adopters)
+{x_Calliandra_intervention_update_endline$milk_yield_5[i]<-x_Calliandra_intervention_update_endline$milk_yield_5[i]+0.156
+ x_Calliandra_intervention_update_endline$milk_yield_6[i]<-x_Calliandra_intervention_update_endline$milk_yield_6[i]+0.156
+ x_Calliandra_intervention_update_endline$DE_perc[i]<-x_Calliandra_intervention_update_endline$DE_perc[i]+0.14
+ x_Calliandra_intervention_update_endline$perc_crude_protein_diet[i]<-x_Calliandra_intervention_update_endline$perc_crude_protein_diet[i]-0.07
+ x_Calliandra_intervention_update_endline$feed_prod_CO2[i]<-x_Calliandra_intervention_update_endline$feed_prod_CO2[i]- (0.726 *
+                                        sum(x_Calliandra_intervention_update_endline[i,paste0("N_animal_type_",1:12)]))
+}
+
+## Feed intervention 2 - Balanced diets
+
+x_balance_intervention_update_endline <- GHG_sim_update_endline$x
+for (i in balance_adopters)
+{x_balance_intervention_update_endline$milk_yield_5[i]<-x_balance_intervention_update_endline$milk_yield_5[i]+0.75
+ x_balance_intervention_update_endline$milk_yield_6[i]<-x_balance_intervention_update_endline$milk_yield_6[i]+0.75
+ x_balance_intervention_update_endline$feed_prod_CO2[i]<-x_balance_intervention_update_endline$feed_prod_CO2[i]-(328.5 * 
+                    sum(x_balance_intervention_update_endline[i,paste0("N_animal_type_",1:12)]))
+}
+
+```
+
+```{r, eval=FALSE}
+# run model for modified x file (endline)
+breed_end_update <- run_model(x_breed_intervention_update_endline, ghg_emissions)
+nobull_end_update <- run_model(x_nobull_intervention_update_endline, ghg_emissions)
+lessReplace_end_update <- run_model(x_lessReplace_intervention_update_endline, ghg_emissions)
+Calliandra_end_update <- run_model(x_Calliandra_intervention_update_endline, ghg_emissions)
+balance_end_update <- run_model(x_balance_intervention_update_endline, ghg_emissions)
+
+breed_end_update$y[,"CO2eq_per_milk"]<-breed_end_update$y$pc_on_farm/breed_end_update$y$pc_milk_yield
+nobull_end_update$y[,"CO2eq_per_milk"]<-nobull_end_update$y$pc_on_farm/nobull_end_update$y$pc_milk_yield
+lessReplace_end_update$y[,"CO2eq_per_milk"]<-lessReplace_end_update$y$pc_on_farm/lessReplace_end_update$y$pc_milk_yield
+Calliandra_end_update$y[,"CO2eq_per_milk"]<-Calliandra_end_update$y$pc_on_farm/Calliandra_end_update$y$pc_milk_yield
+balance_end_update$y[,"CO2eq_per_milk"]<-balance_end_update$y$pc_on_farm/balance_end_update$y$pc_milk_yield
+
+breed_end_update$call<-"multi-scenario call"
+save_temperature_scenarios(breed_end_update,"results","End_update_Breed_intervention")
+nobull_end_update$call<-"multi-scenario call"
+save_temperature_scenarios(nobull_end_update,"results","End_update_Nobull_intervention")
+lessReplace_end_update$call<-"multi-scenario call"
+save_temperature_scenarios(lessReplace_end_update,"results","End_update_LessReplace_intervention")
+Calliandra_end_update$call<-"multi-scenario call"
+save_temperature_scenarios(Calliandra_end_update,"results","End_update_Calliandra_intervention")
+balance_end_update$call<-"multi-scenario call"
+save_temperature_scenarios(balance_end_update,"results","End_update_Balance_intervention")
+```
+
+```{r, include=FALSE}
+breed_end_update<-load_temperature_scenarios("results","End_update_Breed_intervention")
+class(breed_end_update)<-"mcSimulation"
+nobull_end_update<-load_temperature_scenarios("results","End_update_Nobull_intervention")
+class(nobull_end_update)<-"mcSimulation"
+lessReplace_end_update<-load_temperature_scenarios("results","End_update_LessReplace_intervention")
+class(lessReplace_end_update)<-"mcSimulation"
+Calliandra_end_update<-load_temperature_scenarios("results","End_update_Calliandra_intervention")
+class(Calliandra_end_update)<-"mcSimulation"
+balance_end_update<-load_temperature_scenarios("results","End_update_Balance_intervention")
+class(balance_end_update)<-"mcSimulation"
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S8.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
+
+# make data.frame with all intervention results
+
+interventions_end_update<-rbind(cbind(breed_end_update$y[breed_adopters,],
+                           intervention="breed",
+                           pc_milk_baseline=GHG_sim_update$y$pc_milk_yield[breed_adopters],
+                           CO2_intensity_baseline=GHG_sim_update$y$CO2eq_per_milk[breed_adopters]),
+                     cbind(nobull_end_update$y[nobull_adopters,],
+                           intervention="nobull",
+                           pc_milk_baseline=GHG_sim_update$y$pc_milk_yield[nobull_adopters],
+                           CO2_intensity_baseline=GHG_sim_update$y$CO2eq_per_milk[nobull_adopters]),
+                     cbind(lessReplace_end_update$y[lessReplace_adopters,],
+                           intervention="lessReplace",
+                           pc_milk_baseline=GHG_sim_update$y$pc_milk_yield[lessReplace_adopters],
+                           CO2_intensity_baseline=GHG_sim_update$y$CO2eq_per_milk[lessReplace_adopters]),
+                     cbind(Calliandra_end_update$y[Calliandra_adopters,],
+                           intervention="Calliandra",
+                           pc_milk_baseline=GHG_sim_update$y$pc_milk_yield[Calliandra_adopters],
+                           CO2_intensity_baseline=GHG_sim_update$y$CO2eq_per_milk[Calliandra_adopters]),
+                     cbind(balance_end_update$y[balance_adopters,],
+                           intervention="balance",
+                           pc_milk_baseline=GHG_sim_update$y$pc_milk_yield[balance_adopters],
+                           CO2_intensity_baseline=GHG_sim_update$y$CO2eq_per_milk[balance_adopters]))
+
+normal_end_update<-cbind(GHG_sim_update_endline$y,
+                           pc_milk_baseline=GHG_sim_update$y$pc_milk_yield,
+                           CO2_intensity_baseline=GHG_sim_update$y$CO2eq_per_milk)
+normal_end_update[,"CO2eq_per_milk"]<-normal_end_update$pc_on_farm/normal_end_update$pc_milk_yield
+
+interventions_end_update[,"milk_change"]<-interventions_end_update$pc_milk_yield-interventions_end_update$pc_milk_baseline
+interventions_end_update[,"emissions_intensity_change"]<-interventions_end_update$CO2eq_per_milk-interventions_end_update$CO2_intensity_baseline
+interventions_end_update[,"milk_change_relative"]<-(interventions_end_update$pc_milk_yield-interventions_end_update$pc_milk_baseline)/interventions_end_update$pc_milk_baseline*100
+interventions_end_update[,"emissions_intensity_change_relative"]<-
+  (interventions_end_update$CO2eq_per_milk-interventions_end_update$CO2_intensity_baseline)/interventions_end_update$CO2_intensity_baseline*100
+
+normal_end_update[,"milk_change"]<-normal_end_update$pc_milk_yield-normal_end_update$pc_milk_baseline
+normal_end_update[,"emissions_intensity_change"]<-normal_end_update$CO2eq_per_milk-normal_end_update$CO2_intensity_baseline
+normal_end_update[,"milk_change_relative"]<-(normal_end_update$pc_milk_yield-normal_end_update$pc_milk_baseline)/normal_end_update$pc_milk_baseline*100
+normal_end_update[,"emissions_intensity_change_relative"]<-
+  (normal_end_update$CO2eq_per_milk-normal_end_update$CO2_intensity_baseline)/normal_end_update$CO2_intensity_baseline*100
+
+interventions_normal_milk_end_update<-interventions_end_update[which(interventions_end_update$pc_milk_baseline>500),]
+normal_normal_milk_end_update<-normal_end_update[which(normal_end_update$pc_milk_yield>500),]
+normal_plot_update<-rbind(cbind(normal_normal_milk_end_update,intervention="breed"),
+                   cbind(normal_normal_milk_end_update,intervention="nobull"),
+                   cbind(normal_normal_milk_end_update,intervention="lessReplace"),
+                   cbind(normal_normal_milk_end_update,intervention="Calliandra"),
+                   cbind(normal_normal_milk_end_update,intervention="balance"))
+
+intervention_colors<-palette.colors(n = 6, palette = "Okabe-Ito", alpha=1, recycle = FALSE)
+
+normal_plot_update[,"version"]<-"Precise milk yields"
+normal_plot[,"version"]<-"Initial uncertainty"
+
+normal_stuff<-rbind(normal_plot,normal_plot_update)
+
+interventions_normal_milk_end_update[,"version"]<-"Precise milk yields"
+interventions_normal_milk_end[,"version"]<-"Initial uncertainty"
+
+interventions_normal_stuff<-rbind(interventions_normal_milk_end,interventions_normal_milk_end_update)
+
+interventions_scatter_end_update<-ggplot(data=normal_stuff,
+                        aes(x=milk_change,y=emissions_intensity_change)) +
+  geom_point(color="gray",size=1) +
+  geom_point(data=interventions_normal_stuff,
+                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
+  scale_color_manual(values= c("breed" = intervention_colors[1],
+                              "nobull" = intervention_colors[2],
+                              "lessReplace" = intervention_colors[3],
+                              "Calliandra" = intervention_colors[4],
+                              "balance" = intervention_colors[6]),
+                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+  theme_bw() +
+  ggtitle("Intervention impact on milk yield and emissions") + 
+  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+       x="Change in per head milk yield (kg)",
+       color="Intervention")  +
+  facet_grid(cols=vars(version),
+             rows=vars(factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet")))) +
+  theme(legend.position = "none")
+
+interventions_scatter_end_update
+
+ggsave("figures/Fig_S8_intervention_impacts_endline.png", width=9,height=12)
+
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S9.
+*** *Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
+the initial sampling time, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
+The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity.*"}
+
+interventions_plot_update<-interventions_normal_milk_end_update
+none_plot_update<-cbind(normal_normal_milk_end_update ,intervention="none", version="Precise milk yields")
+
+int_none_plot_update<-rbind(interventions_plot_update,none_plot_update)
+
+decreased_emissions_probability_update[,"version"]<-"Precise milk yields"
+decreased_emissions_probability[,"version"]<-"Initial uncertainty"
+decreased_emissions_probability_stuff<-rbind(decreased_emissions_probability[,c("Base_yield","new_yield","probability","version")],
+                                             decreased_emissions_probability_update)
+
+int_none_plot_update[,"version"]<-"Precise milk yields"
+int_none_plot[,"version"]<-"Initial uncertainty"
+int_none_plot_stuff<-rbind(int_none_plot[,which(!colnames(int_none_plot)=="confidence_level")],int_none_plot_update)
+
+interventions_probability_update<-ggplot(data = decreased_emissions_probability_stuff, aes(Base_yield, new_yield)) +
+  geom_contour_filled(aes(z=probability)) +
+  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  labs(y=expression(Milk~yield~at~endline~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~at~baseline~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity") +
+  stat_contour(aes(z=probability),breaks = c(0.9),col="black") + theme_bw() +
+  geom_point(data=int_none_plot_stuff,aes(x=pc_milk_baseline,y=pc_milk_yield), size=0.8, color="coral1") +
+  ylim(c(0,NA)) + xlim(c(0,NA)) +
+  facet_grid(cols=vars(version),
+             rows=vars(factor(intervention,
+                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
+                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))))
+
+interventions_probability_update
+
+ggsave("figures/Fig_S9_probability_of_lower_emissions_intensity_interventions.png", width=7,height=12)
+
+```
+
+```{r, fig.dim = c(8, 12), fig.cap="***Figure S10.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
+Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
+
+# get closest probability value
+
+for(i in 1:nrow(int_none_plot_update))
+  {decreased_emissions_probability_update[,"diff"]<-
+    abs(decreased_emissions_probability_update$Base_yield-int_none_plot_update$pc_milk_baseline [i]) +
+    abs(decreased_emissions_probability_update$new_yield-int_none_plot_update$pc_milk_yield[i])
+  int_none_plot_update[i,"confidence_level"]<-decreased_emissions_probability_update$probability[which(decreased_emissions_probability_update$diff==min(decreased_emissions_probability_update$diff))][1]
+}
+
+confidence_level_analysis_update<-data.frame(confidence=seq(0.25,1,0.01))
+
+for(i in 1:nrow(confidence_level_analysis_update))
+{# compute Type 1 errors for all interventions
+  for(ints in c("breed", "nobull", "lessReplace", "Calliandra", "balance"))
+    confidence_level_analysis_update[i,ints]<-
+      length(which(int_none_plot_update$intervention==ints &
+                     int_none_plot_update$confidence_level <
+                     confidence_level_analysis_update$confidence[i])) /
+      length(which(int_none_plot_update$intervention==ints)) * 100
+
+  confidence_level_analysis_update$Type2error[i]<-
+    length(which(int_none_plot_update$intervention=="none" &
+                   int_none_plot_update$confidence_level>confidence_level_analysis_update$confidence[i])) /
+    length(which(int_none_plot_update$intervention=="none")) * 100
+
+}
+
+confidence_level_analysis_update$confidence<-confidence_level_analysis_update$confidence*100
+confidence_plot_update<-melt(confidence_level_analysis_update,id.vars = "confidence")
+confidence_plot_update$variable <- factor(confidence_plot_update$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
+                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
+                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
+                             "Type I error - failure to reward 'Fewer replacement males' adopters",
+                             "Type I error - failure to reward 'Calliandra feed' adopters",
+                             "Type I error - failure to reward 'Balanced diet' adopters",
+                             "Type II error - rewarding non-adopters"))
+
+confidence_plot_update[,"version"]<-"Precise milk yields"
+confidence_plot[,"version"]<-"Initial uncertainty"
+confidence_plot_stuff<-rbind(confidence_plot,confidence_plot_update)
+confidence_plot_stuff$variable <- factor(confidence_plot_stuff$variable,
+                                         levels = c("Type I error - failure to reward 'Breed upgrade' adopters",
+                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
+                             "Type I error - failure to reward 'Fewer replacement males' adopters",
+                             "Type I error - failure to reward 'Calliandra feed' adopters",
+                             "Type I error - failure to reward 'Balanced diet' adopters",
+                             "Type II error - rewarding non-adopters"),
+                  labels = c("Type I - 'Breed upgrade'",
+                             "Type I - 'Unproductive bulls'",
+                             "Type I - 'Replacement males'",
+                             "Type I - 'Calliandra feed'",
+                             "Type I - 'Balanced diet'",
+                             "Type II "))
+
+ggplot(confidence_plot_stuff, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + theme_bw() +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")+
+  facet_grid(cols=vars(version),
+             rows=vars(variable))
+
+ggsave("figures/Fig_S10_erorr_types.png", width=7,height=12)
+
+```
+
+new sensitivity analysis
+
+```{r, fig.dim = c(12, 4), fig.cap="***Figure S11.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
+This analysis is based on a*"}
+
+pls_GHG<-dairy_households
+pls_GHG$x<-pls_GHG$x[,which(!colnames(pls_GHG$x)=="Scenario")]
+pls_out<-plsr.mcSimulation(object=pls_GHG, resultName = "CO2eq_per_milk")
+legend_table<-read.csv("data/legend.csv", fileEncoding = "UTF-8-BOM")
+PLS_initial<-plot_pls(pls_out,input_table=legend_table) + ggtitle("Initial uncertainty")
+
+pls_GHG_update<-dairy_households_update
+pls_GHG_update$x<-pls_GHG_update$x[,which(!colnames(pls_GHG_update$x)=="Scenario")]
+pls_out_update<-plsr.mcSimulation(object=pls_GHG_update, resultName = "CO2eq_per_milk")
+
+PLS_update<-plot_pls(pls_out_update,input_table=legend_table) + ggtitle("Precise milk yields")
+
+PLS_initial + PLS_update
+
+ggsave("figures/Fig_S11_PLS_dairy_households_update.png", width=12,height=4)
+```
+
 # References

### 2022-06-19 Manuscript_figures.Rmd
another update

@@
 ggsave("figures/Fig_3_scatter_and_surface_per_head_milk_emissions_intensity.png",width=10,height=6)
+
+```
+
+Extracting the LOESS regression from the plot and empirically determining minimum milk yield that would be outside the confidence interval
+
+```{r, warning=FALSE}
+
+prediction_data<-ggplot_build(plot1)$data[[2]]
+
+valuecrowd<-approx(prediction_data$x, y = prediction_data$ymax, xout= seq(100,10000,10), method="linear")
+
+for(i in 1:nrow(prediction_data))
+{
+  closest<-min(which(valuecrowd$x>prediction_data$x[i] & valuecrowd$y<prediction_data$ymin[i]))
+  prediction_data[i,"min_more_milk_loess"]<-valuecrowd$x[closest]
+}
+
 ```
@@
 
-ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield, z =
-                                                      probability)) + geom_contour_filled() +
+ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield)) + geom_contour_filled(aes(z =
+                                                      probability)) +
   ggtitle("Probability of lower emissions intensity on farm B") + 
@@
        fill="Probability of \nlower emissions \nintensity on farm B") +
-  stat_contour(breaks = c(0.9),col="black") + theme_bw()
+  stat_contour(aes(z = probability),breaks = c(0.9),col="black") + 
+  theme_bw() +
+  scale_x_continuous(limits=c(min(decreased_emissions_probability$Base_yield),NA),expand = c(0, 0)) +
+  scale_y_continuous(limits=c(min(decreased_emissions_probability$new_yield),NA),expand = c(0, 0))
 
@@
 
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
+5 for more details).
+The red line indicates a confidence threshold that is based on the 95% confidence interval of the loess regression line shown in Fig.
+3a.
+Using this threshold for predicting changes in emissions intensity based on milk yield data would likely lead to a lot of inaccurate classifications.*"}
+
+ggplot(data = decreased_emissions_probability, aes(Base_yield, new_yield)) + geom_contour_filled(aes(z =
+                                                      probability)) +
+  ggtitle("Probability of lower emissions intensity on farm B") + 
+  labs(y=expression(Milk~yield~of~farm~B~(kg~head^{-1}~year^{-1})),
+       x=expression(Milk~yield~of~farm~A~(kg~head^{-1}~year^{-1})),
+       fill="Probability of \nlower emissions \nintensity on farm B") +
+  stat_contour(aes(z = probability),breaks = c(0.9),col="black") + 
+  geom_smooth(data=prediction_data, aes(x=x,y=min_more_milk_loess),col="red")  +  geom_abline(slope=1,intercept=0,linetype = "dashed")  +
+  theme_bw() +
+  scale_x_continuous(limits=c(min(decreased_emissions_probability$Base_yield),NA),expand = c(0, 0)) +
+  scale_y_continuous(limits=c(min(decreased_emissions_probability$new_yield),NA),expand = c(0, 0))
+
+ggsave("figures/Fig_S2_probability_of_lower_emissions_intensity.png", width=7,height=5)
+
+```
+
 Now a description of the 5 interventions
 
-```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Observed distributions of milk yields for all cattle breeds encountered among dairy cows in a survey of livestock-keeping smallholder farms in Kenya.*"}
+```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S3.*** *Observed distributions of milk yields for all cattle breeds encountered among dairy cows in a survey of livestock-keeping smallholder farms in Kenya.*"}
 # breed intervention
@@
 
-ggsave("figures/Fig_S2_breeds.png", width=9,height=4)
+ggsave("figures/Fig_S3_breeds.png", width=9,height=4)
 
@@
 
-```{r}
+```{r, include=FALSE}
 breed_res<-load_temperature_scenarios("results","Breed_intervention")
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per head milk production and greenhouse gas emissions intensity.*"}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S4.*** *Intervention impact on per head milk production and greenhouse gas emissions intensity.*"}
 
@@
                               "balance" = intervention_colors[6]),
-                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+                    labels = c("Breed change", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
   facet_wrap(~factor(intervention,
                      levels=c("breed","nobull","lessReplace","Calliandra","balance"),
-                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
+                     labels=c("Breed change", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
   theme(legend.position = "none")
@@
 
-ggsave("figures/Fig_6a_intervention_impacts.png", width=9,height=6)
+ggsave("figures/Fig_S4_intervention_impacts.png", width=9,height=6)
 
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S3.*** *Influence of the breed composition of farm-scale dairy cow populations on changes in milk yield and emissions intensity of milk production in response to a breed upgrade intervention*."}
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S5.*** *Influence of the breed composition of farm-scale dairy cow populations on changes in milk yield and emissions intensity of milk production in response to a breed change intervention*."}
 
@@
 
-ggsave("figures/Fig_S3_breed_diagnostics.png",width=8,height=5)
-
-```
-
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S4.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}
+ggsave("figures/Fig_S5_breed_diagnostics.png",width=8,height=5)
+
+```
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S6.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}
 
@@
 
-ggsave("figures/Fig_S4_nobull_diagnostics.png",width=8,height=5)
-
-```
-
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S5.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'fewer replacement males' intervention*."}
+ggsave("figures/Fig_S6_nobull_diagnostics.png",width=8,height=5)
+
+```
+
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S7.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'fewer replacement males' intervention*."}
 
@@
 
-ggsave("figures/Fig_S5_lessReplace_diagnostics.png",width=8,height=5)
+ggsave("figures/Fig_S7_lessReplace_diagnostics.png",width=8,height=5)
 
@@
 
-interventions_summary_milk<-data.frame(Intervention=c("Breed upgrade",
+interventions_summary_milk<-data.frame(Intervention=c("Breed change",
                                                  "Retiring unproductive bulls",
@@
 
-```{r}
+```{r, include=FALSE}
 
@@
 class(GHG_simulation_scenarios_time2)<-"mcSimulation"
+
+```
+
+```{r}
 
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S8.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
                               "balance" = intervention_colors[6]),
-                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
+                    labels = c("Breed change", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
   theme_bw() +
@@
                      levels=c("breed","nobull","lessReplace","Calliandra","balance"),
-                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
+                     labels=c("Breed change", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
   theme(legend.position = "none")
@@
 
-ggsave("figures/Fig_7a_intervention_impacts_endline.png", width=9,height=6)
+ggsave("figures/Fig_S8_intervention_impacts_endline.png", width=9,height=6)
 
@@
 
-interventions_end_summary_milk<-data.frame(Intervention=c("Breed upgrade",
+interventions_end_summary_milk<-data.frame(Intervention=c("Breed change",
                                                  "Retiring unproductive bulls",
@@
                      levels=c("breed","nobull","lessReplace","Calliandra","balance","none"),
-                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet", "No intervention"))  )
+                     labels=c("Breed change", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet", "No intervention"))  ) +
+  scale_x_continuous(limits=c(min(decreased_emissions_probability$Base_yield),max(decreased_emissions_probability$Base_yield)),expand = c(0, 0)) +
+  scale_y_continuous(limits=c(min(decreased_emissions_probability$new_yield),max(decreased_emissions_probability$new_yield)),expand = c(0, 0))
 
@@
 confidence_plot$variable <- factor(confidence_plot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
-                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
+                  labels = c("Type I error - failure to reward 'Breed change' adopters",
                              "Type I error - failure to reward 'Retire unproductive bulls' adopters",
@@
 GHG_sim_update$y[,"CO2eq_per_milk"]<-GHG_sim_update$y$pc_on_farm/GHG_sim_update$y$pc_milk_yield
-dir.create("results")
 save_temperature_scenarios(GHG_sim_update,"results","Update_better_milk_yield_estimates")
@@
 GHG_sim_update_endline$y[,"CO2eq_per_milk"]<-GHG_sim_update_endline$y$pc_on_farm/GHG_sim_update_endline$y$pc_milk_yield
-dir.create("results")
-save_temperature_scenarios(GHG_sim_update_endline,"results","Update__endline_better_milk_yield_estimates")
-
-```
-```{r}
+save_temperature_scenarios(GHG_sim_update_endline,"results","Update_endline_better_milk_yield_estimates")
+
+```
+```{r, include=FALSE}
 GHG_sim_update<-load_temperature_scenarios("results","Update_better_milk_yield_estimates")
 class(GHG_sim_update)<-"mcSimulation"
-GHG_sim_update_endline<-load_temperature_scenarios("results","Update__endline_better_milk_yield_estimates")
+GHG_sim_update_endline<-load_temperature_scenarios("results","Update_endline_better_milk_yield_estimates")
 class(GHG_sim_update_endline)<-"mcSimulation"
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S7.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels, under the assumption that errors in milk yield estimates have been reduced to 15%.
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S9.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels, under the assumption that errors in milk yield estimates have been reduced to 15%.
 Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emissions intensity.
@@
        fill="Probability of \nlower emissions \nintensity on farm B") +
-  stat_contour(breaks = c(0.9),col="black") + theme_bw() + facet_wrap(~version)
+  stat_contour(breaks = c(0.9),col="black") + theme_bw() + facet_wrap(~version) +
+  scale_x_continuous(limits=c(min(dec_emiss_update_effects$Base_yield),max(dec_emiss_update_effects$Base_yield)),expand = c(0, 0)) +
+  scale_y_continuous(limits=c(min(dec_emiss_update_effects$new_yield),max(dec_emiss_update_effects$new_yield)),expand = c(0, 0))
 
@@
 
-ggsave("figures/Fig_S7_updated_probability_of_lower_emissions_intensity.png", width=10,height=5)
+ggsave("figures/Fig_S9_updated_probability_of_lower_emissions_intensity.png", width=10,height=5)
 
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S8.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S10.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
                      levels=c("breed","nobull","lessReplace","Calliandra","balance"),
-                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet")))) +
+                     labels=c("Breed change", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet")))) +
   theme(legend.position = "none")
@@
 
-ggsave("figures/Fig_S8_intervention_impacts_endline.png", width=9,height=12)
-
-```
-
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S9.
+ggsave("figures/Fig_S10_intervention_impacts_endline.png", width=9,height=12)
+
+```
+
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S11.
 *** *Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
@@
                      levels=c("breed","nobull","lessReplace","Calliandra","balance"),
-                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))))
+                     labels=c("Breed change", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))))
 
@@
 
-ggsave("figures/Fig_S9_probability_of_lower_emissions_intensity_interventions.png", width=7,height=12)
-
-```
-
-```{r, fig.dim = c(8, 12), fig.cap="***Figure S10.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
+ggsave("figures/Fig_S11_probability_of_lower_emissions_intensity_interventions.png", width=7,height=12)
+
+```
+
+```{r, fig.dim = c(8, 12), fig.cap="***Figure S12.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
 Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
@@
 confidence_plot_update$variable <- factor(confidence_plot_update$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
-                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
+                  labels = c("Type I error - failure to reward 'Breed change' adopters",
                              "Type I error - failure to reward 'Retire unproductive bulls' adopters",
@@
 confidence_plot_stuff$variable <- factor(confidence_plot_stuff$variable,
-                                         levels = c("Type I error - failure to reward 'Breed upgrade' adopters",
+                                         levels = c("Type I error - failure to reward 'Breed change' adopters",
                              "Type I error - failure to reward 'Retire unproductive bulls' adopters",
@@
                              "Type II error - rewarding non-adopters"),
-                  labels = c("Type I - 'Breed upgrade'",
+                  labels = c("Type I - 'Breed change'",
                              "Type I - 'Unproductive bulls'",
@@
                              "Type I - 'Balanced diet'",
-                             "Type II "))
+                             "Type II"))
 
@@
 
-ggsave("figures/Fig_S10_erorr_types.png", width=7,height=12)
+ggsave("figures/Fig_S12_erorr_types.png", width=7,height=12)
+
+```
+
+Comparing the errors is easier in a line plot.
+
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 11.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
+Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
+
+ggplot(confidence_plot_stuff, aes(x=confidence,y=value, color=variable)) + geom_smooth(stat="identity", fill="darkolivegreen") +
+  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)",
+       color="Classification errors") +
+  facet_wrap(~version,nrow=2) +
+  scale_color_manual(values= c("Type I - 'Breed change'" = intervention_colors[1],
+                              "Type I - 'Unproductive bulls'" = intervention_colors[2],
+                              "Type I - 'Replacement males'" = intervention_colors[3],
+                              "Type I - 'Calliandra feed'" = intervention_colors[4],
+                              "Type I - 'Balanced diet'" = intervention_colors[6],
+                              "Type II" = "blue"),
+                    labels = c("Breed change", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet", "Type II error")) + theme_bw()
+
+ggsave("figures/Fig_11_erorr_types.png", width=7,height=7)
 
@@
 
-```{r, fig.dim = c(12, 4), fig.cap="***Figure S11.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
+```{r, fig.dim = c(12, 4), fig.cap="***Figure S13.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
 This analysis is based on a*"}
@@
 
-ggsave("figures/Fig_S11_PLS_dairy_households_update.png", width=12,height=4)
+ggsave("figures/Fig_S13_PLS_dairy_households_update.png", width=12,height=4)
 ```

### 2022-06-20 Manuscript_figures.Rmd
first (almost) complete draft

@@
 library(patchwork)
+library(DT)
 
@@
 options(scipen=999)
+options(DT.options = list(scrollY="300px",scrollX="300px", pageLength = 100, autoWidth = TRUE))
+
 ```
@@
 library(patchwork)
+library(DT)
 ```
@@
 Here, we express this uncertainty by defining distributions for each variable.
-The following table (**data/livestock_ghg_input_table.csv**) shows all the variables, as well as examples of distributions that may describe their quantities.
+The following table (Table S1) shows all the variables, as well as examples of distributions that may describe their quantities.
 Distributions are defined according to the distribution shape, as well as upper and lower bounds of 90% confidence intervals.
@@
 * posnorm - positive normal distribution; this is a normal distribution that is truncated at 0.
+
+***Table S1*** *Model input parameters for the Monte Carlo simulation, including the variable names used in the greenhouse gas emissions equation, a descriptive label, lower and upper bounds of 90% confidence intervals for the values, and the shape of the probability distribution that is used to describe the data.
+The IPCC_default column shows values used for producing (spuriously) the precise scenario.
+Note that some values are overwritten when producing farm-scale scenarios.
+Data can be downloaded in csv or Excel format or copied to the clipboard by pressing the respective buttons.*
+
+```{r}
+
+input_parameters<-read.csv("data/livestock_ghg_input_table.csv")[,c(7,2,3,5,6,8)]
+input_parameters<-input_parameters[which(!is.na(input_parameters$lower)),]
+
+input_parameters %>%
+  datatable(extensions = 'Buttons',
+            options = list(dom = 'Blfrtip',
+                           buttons = c('copy', 'csv', 'excel'),
+                           lengthMenu = list(c(10,25,50,-1),
+                                             c(10,25,50,"All")),
+                           list(scrollX = "300px"),
+                           autoWidth = TRUE),
+            rownames=FALSE)
+```
 
@@
 
-Here's an example `scenarios` file:
-
-```{r}
-read.table("data/scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)   %>%
-    kableExtra::kbl(digits = 2)  %>%
-    kableExtra::kable_classic_2(full_width = F)
-
-```
-
-The first row ("Runs" in the first column) indicates the number of runs.
-Each subsequent row indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table.
+Each row in this table indicates which value from the input table should be modified, so the string in the first column has to correspond to a variable name in the input table (additionally, the number of runs per scenario can be specified by a row with 'Runs' in the 'Variable' column, but this was not done here).
 The second column indicates which values should be changed, with options being "lower", "upper", "distribution" and "both".
@@
 
-read.table("data/farm_scenarios.csv", sep = ",", fileEncoding = "UTF-8-BOM",header=TRUE)[1:20,1:5]  %>%
-    kableExtra::kbl(digits = 2)  %>%
-    kableExtra::kable_classic_2(full_width = F)
-
+```
+
+***Table S2*** *Model input parameters for producing farm-scale scenarios for the Monte Carlo simulation, listing all variables that need to be updated.
+Data can be downloaded in csv or Excel format or copied to the clipboard by pressing the respective buttons.*
+
+```{r}
+
+scenario_parameters<-read.csv("data/farm_scenarios.csv")
+
+scenario_parameters %>%
+  datatable(extensions = 'Buttons',
+            options = list(dom = 'Blfrtip',
+                           buttons = c('copy', 'csv', 'excel'),
+                           lengthMenu = list(c(10,25,50,-1),
+                                             c(10,25,50,"All")),
+                           list(scrollX = "300px")),
+            rownames=FALSE)
 ```
@@
 Now we can run the Monte Carlo simulation with these scenarios.
+
+The basic parameters for the simulations are shown in Table S1. Note that some of these parameters are overwritten when producing farm-scale scenarios.
+
+Table S2 contains the scenario data derived from results of the farm survey.
+In running the Monte Carlo with the `scenario_mc` function, the initial values for all variables listed in this table (Table S1) are overwritten by those shown here (in Table S2) to generate farm-specific scenarios.
 
@@
                     paste0("CO","$_2$","-eq kg","$^{-1}$")),
-      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
+      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
   index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
+```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
 5 for more details).
-The red line indicates a confidence threshold that is based on the 95% confidence interval of the loess regression line shown in Fig.
+The red line indicates a confidence threshold that is based on the 95% confidence interval of the LOESS regression line shown in Fig.
 3a.
@@
 kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
-      caption = "<p style='color:black'><i><b>Table S2.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
+      caption = "<p style='color:black'><i><b>Table S4.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
 column_spec(2:4, width = "5cm")
@@
          #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
-      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
+      caption = "<p style='color:black'><i><b>Table S5.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
 Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.
@@
 kable(intervention_end_summary, booktabs = TRUE,digits=0,
-      caption = "<p style='color:black'><i><b>Table S4.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
+      caption = "<p style='color:black'><i><b>Table S6.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
 Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures.

### 2022-06-22 Manuscript_figures.Rmd
mainly emissions intensity -> emission intensity

@@
 
-```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "***Figure 1.*** *Estimated total emissions from livestock (a) and emissions intensity of milk production (b) for a smallholder dairy farm in Kenya.
+```{r, warning=FALSE, message=FALSE,fig.dim = c(8, 4), fig.cap = "***Figure 1.*** *Estimated total emissions from livestock (a) and emission intensity of milk production (b) for a smallholder dairy farm in Kenya.
 Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.
@@
   ylab("Relative probability") +
-    xlab(expression(Emissions~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emissions intensity")
+    xlab(expression(Emission~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emission intensity")
 
@@
 
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livestock (a) and emissions intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms).
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livestock (a) and emission intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms).
 Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
@@
   ylab("Relative probability") + theme_bw() +
-    xlab(expression(Emissions~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emissions intensity")
+    xlab(expression(Emission~intensity~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("b) Emission intensity")
 
@@
 We can now look at summary statistics of total and per head emissions from cattle across livestock-keeping smallholder farms in Kenya.
-To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per head milk yields of >500 kg of milk.
+To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emission intensity, we also look at summary statistics across only the farms with per head milk yields of >500 kg of milk.
 
@@
                                                                     c(0,0.05,0.5,0.95,1))/1000,1),
-                                    "Emissions intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
+                                    "Emission intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
                                                                    c(0,0.05,0.5,0.95,1)),1)),
@@
                                                                     c(0,0.05,0.5,0.95,1))/1000,1),
-                                    "Emissions intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
+                                    "Emission intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
                                                                    c(0,0.05,0.5,0.95,1)),1)))
@@
 column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
-  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per head" = 1, "Emissions intensity" = 1))
+  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per head" = 1, "Emission intensity" = 1))
 
@@
 
-We're interested in the suitability of total per head milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.
+We're interested in the suitability of total per head milk production per year as an indicator of the greenhouse gas emission intensity of milk production.
 
@@
 
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per head (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b).
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per head (x-axis) and emission intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b).
 The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig.
@@
   geom_rug(color="lightskyblue3") + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+  labs(y=expression(Emission~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
        x=expression(Milk~yield~per~head~(kg~head^{-1}~year^{-1})))+
@@
   scale_fill_viridis_c() + 
-  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
+  labs(y=expression(Emission~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
        x=expression(Milk~yield~per~head~(kg~head^{-1}~year^{-1})),
@@
 
-ggsave("figures/Fig_3_scatter_and_surface_per_head_milk_emissions_intensity.png",width=10,height=6)
+ggsave("figures/Fig_3_scatter_and_surface_per_head_milk_emission_intensity.png",width=10,height=6)
 
@@
 
-According to these distributions, the emissions intensity at a particular per head milk yield can be described by a slice through the probability surface.
-
-```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-head milk production among livestock-keeping smallholder farms in Kenya.
+According to these distributions, the emission intensity at a particular per head milk yield can be described by a slice through the probability surface.
+
+```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emission intensity of milk production at specific levels of per-head milk production among livestock-keeping smallholder farms in Kenya.
 Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
@@
   ylab("Relative probability") +
-    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~head~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")
+    xlab(expression(Emission~intensity~at~2000~kg~milk~per~head~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emission intensity distribution")
 
@@
   ylab("Relative probability") +
-    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~head~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")
+    xlab(expression(Emission~intensity~at~specific~milk~yield~per~head~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")
 
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels.
-Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emissions intensity.
+```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels.
+Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emission intensity.
 This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}
@@
                                                       probability)) +
-  ggtitle("Probability of lower emissions intensity on farm B") + 
+  ggtitle("Probability of lower emission intensity on farm B") + 
   labs(y=expression(Milk~yield~of~farm~B~(kg~head^{-1}~year^{-1})),
        x=expression(Milk~yield~of~farm~A~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity on farm B") +
+       fill="Probability of \nlower emission \nintensity on farm B") +
   stat_contour(aes(z = probability),breaks = c(0.9),col="black") + 
@@
 
-ggsave("figures/Fig_5_probability_of_lower_emissions_intensity.png", width=7,height=5)
-
-```
-
-```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
+ggsave("figures/Fig_5_probability_of_lower_emission_intensity.png", width=7,height=5)
+
+```
+
+```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
 5 for more details).
@@
 3a.
-Using this threshold for predicting changes in emissions intensity based on milk yield data would likely lead to a lot of inaccurate classifications.*"}
+Using this threshold for predicting changes in emission intensity based on milk yield data would likely lead to a lot of inaccurate classifications.*"}
 
@@
 
-ggsave("figures/Fig_S2_probability_of_lower_emissions_intensity.png", width=7,height=5)
+ggsave("figures/Fig_S2_probability_of_lower_emission_intensity.png", width=7,height=5)
 
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S4.*** *Intervention impact on per head milk production and greenhouse gas emissions intensity.*"}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S4.*** *Intervention impact on per head milk production and greenhouse gas emission intensity.*"}
 
@@
 interventions[,"milk_change"]<-interventions$pc_milk_yield-interventions$pc_milk_baseline
-interventions[,"emissions_intensity_change"]<-interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline
+interventions[,"emission_intensity_change"]<-interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline
 interventions[,"milk_change_relative"]<-(interventions$pc_milk_yield-interventions$pc_milk_baseline)/interventions$pc_milk_baseline*100
-interventions[,"emissions_intensity_change_relative"]<-
+interventions[,"emission_intensity_change_relative"]<-
   (interventions$CO2eq_per_milk-interventions$CO2_intensity_baseline)/interventions$CO2_intensity_baseline*100
@@
 
-intervention_change <- merge(data.frame(aggregate(interventions_normal_milk$emissions_intensity_change,
+intervention_change <- merge(data.frame(aggregate(interventions_normal_milk$emission_intensity_change,
                                                   by = list(interventions_normal_milk$intervention),
@@
 colnames(intervention_change) <-
-  c("intervention", "emissions_intensity_change", "milk_change")
+  c("intervention", "emission_intensity_change", "milk_change")
 
@@
 interventions_scatter<-ggplot(data=interventions_normal_milk,
-                        aes(x=milk_change,y=emissions_intensity_change)) +
+                        aes(x=milk_change,y=emission_intensity_change)) +
   geom_point(aes(color=intervention),size=0.9) +
@@
   ggtitle("Intervention impact on milk yield and emissions") + 
-  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
+  labs(y=expression(Change~'in'~emission~intensity~(kg~CO[2]*'-eq'~kg^-1)),
        x="Change in per head milk yield (kg)",
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per head milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per head milk production and relative greenhouse gas emission intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}
 
 interventions_scatter_relative<-ggplot(data=interventions_normal_milk,
-                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
+                        aes(x=milk_change_relative,y=emission_intensity_change_relative)) +
   geom_point(aes(color=intervention),size=0.9) +
@@
   ggtitle("Intervention impact on milk yield and emissions") + 
-  labs(y="Relative change in emissions intensity (%)",
+  labs(y="Relative change in emission intensity (%)",
        x="Relative change in per head milk yield (%)",
@@
 
-The herd structure interventions showed potential for causing major changes in milk yields and emissions intensity on particular farms.
+The herd structure interventions showed potential for causing major changes in milk yields and emission intensity on particular farms.
 Plotting these impacts against the initial sets of relevant herd attributes serves to illustrate which farms are particularly responsive to these interventions (Fig.
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S5.*** *Influence of the breed composition of farm-scale dairy cow populations on changes in milk yield and emissions intensity of milk production in response to a breed change intervention*."}
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S5.*** *Influence of the breed composition of farm-scale dairy cow populations on changes in milk yield and emission intensity of milk production in response to a breed change intervention*."}
 
@@
                                               "milk_change_relative",
-                                              "emissions_intensity_change_relative",
+                                              "emission_intensity_change_relative",
                                               paste0("breed_",1:8),
@@
                              "milk_change_relative",
-                             "emissions_intensity_change_relative",
+                             "emission_intensity_change_relative",
                              "unproductive_bulls",
@@
 
-metrics <- c("Milk yield per head", "Emissions intensity of milk production")
-names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")
+metrics <- c("Milk yield per head", "Emission intensity of milk production")
+names(metrics) <- c("milk_change_relative","emission_intensity_change_relative")
 
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S6.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S6.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emission intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}
 
@@
                                               "milk_change_relative",
-                                              "emissions_intensity_change_relative",
+                                              "emission_intensity_change_relative",
                                               "unproductive_bulls",
@@
                              "milk_change_relative",
-                             "emissions_intensity_change_relative"))
+                             "emission_intensity_change_relative"))
 
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S7.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emissions intensity of milk production in response to a 'fewer replacement males' intervention*."}
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S7.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emission intensity of milk production in response to a 'fewer replacement males' intervention*."}
 
@@
                                               "milk_change_relative",
-                                              "emissions_intensity_change_relative",
+                                              "emission_intensity_change_relative",
                                               "replacement_males",
@@
     round(quantile(intervention_structures[which(intervention_structures$intervention==ints[i]),
-                                           "emissions_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
+                                           "emission_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
                    na.rm=TRUE),1)}
@@
 </i></p>") %>%
-  pack_rows(index = c("Impact on milk yield per head (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
+  pack_rows(index = c("Impact on milk yield per head (%)" = 5, "Impact on emission intensity (%)" = 5), background="lightgray") %>%
 column_spec(2:6, width = "3cm")
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S8.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S8.*** *Intervention impacts on milk production per head and greenhouse gas emission intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
 interventions_end[,"milk_change"]<-interventions_end$pc_milk_yield-interventions_end$pc_milk_baseline
-interventions_end[,"emissions_intensity_change"]<-interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline
+interventions_end[,"emission_intensity_change"]<-interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline
 interventions_end[,"milk_change_relative"]<-(interventions_end$pc_milk_yield-interventions_end$pc_milk_baseline)/interventions_end$pc_milk_baseline*100
-interventions_end[,"emissions_intensity_change_relative"]<-
+interventions_end[,"emission_intensity_change_relative"]<-
   (interventions_end$CO2eq_per_milk-interventions_end$CO2_intensity_baseline)/interventions_end$CO2_intensity_baseline*100
@@
 normal_end[,"milk_change"]<-normal_end$pc_milk_yield-normal_end$pc_milk_baseline
-normal_end[,"emissions_intensity_change"]<-normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline
+normal_end[,"emission_intensity_change"]<-normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline
 normal_end[,"milk_change_relative"]<-(normal_end$pc_milk_yield-normal_end$pc_milk_baseline)/normal_end$pc_milk_baseline*100
-normal_end[,"emissions_intensity_change_relative"]<-
+normal_end[,"emission_intensity_change_relative"]<-
   (normal_end$CO2eq_per_milk-normal_end$CO2_intensity_baseline)/normal_end$CO2_intensity_baseline*100
@@
 interventions_scatter_end<-ggplot(data=normal_plot,
-                        aes(x=milk_change,y=emissions_intensity_change)) +
+                        aes(x=milk_change,y=emission_intensity_change)) +
   geom_point(color="gray",size=1) +
   geom_point(data=interventions_normal_milk_end,
-                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
+                        aes(x=milk_change,y=emission_intensity_change, color=intervention),size=0.8) +
   scale_color_manual(values= c("breed" = intervention_colors[1],
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per head and greenhouse gas emission intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
 interventions_scatter_relative_end<-ggplot(data=normal_plot,
-                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
+                        aes(x=milk_change_relative,y=emission_intensity_change_relative)) +
   geom_point(color="gray",size=1) +
   geom_point(data=interventions_normal_milk_end,
-                        aes(x=milk_change_relative,y=emissions_intensity_change_relative, color=intervention),size=0.8) +
+                        aes(x=milk_change_relative,y=emission_intensity_change_relative, color=intervention),size=0.8) +
   scale_color_manual(values= c("breed" = intervention_colors[1],
@@
     round(quantile(interventions_normal_milk_end[which(interventions_normal_milk_end$intervention==ints[i]),
-                                           "emissions_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
+                                           "emission_intensity_change_relative"],c(1,0.95,0.5,0.05,0),
                    na.rm=TRUE),1)}
@@
                                       "Strongest effect")]<-
-  round(quantile(normal_normal_milk_end[,"emissions_intensity_change_relative"],
+  round(quantile(normal_normal_milk_end[,"emission_intensity_change_relative"],
                  c(1,0.95,0.5,0.05,0),
@@
 At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
-  pack_rows(index = c("Impact on milk yield per head (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
+  pack_rows(index = c("Impact on milk yield per head (%)" = 6, "Impact on emission intensity (%)" = 6), background="lightgray") %>%
 column_spec(2:6, width = "3cm")
@@
 ```{r, fig.dim = c(8, 4), fig.cap="***Figure 8.
-*** *Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
+*** *Intervention impacts on milk yield and greenhouse gas emission intensity at an endline sampling point vs.
 the initial sampling time.
-The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity.*"}
+The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emission intensity.*"}
 
@@
   geom_contour_filled(aes(z=probability)) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  ggtitle("Probability of lower emission intensity on higher-productivity farm") + 
   labs(y=expression(Milk~yield~at~endline~(kg~head^{-1}~year^{-1})),
@@
 
-ggsave("figures/Fig_8_probability_of_lower_emissions_intensity_interventions.png", width=7,height=5)
-
-```
-
-```{r, fig.dim = c(4, 6), fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred.
+ggsave("figures/Fig_8_probability_of_lower_emission_intensity_interventions.png", width=7,height=5)
+
+```
+
+```{r, fig.dim = c(4, 6), fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred.
 Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
@@
 
-```{r, fig.dim = c(8, 5), fig.cap="***Figure 10.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.*"}
+```{r, fig.dim = c(8, 5), fig.cap="***Figure 10.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.*"}
 
@@
 
-The sensitivity analysis indicated that uncertainty about milk yields is a key driver of output variation, with unfavorable implications for using this metric as an indicator of emissions intensity.
+The sensitivity analysis indicated that uncertainty about milk yields is a key driver of output variation, with unfavorable implications for using this metric as an indicator of emission intensity.
 Published evidence suggests that, with some additional effort, this uncertainty can be reduced to a measurement error of approximately 15%.
-We can explore the impact of such measures on the suitability of milk yield per head as an indicator of emissions intensity.
+We can explore the impact of such measures on the suitability of milk yield per head as an indicator of emission intensity.
 
@@
 
-```{r, fig.dim = c(8, 4), fig.cap="***Figure S9.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels, under the assumption that errors in milk yield estimates have been reduced to 15%.
-Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emissions intensity.
+```{r, fig.dim = c(8, 4), fig.cap="***Figure S9.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels, under the assumption that errors in milk yield estimates have been reduced to 15%.
+Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emission intensity.
 This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}
@@
                                                       probability)) + geom_contour_filled() +
-  ggtitle("Probability of lower emissions intensity on farm B - impact on added precision for milk yield measurements") + 
+  ggtitle("Probability of lower emission intensity on farm B - impact on added precision for milk yield measurements") + 
   labs(y=expression(Milk~yield~of~farm~B~(kg~head^{-1}~year^{-1})),
        x=expression(Milk~yield~of~farm~A~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity on farm B") +
+       fill="Probability of \nlower emission \nintensity on farm B") +
   stat_contour(breaks = c(0.9),col="black") + theme_bw() + facet_wrap(~version) +
@@
 
-ggsave("figures/Fig_S9_updated_probability_of_lower_emissions_intensity.png", width=10,height=5)
+ggsave("figures/Fig_S9_updated_probability_of_lower_emission_intensity.png", width=10,height=5)
 
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure S10.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times.
+```{r, fig.dim = c(8, 8), fig.cap="***Figure S10.*** *Intervention impacts on milk production per head and greenhouse gas emission intensity, in comparison to random changes between two evaluation times.
 The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}
@@
 interventions_end_update[,"milk_change"]<-interventions_end_update$pc_milk_yield-interventions_end_update$pc_milk_baseline
-interventions_end_update[,"emissions_intensity_change"]<-interventions_end_update$CO2eq_per_milk-interventions_end_update$CO2_intensity_baseline
+interventions_end_update[,"emission_intensity_change"]<-interventions_end_update$CO2eq_per_milk-interventions_end_update$CO2_intensity_baseline
 interventions_end_update[,"milk_change_relative"]<-(interventions_end_update$pc_milk_yield-interventions_end_update$pc_milk_baseline)/interventions_end_update$pc_milk_baseline*100
-interventions_end_update[,"emissions_intensity_change_relative"]<-
+interventions_end_update[,"emission_intensity_change_relative"]<-
   (interventions_end_update$CO2eq_per_milk-interventions_end_update$CO2_intensity_baseline)/interventions_end_update$CO2_intensity_baseline*100
@@
 normal_end_update[,"milk_change"]<-normal_end_update$pc_milk_yield-normal_end_update$pc_milk_baseline
-normal_end_update[,"emissions_intensity_change"]<-normal_end_update$CO2eq_per_milk-normal_end_update$CO2_intensity_baseline
+normal_end_update[,"emission_intensity_change"]<-normal_end_update$CO2eq_per_milk-normal_end_update$CO2_intensity_baseline
 normal_end_update[,"milk_change_relative"]<-(normal_end_update$pc_milk_yield-normal_end_update$pc_milk_baseline)/normal_end_update$pc_milk_baseline*100
-normal_end_update[,"emissions_intensity_change_relative"]<-
+normal_end_update[,"emission_intensity_change_relative"]<-
   (normal_end_update$CO2eq_per_milk-normal_end_update$CO2_intensity_baseline)/normal_end_update$CO2_intensity_baseline*100
@@
 interventions_scatter_end_update<-ggplot(data=normal_stuff,
-                        aes(x=milk_change,y=emissions_intensity_change)) +
+                        aes(x=milk_change,y=emission_intensity_change)) +
   geom_point(color="gray",size=1) +
   geom_point(data=interventions_normal_stuff,
-                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
+                        aes(x=milk_change,y=emission_intensity_change, color=intervention),size=0.8) +
   scale_color_manual(values= c("breed" = intervention_colors[1],
@@
 ```{r, fig.dim = c(8, 8), fig.cap="***Figure S11.
-*** *Intervention impacts on milk yield and greenhouse gas emissions intensity at an endline sampling point vs.
+*** *Intervention impacts on milk yield and greenhouse gas emission intensity at an endline sampling point vs.
 the initial sampling time, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
-The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emissions intensity.*"}
+The colored background surface indicates the probability that increases in milk productivity are associated with decreasing emission intensity.*"}
 
@@
   geom_contour_filled(aes(z=probability)) +
-  ggtitle("Probability of lower emissions intensity on higher-productivity farm") + 
+  ggtitle("Probability of lower emission intensity on higher-productivity farm") + 
   labs(y=expression(Milk~yield~at~endline~(kg~head^{-1}~year^{-1})),
        x=expression(Milk~yield~at~baseline~(kg~head^{-1}~year^{-1})),
-       fill="Probability of \nlower emissions \nintensity") +
+       fill="Probability of \nlower emission \nintensity") +
   stat_contour(aes(z=probability),breaks = c(0.9),col="black") + theme_bw() +
@@
 
-ggsave("figures/Fig_S11_probability_of_lower_emissions_intensity_interventions.png", width=7,height=12)
-
-```
-
-```{r, fig.dim = c(8, 12), fig.cap="***Figure S12.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
+ggsave("figures/Fig_S11_probability_of_lower_emission_intensity_interventions.png", width=7,height=12)
+
+```
+
+```{r, fig.dim = c(8, 12), fig.cap="***Figure S12.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
 Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
@@
 
-```{r, fig.dim = c(8, 8), fig.cap="***Figure 11.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emissions intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
+```{r, fig.dim = c(8, 8), fig.cap="***Figure 11.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
 Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
@@
 
-```{r, fig.dim = c(12, 4), fig.cap="***Figure S13.*** *Strength of the relationship of Monte Carlo model input variables with the emissions intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
+```{r, fig.dim = c(12, 4), fig.cap="***Figure S13.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
 This analysis is based on a*"}

### 2022-06-23 Manuscript_figures.Rmd
first really complete draft

@@
 ---
-title: "Accounting for uncertainty in simulating greenhouse gas emissions from dairy cattle in Kenya"
+title: |
+  <center> Uncertainty limits the use of simple </center>
+  <center> indicators to evaluate agricultural emission reductions </center>
+  <center> – </center>
+  <center> evidence from Kenyan smallholder dairy production </center>
+  <br>
+  <center><i> - supplementary materials - </i></center>
+  <br>
+
 author: "Eike Luedeling, Cory Whitney, Andreas Wilkes, Todd Rosenstock"
@@
 
+# Introduction
+
+This document contains supplementary materials for the article "Uncertainty limits the use of simple indicators to evaluate carbon footprints – a case study of smallholder dairy production in Kenya".
+Besides supplementary illustrations that add nuance to the presentation of results provided in the main article, this document also includes reproducible code to run all analyses and produce all figures contained in the main article.
+These figures are therefore also repeated here.
+To improve legibility of the document, we have hidden all code.
+Clicking the gray `Code` buttons on the right will make the code appear (or disappear again).
+All input data is presented in scrollable tables, and they can also be downloaded by clicking the `Copy` (to clipboard), `CSV` or `Excel` buttons above the respective tables.
+
+The code below contains some calls of the functions `save_temperature_scenarios` and `load_temperature_scenarios`, which are contained in the `chillR` package and were originally designed to process temperature data.
+We use them here to save and load the results of time-consuming data processing steps that include the generation of random numbers.
+These steps serve two purposes: 1) to save time when producing this document (we suppressed evaluation of the respective code chunks); 2) to ensure that the simulation outputs are identical between the results presented in the main article and these supplementary materials.
+
 # Setting up
 
-Note that the code for all operations is hidden, so that it doesn't clutter the document. You can make it appear by clicking the code button on the right.
-
-We first need to install and load all required packages.
+We first need to install and load all required packages. We also need to set some options needed to produce this document, which is generated with the rmarkdown package.
 
@@
 ```{r, warning=FALSE, eval=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+
 devtools::install_github("CWWhitney/uncertainty")
@@
 library(DT)
-```
-
-## Load emission calculation function
+options(scipen=999)
+options(DT.options = list(scrollY="300px",scrollX="300px", pageLength = 100, autoWidth = TRUE))
+```
+
+# Implementing the emission calculation function
 
@@
 The document can be found on [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/2006gl/pdf/4_Volume4/V4_10_Ch10_Livestock.pdf).
-The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021], also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
+The chapter is part of the book *Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories* [@ipcc_good_2021], which is also on the [the IPCC website](https://www.ipcc-nggip.iges.or.jp/public/gp/english/).
 Many variables are based on the work done in *Variation in the carbon footprint of milk production on smallholder dairy farms in central Kenya* [@wilkes_variation_2020] and in *Methods and guidance to support MRV of livestock emissions* [@wilkes_methods_2019].
@@
 
-The equations provided in these documents, in particular in the 2006 guideline document [@dong_ipcc_2006] were converted into an R function called `ghg_emissions`, which is defined here (hit the `Code` button on the right).
+The equations provided in these documents, in particular in the 2006 guideline document [@dong_ipcc_2006] were converted into an R function called `ghg_emissions`, which is defined here.
 
@@
 * const - constants (upper and lower bounds are the same)
-* posnorm - positive normal distribution; this is a normal distribution that is truncated at 0.
+* posnorm - positive normal distribution; this is a normal distribution that is truncated at 0
+* tnorm_0_1 - a normal distribution that is truncated at 0 and 1; this is useful for drawing random numbers that represent probabilities (where values <0 and >1 are impossible).
+
+<br>
 
 ***Table S1*** *Model input parameters for the Monte Carlo simulation, including the variable names used in the greenhouse gas emissions equation, a descriptive label, lower and upper bounds of 90% confidence intervals for the values, and the shape of the probability distribution that is used to describe the data.
-The IPCC_default column shows values used for producing (spuriously) the precise scenario.
+The IPCC_default column shows values used for producing the (spuriously) precise scenario described below.
 Note that some values are overwritten when producing farm-scale scenarios.
-Data can be downloaded in csv or Excel format or copied to the clipboard by pressing the respective buttons.*
+Data can be downloaded in csv or Excel format or copied to the clipboard by hitting the respective buttons.*
 
@@
 
-input_parameters<-read.csv("data/livestock_ghg_input_table.csv")[,c(7,2,3,5,6,8)]
+input_table <- "data/livestock_ghg_input_table.csv"
+input_parameters<-read.csv(input_table)[,c(7,2,3,5,6,8)]
 input_parameters<-input_parameters[which(!is.na(input_parameters$lower)),]
@@
 The decisionSupport package [@R-decisionSupport], which we use to run the simulations, also accommodates other distribution types, but we did not consider these appropriate here.
-The positive normal distribution (and other distributions such as normal or lognormal distributions) are defined based on the 5% and 95% quantiles.
-Note that it isn't always possible to fit a positive normal distribution based on such specifications, especially if the 5% quantile is set so close to zero that the remaining 5% of the data do not **fit** in the space between zero and the lower bound.
-In such cases, decisionSupport [@R-decisionSupport] returns a warning message, but it still does its best to generate the requested distribution.
-It is worth being on the lookout for the respective warning messages and try to adjust the bounds of the distribution to **possible** values.
-
-```{r load_data, include=FALSE}
-
-input_table <- "data/livestock_ghg_input_table.csv"
-
-options(knitr.kable.NA = '', encoding = "Windows-1252")
- read.csv(input_table)[,c(1:2,4,7,5:6)]   %>%
-    kableExtra::kbl(digits = 2)  %>%
-    kableExtra::kable_classic_2(full_width = F) %>% 
- scroll_box(height = "300px")
-
-```
+The positive normal distribution (and other distributions such as normal or lognormal distributions) are defined based on their 5% and 95% quantiles.
+Note that it isn't always possible to fit a positive normal distribution, or a normal distribution truncated at 0 and 1, based on such specifications, especially if the 5% or 95% quantiles are set so close to the truncation points that the remaining 5% of the data do not *fit* in the space between these points and the specified lower or upper bound.
+Strictly speaking, a normal distribution always includes a certain chance of values beyond any truncation point, but it is often reasonable to exclude values in the low-probability tail ends of such a distribution).
+In cases where decisionSupport [@R-decisionSupport] struggles to fit the specified distribution, it returns a warning message, but it still does its best to generate the requested distribution.
+It is worth being on the lookout for the respective warning messages and try to adjust the bounds of the distributions to *possible* values.
 
@@
 With the model equation and the input table defined above, we can run a Monte Carlo simulation that returns a probability distribution for the greenhouse gas emissions resulting from the particular herd that is specified in the input table.
-We use the functions provided by the decisionSupport package [@R-decisionSupport] in the R programming language [@R-base].
+We use the functions provided by the decisionSupport package [@R-decisionSupport] for the R programming language [@R-base].
 
@@
 
-The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019].
+The information for the household is extracted from a spreadsheet that contains all the survey responses [@wilkes_variation_2020; @wilkes_methods_2019], which is provided in Table S2.
+
+<br>
+
+***Table S2*** *Results of the farm survey by Wilkes and colleagues [@wilkes_variation_2020; @wilkes_methods_2019], which were used to develop farm-scale scenarios for probabilistic simulations.
+Data can be downloaded in csv or Excel format or copied to the clipboard by hitting the respective buttons.*
+
+```{r}
+# import the data from the survey and ensure that the .csv encoding can be read by R
+Kenyaherds<-read.table("data/Kenya_baseline_individual_cow_wNotes.csv", sep = ",",
+                       fileEncoding = "UTF-8-BOM", header=TRUE)
+
+Kenyaherds %>%
+  datatable(extensions = 'Buttons',
+            options = list(dom = 'Blfrtip',
+                           buttons = c('copy', 'csv', 'excel'),
+                           lengthMenu = list(c(10,25,50,-1),
+                                             c(10,25,50,"All")),
+                           list(scrollX = "300px"),
+                           autoWidth = TRUE),
+            rownames=FALSE)
+```
 
@@
 ```{r}
-# import the data from the survey and ensure that the .csv encoding can be read by R
-
-Kenyaherds<-read.table("data/Kenya_baseline_individual_cow_wNotes.csv", sep = ",",
-                       fileEncoding = "UTF-8-BOM", header=TRUE)
-
 # count the number of animals for each type
@@
 
-***Table S2*** *Model input parameters for producing farm-scale scenarios for the Monte Carlo simulation, listing all variables that need to be updated.
-Data can be downloaded in csv or Excel format or copied to the clipboard by pressing the respective buttons.*
+The probabilistic scenarios that were generated for each farm are presented in Table S3 below.
+
+<br>
+
+***Table S3*** *Model input parameters for producing probabilistic farm-scale scenarios for the Monte Carlo simulation, listing all variables that need to be updated.
+Data can be downloaded in csv or Excel format or copied to the clipboard by hitting the respective buttons.*
 
@@
 ```
-
-This is the resulting file (only the first 3 farms are shown, and only the first 20 rows):
 
@@
 
+The precise scenarios that were generated for each farm are presented in Table S4 below.
+
+<br>
+
+***Table S4*** *Model input parameters for producing precise farm-scale scenarios for the Monte Carlo simulation, listing all variables that need to be updated.
+All values in this scenario are constants, so that no randomness is introduced into the calculations.
+Data can be downloaded in csv or Excel format or copied to the clipboard by hitting the respective buttons.*
+
+```{r}
+
+precise_parameters<-read.csv("data/farm_scenarios_precise.csv")
+
+precise_parameters %>%
+  datatable(extensions = 'Buttons',
+            options = list(dom = 'Blfrtip',
+                           buttons = c('copy', 'csv', 'excel'),
+                           lengthMenu = list(c(10,25,50,-1),
+                                             c(10,25,50,"All")),
+                           list(scrollX = "300px")),
+            rownames=FALSE)
+```
+
 ## Running the simulation
@@
 
-Table S2 contains the scenario data derived from results of the farm survey.
-In running the Monte Carlo with the `scenario_mc` function, the initial values for all variables listed in this table (Table S1) are overwritten by those shown here (in Table S2) to generate farm-specific scenarios.
+Table S3 contains the scenario data derived from results of the farm survey.
+In running the Monte Carlo with the `scenario_mc` function, the initial values for all variables listed in this table (Table S1) are overwritten by those shown here (in Table S3) to generate farm-specific scenarios.
 
@@
 
-Now we look at the overall population.
+Figure 2 below shows total emissions and emission intensity for the overall population of all farms, with Figure S1 illustrating the distribution of emissions per head of cattle.
 
@@
                     paste0("CO","$_2$","-eq kg","$^{-1}$")),
-      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
+      caption = "<p style='color:black'><i><b>Table S5.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
   index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
@@
 
-Here's the relationship between these two variables, for farms with per-head milk yields of >500 kg per year.
+Figure 3 shows the relationship between these two variables, for farms with per-head milk yields of >500 kg per year.
 
@@
 
-Extracting the LOESS regression from the plot and empirically determining minimum milk yield that would be outside the confidence interval
-
-```{r, warning=FALSE}
+We'll add a cursory analysis below of the impact of simply using the LOESS regression line in Fig.
+3a, plus its confidence interval, for an assessment of emission intensity.
+To evaluate this, we need to extract the respective information from the plot.
+The result is shown in Fig.
+S2.
+
+```{r, warning=FALSE, message=FALSE}
 
@@
 According to these distributions, the emission intensity at a particular per head milk yield can be described by a slice through the probability surface.
+We can extract such slices using the `varslice_resample` function from the `uncertainty` package.
+Note that this function doesn't (yet) have the capability for truly random sampling of probabilities along this slice.
+Instead, it extracts probabilities at 1000 regularly spaced points along the slice (in this case 1000).
+Random samples are then drawn from the population of sampling values, with the extracted probabilities specifying the odds for the randomization.
+The empirical nature of this process does not compromise the analysis in a meaningful way, yet we mention this constraint here for full disclosure.
 
@@
 By resampling from these distributions, we can determine the likelihood that a value of one of these distributions is greater than one of the other.
-We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per head.
-
-```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels.
+This simply requires a lot of pairwise comparisons between the values of the two distributions, with the percentage of comparisons that result in a particular outcome (A greater, equal or less than B) indicating the respective probability.
+
+We can systematically expand this logic to include more comparisons that span the range of values contained in the dataset.
+In this case, we sample the overall probability distribution (Fig.
+3b) at intervals of 100 kg milk yield per head.
+The results can be plotted as a three-dimensional surface illustrating the relationship between two farm-scale estimates of milk yield per head and the probability with which the difference between the values allows judgement on one of the farms having a lower emission intensity than the other (Fig.
+5).
+
+```{r, fig.dim = c(8, 4), message=FALSE, fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels.
 Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emission intensity.
@@
 
-```{r, warning=FALSE, fig.dim = c(8, 4), fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
+For illustrative purposes, we can plot the confidence implied by using the LOESS regression shown in Fig.
+3a, including its 90% confidence interval, for deciding on whether emission reductions have occurred (Fig.
+S2).
+
+```{r, warning=FALSE, fig.dim = c(8, 4), message=FALSE,  fig.cap="***Figure S2.*** *Probability of a particular farm B featuring lower emission intensity of milk production than another farm A, based on observed per-head milk production levels (see caption to Fig.
 5 for more details).
@@
 
-Now a description of the 5 interventions
+The regression-based decision on which of two farms has lower emission intensity is located at an empirically derived probability level between approximately 50 and 60% This indicates that this classification threshold would constitute a strongly overconfident evaluation instrument, i.e.
+it would imply making inaccurate decisions on relative emissions between farms in a large number of cases.
+
+# Farm-scale interventions
+
+We simulated the impact of 5 farm-scale interventions:
+
+* Breed change - replacing dairy cows of low-productivity breeds with Holstein-Friesian cows
+* Retiring unproductive bulls - removing all bulls older than 36 months from the herds
+* Fewer replacement males - eliminating replacement males (i.e. uncastrated young males) on 23% of farms
+* Calliandra feed - supplementing animal feed with Calliandra leaves
+* Balanced diet - reducing the amount of concentrate fed to all cattle
+
+To support generation of the Breed change intervention, we compared milk yields across the array of dairy cow breeds that was encountered during the farm survey.
+Results of this comparison and the resulting intervention factors are shown in Fig.
+S3 and Tab.
+S6.
 
@@
 kable(breed_milk_summary, col.names = c("Breed", "Number of dairy cows", "Daily milk yield (kg)", "Intervention factor"),
-      caption = "<p style='color:black'><i><b>Table S4.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
+      caption = "<p style='color:black'><i><b>Table S6.</b> Mean milk yield across all dairy cow breeds encountered during a farm survey in Kenya</i></p>") %>%
 column_spec(2:4, width = "5cm")
@@
 
-\
+To simulate the impacts of these interventions, we first extracted the input table from the Monte Carlo simulation presented above.
+We then modified these inputs to simulate the changes that are expected on farms that adopt the proposed interventions.
 
@@
 
-To run the interventions, we need a function to run the model with a specific set of x values.
+Intervention impacts were simulated by running the greenhouse gas emissions equation with the inputs specified in these modified simulation input tables.
+To do this, we created a function (`run_model`) that runs the model with a specific set of x values, without drawing new random numbers.
 
@@
 ```
+
+We now use this functions to simulate intervention outcomes for all farms. These are saved and loaded again (see reasoning for this in Section 1 above).
 
@@
 
-```{r, include=FALSE}
+```{r}
 breed_res<-load_temperature_scenarios("results","Breed_intervention")
@@
 ```
+
+# Impacts of farm-scale interventions
+
+## Impacts compared to a counterfactual scenario
+
+We now evaluate the outputs of the intervention impact simulations.
+For this, we first make a table that contains the outcomes of all interventions, subtract from this the outcomes of the counterfactual (i.e.
+the original simulation, without the interventions) and illustrate the results for each intervention (Fig.
+S4).
 
@@
 
+Since the importance of these impacts is difficult to appreciate in this format, we also calculated the relative changes that resulted from the interventions (Fig. 6).
+
 ```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per head milk production and relative greenhouse gas emission intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}
@@
 
-The herd structure interventions showed potential for causing major changes in milk yields and emission intensity on particular farms.
-Plotting these impacts against the initial sets of relevant herd attributes serves to illustrate which farms are particularly responsive to these interventions (Fig.
-S3).
+## Characteristics of responsive farms
+
+Intervention impacts differed strongly between farms, particularly for the Breed change and herd structure (Retire unproductive males, and Fewer replacement males) interventions.
+We therefore evaluated the response of milk yield per head and emission intensity to particular farm characteristics that were targeted by the interventions.
+
+For the Breed Change intervention, the obvious farm characteristic of interest is the breed composition of the dairy cows in the herd (Fig. S5).
 
@@
 
+The Retire unproductive bulls intervention was obviously particularly effective on farms where a large percentage of animals fell within this animal type category (Fig. S6).
+
 ```{r, fig.dim = c(8, 4), fig.cap="***Figure S6.*** *Influence of the percentage of unproductive bulls in a cattle herd on changes in milk yield and emission intensity of milk production in response to a 'retiring unproductive bulls' intervention*."}
@@
 
+Success of the Fewer replacement males intervention depended primarily on the percentage of replacement males in the herd (Fig. S7).
+
 ```{r, fig.dim = c(8, 4), fig.cap="***Figure S7.*** *Influence of the percentage of replacement males in a cattle herd on changes in milk yield and emission intensity of milk production in response to a 'fewer replacement males' intervention*."}
@@
 
-Summary statistics of interventions for the counterfactual simulation
+Table S7 shows a summary of the impact of interventions on the milk yield per head and emission intensity across the whole population of farms, compared to the counterfactual simulation.
 
@@
          #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
-      caption = "<p style='color:black'><i><b>Table S5.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
+      caption = "<p style='color:black'><i><b>Table S7.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
 Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.
@@
 
-\
-
-These are noticeable changes, but can they be detected against the random variation that all farms experience?
-We'll now take another Monte Carlo sample, apply the same interventions to the same farms and evaluate the results again.
+## Intervention impacts in the face of random variation
+
+Comparison of intervention impacts against a counterfactual scenario indicated that all interventions may generate noticeable changes.
+In a more realistic setting that is affected by random effects, however, impacts may be less clearly discernible.
+We evaluate such a setting by running another Monte Carlo simulation and applying the same interventions to the same farms.
+The code for this is largely identical to what was already shown above.
 
@@
 
-```{r, include=FALSE}
+```{r}
 
@@
 
-```{r, include=FALSE}
+```{r}
 breed_end<-load_temperature_scenarios("results","End_Breed_intervention")
@@
 ```
+
+As shown in Figure S8, and in relative terms in Figure 7, intervention impacts are much less clearly visible when both scenarios used in the comparison are subjected to random noise.
+The illustrations show considerable overlap between the point cloud representing non-adopters (gray) and those representing adopters of the various interventions.
 
@@
 
-Summary statistics of interventions for the realistic-endline simulation
+The summary statistics of the intervention impacts still indicate that interventions had certain impacts, but these impacts are now much more difficult to see.
+In the case of the feed interventions, outcomes for the innovation-adopting farms are difficult to distinguish from those of non-adopting farms (Table S8).
 
@@
 kable(intervention_end_summary, booktabs = TRUE,digits=0,
-      caption = "<p style='color:black'><i><b>Table S6.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
+      caption = "<p style='color:black'><i><b>Table S8.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya.
 Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures.
@@
 
-\
-
-```{r, fig.dim = c(8, 4), fig.cap="***Figure 8.
+# Detectability of intervention impacts
+
+## Probability of correctly associating milk yield changes with emission intensity changes
+
+The difficulty of detecting intervention impacts against the backdrop of real-world variability can be illustrated by plotting the differences between the outcomes of a baseline scenarios (Monte Carlo simulation without interventions) and an endline scenario (second Monte Carlo simulation with interventions) against the three-dimensional difference probability surface (Fig.
+8).
+The point cloud shows only the adopting farms Strikingly, most farms are located near the identity line, and far away from areas of the plot that would be associated with a high probability that emission reductions have occurred.
+
+```{r, fig.dim = c(8, 4), message=FALSE, warning=FALSE, fig.cap="***Figure 8.
 *** *Intervention impacts on milk yield and greenhouse gas emission intensity at an endline sampling point vs.
@@
 
-```{r, fig.dim = c(4, 6), fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred.
+## Classification errors
+
+The difficulty of confidently associating changes in milk yield per head with emission intensity changes implies that the use of such an indicator would lead to considerable classification errors.
+By extracting the respective probability from the probability surface for all farms, we can systematically evalute the size of Type I and Type II errors that would result from the choice of particular levels of *required confidence* in an emissions reduction scheme.
+
+```{r, fig.dim = c(4, 6), warning=FALSE, message=FALSE, fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred.
 Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
@@
 
-and now for the sensitivity analysis
+# Important uncertainties
+
+Among all the uncertain variables used as inputs to the emissions model, some are likely to be more influential than others.
+We evaluated the impact of variation in each input variable on variation of the emission intensity of milk production.
+To achieve this, we implemented a Partial Least Squares regression (Fig.
+10).
 
@@
 
+This analysis revealed that milk yields of the various dairy cow types, the digestible energy content of animal feed and several other variables showed important correlations with the outcome metric.
+
 # Narrowing key uncertainties
@@
 Published evidence suggests that, with some additional effort, this uncertainty can be reduced to a measurement error of approximately 15%.
-We can explore the impact of such measures on the suitability of milk yield per head as an indicator of emission intensity.
+We now explore the impact of such measures on the suitability of milk yield per head as an indicator of emission intensity.
+
+We'll first look a the implications of this added precision on milk yields on the change-detection probability surface (Fig. S9).
 
@@
 
-We need to implement the same interventions as for the initial simulation.
+To evaluate the impact of added precision on milk yields on intervention outcomes, we need to implement the same interventions as for the initial simulation.
 
@@
 
-```{r, include=FALSE}
+```{r}
 breed_end_update<-load_temperature_scenarios("results","End_update_Breed_intervention")
@@
 
+We can now compare the implications of added precision on the farm's location on the change-detection probability surface (Fig. S11).
+
 ```{r, fig.dim = c(8, 8), fig.cap="***Figure S11.
@@
 ```
+
+The distribution has narrowed somewhat compared to the original simulation.
+We can now evaluate the implications of this on our ability to select a required confidence level for distinguishing adopters from non-adopters (Fig.
+S12).
 
@@
 
-Comparing the errors is easier in a line plot.
+Comparing the errors is easier in a line graph that directly convenes all the error probabilities in one plot.
 
@@
 
-new sensitivity analysis
-
-```{r, fig.dim = c(12, 4), fig.cap="***Figure S13.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.
-This analysis is based on a*"}
+Even though the error probabilities have dropped somewhat, we still see a considerable risk of wrong classifications.
+Another PLS regression analysis provides insights on what next precision-enhancement steps would be useful in gaining clarity which farms have taken steps to reduce their emission intensity.
+
+```{r, fig.dim = c(10, 10), fig.cap="***Figure S13.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression, for the initial uncertainty scenario (a) and a scenario where error estimates for milk yield measurement have been reduced to 15% (b).
+This analysis is based on the outputs of Monte Carlo simulations with 4140 runs, with scenarios based on farm characteristics oberved on smallholder livestock farms in Kenya.*"}
 
@@
 legend_table<-read.csv("data/legend.csv", fileEncoding = "UTF-8-BOM")
-PLS_initial<-plot_pls(pls_out,input_table=legend_table) + ggtitle("Initial uncertainty")
+PLS_initial<-plot_pls(pls_out,input_table=legend_table) + ggtitle("a) Initial uncertainty")
 
@@
 
-PLS_update<-plot_pls(pls_out_update,input_table=legend_table) + ggtitle("Precise milk yields")
-
-PLS_initial + PLS_update
-
-ggsave("figures/Fig_S13_PLS_dairy_households_update.png", width=12,height=4)
+PLS_update<-plot_pls(pls_out_update,input_table=legend_table) + ggtitle("b) Precise milk yields")
+
+PLS_initial + PLS_update + plot_layout(ncol = 1, guides='collect') &
+  theme(legend.position='bottom')
+
+ggsave("figures/Fig_S13_PLS_dairy_households_update.png", width=10,height=10)
 ```

### 2022-06-27 Luedeling_et_al_2022_supplementary_materials.Rmd
new html doc - supplement

--- 2022-06-23 Manuscript_figures.Rmd
@@
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
- geom_density(alpha=.2, fill="#FF6666") +
+ geom_density(alpha=.5, fill="#FF6666") +
   ylab("Relative probability") +
@@
 two_slices<-ggplot(data=slices, aes(Values, fill=Level)) +
- geom_density(alpha=.2) +
+ geom_density(alpha=.5) +
   ylab("Relative probability") +
@@
 ```{r, fig.dim = c(4, 6), warning=FALSE, message=FALSE, fig.cap="***Figure 9.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred.
-Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
+Type I errors indicate the probability of recognizing households for reducing emissions, even though they did not adopt an innovation, whereas the Type II error represents the percentage of households that made changes, but are not recognized for these.*"}
 
@@
 for(i in 1:nrow(confidence_level_analysis))
-{# compute Type 1 errors for all interventions
+{# compute Type 2 errors for all interventions
   for(ints in c("breed", "nobull", "lessReplace", "Calliandra", "balance"))
@@
 
-  confidence_level_analysis$Type2error[i]<-
+  confidence_level_analysis$Type1error[i]<-
     length(which(int_none_plot$intervention=="none" &
@@
 confidence_plot<-melt(confidence_level_analysis,id.vars = "confidence")
-confidence_plot$variable <- factor(confidence_plot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
-                  labels = c("Type I error - failure to reward 'Breed change' adopters",
-                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
-                             "Type I error - failure to reward 'Fewer replacement males' adopters",
-                             "Type I error - failure to reward 'Calliandra feed' adopters",
-                             "Type I error - failure to reward 'Balanced diet' adopters",
-                             "Type II error - rewarding non-adopters"))
+confidence_plot$variable <- factor(confidence_plot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type1error"),
+                  labels = c("Type II error - failure to reward 'Breed change' adopters",
+                             "Type II error - failure to reward 'Retire unproductive bulls' adopters",
+                             "Type II error - failure to reward 'Fewer replacement males' adopters",
+                             "Type II error - failure to reward 'Calliandra feed' adopters",
+                             "Type II error - failure to reward 'Balanced diet' adopters",
+                             "Type I error - rewarding non-adopters"))
 
@@
 ```{r, fig.dim = c(8, 12), fig.cap="***Figure S12.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
-Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
+Type I errors indicate the probability of recognizing households for reducing emissions, even though they did not adopt an innovation, whereas the Type II error represents the percentage of households that made changes, but are not recognized for these.*"}
 
@@
 for(i in 1:nrow(confidence_level_analysis_update))
-{# compute Type 1 errors for all interventions
+{# compute Type 2 errors for all interventions
   for(ints in c("breed", "nobull", "lessReplace", "Calliandra", "balance"))
@@
 
-  confidence_level_analysis_update$Type2error[i]<-
+  confidence_level_analysis_update$Type1error[i]<-
     length(which(int_none_plot_update$intervention=="none" &
@@
 confidence_plot_update<-melt(confidence_level_analysis_update,id.vars = "confidence")
-confidence_plot_update$variable <- factor(confidence_plot_update$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
-                  labels = c("Type I error - failure to reward 'Breed change' adopters",
-                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
-                             "Type I error - failure to reward 'Fewer replacement males' adopters",
-                             "Type I error - failure to reward 'Calliandra feed' adopters",
-                             "Type I error - failure to reward 'Balanced diet' adopters",
-                             "Type II error - rewarding non-adopters"))
+confidence_plot_update$variable <- factor(confidence_plot_update$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type1error"),
+                  labels = c("Type II error - failure to reward 'Breed change' adopters",
+                             "Type II error - failure to reward 'Retire unproductive bulls' adopters",
+                             "Type II error - failure to reward 'Fewer replacement males' adopters",
+                             "Type II error - failure to reward 'Calliandra feed' adopters",
+                             "Type II error - failure to reward 'Balanced diet' adopters",
+                             "Type I error - rewarding non-adopters"))
 
@@
 confidence_plot_stuff$variable <- factor(confidence_plot_stuff$variable,
-                                         levels = c("Type I error - failure to reward 'Breed change' adopters",
-                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
-                             "Type I error - failure to reward 'Fewer replacement males' adopters",
-                             "Type I error - failure to reward 'Calliandra feed' adopters",
-                             "Type I error - failure to reward 'Balanced diet' adopters",
-                             "Type II error - rewarding non-adopters"),
-                  labels = c("Type I - 'Breed change'",
-                             "Type I - 'Unproductive bulls'",
-                             "Type I - 'Replacement males'",
-                             "Type I - 'Calliandra feed'",
-                             "Type I - 'Balanced diet'",
-                             "Type II"))
+                                         levels = c("Type II error - failure to reward 'Breed change' adopters",
+                             "Type II error - failure to reward 'Retire unproductive bulls' adopters",
+                             "Type II error - failure to reward 'Fewer replacement males' adopters",
+                             "Type II error - failure to reward 'Calliandra feed' adopters",
+                             "Type II error - failure to reward 'Balanced diet' adopters",
+                             "Type I error - rewarding non-adopters"),
+                  labels = c("Type II - 'Breed change'",
+                             "Type II - 'Unproductive bulls'",
+                             "Type II - 'Replacement males'",
+                             "Type II - 'Calliandra feed'",
+                             "Type II - 'Balanced diet'",
+                             "Type I"))
 
@@
 ```{r, fig.dim = c(8, 8), fig.cap="***Figure 11.*** *Probability of Type I and Type II errors incurred by choosing specific levels of confidence that emission intensity changes have actually occurred, comparing initial uncertainty with improved measurements of milk yields (assuming a reduction of measurement errors to 15%).
-Type I errors indicate the probability of failing to recognize households that implemented interventions, whereas the Type II error represents the percentage of households that did not make any changes, but are nevertheless rewarded due to random effects.*"}
+Type I errors indicate the probability of recognizing households for reducing emissions, even though they did not adopt an innovation, whereas the Type II error represents the percentage of households that made changes, but are not recognized for these.*"}
 
@@
   facet_wrap(~version,nrow=2) +
-  scale_color_manual(values= c("Type I - 'Breed change'" = intervention_colors[1],
-                              "Type I - 'Unproductive bulls'" = intervention_colors[2],
-                              "Type I - 'Replacement males'" = intervention_colors[3],
-                              "Type I - 'Calliandra feed'" = intervention_colors[4],
-                              "Type I - 'Balanced diet'" = intervention_colors[6],
-                              "Type II" = "blue"),
-                    labels = c("Breed change", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet", "Type II error")) + theme_bw()
+  scale_color_manual(values= c("Type II - 'Breed change'" = intervention_colors[1],
+                              "Type II - 'Unproductive bulls'" = intervention_colors[2],
+                              "Type II - 'Replacement males'" = intervention_colors[3],
+                              "Type II - 'Calliandra feed'" = intervention_colors[4],
+                              "Type II - 'Balanced diet'" = intervention_colors[6],
+                              "Type I" = "blue"),
+                    labels = c("Breed change", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet", "Type I error")) + theme_bw()
 

### 2022-06-27 Luedeling_et_al_2022_supplementary_materials.Rmd
final commit - need to merge branches now

@@
 
-```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livestock (a) and emission intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms).
-Plausible estimates of both metrics are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
+```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure 2.*** *Total greenhouse gas emissions from livest3ck (a) and emission intensity of milk production (b) across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms, including 383 milk-producing farms).
+Plausible estimates of both metrics are generated through Monte Carlo simulation, with uncertainty introduced by specifying inputs as probability distributions.*"}
 
@@
 ```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per head (x-axis) and emission intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b).
-The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig.
-4b).*"}
+The vertical lines in (b) highlight slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^.
+These productivity levels are used to illustrate the slicing methodology in Fig.
+4b.*"}
 
@@
 confidence_plot<-melt(confidence_level_analysis,id.vars = "confidence")
-confidence_plot$variable <- factor(confidence_plot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type1error"),
-                  labels = c("Type II error - failure to reward 'Breed change' adopters",
-                             "Type II error - failure to reward 'Retire unproductive bulls' adopters",
-                             "Type II error - failure to reward 'Fewer replacement males' adopters",
-                             "Type II error - failure to reward 'Calliandra feed' adopters",
-                             "Type II error - failure to reward 'Balanced diet' adopters",
-                             "Type I error - rewarding non-adopters"))
-
-ggplot(confidence_plot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
+confidence_plot_fig<-confidence_plot
+confidence_plot_fig$variable <- factor(confidence_plot_fig$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type1error"),
+                  labels = c("Type II error - 'Breed change'",
+                             "Type II error - 'Unproductive bulls'",
+                             "Type II error - 'Replacement males'",
+                             "Type II error - 'Calliandra feed'",
+                             "Type II error - 'Balanced diet'",
+                             "Type I error"))
+
+ggplot(confidence_plot_fig, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
   labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")
@@
 
-```{r, fig.dim = c(8, 5), fig.cap="***Figure 10.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression.*"}
+```{r, fig.dim = c(8, 5), fig.cap="***Figure 10.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-in-the-Projection (VIP) score of a Partial Least Squares regression.
+We show only variables with VIP >0.8, a threshold that is often used for considering variables important.*"}
 
@@
 
-plot_pls(pls_out,input_table=legend_table)
+plot_pls(pls_out,input_table=legend_table, cut_off_line = 0.8)
 ggsave("figures/Fig_10_PLS_dairy_households.png", width=8,height=5)
@@
 confidence_plot_update<-melt(confidence_level_analysis_update,id.vars = "confidence")
-confidence_plot_update$variable <- factor(confidence_plot_update$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type1error"),
-                  labels = c("Type II error - failure to reward 'Breed change' adopters",
-                             "Type II error - failure to reward 'Retire unproductive bulls' adopters",
-                             "Type II error - failure to reward 'Fewer replacement males' adopters",
-                             "Type II error - failure to reward 'Calliandra feed' adopters",
-                             "Type II error - failure to reward 'Balanced diet' adopters",
-                             "Type I error - rewarding non-adopters"))
 
@@
 confidence_plot_stuff$variable <- factor(confidence_plot_stuff$variable,
-                                         levels = c("Type II error - failure to reward 'Breed change' adopters",
-                             "Type II error - failure to reward 'Retire unproductive bulls' adopters",
-                             "Type II error - failure to reward 'Fewer replacement males' adopters",
-                             "Type II error - failure to reward 'Calliandra feed' adopters",
-                             "Type II error - failure to reward 'Balanced diet' adopters",
-                             "Type I error - rewarding non-adopters"),
+                                         levels = c("breed",
+                                                    "nobull",
+                                                    "lessReplace",
+                                                    "Calliandra",
+                                                    "balance",
+                                                    "Type1error"),
                   labels = c("Type II - 'Breed change'",
@@
                               "Type I" = "blue"),
-                    labels = c("Breed change", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet", "Type I error")) + theme_bw()
+                    labels = c("Type II error - Breed change", "Type II error - No bulls", "Type II error - Fewer males", "Type II error - Calliandra feed", "Type II error - Balanced diet", "Type I error")) + theme_bw()
 
@@
 ```{r, fig.dim = c(10, 10), fig.cap="***Figure S13.*** *Strength of the relationship of Monte Carlo model input variables with the emission intensity of milk production, as indicated by the Variable-Importance-Score of a Partial Least Squares regression, for the initial uncertainty scenario (a) and a scenario where error estimates for milk yield measurement have been reduced to 15% (b).
-This analysis is based on the outputs of Monte Carlo simulations with 4140 runs, with scenarios based on farm characteristics oberved on smallholder livestock farms in Kenya.*"}
+This analysis is based on the outputs of Monte Carlo simulations with 4140 runs, with scenarios based on farm characteristics oberved on smallholder livestock farms in Kenya.
+We show only variables with VIP >0.8, a threshold that is often used for considering variables important.*"}
 
@@
 legend_table<-read.csv("data/legend.csv", fileEncoding = "UTF-8-BOM")
-PLS_initial<-plot_pls(pls_out,input_table=legend_table) + ggtitle("a) Initial uncertainty")
+PLS_initial<-plot_pls(pls_out,input_table=legend_table,cut_off_line = 0.8) + ggtitle("a) Initial uncertainty")
 
@@
 
-PLS_update<-plot_pls(pls_out_update,input_table=legend_table) + ggtitle("b) Precise milk yields")
+PLS_update<-plot_pls(pls_out_update,input_table=legend_table,cut_off_line = 0.8) + ggtitle("b) Precise milk yields")
 

### 2022-06-27 preliminary_versions/210527_modeling_thoughts.Rmd
cleaning up

---
title: "Model building thoughts"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# kernel modeling

Basically ready to be supplied with real data. Currently, the file chill-yield-vignette.Rmd contains the code. Text needs to be adjusted - reflects the chill-yield problem still.

# Model updating

## Problem description

The first version of the model assumed certain factors as constant and others as variable. We questioned this selection and decided to update (right?). The MC simulation can account for all kinds of uncertainties, and we should make full use of this capability. Having said this, we also want to stick closely to the *official* equation that Joshua's first model was based on.

The original input table for Joshua's model is contained in the livestock_ghg_input_table.csv file.

Todd has already made a more comprehensive list of all the possible variables (based on the official equation) - livestock-uncertainty-codebook.csv. We now need to decide which ones we want to assume constant, and which ones should be considered as uncertain variables.

## Tasks ahead

1. Decision which variables we want uncertainty for
2. Take a look at current rationale for deciding on precise values
3. Add uncertainty ranges to all these and clearly define as inputs to the model
4. Run simulation
5. Identify most important factors that determine GHG emissions (VIP analysis)
6. Based on outputs, produce a "chill-kernel" prediction model for GHG emissions

### 2022-06-27 preliminary_versions/SOC_uncertainty markdown file.Rmd
still cleaning up

---
title: "Uncertainty analysis for estimating soil organic carbon stocks"
author: "Joshua Wafula, Ermias Betemariam"
date: "10/22/2020"
output:
  html_document: default
  word_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# Call data manipulation functions
library(dplyr)
library(magrittr)
library(decisionSupport)

# create work station
soil<- read.csv("data.csv")
soil<- dplyr::select(soil,soc,depth,bdensity,cfragments)
input_table<- read.csv("MC simulation data.csv")
input_table<-dplyr::select(input_table,-median)
results_folder <- "MC results"
```

## Introduction

Uncertainties in soil organic carbon (SOC) stock assessments are rarely quantified even though they are critical in determining the significance of the results. Most studies generally focus on a single variable involved in the SOC stock calculation (SOC concentration, sampling depth, bulk density and rock fragment content) or on a single scale, rather than using an integrated approach (i.e. taking into account interactions between variables)(Goidts et al. 2009). We used the error propagation method to quantify the relative contribution of each variable and interaction involved to the final SOC stock variability. Monte Carlo simulations were used to cross-check the results. We used a data set for a study that was conducted around Serowe township in Botswana. Soil composite samples were collected from selected plot centers and at four different corners of each plot at 30 cm depth. The carbon content (by weight %) was determined using the Walkley-Black method (van Reeuwijk, 2002) in the soil laboratory of the Institute for Geo-information Science and Earth Observation (ITC) in the Netherlands. The air-dried soil samples were crushed and then passed through a 2 mm sieve prior to laboratory analysis. Bulk density was determined using the core sampling technique (Stone, 1991).

## Methods

Soil organic C stocks (Mg C ha-1) was calculated as: SOC stock = C/100* Bulk Density (BD)* 100, where SOC stock is given in Mg C ha-1; C is soil organic carbon concentration of fine soil (fraction < 2 mm) determined in the laboratory (g kg-1). BD is in g m-3 and 100 is used to convert the units to Mg C ha-1.

```{r, eval=FALSE}
# calculate soil carbon stocks (soc x depth x bdensity*(1-cfragments/100)
socstock=(soil$soc*soil$depth*soil$bdensity*(1-soil$cfragments/100))
socstock_mean<-mean(socstock)
```

### a) Error propagation

To propagate the errors from the variables used to calculate SOC stocks, we first determined the the absolute error for each of the variables. We did this by calculating the margins of error with a desired confidence level of 90% (Z score).

```{r, eval=FALSE}
# calculate absolute uncertainty. we have 75 data points
z_score<-1.65 # desired confidence level is 90% 
soc_abs_unc<-z_score*(sd(soil$soc)/sqrt(75))
dep_abs_unc<-z_score*(sd(soil$depth)/sqrt(75))
bdensity_abs_unc<-z_score*(sd(soil$bdensity)/sqrt(75))
cfr_abs_unc<-z_score*(sd(soil$cfragments)/sqrt(75))
```

We converted the absolute errors into relative errors (%) to enable propagation of the errors when multiplying/dividing variables.

```{r, eval=FALSE}
# calculate relative uncertainty
soc_rel_unc<-100*soc_abs_unc/soil$soc
dep_rel_unc<-100*dep_abs_unc/soil$depth
bdensity_rel_unc<-100*bdensity_abs_unc/soil$bdensity
cfr_rel_unc<-100*cfr_abs_unc/soil$cfragments
```

We then computed the square root of the sum of the square of each of the SOC stock input variables to get the propagated relative error (Palmer, 2003). We converted the resultant relative error to absolute error.

```{r, eval=FALSE}
# uncertainty propagation
socstock_rel_unc<-sqrt(soc_rel_unc^2+dep_rel_unc^2+bdensity_rel_unc^2+cfr_rel_unc^2)
socstock_abs_unc<-socstock*socstock_rel_unc/100
socstock_abs_unc_mean<-mean(socstock_abs_unc)
```

### b) Monte Carlo simulations using decisionSupport package

We used the `decisionSupport` function in the R package to run a Monte Carlo simulation of SOC stocks.The function requires two inputs:

1. An `input_table` (in .csv format) specifying the names and probability distributions for all variables used in the analysis. The variable distributions are described by a 90% confidence interval, which are specified by lower (5% quantile) and upper (95% quantile) bounds, as well as the shape of the distribution. `const` describe variables that are constant throughout, `norm` describe variables with normal distribution, `tnorm_0_1` describe variables with a truncated normal distribution that can only have values between 0 and 1 (useful for probabilities) and `posnorm` describe variables with normal distribution truncated at 0 (only positive values allowed). For this analysis, we used the calculated margins of error to estimate the lower and upper bounds for the mean values of SOC stocks input variables. This enabled us to generate the `input_table` below.

```{r, echo= FALSE}
knitr::kable(input_table, digits = 2, caption = "Monte Carlo input data")
```

2. An `R_function` that calculates SOC stocks based on the variables named in `input_table`. For this analysis, the function `soil_carbon_stocks` describing interactions between SOC stock variables (SOC concentration, sampling depth, bulk density and rock fragment content) was used.

```{r, eval=FALSE}
soil_carbon_stocks<-function(x, varnames){
# calculate soil carbon stocks (soc x depth x bdensity*(1-cfragments/100)  
carbon_stocks.1<-(sampling_depth*soil_organic_cab_conc.1*bulk_density.1*(1-prop_rock_frag_cont.1/100))
  return(carbon_stocks)
}
```

To run the simulation, the `soil_carbon_stocks` function, along with the data from the `input_table`, were fed into the Monte Carlo simulation (MC) function (Luedeling and Whitney, 2018) to conduct the full analysis. Below is the code we used to perform the Monte Carlo simulation with 10,000 model runs. We also applied Partial Least Squares (PLS) regression analysis to the MC simulation results to generate the Variable-Importance-in-the-Projection (VIP) statistic for input parameters (Luedeling & Gassner, 2012).The VIP statistic represents the direction and strength of each input variables relationship with the output variable (Wold et al. 2001). This helped us to identify the main sources of error when calculating SOC stocks.

```{r, echo = TRUE}
soil_carbon_stocks<-function(x, varnames){
make_variables<-function(est,n=1)
{ x<-random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
make_variables(estimate_read_csv(input_table))
decisionSupport(inputFilePath = input_table, #input file with estimates
                outputPath = results_folder, #output folder
                welfareFunction = soil_carbon_stocks, #the function created above
                numberOfModelRuns = 10000,
                functionSyntax = "plainNames", 
                write_table = TRUE,)}
```

## Results and Discusion

The error propagation method projected a mean SOC stock of 21.26 Mg C ha-1 with a computed margin of error of 3.63 Mg C ha-1. MC simulation approach had a mean of 22 Mg C ha-1 with 90% confidence that the actual value was between the range of 20 to 24 Mg C ha-1 (i.e about 2 Mg C ha-1 margin of error). This outcome points to an overestimation of about 8% when using the error propagation method as compared to MC simulation method. This trend has also been reported by Jones (1989) (overestimation range of between 1.8 to 16.9%). Despite this overestimation, the error propagation method is useful to identify the contribution of each term of various types of equations to the final variability observed on the result, provided that uncertainties in individual terms and interaction can be estimated. The main source of uncertainty in the SOC stock was the variability of SOC concentration (due to errors from the laboratory and to the high SOC spatial variability) (Fig.1). This outcome shows that When assessing SOC stock at the landscape scale, one should focus on the precision of SOC analyses from the laboratory.

![Figure 1: Important variables (determined by VIP analysis of PLS regression models) prejections, depicting the main sources of errors. The results were produced through MC simulation (10,000 model runs) of SOC stock calculations.](MC results/output_1_PLS_VIP.png)

## Conclusion

Uncertainties are difficult to quantify and to identify because they stem from complex interactions between the variables involved in SOC stocks (i.e. SOC concentration, bulk density, sampling depth and rock fragment content). However, the concept of error propagation can be directly applied to the mathematical expression of SOC stock, since the random error in the SOC stock comes from the propagation of the random errors in each variable used in the SOC stock equation. Given that these individual random errors and interaction are first estimated, their propagation can be assessed by different methods, including the Monte Carlo simulation method (MC), which generates numerous SOC stocks using random values for each individual variable (stochastic approach). The advantage of the error propagation method is to give an explicit equation for the final SOC stock variability (i.e. the solution comes in an analytical form), which allows one to quantify the relative contribution of the individual sources of uncertainties as well as their interaction (Heuvelink, 1998). However, when more complex functions than the SOC stock are considered, the MC method is easier to implement even if a large number of runs might be needed.

## References

Goidts, E., Van Wesemael, B., & Crucifix, M. (2009). Magnitude and sources of uncertainties in soil organic carbon (SOC) stock assessments at various scales. European Journal of Soil Science. https://doi.org/10.1111/j.1365-2389.2009.01157.x

Heuvelink, G.B.M. 1998. Uncertainty analysis in environmental modeling under a change of spatial scale. Nutrient Cycling in Agroecosystems, 50, 255-264.

Jones, L. 1989. Some results comparing Monte Carlo simulation and ???rst order Taylor series approximation for steady groundwater ???ow. Stochastic Hydrology & Hydraulics, 3, 179-190

Luedeling, E., Whitney, C.W., 2018. R package decisionSupport: Controlled burns in conifer forests. Retrieved from https://cran.r-project.org/web/packages/decisionSupport/vignettes/wildfire_example.html

Luedeling, E., Gassner, A., 2012. Partial Least Squares Regression for analyzing walnut phenology in California. Agric. For. Meteorol. https://doi.org/10.1016/j.agrformet.2011.10.020

van Reeuwijk, L.P., 2002. Procedures for Soil Analysis. Wageningen, The Netherlands

Palmer, M. (2003). Propagation of Uncertainty through Mathematical Operations. MIT Module.

Stone, J. A. (1991). Core sampling technique for bulk density and porosity determination on a clay loam soil. Soil and Tillage Research. https://doi.org/10.1016/0167-1987(91)90032-S

Wold, S., Sj??str??m, M., Eriksson, L., 2001. PLS-regression: A basic tool of chemometrics, in: Chemometrics and Intelligent Laboratory Systems. pp. 109130. https://doi.org/10.1016/S0169-7439(01)00155-1

### 2022-06-27 preliminary_versions/SOC_uncertainty.R

@@ -1,0 +1,68 @@
+# Uncertainty analysis for estimating soil organic carbon stocks
+
+# call data manipulation functions
+library(dplyr)
+library(magrittr)
+
+# create work station
+soil<- read.csv("C:/Users/jwafula/Dropbox/JFPE submission/GHG uncertainity/data/data.csv")
+soil<- dplyr::select(soil,soc,depth,bdensity,cfragments)
+
+# calculate soil carbon stocks (soc x depth x bdensity*(1-cfragments/100)
+socstock=(soil$soc*soil$depth*soil$bdensity*(1-soil$cfragments/100))
+socstock_mean<-mean(socstock)
+
+#a) Uncertainty analysis using error propagation method
+
+# calculate absolute uncertainty
+z_score<-1.65 # desired confidence level is 90% 
+soc_abs_unc<-z_score*(sd(soil$soc)/sqrt(75))
+dep_abs_unc<-z_score*(sd(soil$depth)/sqrt(75))
+bdensity_abs_unc<-z_score*(sd(soil$bdensity)/sqrt(75))
+cfr_abs_unc<-z_score*(sd(soil$cfragments)/sqrt(75))
+
+# calculate relative uncertainty
+soc_rel_unc<-100*soc_abs_unc/mean(soil$soc)
+dep_rel_unc<-100*dep_abs_unc/mean(soil$depth)
+bdensity_rel_unc<-100*bdensity_abs_unc/mean(soil$bdensity)
+cfr_rel_unc<-100*cfr_abs_unc/mean(soil$cfragments)
+
+# uncertainty propagation
+socstock_rel_unc<-sqrt(soc_rel_unc^2+dep_rel_unc^2+bdensity_rel_unc^2+cfr_rel_unc^2)
+socstock_abs_unc<-socstock_mean*socstock_rel_unc/100
+
+# (b) MC simulation approach using decisionSupport package
+
+library(decisionSupport)
+
+mean_soc<-mean(soil$soc)
+mean_depth<-mean(soil$depth)
+mean_bdensity<-mean(soil$bdensity)
+mean_cfragments<-mean(soil$cfragments)
+
+input_table <- "MC simulation data.csv"
+results_folder <- "MC results"
+
+# Generate a function for testing the model function 'line by line' 
+# by taking a single random sample of the provided estimates
+make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
[…]

### 2022-06-27 preliminary_versions/Uncertainty in Kenyan Livestock Emission Reductions.R

@@ -1,0 +1,689 @@
+# Estimating Uncertainty in Kenyan Livestock Emission Reductions
+
+# Load data
+ghg_variables<- read.csv("C:/Users/jwafula/ghg-uncertainty/Kenya_baseline_individual_cow_wNotes.csv")
+
+# (a) MC simulation approach using decisionSupport package
+
+# Calculate mean values for the variables to help generate the MC simulation table
+mean_Cfi<-mean(ghg_variables$Cfi)
+mean_LW<-mean(ghg_variables$live_weight_LW)
+mean_Ca<-mean(ghg_variables$Ca)
+mean_C<-mean(ghg_variables$C)
+mean_milk_yield<-mean(ghg_variables$milk_yield)
+mean_milk_fat<-mean(ghg_variables$milk_fat)
+mean_Cp<-mean(ghg_variables$Cp)
+mean_MW<-mean(ghg_variables$mature_weight_MW)
+mean_WG<-mean(ghg_variables$weight_gain_WG)
+mean_dig_energy<-mean(ghg_variables$dig_energy)
+mean_bo<-mean(ghg_variables$bo)
+mean_mms_pasture<-mean(ghg_variables$mms_pasture)
+mean_mms_spread<-mean(ghg_variables$mms_spread)
+mean_mms_drylot<-mean(ghg_variables$mms_drylot)
+mean_mms_solidstore<-mean(ghg_variables$mms_solidstore)
+mean_mms_composted<-mean(ghg_variables$mms_composted)
+mean_mms_liquid<-mean(ghg_variables$mms_liquid)
+mean_mms_biogas<-mean(ghg_variables$mms_biogas)
+mean_mms_burn<-mean(ghg_variables$mms_burn)
+mean_feed_kgCO2<-mean(ghg_variables$feed_kgCO2)
+mean_feed_transport_kgCO2<-mean(ghg_variables$feed_transport_kgCO2)
+mean_fpcm<-mean(ghg_variables$kg_FPCM_year)
+  
+# Calculate absolute uncertainty to help get upper and lower bounds of variables in the MC input table
[…]
